### 2012-01-31 paper_jasa.tex
first commit

%\documentclass{article}
\documentclass{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}

\graphicspath{{images/}}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.4in}
\setlength{\textheight}{9in}
\evensidemargin \oddsidemargin

\newtheorem{thm}{Theorem}[section]
\newtheorem{dfn}{Definition}[section]
\newtheorem{cor}{Corollary}[thm]
\newtheorem{con}{Conjecture}[thm]
%\setlength{\parindent}{0in}   % for no indent

%\topmargin -0.10in   % when making pdf
%\textheight 9.15in  % when making pdf

\begin{document}

% Article top matter
\title{Validation of Visual Statistical Inference, with Application to Linear Models}
\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
        Department of Statistics, Iowa State University}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract} 
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently there were no formal visual methods in place for determining statistical significance of findings. This changed, when \citet{buja:2009} conceptually introduced two protocols for formal tests of visual findings. In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against classical statistical testing  of the significance of regression model parameters. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results equivalent to the uniformly most powerful (UMP) test and for some scenarios yields better power than the UMP test.
\end {abstract}

%\begin{multicols}{2}
%\twocolumn

\section{Introduction} Any statistical analysis must include some statistical graphics. For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics. Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.

The scientific foundation of the graphical methods for data analysis is well established by \cite{cleveland:1984}. In recent years we have seen several major advances in statistical graphics. A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots. We have modern computing systems like R and SAS that facilitate easy production of high quality statistical plots. \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}. \citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.

In this paper we present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.

Section \ref{sec:visual_test} describes the basic ideas of visual inference, Section \ref{sec:regression} applies these ideas to the case of inference for regression analysis. In Section \ref{sec:simulation} we outline the setup for the human-subject study and  and present results.

%The paper concludes by describing the plans for future experiments.

\begin{table*}[hbtp]
\caption{Comparison of visual inference with existing inference technique}
\centering 
\begin{tabular}{llll} 
\hline
  & Mathematical Inference &  Visual Inference \\ %[0.5ex] % inserts table %heading 
\hline
  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
 & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.15}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\

 & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Null Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.15}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
 & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
\hline 
\end{tabular}
\label{tbl:compare}
\end{table*}

\section{Visual Statistical Inference} \label{sec:visual_test}

%This section gives the background to the lineup protocol described by \citet{buja:2009}.
This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference. Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.

Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$. 
% In hypothesis testing terminology, the parameter space $\Theta$ of a population parameter $\theta$, can be partitioned into $\Theta_0$ and $\Theta_0^c$. We test $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.

%\subsection{Visual Statistic}

Unlike classical hypothesis testing the statistic in visual inference is not a single value, but a plot that is appropriately chosen to describe the parameter of interest, $\theta$. When the alternative hypothesis is true, it is expected that the plot of the observed data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
%% As per the grammar of graphics presented by \cite{wilkinson:1999} and its application in R by \cite{hadley:2009}  it is possible to obtain some graphs associated with an arbitrary hypothesis.

%%\subsection{Lineup Plot} 
%
%%To compare whether the visible pattern in the test statistic is due to randomness or not we obtain the lineup plot defined as follows.
%
\begin{dfn}\label{dfn:lplot}
A lineup plot is a layout of $m$ visual statistics, consisting of 
\begin{itemize}\itemsep-3pt
\item $m-1$ plots simulated from the model specified by $H_0$  (null plots) and 
\item the test statistic produced by plotting the observed data, possibly arising from $H_1$.
\end{itemize}
\end{dfn}

If $H_1$ is true, the test statistic is expected to be the plot that is most different from the other plots in the lineup plot. A careful visual inspection should reveal the differences in the feature shown by the test statistic under null and alternative hypothesis. {\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.} The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.

%Table \ref{tbl:compare}, derived from \citet{buja:2009}, describes the procedure in comparison to traditional testing.

%\subsection{Type-I error and P-value}

Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$. Thus we have type-I error probability of $1/m$.

The lineup plot can be evaluated by one or more individuals. When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.

If $N$ individuals evaluate a lineup plot independently, we count the number of successful evaluations as $U \sim \text{Binom} (N,\frac{1}{m})$ and report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations. %Notice that when $N=1$, this $p$-value is $\frac1m$.

%\subsection{Power}

For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually. This should be reflected in the power of the test. We can assess power therefore both empirically based on experimental data and through theoretical considerations. Next, we will develop power theoretically and then %It is feasible to assess power theoretically and 
relate it to the empirical results.

\begin{dfn} \label{dfn:power}
For a lineup of $m$ plots the power of $\theta$ is defined as 
    \begin{equation*}
      \text{Power}(\theta)= 
        \begin{cases} 
              \text{Type-I error}=\frac1m & \text{if $\theta \in \Theta_0$,} \\
              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
        \end{cases}
    \end{equation*}
\end{dfn}

Power has a lower limit of $\frac1m$ since the probability that a person will randomly pick the test statistic under $H_0$ is $\frac1m$. In general,  $Pr(\text{Reject } H_0)$ is determined by the type of test being conducted. Theoretical power for the regression parameters is derived in the next section.

\section{Inference for a Regression Model} \label{sec:regression}

Consider a linear regression model 
\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.

Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0, k=1, ...,p$. If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the observed data, the resulting residual plot shows two groups of residuals. To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic. If $\beta_k\ne 0$ the boxplots show a vertical displacement. (Table \ref{tbl:stat_multiple} describes visual statistics for testing other hypotheses related to regression model \ref{multi}.)

A lineup including this test statistic is shown in Figure \ref{fig:test_category}. The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma^2})$. The test statistic, the plot containing the observed data, is randomly placed among these null plots. If the test statistic is identifiable the null hypothesis is rejected with a $p$-value of at most 0.05.

\begin{figure}[hbt]
%\begin{figurehere}
   \centering
       \scalebox{0.65}{\includegraphics{test_category.pdf}}
       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$. When the alternative hypothesis is true the observed plot should show a vertical displacement between box plots. Can you identify the observed plot?}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

\begin{table*}[hbtp] 
\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
\hline\hline 
Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
$H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
$H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
$H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
$H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
$H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
$H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
$H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
$H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

%\begin{table*}[hbtp] 
%	\caption{Test Statistics for Testing Hypothesis Related To The Model $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2+ \epsilon $ } 
%	\centering 
%		\begin{tabular}{m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
%			\hline\hline 
%				Null Hypothesis & Statistic Name & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
%			\hline 
%
%		\end{tabular} 
%	\label{tbl:stat_category} 
%\end{table*}

%\subsection{Expected power}

% Di: This is specific for the regression model
%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
Now consider estimating the power of the visual test. We have the estimate of the $\beta$ with a $p$-value $p_B$. The distribution of $p_B$ is a non-central $t$ distribution under $H_1$ and uniform under $H_0$. 
In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot. This leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as \begin{equation}\label{power_exp} Power(\beta)=Pr(p_{B} < p_0) \end{equation}

Figure \ref{fig:power_expected} shows the power of UMP test and expected power of visual test obtained from equation \eqref{power_exp}. Notice that the expected power of visual test is almost as good as the power of UMP test.

\begin{figure}[hbtp]
   \centering
       \scalebox{.45}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.}
       \label{fig:power_expected}
\end{figure}

% \pagebreak

We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}. Suppose, we have responses from $N$ independent observers with $u$ identifications of the observed plot. This gives an estimated power of
\begin{equation}\label{power_est} Power(\beta)=\frac{u}{N} \hspace{1cm} 0 \le u \le N \end{equation}

Suppose each of $N$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{ij}$. Let $Y_{ij} = 1$, if subject $j$ correctly identifies the test statistic on lineup $i$, and 0 otherwise.
We model $\pi_{ij} = E(Y_{ij})$ as  a mixed effects logistic regression

\begin{equation}
g(\pi_{ij}) = X_{ij}B  + Z_{ij} \tau_j 
\label{mixed} 
\end{equation}
%
where $\tau_j$ is random effects coefficient vector of length $q$ for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and link function $g(\mu)=log(\frac{\mu}{1-\mu}); 0 \le \mu \le 1$. The covariates could be demographic information of people such as age, gender, education level etc. as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.

From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
%Thus we have the overall probability of success ($\pi$ or power) 
for specified sample size ($n$) and  variance ($\sigma$)  as

\begin{equation}\label{eqn:power} 
Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
\end{equation}

\section{Simulation Experiment} \label{sec:simulation}

\subsection{Experiment with a discrete covariate}

The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \eqref{multi}). Data is simulated for different values of $\beta_2 (=0, 1, 3, 5, 7, 8, 10, 16)$, with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The set of $\beta_2$ values was chosen so that estimates of the power should produce reasonable power curves, comparable to the theoretical UMP test. We fixed the values of $\beta_0 = 5$,  $\beta_1=15$ and values for $X_1$ were generated as a random sample from a Poisson(30) distribution. Data sets with different combinations of $\beta_2$,  $n$ and $\sigma$ were generated with frequencies shown in Table \ref{tbl:experiment_params}. Three replicates of each level were generated. These produced 60 different ``observed data sets''.

%\begin{table}[hbtp] % test table
%\caption{Values of parameters considered for survey experiment} % title name of the table
%\centering
%  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
%    \hline
%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&\multicolumn{4}{c|}{Rank}&\multirow{2}{*}{Total}\\
%    \arrayrulecolor{kugray5}
%    \arrayrulecolor{black}
%    \cline{2-5}
%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&A&B&C&Other&\\
%    \hline
%    type 2&8&14&5&2&29\\
%    \hline
%  \end{tabular}
%\end{table}

\begin{table}[hbtp]
\caption{Values of parameters considered for three survey experiments} % title name of the table
\centering
% centering table
\begin{tabular}{c c l l l}
% creating 10 columns
\hline\hline

% inserting double-line
& & \multicolumn{3}{c}{values for slope ($\beta$)} \\
 \cline{3-5}

\raisebox{1.5ex}  {Sample size ($n$)} &   \raisebox{1.5ex} {$\sigma$} &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3}
\\ [0.5ex]
\hline
% inserts single-line

% Entering 1st row
&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
\raisebox{1.5ex}{100} &12
& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]

% Entering 2rd row
&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
\raisebox{1.5ex}{300} & 12
& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
% [1ex] adds vertical space
\hline
% inserts single-line
\end{tabular}
\label{tbl:experiment_params}
\end{table}

For added control, to ensure a signal in the simulated observed data a blocking structure was used to filter data sets. A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated. The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values. Additional control was applied to the 19 null plots. Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.

\subsection{Experiment with a continuous covariate}

This experiment is designed for the same regression model (Equation \eqref{multi}) with $p=1$ and covariate $X_1$ being continuous. Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ while generating observed data. For each combination of parameter, we generate at least three different observed data sets constituting a total of 65 lineups.

\subsection{Experiment with contaminated data}

The first two simulation experiments are based on normal error model which satisfies the conditions for uniformly most powerful (UMP) test. To study how the proposed visual inference work for a scenario where the data are not really coming from a normal model we design a contaminated data experiment. To generate the contaminated data set of size $n=100$ we used the following model

\[
  Y_i = \left\{
  \begin{array}{l l}
    5+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad \text{ $i$ =1,...,85}\\
    15+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=86,...,n\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. One such plot of the contaminated data for $\sigma$=5 and $\beta$=5 is shown in figure \ref{fig:cont_dat}. For this experiment we only consider sample size $n=100$ and error standard deviation $\sigma=5$.

For this experiment we generate five replications of observed data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.

\begin{figure}[hbt]
%\begin{figurehere}
   \centering
       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \eqref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
}
       \label{fig:cont_dat}
\end{figure}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
       \label{fig:power_observed}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
       \label{fig:power_diff_exp}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.}
       \label{fig:power_observed_exp3}
\end{figure*}

Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk. Each participant was shown a sequence of 10 lineup plots. Participants are asked to select the plot with the biggest vertical difference, give a reason for their choice, and determine a level of  confidence for their decision. Gender, age, education and geographic location of each participant are also collected.
 In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.  The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{power_est} along with 95\% confidence interval calculated using Fisher's exact method.

\begin{figure}[hbtp]
   \centering
       \scalebox{0.40}{\includegraphics{power_model.pdf}}
       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
       \label{fig:power_model}
\end{figure}

We fit model \eqref{mixed} to the survey data obtained from the simulation experiment. The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}. Model \ref{mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}. The plot includes 20 randomly selected subject-specific power curves. Notice that the power curve estimated for one subject is above the UMP test power curve.

\begin{figure}[hbtp]
   \centering
       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
       \label{fig:power_subject}
\end{figure}

\section{Conclusions}

The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods. We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to traditional inference.  The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where traditional tests cannot be used. For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots. Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.

\paragraph{Acknowledgement:}

This work was funded by National Science Foundation grant DMS 1007697.

%\bibliographystyle{plain}
%\bibliographystyle{plainnat}
\bibliographystyle{asa}
%\bibliographystyle{ieeetr}
\bibliography{references}

%\end{multicols}

\end{document}  %End of document.

\begin {abstract} 
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently there were no formal visual methods in place for determining statistical significance of findings. This changed, when Buja et al.(2009) conceptually introduced two protocols for formal tests of visual findings. In this paper we take this a step further by comparing the lineup protocol (Buja et al.2009) against classical statistical testing  of the significance of regression model parameters. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results equivalent to the uniformly most powerful (UMP) test and for some scenarios yields better power than the UMP test.
\end {abstract}

\end{document}

### 2012-02-01 paper_jasa.tex
first version of tex added by mahbub

@@
 
-The scientific foundation of the graphical methods for data analysis is well established by \cite{cleveland:1984}.
+The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
 In recent years we have seen several major advances in statistical graphics.
+Modern computing systems like R and SAS tease the production of high quality statistical plots.
 A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
-We have modern computing systems like R and SAS that facilitate easy production of high quality statistical plots.
 \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
+
 \citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
@@
 
-In this paper we present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
+In this paper we compare graphical inference to classical statistical inference for a set of regression type example.
+We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
 
@@
 When the alternative hypothesis is true, it is expected that the plot of the observed data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
-%% As per the grammar of graphics presented by \cite{wilkinson:1999} and its application in R by \cite{hadley:2009}  it is possible to obtain some graphs associated with an arbitrary hypothesis.
+We will call a plot with this property a {\it visual statistic} for $\theta$.
 
-%%\subsection{Lineup Plot} 
-%
-%%To compare whether the visible pattern in the test statistic is due to randomness or not we obtain the lineup plot defined as follows.
-%
 \begin{dfn}\label{dfn:lplot}
@@
 
-If $H_1$ is true, the test statistic is expected to be the plot that is most different from the other plots in the lineup plot.
-A careful visual inspection should reveal the differences in the feature shown by the test statistic under null and alternative hypothesis.
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.} The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.
-
-%Table \ref{tbl:compare}, derived from \citet{buja:2009}, describes the procedure in comparison to traditional testing.
-
-%\subsection{Type-I error and P-value}
+The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.
+If $H_1$ is true, the most different plot is expected to be the test statistic, i.e, the plot of the data.
+{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual insfection follows, and the observer is asked to point out the plot most different from the lineup.
 
@@
 
-If $N$ individuals evaluate a lineup plot independently, we count the number of successful evaluations as $U \sim \text{Binom} (N,\frac{1}{m})$ and report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
+If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
 %Notice that when $N=1$, this $p$-value is $\frac1m$.
@@
 
-Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0, k=1, ...,p$.
+Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
 If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the observed data, the resulting residual plot shows two groups of residuals.
@@
 If $\beta_k\ne 0$ the boxplots show a vertical displacement.
-(Table \ref{tbl:stat_multiple} describes visual statistics for testing other hypotheses related to regression model \ref{multi}.)
+
+In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model (e.g.
+Cook \& ...
+).
+Table \ref{tbl:stat_multiple} shows an overview of some commonly used charts and their corresponding hypothesis tests.
+Note that some charts can be associated with several different testing situations.
+In our paper, we pick the example fo situation 2 and 3 as candidates.
 
@@
 \centering 
-\begin{tabular}{m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
+\begin{tabular}{m{0.5cm}m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
 \hline\hline 
-Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
+Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
 \hline 
-$H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
+1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
 For lineup plot, we simulate data from fitted null model.
 \\
-$H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots.
+2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-$H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-$H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid.
+4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid.
 For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
-$H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
+5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
 For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-$H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
+6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
 For lineup plot, we simulate data from standard normal.
 \\
-$H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
+7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
 For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
 \\
-$H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data.
+8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data.
 For lineup plot, we simulate data from fitted model.
 \\[1ex] % [1ex] adds vertical space
-Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
+9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
 \hline 
@@
 Now consider estimating the power of the visual test.
-We have the estimate of the $\beta$ with a $p$-value $p_B$.
+For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$.
 The distribution of $p_B$ is a non-central $t$ distribution under $H_1$ and uniform under $H_0$.
@@
 
-Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot.
-This leads to the decision to reject $H_0$ when $p_{B} < p_0$.
-Thus we have the expected power as \begin{equation}\label{power_exp} Power(\beta)=Pr(p_{B} < p_0) \end{equation}
+Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot. This leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as
+
+\begin{equation}\label{power_exp} 
+   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for} \beta \ne 0
+\end{equation}
 
@@
 Notice that the expected power of visual test is almost as good as the power of UMP test.
+In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution.
+For infinitely many null plots we have the following theorem.
+
+\begin{thm}
+ For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$, where $F_{p_B}(.)$ is the distribution function of a non-central $t$ distribution. 
+ \end{thm}
+
+{ \em proof}
+
+Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
+\begin{eqnarray*}
+F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
+  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
+  &=& \left[ x^{m-1}\right]^1_{1-t}\\
+  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
+\end{eqnarray*}
+
+Again $p_0 \in [0,1]$ and hence $F{p_0}(t) =0$ for $t \le 0$ and  $F{p_0}(t) =1$ for $t \ge 1$. Thus expected power in equation \ref{power_exp} be
+\begin{eqnarray*}
+ Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
+  &=& 1- \int_{-\infty}^{\infty}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
+  &=& 1-  \int_{-\infty}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  = 1-  \int_{0}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  \\
+  & \rightarrow &1-  \int_{0}^{\infty}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
+\end{eqnarray*}
 

### 2012-02-02 paper_jasa.tex
updated reference
updated

@@
 Now consider estimating the power of the visual test.
-For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$.
-The distribution of $p_B$ is a non-central $t$ distribution under $H_1$ and uniform under $H_0$.
+For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$ and its distribution function be  $F_{p_B}(.)$ .
+As noted by \cite{Ruppert:2007} $F_{p_B}(.)$ is a right skewed distribution under $H_1$ and uniform under $H_0$.
 In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
@@
 \begin{equation}\label{power_exp} 
-   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for} \beta \ne 0
+   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for}  \quad \beta \ne 0
 \end{equation}
@@
 \begin{thm}
- For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$, where $F_{p_B}(.)$ is the distribution function of a non-central $t$ distribution. 
+ For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$.
 \end{thm}
@@
 
-Again $p_0 \in [0,1]$ and hence $F{p_0}(t) =0$ for $t \le 0$ and  $F{p_0}(t) =1$ for $t \ge 1$. Thus expected power in equation \ref{power_exp} be
+Thus expected power in equation \ref{power_exp} be
 \begin{eqnarray*}

### 2012-02-02 paper_jasa.tex
update

@@
  Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
-  &=& 1- \int_{-\infty}^{\infty}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
-  &=& 1-  \int_{-\infty}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  = 1-  \int_{0}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  \\
-  & \rightarrow &1-  \int_{0}^{\infty}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
+  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
+  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt }    \\
+  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
 \end{eqnarray*}

### 2012-02-02 paper_jasa.tex
added Tukey reference

@@
 \graphicspath{{images/}}
+
+\usepackage{color}
+\newcommand{\blue}[1]{{\color{blue} #1}}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 \section{Introduction} Any statistical analysis must include some statistical graphics.
-For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
+For exploratory data analysis \cite{tukey:eda}, statistical graphics play an invaluable role in model checking and diagnostics.
 Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.
@@
 In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS tease the production of high quality statistical plots.
+Modern computing systems like R and SAS ease the production of high quality statistical plots.
 A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
 \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
+\blue{Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
 
-\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
+\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing and quantification of discoveries made from statistical graphics.
 This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.

### 2012-02-02 paper_jasa.tex
couple more comments

@@
 
-The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.
+The $(m-1)$ null plots show samples drawn from the sampling distribution under $H_0$. %of the test statistic assuming that the null hypothesis is true. 
 If $H_1$ is true,  the most different plot is expected to be the test statistic, i.e, the plot of the data.
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual insfection follows, and the observer is asked to point out the plot most different from the lineup.
+A careful visual inspection follows, and the observer is asked to point out the plot most different from the lineup.
+{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}
 
-Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$. Thus we have type-I error probability of $1/m$.
+Since the lineup plot consists of $m$ plots, the probability of choosing any one of them at random is $1/m$. Thus we have a type-I error probability of $1/m$.
+
+When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
 
 The lineup plot can be evaluated by one or more individuals.
-When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
-
 If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
@@
 
+\blue{'successful' evaluation - that implies that we want to reject $H_0$. We need to phrase this more carefully. }
 %\subsection{Power}
 
-For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
+\blue{This gives us a tool to evaluate different types of graphics with respect to their efficiency:} for two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
 This should be reflected in the power of the test.

### 2012-02-02 paper_jasa.tex
couple more comments

@@
 
-\blue{This gives us a tool to evaluate different types of graphics with respect to their efficiency:} for two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
+\blue{This gives us a tool to evaluate different types of graphics with respect to their efficiency:} for two different visual test statistics of the same observed data, the one  is \blue{more efficient}, in which a specific pattern is more easily distinguishable visually.
 This should be reflected in the power of the test.

### 2012-02-02 paper_jasa.tex
up to theorem 3.1

@@
 For a lineup of $m$ plots the power of $\theta$ is defined as 
+\blue{shouldn't we define power in terms of the test statistic rather than the parameter?}
     \begin{equation*}
@@
 \end{equation}
-where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.
+where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates $X_j, j=1,..,p$, can be continuous or discrete.
 
 Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the observed data, the resulting residual plot shows two groups of residuals.
-To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic.
+If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the observed data, the resulting residual plot shows two groups of residuals \blue{with a difference in means reflecting the strength of the signal}.
+As a visual test statistic we use side-by-side boxplots of the residuals conditioned on the two levels of $X_k$.
 If $\beta_k\ne 0$ the boxplots show a vertical displacement.
-
-In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model (e.g.
-Cook \& ...
-).
-Table \ref{tbl:stat_multiple} shows an overview of some commonly used charts and their corresponding hypothesis tests.
-Note that some charts can be associated with several different testing situations.
-In our paper, we pick the example fo situation 2 and 3 as candidates.
 
@@
 If the test statistic is identifiable the null hypothesis is rejected with a $p$-value of at most 0.05.
+
+In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model (e.g.
+\citep{cook:99}).
+Table \ref{tbl:stat_multiple} shows an overview of some commonly used charts and their corresponding hypothesis tests.
+Note that some charts can be associated with several different testing situations.
+In our paper, we pick  situations 2 and 3 as candidates for further discussion.
 
@@
 As noted by \cite{Ruppert:2007} $F_{p_B}(.)$ is a right skewed distribution under $H_1$ and uniform under $H_0$.
+\blue{you were using that this is a non-central $t$ distribution before ...
+where did that go?}
+
  In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
-Suppose  $p_0=\min_i(p_{0,i})$.
-Thus $p_0 \sim \text{Beta}(1,m-1)$.
-
-Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot. This leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as
+Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot.
+This leads to the decision to reject $H_0$ when $p_{B} < p_0 := \min_{i} p_{0,i}$ with  $p_0 \sim \text{Beta}(1,m-1)$.
+Thus we have the expected power as
 

### 2012-02-02 paper_jasa.tex
up to theorem 3.1
some change

--- 2012-02-02 paper_jasa.tex
@@
 \graphicspath{{images/}}
-
-\usepackage{color}
-\newcommand{\blue}[1]{{\color{blue} #1}}
-\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 %\textheight 9.15in  % when making pdf
+
+%testing
 
@@
 \section{Introduction} Any statistical analysis must include some statistical graphics.
-For exploratory data analysis \cite{tukey:eda}, statistical graphics play an invaluable role in model checking and diagnostics.
+For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
 Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.
@@
 In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS ease the production of high quality statistical plots.
+Modern computing systems like R and SAS tease the production of high quality statistical plots.
 A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
 \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
-\blue{Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
 
-\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing and quantification of discoveries made from statistical graphics.
+\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
 This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
@@
 
-The $(m-1)$ null plots show samples drawn from the sampling distribution under $H_0$. %of the test statistic assuming that the null hypothesis is true. 
+The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.
 If $H_1$ is true, the most different plot is expected to be the test statistic, i.e, the plot of the data.
-A careful visual inspection follows, and the observer is asked to point out the plot most different from the lineup.
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}
+{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual insfection follows, and the observer is asked to point out the plot most different from the lineup.
 
-Since the lineup plot consists of $m$ plots, the probability of choosing any one of them at random is $1/m$. Thus we have a type-I error probability of $1/m$.
+Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$. Thus we have type-I error probability of $1/m$.
 
+The lineup plot can be evaluated by one or more individuals.
 When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
 
-The lineup plot can be evaluated by one or more individuals.
 If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
@@
 
-\blue{'successful' evaluation - that implies that we want to reject $H_0$. We need to phrase this more carefully. }
 %\subsection{Power}
 
-\blue{This gives us a tool to evaluate different types of graphics with respect to their efficiency:} for two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
+For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
 This should be reflected in the power of the test.

### 2012-02-03 paper_jasa.tex
distribution of p-value added

@@
 Now consider estimating the power of the visual test.
-For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$ and its distribution function be  $F_{p_B}(.)$ .
+Suppose $F_{|t|}(.)$ be the distribution function of an absolute value of $t$.
+For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$ and from definition we have
+
+$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow t_{obs}=F_{|t|}^{-1}(1-p_B)$$
+
+Now suppose $F_{|t|}(.,\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
+\begin{eqnarray*}\label{dist_p}
+F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p)\\
+  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) )\\
+  &=& 1-Pr(t_{obs} \le F_{|t|}^{-1}(1-p))\\
+  &=&\left\{ \begin{array}{ll}
+          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
+          1-F_{|t|}( F_{|t|}^{-1}(1-p),\delta) &\mbox{ ; under $H_1$} 
+       \end{array} \right.     
+\end{eqnarray*}
+
 As noted by \cite{Ruppert:2007} $F_{p_B}(.)$ is a right skewed distribution under $H_1$ and uniform under $H_0$.

### 2012-02-03 paper_jasa.tex
included m in proof to theorem 3.1

@@
 
-{ \em proof}
+\begin{proof}
 
@@
   &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt }    \\
-  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
+ &=& 1 - \int_{0}^{1}(t) f_{p_{B} }(t)dt + \int_{0}^{1} (1-t)^{m-1} f_{p_{B} }(t)dt \\
+ &=& E\left[(1-p_B)^{m-1}\right] \\
+%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
 \end{eqnarray*}
+\end{proof}
 

### 2012-02-03 paper_jasa.tex
some updates for \blue command
blue is back
Theorem 3.1 is meaningful now

--- 2012-02-03 paper_jasa.tex
@@
 \graphicspath{{images/}}
+\usepackage{color}
+\newcommand{\blue}[1]{{\color{blue} #1}}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
+\blue {Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
 
@@
 
-Now suppose $F_{|t|}(.,\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
-\begin{eqnarray*}\label{dist_p}
-F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p)\\
-  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) )\\
-  &=& 1-Pr(t_{obs} \le F_{|t|}^{-1}(1-p))\\
+Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
+\begin{eqnarray}\label{dist_p}
+F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
+  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
+  &=& 1-Pr(t_{obs} \le F_{|t|}^{-1}(1-p)) \nonumber \\
   &=&\left\{ \begin{array}{ll}
           1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-          1-F_{|t|}( F_{|t|}^{-1}(1-p),\delta) &\mbox{ ; under $H_1$} 
+          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
        \end{array} \right.     
-\end{eqnarray*}
+\end{eqnarray}
 
-As noted by \cite{Ruppert:2007} $F_{p_B}(.)$ is a right skewed distribution under $H_1$ and uniform under $H_0$.
+Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
+As noted by \cite{Ruppert:2007} the density $f_{p_B}(.)$ derived from equation \ref{dist_p}  is a right skewed distribution under $H_1$.
 In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
@@
 \begin{thm}
- For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$.
+ Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
 \end{thm}
 
-\begin{proof}
+{ \em proof}
 
@@
   &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
- &=& 1 - \int_{0}^{1}(t) f_{p_{B} }(t)dt + \int_{0}^{1} (1-t)^{m-1} f_{p_{B} }(t)dt \\
- &=& E\left[(1-p_B)^{m-1}\right] \\
+  &=& E(1-p_B)^{m-1}   \\
 %  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
 \end{eqnarray*}
-\end{proof}
 

### 2012-02-03 paper_jasa.tex
new update

@@
 %\textheight 9.15in  % when making pdf
-
-%testing
 
@@
 
-%The paper concludes by describing the plans for future experiments.
-
 \begin{table*}[hbtp]
@@
 
-%This section gives the background to the lineup protocol described by \citet{buja:2009}.
 This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
@@
 Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-% In hypothesis testing terminology, the parameter space $\Theta$ of a population parameter $\theta$, can be partitioned into $\Theta_0$ and $\Theta_0^c$.
-We test $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
 
@@
 We can assess power therefore both empirically based on experimental data and through theoretical considerations.
-Next, we will develop power theoretically and then %It is feasible to assess power theoretically and
-relate it to the empirical results.
+Next, we will develop power theoretically and then relate it to the empirical results.
 
@@
 
-%\begin{table*}[hbtp] 
-%	\caption{Test Statistics for Testing Hypothesis Related To The Model $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2+ \epsilon $ } 
-%	\centering 
-%		\begin{tabular}{m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
-%			\hline\hline 
-%				Null Hypothesis & Statistic Name & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
-%			\hline 
-%
-%		\end{tabular} 
-%	\label{tbl:stat_category} 
-%\end{table*}
-
 %\subsection{Expected power}
@@
 In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution.
-For infinitely many null plots we have the following theorem.
+Theorem \ref{thm:power} helps to compute the expected power of visual inference.
 
-\begin{thm}
+\begin{thm}\label{thm:power}
  Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
@@
 \end{eqnarray*}
+
+% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
 
@@
 These produced 60 different ``observed data sets''.
-
-%\begin{table}[hbtp] % test table
-%\caption{Values of parameters considered for survey experiment} % title name of the table
-%\centering
-%  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-%    \hline
-%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&\multicolumn{4}{c|}{Rank}&\multirow{2}{*}{Total}\\
-%    \arrayrulecolor{kugray5}
-%    \arrayrulecolor{black}
-%    \cline{2-5}
-%    \multicolumn{1}{|>{\columncolor{kugray5}}c|}{}&A&B&C&Other&\\
-%    \hline
-%    type 2&8&14&5&2&29\\
-%    \hline
-%  \end{tabular}
-%\end{table}
 

### 2012-02-03 paper_jasa.tex
Density of p-value added

@@
 Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007} the density $f_{p_B}(.)$ derived from equation \ref{dist_p}  is a right skewed distribution under $H_1$.
+As noted by \cite{Ruppert:2007} the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution under $H_1$.
 In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.

### 2012-02-05 paper_jasa.tex
density derivation updated
density derivation edited

@@
 Suppose $F_{|t|}(.)$ be the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$ and from definition we have
+For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
+By definition we have
 
-$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow t_{obs}=F_{|t|}^{-1}(1-p_B)$$
+$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
 
@@
   &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-  &=& 1-Pr(t_{obs} \le F_{|t|}^{-1}(1-p)) \nonumber \\
+  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
   &=&\left\{ \begin{array}{ll}

### 2012-02-09 paper_jasa.tex
tweaking plots

@@
    \centering
-       \scalebox{0.65}{\includegraphics{test_category.pdf}}
+       \scalebox{0.95}{\includegraphics{test_category.pdf}}
        \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$.
-When the alternative hypothesis is true the observed plot should show a vertical displacement between box plots.
-Can you identify the observed plot?}
+When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots.
+Which plot has the biggest vertical displacement between the boxplots?}
        \label{fig:test_category}
@@
    \centering
-       \scalebox{.45}{\includegraphics{power_expected.pdf}}
+       \scalebox{.65}{\includegraphics{power_expected.pdf}}
        \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.}

### 2012-02-11 paper_jasa.tex
new figure added

@@
 Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007} the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution under $H_1$.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
 In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
@@
    \centering
+       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+       \label{fig:power_loess1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
+       \label{fig:power_loess2}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
        \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}

### 2012-02-11 power.tex
first commit - some consideration on power, that should help with theorem 3.1

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Brief Article}
\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
%\maketitle
\section{Some Power Considerations}

Let us think of a lineup as a head-to-head comparison of the data plot and $m-1$ null plots. Let the data plot have a $p$ value of $p_B$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
We know that for each comparison the probability that the data plot has the smaller $p$ value is 
\begin{eqnarray*}
P(p_B < p_{0,i}) &=& 1 - P(p_{i,0} \le p_B) = 1- \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt = 1 - \int_0^1 t f_{p_B}(t) dt = 1 - E[p_B],
\end{eqnarray*}
which, in particular, is independent of $p_{0,i}$ for all $i$.

Let us now define $X$ as the number of null plots in a lineup, for which the $p$ value $p_{0,i}$ is smaller than $p_B$ (i.e. under the assumption that the onlooker is able to identify the chart with the smallest $p$ value, they would pick any of the $X$  null plots over the data plot). 
Then $X \sim B_{m-1, E[p_B]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
\[
P(X=0) = \left(1 - E[p_B] \right)^{m-1} = P(p_B \le p_0),
\]
where $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$

\section{Expected number of choices}
Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
\[
E[X] = (m-1)(1-E[p_B]).
\]
With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
This should allow us to infer some of the signal strength $p_B$ as
\[
E[p_B] = 1 - E[X]/(m-1),
\]
i.e. by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.

Now assume that the same lineup plot is evaluated by $K$ independent observers. Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$

For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
\[
n_m\sim \left \{ 
\begin{array}{cl}
B_{K, (1-E[p_B])^{m-1} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]

\end{document}

### 2012-02-11 power.tex
added figure for probability to pick dataplot versus size of lineup

@@
 \]
+Figure \ref{power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
+Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e.
+reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=2in]{images/powerplot.pdf} 
+   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$.
+The grey line shows probability of picking the data plot under $H_0$.}
+   \label{power}
+\end{figure}
 

### 2012-02-11 power.R
code for plots

@@ -1,0 +1,9 @@
+
+library(ggplot2)
+
+powerdf <- data.frame(expand.grid(m = 5:50,
+pb=c(seq(0.05,0.5, by=0.05), seq(0.25, 0.5, by=0.05))))
+
+powerdf$power <- with(powerdf, (1-pb)^(m-1))
+
+qplot(m, power, geom="line", group=pb, data=powerdf, colour=pb) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B]))

### 2012-02-11 power.R
first commit

@@ -7,3 +7,4 @@ pb=c(seq(0.05,0.5, by=0.05), seq(0.25, 0.5, by=0.05))))
 powerdf$power <- with(powerdf, (1-pb)^(m-1))
 
 qplot(m, power, geom="line", group=pb, data=powerdf, colour=pb) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B]))
+ggsave(file="powerplot.pdf")

### 2012-02-11 power.tex
added a few notes

@@
    \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$.
-The grey line shows probability of picking the data plot under $H_0$.}
+The grey line shows probability of picking the data plot under $H_0$.
+We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$.
+(Mahbub, is that what we see in practice?)}
    \label{power}

### 2012-02-12 paper_jasa.tex
Heike's writeup added with supporting figures

@@
 
+% -------- Heike's writeup added here
+
+\subsection{Some Power Considerations}
+
+Let us think of a lineup as a head-to-head comparison of the data plot and $m-1$ null plots.
+Let the data plot have a $p$ value of $p_B$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
+We know that for each comparison the probability that the data plot has the smaller $p$ value is 
+\begin{eqnarray*}
+P(p_B < p_{0,i}) &=& 1 - P(p_{i,0} \le p_B) = 1- \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
+&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt = 1 - \int_0^1 t f_{p_B}(t) dt = 1 - E[p_B],
+\end{eqnarray*}
+which, in particular, is independent of $p_{0,i}$ for all $i$.
+
+Let us now define $X$ as the number of null plots in a lineup, for which the $p$ value $p_{0,i}$ is smaller than $p_B$ (i.e.
+under the assumption that the onlooker is able to identify the chart with the smallest $p$ value, they would pick any of the $X$  null plots over the data plot).
+Then $X \sim B_{m-1, E[p_B]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
+\[
+P(X=0) = \left(1 - E[p_B] \right)^{m-1} = P(p_B \le p_0),
+\]
+where $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$
+
+\subsection{Expected number of choices}
+Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
+\[
+E[X] = (m-1)(1-E[p_B]).
+\]
+With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
+This should allow us to infer some of the signal strength $p_B$ as
+\[
+E[p_B] = 1 - E[X]/(m-1),
+\]
+i.e.
+by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
+
+Now assume that the same lineup plot is evaluated by $K$ independent observers.
+Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots.
+Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$
+
+For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
+\[
+n_m\sim \left \{ 
+\begin{array}{cl}
+B_{K, (1-E[p_B])^{m-1}} & \text { under } H_1\\
+B_{K, 1/m} & \text { under } H_0
+\end{array}
+\right .
+\]
+Figure \ref{fig:pval_power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
+Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e.
+reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
+   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$.
+The grey line shows probability of picking the data plot under $H_0$.
+We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$.
+(Mahbub, is that what we see in practice?) Yes, we see this pattern in our experiments as shown in figure \ref{fig:pval_pcorrect}}
+   \label{fig:pval_power}
+\end{figure}
+
+\subsection{Other methods to obtain power}
+
 We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
@@
    \centering
+       \scalebox{0.7}{\includegraphics{power_loess_exp3.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.}
+       \label{fig:power_loess3}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
        \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
@@
        \label{fig:power_observed_exp3}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
+       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
+After p-value exceed 1.5 it is very unlikely to identify the observed plot.
+The theoretical justification of this is shown in figure \ref{fig:pval_power}}
+       \label{fig:pval_pcorrect}
 \end{figure*}

### 2012-02-12 paper_jasa.tex
question for Di/Heike

@@
    \centering
-       \scalebox{0.7}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.}
+       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
+Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
+Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
        \label{fig:power_loess3}
@@
        \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.}
+       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+}
        \label{fig:power_observed_exp3}
@@
        \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
-After p-value exceed 1.5 it is very unlikely to identify the observed plot.
+After p-value exceeds 1.5 it is very unlikely to identify the observed plot.
 The theoretical justification of this is shown in figure \ref{fig:pval_power}}

### 2012-02-14 power.R
some updates

@@ -8,3 +8,19 @@ powerdf$power <- with(powerdf, (1-pb)^(m-1))
 
 qplot(m, power, geom="line", group=pb, data=powerdf, colour=pb) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B]))
 ggsave(file="powerplot.pdf")
+
+
+powerdf$EX <- with(powerdf, (m-1)*(1-pb))
+
+
+qplot(m, EX, geom="line", group=pb, data=powerdf, colour=pb) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B]))
+
+
+
+library(nullabor)
+n <- 100
+x <- rnorm(n)
+y <- rnorm(n) + 0.15*x
+summary(lm(y~x))
+df <- data.frame(x=x, y=y)
+qplot(x, y, data = df)  %+%  lineup(null_permute("y")) + facet_wrap(~ .sample) + geom_smooth(method="lm")

### 2012-02-14 power.tex
corrected typos
some updates

@@
 Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$
+Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$, and $(n_1, n_2, ..., n_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$.
+We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
 
@@
 \begin{array}{cl}
-B_{K, (1-E[p_B])^{m-1} & \text { under } H_1\\
+B_{K, (1-E[p_B])^{m-1}} & \text { under } H_1\\
 B_{K, 1/m} & \text { under } H_0
@@
 \end{figure}
-
+Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
+\[
+\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
+\]
+Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as
+\[
+\widehat{E[p_B]}  = 1 - 1/(m-1) \cdot \sum_i I (n_i > n_m),
+\]
+where $I(x) = 1$ for $x$ is true and $I(x) = 0$ if $x$ is false.
 \end{document}

### 2012-02-14 power.tex
strength of dataplot

@@
 \end{figure}
+Based on $K$ observers, we can assume that $n_m$ votes for the dataplot represent support for the data over any of the nullplots, i.e. $\pi_m = P(X=0)$.
+
 Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
@@
 \]
-Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as
-\[
-\widehat{E[p_B]}  = 1 - 1/(m-1) \cdot \sum_i I (n_i > n_m),
-\]
-where $I(x) = 1$ for $x$ is true and $I(x) = 0$ if $x$ is false.
+%Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as
+%\[
+%\widehat{E[p_B]}  = 1 - 1/(m-1) \cdot \sum_i I (n_i > n_m),
+%\]
+%where $I(x) = 1$ for $x$ is true and $I(x) = 0$ if $x$ is false.
 \end{document}

### 2012-02-19 power.tex
fixed the 1- bug
two methods of computing pB

@@
 Let us think of a lineup as a head-to-head comparison of the data plot and $m-1$ null plots.
-Let the data plot have a $p$ value of $p_B$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-We know that for each comparison the probability that the data plot has the smaller $p$ value is 
+Let the data plot have a $p$-value of $p_B$, and the null plots $p$-values of $p_{0, i}$ with $1 \le i \le m-1$.
+We know that for each comparison the probability that the data plot has the smaller $p$-value is 
 \begin{eqnarray*}
@@
 
-Let us now define $X$ as the number of null plots in a lineup, for which the $p$ value $p_{0,i}$ is smaller than $p_B$ (i.e.
-under the assumption that the onlooker is able to identify the chart with the smallest $p$ value, they would pick any of the $X$  null plots over the data plot).
+Let us now define $X$ as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_B$ (i.e.
+under the assumption that the onlooker is able to identify the chart with the smallest $p$-value, they would pick any of the $X$  null plots over the data plot).
 Then $X \sim B_{m-1, E[p_B]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
-\[
+\begin{equation}\label{p0}
 P(X=0) = \left(1 - E[p_B] \right)^{m-1} = P(p_B \le p_0),
-\]
+\end{equation}
 where $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$
@@
 \[
-E[X] = (m-1)(1-E[p_B]).
+E[X] = (m-1)E[p_B].
 \]
 With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-This should allow us to infer some of the signal strength $p_B$ as
+This allows us to infer some of the signal strength $p_B$ as
 \[
-E[p_B] = 1 - E[X]/(m-1),
+E[p_B] = E[X]/(m-1),
 \]
 i.e.
-by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
+by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer:
 
-Now assume that the same lineup plot is evaluated by $K$ independent observers.
+assume that the same lineup plot is evaluated by $K$ independent observers.
 Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots.
@@
 We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
+
+We can then find an estimate for $X$, the number of null plots with a stronger signal than the data plot, as the number of null plots with more votes $n_i$ than the dataplot $n_m$.
+This gives an estimate for $X$ as
+\[
+\hat{X} = \sum_{i=0}^{m-1} I(n_i \ge n_m),
+\]
+where $I(x) = 1$ if $x$ is true and 0 otherwise.
+
+However, this is not a very precise estimator for $E[p_B]$, since the only possible values for the  estimates are multiples of $1/(m-1)$.
+
+As an alternative, we will make use of the fact that $\pi_m = P(X=0)$, since a vote for the dataplot implies that none of the other plots have a stronger signal.
+
+With $\widehat{\pi_m} = n_m/K$ as an estimate for $\pi_m$, we get an plugin estimate for $E[p_B]$ from eq (\ref{p0}) as
+\[
+\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
+\]
+
+Since $n_m$ has a binomial distribution $n_m \sim B_{m-1, \pi_m}$, we can also derive a variance of the above estimate as
+\[
+Var \ \widehat{p_B} = K^{-2/(m-1)} Var \left[n_m^{1/(m-1)}\right]
+\]
 
@@
 The grey line shows probability of picking the data plot under $H_0$.
-We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$.
-(Mahbub, is that what we see in practice?)}
+We see that for a $p$-value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$.
+}
    \label{power}
 \end{figure}
-Based on $K$ observers, we can assume that $n_m$ votes for the dataplot represent support for the data over any of the nullplots, i.e. $\pi_m = P(X=0)$.
-
-Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
-\[
-\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
-\]
+%Based on $K$ observers, we can assume that $n_m$ votes for the dataplot represent support for the data over any of the nullplots, i.e. $\pi_m = P(X=0)$.
+%
+%Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
+%\[
+%\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
+%\]
 %Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as

### 2012-02-26 power.tex
some changes

@@
 Let the data plot have a $p$-value of $p_B$, and the null plots $p$-values of $p_{0, i}$ with $1 \le i \le m-1$.
-We know that for each comparison the probability that the data plot has the smaller $p$-value is 
+We know that for each comparison the probability that the null plot has the smaller $p$-value is 
 \begin{eqnarray*}
-P(p_B < p_{0,i}) &=& 1 - P(p_{i,0} \le p_B) = 1- \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt = 1 - \int_0^1 t f_{p_B}(t) dt = 1 - E[p_B],
+P(p_{i,0} < p_B) &=& \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
+&=& \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt =  \int_0^1 t f_{p_B}(t) dt =  E[p_B],
 \end{eqnarray*}
@@
 With $\widehat{\pi_m} = n_m/K$ as an estimate for $\pi_m$, we get an plugin estimate for $E[p_B]$ from eq (\ref{p0}) as
-\[
+\begin{equation}\label{signal}
 \widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
-\]
+\end{equation}
 
@@
 \]
+
+Figure \ref{fig:pval_plot_signal} shows a comparison of simulation based $p$-values resulting from model theory ($x$-axis) and graphical signal strength, defined in eq.
+(\ref{signal}) evaluated through experimentation.
+Up to about a $p$-value of 0.15 there is a linear dependency between the two measures.
+Beyond that, the graphical signal strength flattens out.
+This `flattening out' is due to a reduction of the signal to noise ratio and is driven by the size $m$ of the lineup.
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{images/p_val_plot_signal.pdf}}
+       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation (\ref{plot_signal}).}
+       \label{fig:pval_plot_signal}
+\end{figure*}
 
@@
 Figure \ref{power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
-Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e.
-reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.
 
@@
 \end{figure}
-%Based on $K$ observers, we can assume that $n_m$ votes for the dataplot represent support for the data over any of the nullplots, i.e. $\pi_m = P(X=0)$.
-%
-%Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
-%\[
-%\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
-%\]
-%Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as
-%\[
-%\widehat{E[p_B]}  = 1 - 1/(m-1) \cdot \sum_i I (n_i > n_m),
-%\]
-%where $I(x) = 1$ for $x$ is true and $I(x) = 0$ if $x$ is false.
+
+\section{The Flattening out}
+The flattening out in figure \ref{fig:pval_plot_signal} happens, when the signal in the data plot becomes too low to make it distinguishable from a nullplot. 
+Under $H_0$ the $p$-value of the data plot $p_B \sim U[0,1]$.
+If we can assume that an observer is able to pick out the plot corresponding to the minimal $p$-value, we have a distribution of $Beta_{1, m}$ for the picked plot.
+
 \end{document}

### 2012-03-04 paper_jasa.tex
plot signal strength added
trying to clean up and tighten the introduction

@@
 
-The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the the test statistic assuming that the null hypothesis is true.
-If $H_1$ is true, the most different plot is expected to be the test statistic, i.e, the plot of the data.
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual insfection follows, and the observer is asked to point out the plot most different from the lineup.
+The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the test statistic assuming that the null hypothesis is true.
+If $H_1$ is true, \blue{we expect this to be reflected as a feature in the test statistic, i.e. the plot of the data, that makes it visually distinguishable from the null plots.}
+{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual inspection follows, and the observer is asked to point out the plot most different from the lineup.
 
@@
 
-The lineup plot can be evaluated by one or more individuals.
+%The lineup plot can be evaluated by one or more individuals.
 When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
 
-If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
+%If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
 %Notice that when $N=1$, this $p$-value is $\frac1m$.
@@
 
-For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable visually.
+For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable to the observer.
 This should be reflected in the power of the test.
 We can assess power therefore both empirically based on experimental data and through theoretical considerations.
+\blue{
+
+\begin{dfn} \label{dfn:test}
+A \textit{visual test} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot used in the lineup.
+The grammar consists of type and specification of aesthetics necessary for complete reproducibility.
+\end{dfn}
+}
 Next, we will develop power theoretically and then relate it to the empirical results.
@@
 \begin{dfn} \label{dfn:power}
-For a lineup of $m$ plots the power of $\theta$ is defined as 
+\red{HH: I don't think power should be defined in terms of $\theta$ Usually power is considered a property of the test.
+In our case that's the specifics of the plot used for the lineup.
+We might need an additional definition for that - maybe we can phrase that within the framework of the grammar of graphics?
+- that's what I've tried above}
+
+For a lineup of $m$ plots the power of \blue{a visual test}
+%$\theta$
+ is defined as 
     \begin{equation*}
-      \text{Power}(\theta)= 
+      \text{Power}(T_\theta)= 
         \begin{cases} 
@@
 \end{dfn}
+\red{this is a non-standard definition of power - usually power is just the prob to reject under $H_1$. could we back the non-standard definition up with some literature?}
 
-Power has a lower limit of $\frac1m$ since the probability that a person will randomly pick the test statistic under $H_0$ is $\frac1m$.
-In general,  $Pr(\text{Reject } H_0)$ is determined by the type of test being conducted.
+\red{HH: let's try to keep the power discussion as general as possible, and only use regression as an example.}
+
+For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
+
+% HH: I think this is covered by making power a property of the test
+% In general,  $Pr(\text{Reject } H_0)$ is determined by the type of test being conducted.
+
+\red{HH: We need to define p values corresponding to the data shown in plots. }
+\red{second step: the p-values under $H_0$ are uniform.}
+
 Theoretical power for the regression parameters is derived in the next section.
-
 \section{Inference for a Regression Model} \label{sec:regression}
@@
 
-In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model (e.g.
-Cook \& ...
-).
+In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model \citep{cook:99}.
 Table \ref{tbl:stat_multiple} shows an overview of some commonly used charts and their corresponding hypothesis tests.
 Note that some charts can be associated with several different testing situations.
-In our paper, we pick the example fo situation 2 and 3 as candidates.
+In our paper, we pick the examples for situation 2 and 3 as candidates.
 
@@
 The test statistic, the plot containing the observed data, is randomly placed among these null plots.
-If the test statistic is identifiable the null hypothesis is rejected with a $p$-value of at most 0.05.
+If the test statistic is identifiable the null hypothesis is rejected.
+% with a $p$-value of at most 0.05.
 
@@
 Now consider estimating the power of the visual test. 
-Suppose $F_{|t|}(.)$ be the distribution function of an absolute value of $t$.
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
 For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
@@
 This should allow us to infer some of the signal strength $p_B$ as
-\[
+\begin{equation} \label{plot_signal}
 E[p_B] = 1 - E[X]/(m-1),
-\]
+\end{equation}
 i.e.
@@
 
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
+       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \eqref{plot_signal}}
+       \label{fig:pval_plot_signal}
+\end{figure*}
+
 Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.

### 2012-03-04 paper_jasa.tex
tightening paper, restructured section on power considerations for lineup

@@
 
+\setlength\parindent{0pt}
+
 \begin{document}
@@
 
-\section{Introduction} Any statistical analysis must include some statistical graphics.
+\section{Introduction}
+
+Any statistical analysis must include some statistical graphics.
 For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
@@
 \red{HH: We need to define p values corresponding to the data shown in plots. }
-\red{second step: the p-values under $H_0$ are uniform.}
+
+\blue{Let $D$ be a measure for the effect contained in a data set. We will assume that $D$ follow a distribution $F_D$. Let $p_D$ be the $p$-value associated with the data shown in the test statistic.
+
+Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
+Since this data is generated consistently with the null hypothesis, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
+}
+
+\red{move power consideration up:}
+
+Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
+%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
+We know that for each comparison the probability that the data plot has the smaller $p$ value is 
+\begin{eqnarray*}
+P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
+&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
+\end{eqnarray*}
+which, in particular, is independent of $p_{0,i}$ for all $i$.
+
+Let us now define $X$ as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$ (i.e.
+under the assumption that the onlooker is able to identify the chart with the smallest $p$-value, they would pick any of the $X$  null plots over the data plot).
+\red{HH: we need to make this assumption a bit more obvious - this is a crucial assumption, and we need to back it up later in the evaluation section.}
+
+Then $X \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
+\[
+P(X=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
+\]
+
+%\red{pull the results for $P(p_B < p_{i,0})$ and $P(p_B < p_0)$ to the front, including the figure on $1/m$ versus power.
+These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
+
+Figure \ref{fig:pval_power} gives an overview of the probability of picking the data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
+We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$.
+This pattern is also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
+
+Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
+Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
+   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_D$.
+The grey line shows probability of picking the data plot under $H_0$.
+}
+   \label{fig:pval_power}
+\end{figure}
+
+Now assume that the same lineup plot is evaluated by $K$ independent observers.
+W.l.o.g.
+let the last plot be the data plot and plots 1 through $m-1$ the null plots.
+Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, then $\sum_{k=1}^{m} n_k = K$, and $(n_1, n_2, ..., n_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$.
+We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
+
+For the distribution of $n_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
+we get
+\[
+n_m\sim \left \{ 
+\begin{array}{ll}
+B_{K, (1-E[p_D])^{m-1}} & \text { under } H_1\\
+B_{K, 1/m} & \text { under } H_0
+\end{array}
+\right .
+\]
+
+Under $H_1$ we therefore get an estimate for the strength of the signal in the test statistic based on a lineup evaluation as 
+\[
+\widehat{E[p_D]} = 1- \left(\frac{n_m}{K}\right)^{1/(m-1)}.
+\]
+
+\red{We need comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.}
 
@@
 Consider a linear regression model 
-\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i 
+\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
 \end{equation}
@@
 
-A lineup including this test statistic is shown in Figure \ref{fig:test_category}.
-The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma^2})$.
+A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
+The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
 The test statistic, the plot containing the observed data, is randomly placed among these null plots.
@@
 \begin{table*}[hbtp] 
-\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i  $ } 
+\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
 \centering 
@@
 
+\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
 Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
@@
 As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
-In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
+%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
 Suppose  $p_0=\min_i(p_{0,i})$.
@@
 
-Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot. This leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as
+Under the assumption that an observer is able to pick the plot with the smallest $p$-value from a lineup plot,  this leads to the decision to reject $H_0$ when $p_{B} < p_0$.
+Thus we have the expected power as
 
@@
 
-Figure \ref{fig:power_expected} shows the power of UMP test and expected power of visual test obtained from equation \eqref{power_exp}.
-Notice that the expected power of visual test is almost as good as the power of UMP test.
-In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution.
-Theorem \ref{thm:power} helps to compute the expected power of visual inference.
+Figure \ref{fig:power_expected} shows the power of the uniformly most powerful test (UMP) in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
+Notice that the expected power of the visual test is almost as good as the power of UMP test.
+In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
 
-\begin{thm}\label{thm:power}
- Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
-\end{thm}
-
-{ \em proof}
-
-Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
-\begin{eqnarray*}
-F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
-  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
-  &=& \left[ x^{m-1}\right]^1_{1-t}\\
-  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
-\end{eqnarray*}
-
-% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
-
-Thus expected power in equation \ref{power_exp} be
-\begin{eqnarray*}
- Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
-  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
-  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
-  &=& E(1-p_B)^{m-1}   \\
-%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
-\end{eqnarray*}
-
+%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
+%\begin{thm}\label{thm:power}
+% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
+%\end{thm}
+% 
+%{ \em proof}
+%
+%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
+%\begin{eqnarray*}
+%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
+%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
+%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
+%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
+%\end{eqnarray*}
+%
+%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
+%
+%Thus expected power in equation \ref{power_exp} be
+%\begin{eqnarray*}
+% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
+%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
+%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
+%  &=& E(1-p_B)^{m-1}   \\
+%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
+%\end{eqnarray*}
+%
+%
 \begin{figure}[hbtp]
@@
 
-\subsection{Some Power Considerations}
+%\subsection{Some Power Considerations}
 
-Let us think of a lineup as a head-to-head comparison of the data plot and $m-1$ null plots.
-Let the data plot have a $p$ value of $p_B$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-We know that for each comparison the probability that the data plot has the smaller $p$ value is 
-\begin{eqnarray*}
-P(p_B < p_{0,i}) &=& 1 - P(p_{i,0} \le p_B) = 1- \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt = 1 - \int_0^1 t f_{p_B}(t) dt = 1 - E[p_B],
-\end{eqnarray*}
-which, in particular, is independent of $p_{0,i}$ for all $i$.
-
-Let us now define $X$ as the number of null plots in a lineup, for which the $p$ value $p_{0,i}$ is smaller than $p_B$ (i.e.
-under the assumption that the onlooker is able to identify the chart with the smallest $p$ value, they would pick any of the $X$  null plots over the data plot).
-Then $X \sim B_{m-1, E[p_B]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
-\[
-P(X=0) = \left(1 - E[p_B] \right)^{m-1} = P(p_B \le p_0),
-\]
-where $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$
-
-\subsection{Expected number of choices}
-Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
-\[
-E[X] = (m-1)(1-E[p_B]).
-\]
-With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-This should allow us to infer some of the signal strength $p_B$ as
-\begin{equation} \label{plot_signal}
-E[p_B] = 1 - E[X]/(m-1),
-\end{equation}
-i.e.
+%\subsection{Expected number of choices}
+%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
+%\[
+%E[X] = (m-1)(1-E[p_B]).
+%\]
+%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
+%This should allow us to infer some of the signal strength $p_B$ as
+%\begin{equation} \label{plot_signal}
+%E[p_B] = 1 - E[X]/(m-1),
+%\end{equation}
+%i.e.
 by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
-
-Now assume that the same lineup plot is evaluated by $K$ independent observers.
-Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$
-
-For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
-\[
-n_m\sim \left \{ 
-\begin{array}{cl}
-B_{K, (1-E[p_B])^{m-1}} & \text { under } H_1\\
-B_{K, 1/m} & \text { under } H_0
-\end{array}
-\right .
-\]
-Figure \ref{fig:pval_power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
-Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e.
-reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
-   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$.
-The grey line shows probability of picking the data plot under $H_0$.
-We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$.
-(Mahbub, is that what we see in practice?) Yes, we see this pattern in our experiments as shown in figure \ref{fig:pval_pcorrect}}
-   \label{fig:pval_power}
-\end{figure}
+%
 

### 2012-03-08 R_codes/turk1_data_analysis_jasa.R
beta/sigma codes added

@@ -49,7 +49,7 @@ dat <- subset(raw.dat, id %in% included_id)
 
 calculate_ump_power <- function (beta, n, sigma){
 	alpha <- 0.05/2
-	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs derivation of power
+	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs for derivation of power
 	mu <- beta/se_beta
 	t_n <- qt(p=1-alpha,df=n-3)
 	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
@@ -209,6 +209,7 @@ dat_boot_ribbon <-  ddply(dat_boot_pow,.(beta,sample_size,sigma), summarize,
            limit2 = quantile(pow,.975, na.rm=T)
            )
 dat_boot_ribbon <- rbind(dat_boot_ribbon,cbind(beta= -dat_boot_ribbon[,1],dat_boot_ribbon[,-1]))
+
 p <- ggplot() +
      geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val, colour=alpha("red",.3)) +
      geom_ribbon(aes(x=beta,ymin=limit1,ymax=limit2), data=dat_boot_ribbon, fill=alpha("red",.2)) +

### 2012-03-08 power.tex

@@
 
-\section{The Flattening out}
-The flattening out in figure \ref{fig:pval_plot_signal} happens, when the signal in the data plot becomes too low to make it distinguishable from a nullplot. 
-Under $H_0$ the $p$-value of the data plot $p_B \sim U[0,1]$.
+%\section{The Flattening out}
+%The flattening out in figure \ref{fig:pval_plot_signal} happens, when the signal in the data plot becomes too low to make it distinguishable from a nullplot. 
+%Under $H_0$ the $p$-value of the data plot $p_B \sim U[0,1]$.
 If we can assume that an observer is able to pick out the plot corresponding to the minimal $p$-value, we have a distribution of $Beta_{1, m}$ for the picked plot.

### 2012-03-08 paper_jasa.tex
small changes

@@
 
-\subsection{Other methods to obtain power}
+\subsection{Estimating Power from Lineups}
 
@@
 %
-where $\tau_j$ is random effects coefficient vector of length $q$ for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and link function $g(\mu)=log(\frac{\mu}{1-\mu}); 0 \le \mu \le 1$.
+where $\tau_j$ is random effects coefficient vector of length $q$ for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
 The covariates could be demographic information of people such as age, gender, education level etc.

### 2012-03-08 R_codes/turk1_data_analysis_jasa.R

@@ -10,6 +10,8 @@ raw.dat <- read.csv("../data/raw_data_turk1.csv")
 # evaluate 50% of those extremely easy lineups, all his responses are discarded.
 # A lineup with p_value < 0.0002 is considered easy
 
+raw.dat$id <- raw.dat$nick_name
+raw.dat$response <- raw.dat$response_no == raw.dat$plot_location
 
 dp <- ddply(subset(raw.dat, p_value < 0.0002),.(id),summarise,
    percent_correct = sum(response==TRUE)*100/length(response)
@@ -86,6 +88,9 @@ calculate_ump_power <- function (beta, n, sigma){
 }
 
 calculate_ump_power(3,100,5)
+calculate_ump_power(3,300,5)
+
+calculate_ump_power(3/5,300,1)
 
 
 

### 2012-03-09 paper_jasa.tex
parametric vs non-parametric tests

@@
 }
-Next, we will develop power theoretically and then relate it to the empirical results.
+%Next, we will develop power theoretically and then relate it to the empirical results.
+\blue{We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
+we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$.
+In a next step we will then relate the theoretical results to our empirical findings.}
 
 \begin{dfn} \label{dfn:power}
-\red{HH: I don't think power should be defined in terms of $\theta$ Usually power is considered a property of the test.
-In our case that's the specifics of the plot used for the lineup.
-We might need an additional definition for that - maybe we can phrase that within the framework of the grammar of graphics?
-- that's what I've tried above}
-
-For a lineup of $m$ plots the power of \blue{a visual test}
+For a lineup of $m$ plots the power of \blue{a visual test} $T$
 %$\theta$
@@
     \begin{equation*}
-      \text{Power}(T_\theta)= 
+      \text{Power}_T(\theta)= 
         \begin{cases} 
@@
 \end{dfn}
-\red{this is a non-standard definition of power - usually power is just the prob to reject under $H_1$. could we back the non-standard definition up with some literature?}
 
@@
 
-\blue{Let $D$ be a measure for the effect contained in a data set. We will assume that $D$ follow a distribution $F_D$. Let $p_D$ be the $p$-value associated with the data shown in the test statistic.
+\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
+For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
+By definition we have
 
-Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
+$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
+
+Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
+\begin{eqnarray}\label{dist_p}
+F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
+  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
+  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
+  &=&\left\{ \begin{array}{ll}
+          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
+          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
+       \end{array} \right.     
+\end{eqnarray}
+
+Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+
+\blue{Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
 Since this data is generated consistently with the null hypothesis, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
@@
 
-\red{move power consideration up:}
+\red{HH: $p_D$ isn't defined yet}
 
@@
 Under $H_1$ we therefore get an estimate for the strength of the signal in the test statistic based on a lineup evaluation as 
+\begin{equation}\label{eqn:power_estimate}
+\widehat{E[p_D]} = 1- \left(\frac{n_m}{K}\right)^{1/(m-1)}.
+\end{equation}
+
+\blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+This allows us to compute this value independently of the parametric setting.
+We therefore define the {\it signal strength} of a  visual test as the following:
+
+\begin{dfn} \label{dfn:signal}
+The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
+\end{dfn}
+
+In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
 \[
-\widehat{E[p_D]} = 1- \left(\frac{n_m}{K}\right)^{1/(m-1)}.
+\sigma_T^{m-1} = \frac{n_m}{K}.
 \]
 
-\red{We need comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.}
+In a parametric setting, signal strength of a test coincides with the power of the test.
+}
+
+\red{We need to comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.}
 
@@
 %Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
-Now consider estimating the power of the visual test. 
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
+%%Now consider estimating the power of the visual test. 
+%%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
 For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
 By definition we have
-
-$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
-
-\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
-Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
-\begin{eqnarray}\label{dist_p}
-F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
-  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
-  &=&\left\{ \begin{array}{ll}
-          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-       \end{array} \right.     
-\end{eqnarray}
-
-Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
+%%
+%%$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
+%%
+%%\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
+%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
+%%\begin{eqnarray}\label{dist_p}
+%%F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
+%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
+%%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
+%%  &=&\left\{ \begin{array}{ll}
+%%          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
+%%          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
+%%       \end{array} \right.     
+%%\end{eqnarray}
+%%
+%%Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
 As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.

### 2012-03-29 paper_jasa_abstract.tex
abstract added

--- 2012-03-09 paper_jasa.tex
@@
 Until recently there were no formal visual methods in place for determining statistical significance of findings.
-This changed, when \citet{buja:2009} conceptually introduced two protocols for formal tests of visual findings.
-In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against classical statistical testing  of the significance of regression model parameters.
+This changed, when Buja et al.
+(2009) conceptually introduced two protocols for formal tests of visual findings.
+In this paper we take this a step further by comparing the lineup protocol (Buja et al., 2009) against classical statistical testing  of the significance of regression model parameters.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
@@
 \end {abstract}
+
+\end{document}
 

### 2012-04-05 paper_jasa.tex
update comments

@@
 
-\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}
+\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
+\red{MM: This $t$ comes from the idea that a classical $T-test$ is available for which we are describing the visual test.
+Generalizing it is a little bit technical since for all visual test, we may not have a classical $T-test$.
+Thus the concept of $F(.)$ is little confusing too.
+I don't know how deal with this situation here.}\\
 Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
@@
 
-\red{HH: $p_D$ isn't defined yet}
+\red{HH: $p_D$ isn't defined yet} \\
+\red{MM: $p_D$ and $p_B$ are same, right?}
 

### 2012-04-05 paper_jasa.tex
updated t dist figure

@@
   Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
- & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 
@@
 
- & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 
- Null Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.15}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
- & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{1.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+ Null Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
  Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\

### 2012-04-05 R_codes/paper_jasa.R

@@ -1,11 +1,15 @@
 # The complete R codes for JASA Paper
 # Last modified by Mahbub on Feb 10, 2012.
 
+library(ggplot2)
+
+
 # =================== mathematical test t distribution  ===================
 
 t <- seq(-4, 4, by = 0.1)
 density_t <- dt(t, df = 15)
-qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5)
+qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5) + ylab("Density")
+ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
 
 # =================== Turk1 data analysis  ================================
 
@@ -265,7 +269,7 @@ p <- ggplot(pdat) +
      geom_abline(aes(intercept=0,slope=1))
 p 
 
-
+ggsave(p,file="../images/p_val_plot_signal.pdf", height = 4.25, width = 10)
 
 # -------- Hieke's code for power vs m and p-value------------------
 

### 2012-04-05 paper_jasa.tex
updated lineup plot

@@
 
- Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.15}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
+ Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
 
@@
 \\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.

### 2012-04-05 R_codes/paper_jasa.R

@@ -1,5 +1,5 @@
 # The complete R codes for JASA Paper
-# Last modified by Mahbub on Feb 10, 2012.
+# Last modified by Mahbub on April 5, 2012.
 
 library(ggplot2)
 
@@ -10,6 +10,48 @@ t <- seq(-4, 4, by = 0.1)
 density_t <- dt(t, df = 15)
 qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5) + ylab("Density")
 ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
+
+
+# === plotting test stat and lineup for categorical variable
+
+# generating data
+set.seed(29)
+n <- 300
+beta <- 15
+sigma <- 12	
+x1 <- rpois(n,lambda=20)
+x2 <- factor(sample(c("A","B"), size=n, replace=T))
+cnd <- sample(c("A","B"), size=1)
+y <- round(5 + 1.5 * x1 + beta * (x2==cnd) + rnorm(n=n, mean=0, sd=sigma))
+#qplot(x1,y, colour=factor(x2)) + geom_smooth(method="lm", se=F, size=1)
+
+
+# plotting test stat for testing b2=0
+fit <- lm(y~x1 )
+fit.stat <- summary(fit)
+sigma_hat <- fit.stat$sigma
+obs.residuals <- as.vector(resid(fit)) 
+
+qplot(x2, obs.residuals,colour=x2,geom="boxplot", ylab="Residual") + 
+      xlab(expression(X[2]))
+ggsave("../images/stat_category.pdf",height=2,width=2.5)
+
+
+
+# plotting lineup for testing b2=0
+loc <- sample(1:20, size=1)  # location of observed plot
+simdat <- matrix(rnorm(n=20*n,mean=0, sd=sigma_hat),ncol=20)
+simdat[,loc] <- obs.residuals
+pdat <- data.frame(simdat,x2)
+names(pdat) <- c(1:20,"Xk")
+pdat.m <- melt(pdat,id="Xk")
+
+qplot(Xk, value, data=pdat.m ,colour=Xk,,geom="boxplot", ylab="Residual") + 
+     facet_wrap(~variable)+ xlab(expression(X[k]))
+ggsave("../images/test_category.pdf",height=5,width=5)
+
+
+
 
 # =================== Turk1 data analysis  ================================
 

### 2012-04-11 paper_jasa.tex
defined visual test p-value

@@
 
-%If $N$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (N,\frac{1}{m})$ and we report a  $p$-value of at most $Pr(U \ge u)= \sum_{k \ge u}^N {{N \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(N-k)}}$ where $u$ is the observed number of successful evaluations.
  %Notice that when $N=1$, this $p$-value is $\frac1m$.
@@
 
+When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
+Thus we define $p$-value for visual test $T_{\theta}$ as in definition \ref{dfn:pval}.
+\begin{dfn} \label{dfn:pval}
+The $p$-value of A visual test $T_{\theta}$ is at most $Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(K-k)}}$ where $u$ is the observed number of successful evaluations.
+\end{dfn}
 % HH: I think this is covered by making power a property of the test

### 2012-04-11 R_codes/paper_jasa.R
updated code

@@ -20,10 +20,10 @@ n <- 300
 beta <- 15
 sigma <- 12	
 x1 <- rpois(n,lambda=20)
-x2 <- factor(sample(c("A","B"), size=n, replace=T))
+Xk <- factor(sample(c("A","B"), size=n, replace=T))
 cnd <- sample(c("A","B"), size=1)
-y <- round(5 + 1.5 * x1 + beta * (x2==cnd) + rnorm(n=n, mean=0, sd=sigma))
-#qplot(x1,y, colour=factor(x2)) + geom_smooth(method="lm", se=F, size=1)
+y <- round(5 + 1.5 * x1 + beta * (Xk==cnd) + rnorm(n=n, mean=0, sd=sigma))
+#qplot(x1,y, colour=factor(xk)) + geom_smooth(method="lm", se=F, size=1)
 
 
 # plotting test stat for testing b2=0
@@ -32,8 +32,8 @@ fit.stat <- summary(fit)
 sigma_hat <- fit.stat$sigma
 obs.residuals <- as.vector(resid(fit)) 
 
-qplot(x2, obs.residuals,colour=x2,geom="boxplot", ylab="Residual") + 
-      xlab(expression(X[2]))
+qplot(Xk, obs.residuals,colour=Xk,geom="boxplot", ylab="Residual") + 
+      xlab(expression(X[k]))
 ggsave("../images/stat_category.pdf",height=2,width=2.5)
 
 

### 2012-04-11 paper_jasa.tex
modified definition of p-value

@@
 When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-Thus we define $p$-value for visual test $T_{\theta}$ as in definition \ref{dfn:pval}.
+This leads us to obtain the probability $$Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(K-k)}}$$ where $u$ is the observed number of successful evaluations.
+We use this probability to  define $p$-value for visual test $T_{\theta}$ as in definition \ref{dfn:pval}.
 \begin{dfn} \label{dfn:pval}
-The $p$-value of A visual test $T_{\theta}$ is at most $Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(K-k)}}$ where $u$ is the observed number of successful evaluations.
+The $p$-value associated with the decision of a visual test $T_{\theta}$ is given by 
+    \begin{equation*}
+        \begin{cases} 
+              p-value \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
+              p-value > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
+        \end{cases}
+    \end{equation*}
 \end{dfn}

### 2012-04-11 paper_jasa.tex
used p_D

@@
 
-\red{HH: We need to define p values corresponding to the data shown in plots. }
+\red{HH: We need to define p values corresponding to the data shown in plots.  MM: $p_D$ is what we use for this, right?}
 
-\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
-\red{MM: This $t$ comes from the idea that a classical $T-test$ is available for which we are describing the visual test.
-Generalizing it is a little bit technical since for all visual test, we may not have a classical $T-test$.
-Thus the concept of $F(.)$ is little confusing too.
-I don't know how deal with this situation here.}\\
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
+%\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
+
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
+Suppose the associated test statistic in classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
 By definition we have
 
-$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
+$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
 
@@
 \begin{eqnarray}\label{dist_p}
-F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
-  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
+F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
+  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
   &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
@@
 
-Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
 
@@
 
-\red{HH: $p_D$ isn't defined yet} \\
-\red{MM: $p_D$ and $p_B$ are same, right?}
+%\red{HH: $p_D$ isn't defined yet} \\
+%\red{MM: $p_D$ and $p_D$ are same, right?}
 
@@
 
-%\red{pull the results for $P(p_B < p_{i,0})$ and $P(p_B < p_0)$ to the front, including the figure on $1/m$ versus power.
+%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the figure on $1/m$ versus power.
 These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
@@
 %%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_B$.
+For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_D$.
 By definition we have
 %%
-%%$$p_B=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_B)$$
+%%$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
 %%
@@
 %%\begin{eqnarray}\label{dist_p}
-%%F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
-%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
+%%F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
+%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
 %%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
@@
 %%
-%%Thus the density of $p_B$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
 %In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
@@
 %\begin{thm}\label{thm:power}
-% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
+% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
 %\end{thm}
@@
 %  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
-%  &=& E(1-p_B)^{m-1}   \\
-%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
+%  &=& E(1-p_D)^{m-1}   \\
+%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
 %\end{eqnarray*}
@@
 %\[
-%E[X] = (m-1)(1-E[p_B]).
+%E[X] = (m-1)(1-E[p_D]).
 %\]
 %With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-%This should allow us to infer some of the signal strength $p_B$ as
+%This should allow us to infer some of the signal strength $p_D$ as
 %\begin{equation} \label{plot_signal}
-%E[p_B] = 1 - E[X]/(m-1),
+%E[p_D] = 1 - E[X]/(m-1),
 %\end{equation}

### 2012-04-17 paper_jasa.tex
distribution of p-value

@@
 
-Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
+Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the classical test statistic, with non-centrality parameter $\delta$.
+Thus the distribution function of $p-value$ be
 \begin{eqnarray}\label{dist_p}
@@
 Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
 
 \blue{Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
-Since this data is generated consistently with the null hypothesis, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
+Since this data is generated consistently with the null hypothesis, by equation \ref{dist_p}, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
 }

### 2012-04-19 paper_jasa.tex
some intro changes

@@
 
+\blue{Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.}
 %\subsection{Visual Statistic}
 
-Unlike classical hypothesis testing the statistic in visual inference is not a single value, but a plot that is appropriately chosen to describe the parameter of interest, $\theta$.
+For visual inference, the statistic is -- unlike in the classical hypothesis test -- not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
 When the alternative hypothesis is true, it is expected that the plot of the observed data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
 We will call a plot with this property a {\it visual statistic} for $\theta$.
+More formally,
+
+\blue{
+
+\begin{dfn} \label{dfn:test}
+A \textit{visual test (statistic)} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot,  consisting of type and specification of aesthetics necessary for complete reproducibility.
+\end{dfn}
+}
 
@@
 We can assess power therefore both empirically based on experimental data and through theoretical considerations.
-\blue{
 
-\begin{dfn} \label{dfn:test}
-A \textit{visual test} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot used in the lineup.
-The grammar consists of type and specification of aesthetics necessary for complete reproducibility.
-\end{dfn}
-}
 %Next, we will develop power theoretically and then relate it to the empirical results.

### 2012-04-19 paper_jasa.tex
some small questions - Mahbub, could you try to address those? thanks!

@@
 \begin{dfn} \label{dfn:power}
-For a lineup of $m$ plots the power of \blue{a visual test} $T$
-%$\theta$
- is defined as 
+For a lineup of $m$ plots the power of \blue{a visual test} $T$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
     \begin{equation*}
@@
 
-\red{HH: let's try to keep the power discussion as general as possible, and only use regression as an example.}
-
 For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
@@
 When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability $$Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} (\frac{1}{m})^k(1-\frac 1m)^{(K-k)}}$$ where $u$ is the observed number of successful evaluations.
-We use this probability to  define $p$-value for visual test $T_{\theta}$ as in definition \ref{dfn:pval}.
+This leads us to obtain the probability $$Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}$$ where $u$ is the observed number of successful evaluations.
+We use this probability to  define the $p$-value for visual test $T_{\theta}$ as follows:
 \begin{dfn} \label{dfn:pval}
 The $p$-value associated with the decision of a visual test $T_{\theta}$ is given by 
+\red{HH: Why is the $p$ value not just $Pr(U \ge u)$ ?
+(We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept.
+}
+
     \begin{equation*}
         \begin{cases} 
-              p-value \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
-              p-value > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
+              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
+              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
         \end{cases}
@@
 \end{dfn}
-% HH: I think this is covered by making power a property of the test
-% In general,  $Pr(\text{Reject } H_0)$ is determined by the type of test being conducted.
-
-\red{HH: We need to define p values corresponding to the data shown in plots.  MM: $p_D$ is what we use for this, right?}
 
 %\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
+\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
 

### 2012-04-22 paper_jasa.tex
some edits to intro

@@
 \begin{dfn} \label{dfn:pval}
-The $p$-value associated with the decision of a visual test $T_{\theta}$ is given by 
-\red{HH: Why is the $p$ value not just $Pr(U \ge u)$ ?
-(We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept.
+The $p$-value associated with the decision of a visual test $T_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
+where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$. Since $u$ is a discrete value with boundaries 0 and $K$,
+ we can also re-formulate the $p$-value in terms of upper and lower boundaries:
+ \begin{equation*}
+ \begin{array}{rcll}
+ P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=1\\
+( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
+ \end{array}
+ \end{equation*}
  }
+\red{HH: Tried to re-write it to answer the question below.
+Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
 
@@
 
-$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
+$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
 
-Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the classical test statistic, with non-centrality parameter $\delta$.
-Thus the distribution function of $p-value$ be
+%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the classical test statistic.
+%, with non-centrality parameter $\delta$. 
+Then the distribution function of $p$-value $p_D$ under $H_0$ is uniform, since:  
 \begin{eqnarray}\label{dist_p}
 F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
-  &=&\left\{ \begin{array}{ll}
-          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-       \end{array} \right.     
+  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
+  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
+  &=&%\left\{ \begin{array}{ll}
+          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
+ %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
+%       \end{array} \right.     
 \end{eqnarray}
 
-Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
+%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
 As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
@@
 \blue{Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
-Since this data is generated consistently with the null hypothesis, by equation \ref{dist_p}, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
+Since this data is generated consistently with the null hypothesis,  the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$ (by equation \ref{dist_p}).
 }
-
-%\red{HH: $p_D$ isn't defined yet} \\
-%\red{MM: $p_D$ and $p_D$ are same, right?}
 
@@
 %Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-We know that for each comparison the probability that the data plot has the smaller $p$ value is 
+We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
 \begin{eqnarray*}
@@
 
-Let us now define $X$ as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$ (i.e.
-under the assumption that the onlooker is able to identify the chart with the smallest $p$-value, they would pick any of the $X$  null plots over the data plot).
-\red{HH: we need to make this assumption a bit more obvious - this is a crucial assumption, and we need to back it up later in the evaluation section.}
+Let us make the assumption that an onlooker is able to identify  the chart corresponding to the data with the smallest $p$-value.
+Further we will assume that all onlookers  have the same ability in identifying the data plot.
+
+With that, we define $X$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.
 
@@
 We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$.
-This pattern is also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
+\red{The following sentence should be moved back} This pattern is also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
 

### 2012-04-23 paper_jasa.tex
model with subject specific ability

@@
 In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS tease the production of high quality statistical plots.
+Modern computing systems like R and SAS ease the production of high quality statistical plots.
 A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
@@
 \section{Visual Statistical Inference} \label{sec:visual_test} 
+Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
+
+Power of a plot and $p$-value have been derived in Buja and Wickham as ...
+
+for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
 
@@
 
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability $$Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}$$ where $u$ is the observed number of successful evaluations.
-We use this probability to  define the $p$-value for visual test $T_{\theta}$ as follows:
+When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$. This leads us to obtain the probability 
+\begin{equation}\label{eqn:pvalue}
+Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
+\end{equation}
+where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $T_{\theta}$ as follows:
 \begin{dfn} \label{dfn:pval}
 The $p$-value associated with the decision of a visual test $T_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
-where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$. Since $u$ is a discrete value with boundaries 0 and $K$,
+where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
+Since $u$ is a discrete value with boundaries 0 and $K$,
  we can also re-formulate the $p$-value in terms of upper and lower boundaries:
@@
  \begin{array}{rcll}
- P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=1\\
+ P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
 ( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
@@
    \centering
-   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
+   \includegraphics[width=3.5in]{images/powerplot.pdf} 
    \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_D$.
@@
 let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, then $\sum_{k=1}^{m} n_k = K$, and $(n_1, n_2, ..., n_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$.
+Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
+Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$.
 We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
 
-For the distribution of $n_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
+For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
 we get
 \[
-n_m\sim \left \{ 
+N_m\sim \left \{ 
 \begin{array}{ll}
@@
 
-Under $H_1$ we therefore get an estimate for the strength of the signal in the test statistic based on a lineup evaluation as 
+By matching the expected values, we therefore get an expression for the $p$-value in the data under $H_1$  as:
 \begin{equation}\label{eqn:power_estimate}
-\widehat{E[p_D]} = 1- \left(\frac{n_m}{K}\right)^{1/(m-1)}.
+{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
 \end{equation}
@@
 \blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
-This allows us to compute this value independently of the parametric setting.
-We therefore define the {\it signal strength} of a  visual test as the following:
+While this allows us to compute this value independently of the parametric setting, it also means that onlookers might pick out the data plot for a reason not related to the value of $\theta$.
 
-\begin{dfn} \label{dfn:signal}
-The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
-\end{dfn}
+%Picking out the data plot is still a measure of how different the data plot is compared to the null plots, and therefore is a test statistic of the visual test.
 
-In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
-\[
-\sigma_T^{m-1} = \frac{n_m}{K}.
-\]
-
-In a parametric setting, signal strength of a test coincides with the power of the test.
+%This allows us to compute this value independently of the parametric setting. We therefore define the {\it signal strength} of a  visual test as the following:
+%
+%\begin{dfn} \label{dfn:signal}
+%The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
+%\end{dfn}
+%
+%In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
+%\[
+%\sigma_T^{m-1} = \frac{n_m}{K}.
+%\]
+%
+%In a parametric setting, signal strength of a test coincides with the power of the test.
 }
 
-\red{We need to comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.}
+\red{We need to comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.
+$p$ value of data plot and $p$ values from null are negatively correlated (if one is low, the other has to be higher.
+We need to address this.}
+\red{HH: should we re-run some of the lineups with different null plots?}
+
+\blue{
+\paragraph{Subject-specific Power}
+Probability to reject null hypothesis is power.
+This is intricately related to the probability to identify the data plot from the lineup.
+We need to address the problematic of individual specific abilities in identifying the data plot
+}. 
+\red{HH: need to cite some literature from cognitive science here}
+
+\blue{Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$ be the random variable, with $X_{j\ell}= 1$, if  subject $j$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
+We can model $\pi_{j\ell}$ with a logistic regression with subject specific random effects as:
+\[
+\text{logit } \pi_{j\ell} = X_\ell \beta + Z u_j,  
+\]
+where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., J$.
+}
 

### 2012-04-26 paper_jasa.tex
small changes

@@
 \blue{
-\paragraph{Subject-specific Power}
+\paragraph{Subject-specific power}
 Probability to reject null hypothesis is power.
@@
 \]
-where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., J$.
+where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., N_\ell$ for each of $N_\ell$ subjects presented with lineup $\ell$.
 }

### 2012-04-30 paper_jasa.tex
working on intro

--- 2012-04-23 paper_jasa.tex
@@
 \setlength{\textheight}{9in}
-\evensidemargin \oddsidemargin
+\evensidemargin 
+\oddsidemargin
 
@@
 
-\setlength\parindent{0pt}
+%\setlength\parindent{0pt}
 
@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results equivalent to the uniformly most powerful (UMP) test and for some scenarios yields better power than the UMP test.
+Results suggest that the lineup protocol provides results comparable to the uniformly most powerful (UMP) test, out-performs the UMP when data is contaminated, and, there may be some super-individuals who yield better power than the UMP test even in the most difficult trials.
 \end {abstract}
@@
 
-Any statistical analysis must include some statistical graphics.
+Statistical graphics plays a vital role in discovery.
+Finding the unexpected in data, structure that was not previously anticipated or covered by the original hypotheses is one of the strengths of visual analysis.
+This was an area pioneered by \cite{tukey:eda} and is particularly pertinent in today's data-rich world where data mining has become an important activity.
+Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
+
+The foundation of statistical graphics is well-established.
+Work by \cite{cleveland:1984}, repeated and expanded by \cite{heer:2010} provides a framework for evaluating human perception of statistical graphics.
+\cite{wilkinson:1999} has developed a grammar of graphics which has been enhanced by \cite{hadley:2009} that formalizes statistics graphics helping to describes how different plots are related.
+The availability of software systems enabling beautiful data graphics to be tightly coupled with modeling, such as R \citep{R}.
+
+But the strength of the patterns revealed by plots or individual differences among the audience members is something difficult to quantify.
+\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
+This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
+
+%, analysis requires plots of data.
 For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
@@
 
-The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
+%The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
 In recent years we have seen several major advances in statistical graphics.
@@
 
-\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
+%\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
 This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
 
-In this paper we compare graphical inference to classical statistical inference for a set of regression type example.
+In this paper we compare the lineup protocol to classical statistical inference, specifically applied to the regression framework.
 We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.

### 2012-05-04 visual-classical.tex
table of visual vs classical inference with computations

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Brief Article}
\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
%\maketitle
\section{Notation and Definitions for Visual Inference}

\begin{tabular}{lp{4in}}
$H_0$ & the data plot is consistent with the null distribution \\
$H_a$ & \\
test statistic & data plot inserted into a lineup of $m-1$ null plots, that are created consistently with the null hypothesis.\\
sampling distribution & \\
Significance level $\alpha$ & $\alpha \in (0,1)$ is the threshold at which we will reject the null hypothesis, should the $p$-value fall under it, usually $\alpha=0.05$.\\
test procedure & a set of $K$ independent judges evaluates a lineup of size $m$ to come to a conclusion whether to reject $H_0$.\\
$p$-value & probability that $x$ or more of $K$ observers pick the data plot  under $H_0$. \\
Power & probability to reject the null hypothesis\\
Type I error & probability to reject the null hypothesis if it is true\\
Type II error & probability to fail to reject the null hypothesis if it is false\\
\end{tabular}

\section{Calculations}
\paragraph{Significance Level}
Setting a significance level $\alpha$ imposes restrictions on the size of the lineup $m$ and the number of judges $K$. Since we want our test to be able to reject the null hypothesis, if there is evidence against it. Practical considerations therefore put limits on the significance level that we can impose: a lineup of size $m > 20$ is only feasible in very special circumstances, which implies, that for a significance level below 0.05 we will need at least two judges to be able to reject. Generally, with a lineup of size $m$ and $K$ independent judges we will be able to reject at a significance level at or above $\alpha = m^{-K}$.

\paragraph{$p$-value}
Assuming that we are dealing with $K$ independent judges, we can model the number of times that the data plot is being selected under $H_0$ as $X$, where $X \sim B_{K, 1/m}$. Given that $x$ out of the $K$ judges picked the data plot, then  the $p$-value is the probability to observe at least $x$ picks or more:
\[
p\text{-value}  =  P(X \ge x) = 1 - B_{K, 1/m} (x-1)
\]
for $x \ge 1$.
Since the above formula does not lead to informative $p$-values for  $x=0$ we will be reporting a lower boundary in this situation, i.e:
\begin{eqnarray*}
p\text{-value} &\ge& P(X \ge 1) \text{ for } x= 0.\\
\end{eqnarray*}

\paragraph{Power}
We define the power of a lineup as the probability to reject the null hypothesis. Let $X_i$ be the random variable that observer $i$ picks out the data plot. Then $X_i \sim B_{1, p_i}$ and $P(X_i = 1) = p_i$. Let $X = \sum_{i=1}^K X_i$, i.e. $X$ has a Poisson-Binomial distribution.
We can then express the power of the lineup as $P(X \ge x)$ with $x = 0, ..., K$.

$P(X \ge x)$ can be approximated by $Po_\lambda$ with $\lambda= \sum_i p_i$. (Le Cam Inequality, but error is too big: $\sum_i p_i^2$, since $p_i$ are not small). Closed form density (Closed-Form Expression for the Poisson-Binomial Probability Density Function, Martin Lockheed et al, 2010) might be a way out, but simulation should work.

Computing power  is intricately linked with our ability to accurately model an individual's ability to pick the data plot $p_i$.

Under the assumption, that every person has the same ability $p_D$ to pick the data plot from a lineup, we know that the distribution for $X$, the number of participants identifying the data plot given $H_1$  simplifies to  Binomial distribution $B_{K, p}$, and we can estimate $p$ as
\[
\hat{p} = x/K.
\]

In the situation, that we have multiple evaluations of lineups by the same participants, we can estimate and subject-specific abilities of identifying the data plot, and emply a subject-specific random effects to more accurately estimate $p_{ij}$ as:

\begin{equation}
g(p_{ij}) = X_{ij}B  + Z_{ij} \tau_j 
\label{mixed} 
\end{equation}
%
%where $\tau_j$ is a vector of random effects for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$. \end{document}

\end{document}

### 2012-05-08 visual-classical.tex
updated glossary

@@
 \begin{tabular}{lp{4in}}
-$H_0$ & the data plot is consistent with the null distribution \\
-$H_a$ & \\
-test statistic & data plot inserted into a lineup of $m-1$ null plots, that are created consistently with the null hypothesis.\\
-sampling distribution & \\
-Significance level $\alpha$ & $\alpha \in (0,1)$ is the threshold at which we will reject the null hypothesis, should the $p$-value fall under it, usually $\alpha=0.05$.\\
-test procedure & a set of $K$ independent judges evaluates a lineup of size $m$ to come to a conclusion whether to reject $H_0$.\\
-$p$-value & probability that $x$ or more of $K$ observers pick the data plot  under $H_0$. \\
-Power & probability to reject the null hypothesis\\
-Type I error & probability to reject the null hypothesis if it is true\\
-Type II error & probability to fail to reject the null hypothesis if it is false\\
+$H_o$ & Null hypothesis, of the essence that there is nothing interesting to see in this data. \\
+$H_a$ & Alternative hypothesis, of the essence that the patterns seen in the data are inconsistent with randomness, without a need to be very specific. \\
+Test statistic & Plot of the data data, designed to be able to detect interesting as defined by the null and alternative hypotheses\\
+Null plot & Plot of data that is generated in a manner consistent with the null hypothesis. Sometimes the distribution that this data is drawn from is considered to be the null distribution. \\
+Sampling distribution & Distribution of the test statistic, assuming that the null hypothesis is true.
+When a lineup is used to test the null hypothesis a small, finite number of samples is drawn from the null distribution, and plotted in the same manner as the test statistic.
+This forms the sampling distribution upon which the test is made.
+\\
+Significance level, $\alpha$ & $\alpha \in (0,1)$, usually $\alpha=0.05$, is the threshold value at which we will reject the null hypothesis.
+Because a small finite number of null plots are shown along with the data, the significance level will be 1/(\# of plots in the lineup) for one observer.
+For multiple observers, $\alpha$ might be set to be lower than 0.05.
+\\
+%test procedure & a set of $K$ independent judges evaluates a lineup of size $m$ to come to a conclusion whether to reject $H_0$.\\
+$p$-value & By definition, the $p$-value is the probability the data plot is recognizable as different from the null plots, assuming that the null hypothesis is true.
+For multiple observers, this can be estimated by assuming that the chance that the data plot is chosen by one observer is $p=0.05$ (typically), and accounting for the mulitple picks using a Beta distribution.
+\\
+Power & Defined as the probability of rejecting the null hypothesis, when the null hypothesis is not true.
+This is estimated from the trials as the proportion, number of observers who pick the data plot/total number of observers.
+\\
+Type I error & The probability of rejecting the null hypothesis if it is actually true. This is typically controlled at the value $\alpha$. \\
+Type II error & The probability of failing to reject the null hypothesis if it is not true.\\
 \end{tabular}

### 2012-05-08 visual-classical.tex
updated glossary

@@
 
-\begin{tabular}{lp{4in}}
-$H_o$ & Null hypothesis, of the essence that there is nothing interesting to see in this data. \\
-$H_a$ & Alternative hypothesis, of the essence that the patterns seen in the data are inconsistent with randomness, without a need to be very specific. \\
-Test statistic & Plot of the data data, designed to be able to detect interesting as defined by the null and alternative hypotheses\\
+\begin{tabular}{lp{4.5in}}
+$H_o$ & Null hypothesis, of the essence that there is nothing interesting in the population, or population is consistent with randomness. \\
+$H_a$ & Alternative hypothesis, of the essence that the patterns are inconsistent with randomness, without a need to be very specific. \\
+Statistic & Mapping from the random variables to the sample statistic, plot, ideally using a grammar of graphics language, designed to be able to detect interesting as defined by the null and alternative hypotheses.
+\\
+Test statistic & Plot of the observed data.\\
 Null plot & Plot of data that is generated in a manner consistent with the null hypothesis. Sometimes the distribution that this data is drawn from is considered to be the null distribution. \\
@@
 This forms the sampling distribution upon which the test is made.
+Sometimes this is referred to as the null distribution, but it leads to ambiguous language.
 \\
 Significance level, $\alpha$ & $\alpha \in (0,1)$, usually $\alpha=0.05$, is the threshold value at which we will reject the null hypothesis.
-Because a small finite number of null plots are shown along with the data, the significance level will be 1/(\# of plots in the lineup) for one observer.
-For multiple observers, $\alpha$ might be set to be lower than 0.05.
+This is the rate at which we are willing to be wrong in rejecting the null hypothesis.
+Because a small finite number of null plots are shown along with the data, the significance level might commonly be decided to be 1/(\# of plots in the lineup).
 \\
 %test procedure & a set of $K$ independent judges evaluates a lineup of size $m$ to come to a conclusion whether to reject $H_0$.\\
-$p$-value & By definition, the $p$-value is the probability the data plot is recognizable as different from the null plots, assuming that the null hypothesis is true.
-For multiple observers, this can be estimated by assuming that the chance that the data plot is chosen by one observer is $p=0.05$ (typically), and accounting for the mulitple picks using a Beta distribution.
+$P$-value & The $p$-value is the probability that one or more of the null plots is more recognizable as different than the test statistic.
+Or, another way to say this is that if the test statistic is chosen that none of the null plots was more recognizable.\\%, so the $p$-value is 1/(\# of plots shown).
+\\% is chosen, assuming that the null hypothesis is true.
+That is, the chance that the data plot is recognizable as different from the null plots.
 \\
-Power & Defined as the probability of rejecting the null hypothesis, when the null hypothesis is not true.
-This is estimated from the trials as the proportion, number of observers who pick the data plot/total number of observers.
+Power & Defined as the probability of rejecting the null hypothesis, often considered in the context of the alternative hypothesis.
+\\%This is estimated from the trials as the proportion, number of observers who pick the data plot/total number of observers.
 \\
-Type I error & The probability of rejecting the null hypothesis if it is actually true. This is typically controlled at the value $\alpha$. \\
-Type II error & The probability of failing to reject the null hypothesis if it is not true.\\
+Type I error & The probability of rejecting the null hypothesis if it is actually true.
+This is typically controlled at the value $\alpha$.
+This is the probability that the test statistic is chosen, if the null hypothesis is true.\\
+Type II error & The probability of failing to reject the null hypothesis if it is not true.
+This is the probability that the test statistic is not chosen, even though the null hypothesis is not true.\\
 \end{tabular}

### 2012-05-21 paper_jasa.tex
tweaking intro, and then will rearrange paper to get it into shape for submission

@@
 % Article top matter
-\title{Validation of Visual Statistical Inference, with Application to Linear Models}
+\title{Validation of Visual Statistical Inference, Applied to Linear Models}
 \author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
@@
 
-Statistical graphics plays a vital role in discovery.
-Finding the unexpected in data, structure that was not previously anticipated or covered by the original hypotheses is one of the strengths of visual analysis.
-This was an area pioneered by \cite{tukey:eda} and is particularly pertinent in today's data-rich world where data mining has become an important activity.
+Statistical graphics nourishes the discovery process in data analysis: revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally contrasting prevailing hypotheses.
+It is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
 Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
 
-The foundation of statistical graphics is well-established.
-Work by \cite{cleveland:1984}, repeated and expanded by \cite{heer:2010} provides a framework for evaluating human perception of statistical graphics.
-\cite{wilkinson:1999} has developed a grammar of graphics which has been enhanced by \cite{hadley:2009} that formalizes statistics graphics helping to describes how different plots are related.
-The availability of software systems enabling beautiful data graphics to be tightly coupled with modeling, such as R \citep{R}.
+Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years.
+Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{heer:2010} with findings supporting the original results.
+This body of work provides a framework for evaluating new statistical graphics.
+In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots.
+\cite{wilkinson:1999} developed a grammar of graphics that is enhanced by \cite{hadley:2009}.
+These methods provide a mechanism to abstract the way data is mapped to graphical form.
+Finally, technology advances make it simple and easy for everyone to draw plots of data, and particularly the existence of software systems, such as R \citep{R}, enable making beautiful data graphics that can be tightly coupled with statistical modeling.
 

### 2012-05-21 paper_jasa.tex
working on intro still

@@
 
-But the strength of the patterns revealed by plots or individual differences among the audience members is something difficult to quantify.
-\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
+However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance.
+This has also been addressed in recent research.
+\citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics.
 This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
+One of the protocols, the lineup, places the data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
+Figure \ref{fig:test_category} shows an example lineup.
+Which plot do you think is the most different from the others?
+Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
+The null hypothesis typically derives from the task at hand, or the type of plot being made.
+The alternative encompasses all possible antitheses, all types of patterns that might be detected in the data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
+The probability of rejection can be quantified, along with Type I, and Type II error, and power can be defined and estimated.
+
+\begin{figure}[htp]
+%\begin{figurehere}
+   \centering
+       \includegraphics[width=0.95\textwidth]{test_category.pdf}
+       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$.
+When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots.
+Which plot has the biggest vertical displacement between the boxplots?}
+       \label{fig:test_category}
+\end{figure}
+%\end{figurehere}
+
+The protocol has only been informally tested until now.
+This paper describes research where the lineup protocol is compared head to head with the relevant classical test.
+Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
+This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario.
+The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
+The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works.
 
@@
 % with a $p$-value of at most 0.05.
-
-\begin{figure}[hbt]
-%\begin{figurehere}
-   \centering
-       \scalebox{0.95}{\includegraphics{test_category.pdf}}
-       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$.
-When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots.
-Which plot has the biggest vertical displacement between the boxplots?}
-       \label{fig:test_category}
-\end{figure}
-%\end{figurehere}
 

### 2012-05-22 paper_jasa.tex
finished with intro, and have re-titled the following sections

@@
 The alternative encompasses all possible antitheses, all types of patterns that might be detected in the data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
-The probability of rejection can be quantified, along with Type I, and Type II error, and power can be defined and estimated.
+The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.
 
@@
 The protocol has only been informally tested until now.
-This paper describes research where the lineup protocol is compared head to head with the relevant classical test.
+In the work described in this paper, the lineup protocol is compared head to head with the relevant classical test.
 Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
-This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario.
+This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference.
 The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
-The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works.
+The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
+Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
+The results are compared with those of the classical test.
 
@@
 
-In this paper we compare the lineup protocol to classical statistical inference, specifically applied to the regression framework.
+%In this paper we compare the lineup protocol to classical statistical inference, specifically applied to the regression framework.
 We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
 
-Section \ref{sec:visual_test} describes the basic ideas of visual inference, Section \ref{sec:regression} applies these ideas to the case of inference for regression analysis.
-In Section \ref{sec:simulation} we outline the setup for the human-subject study and  and present results.
+The paper is organized as follows.
+Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data.
+Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
+Section \ref{sec:simulation} describes the results of the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent uniformly most powerful (UMP) and an analysis of the resulting data.
+The last section describes the next steps in this research.
+
+\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test} 
+Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
+
+Power of a plot and $p$-value have been derived in Buja and Wickham as ...
+
+for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
+
+This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
+Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
+
+Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
+Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
+
+\blue{Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.}
+%\subsection{Visual Statistic}
 
@@
 \end{table*}
-
-\section{Visual Statistical Inference} \label{sec:visual_test} 
-Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
-
-Power of a plot and $p$-value have been derived in Buja and Wickham as ...
-
-for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
-
-This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
-Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
-
-Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
-Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-
-\blue{Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.}
-%\subsection{Visual Statistic}
 
@@
 Theoretical power for the regression parameters is derived in the next section.
-\section{Inference for a Regression Model} \label{sec:regression}
+
+\section{Application of Visual Inference in Linear Models} \label{sec:regression}
 
@@
 
-\section{Simulation Experiment} \label{sec:simulation}
+\section{Results of Simulation Experiments} \label{sec:simulation}
 

### 2012-05-22 paper_jasa.tex
changed one word
working on vis inf glossary

@@
        \includegraphics[width=0.95\textwidth]{test_category.pdf}
-       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$.
-When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots.
-Which plot has the biggest vertical displacement between the boxplots?}
+       \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
+One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
+Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?
+*** Make the legend text use a subscript}
        \label{fig:test_category}
@@
 The protocol has only been informally tested until now.
-In the work described in this paper, the lineup protocol is compared head to head with the relevant classical test.
+In the work described in this paper, the lineup protocol is compared head to head with the equivalent classical test.
 Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
@@
 \section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test}
+
+Table \ref{tbl:compare} illustrates the lineup protocol in relation to traditional hypothesis testing.
+Both start from the same place, the same null hypothesis.
+The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
+In the lineup, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
+In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
+If it is extreme on this scale then the null hypothesis is rejected.
+In constrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
+If the data plot is selected as the most different, then this results in rejection of the null hypothesis.
+
+\begin{table*}[hbtp]
+\caption{Comparison of visual inference with existing inference technique. *** The plot for the traditional test is wrong, for two sided, p-value is area to the right and left of $|T(y|$.}
+\centering 
+\begin{tabular}{llll} 
+\hline
+  & Traditional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
+\hline
+  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+
+ Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
+
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+
+ Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
+ & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
+ Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
+\hline 
+\end{tabular}
+\label{tbl:compare}
+\end{table*}
+
 Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
@@
 %\subsection{Visual Statistic}
-
-\begin{table*}[hbtp]
-\caption{Comparison of visual inference with existing inference technique}
-\centering 
-\begin{tabular}{llll} 
-\hline
-  & Mathematical Inference &  Visual Inference \\ %[0.5ex] % inserts table %heading 
-\hline
-  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
-
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Null Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
- Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
-\hline 
-\end{tabular}
-\label{tbl:compare}
-\end{table*}
 

### 2012-05-22 paper_jasa.tex
legend subscript updated
tweaking glossary

@@
 The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
-In the lineup, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
+In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
 In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.

### 2012-05-22 paper_jasa.tex
made suggested *** changes

--- 2012-05-22 paper_jasa.tex
@@
 \newcommand{\red}[1]{{\color{red} #1}}
+
+% \usepackage{pifont} % this package is used to print check mark \checkmark
 
@@
 
+%\checkmark
+
 Statistical graphics nourishes the discovery process in data analysis: revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally contrasting prevailing hypotheses.
@@
 Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?
-*** Make the legend text use a subscript}
+*** Make the legend text use a subscript (MM: is it ok now?)}
        \label{fig:test_category}
@@
 \begin{table*}[hbtp]
-\caption{Comparison of visual inference with existing inference technique. *** The plot for the traditional test is wrong, for two sided, p-value is area to the right and left of $|T(y|$.}
+\caption{Comparison of visual inference with existing inference technique.
+*** The plot for the traditional test is wrong, for two sided, p-value is area to the right and left of $|T(y|$.
+(MM: the vertical bar is showing the observed $T$ location not p-value.) }
 \centering 
@@
 
- Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
+ Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{test_category_small.pdf}} \end{center} \end{minipage} \\
  & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

### 2012-05-22 R_codes/paper_jasa.R

@@ -2,6 +2,8 @@
 # Last modified by Mahbub on April 5, 2012.
 
 library(ggplot2)
+library(plyr)
+library(reshape)
 
 
 # =================== mathematical test t distribution  ===================
@@ -23,7 +25,7 @@ x1 <- rpois(n,lambda=20)
 Xk <- factor(sample(c("A","B"), size=n, replace=T))
 cnd <- sample(c("A","B"), size=1)
 y <- round(5 + 1.5 * x1 + beta * (Xk==cnd) + rnorm(n=n, mean=0, sd=sigma))
-#qplot(x1,y, colour=factor(xk)) + geom_smooth(method="lm", se=F, size=1)
+#qplot(x1,y, colour=factor(Xk)) + geom_smooth(method="lm", se=F, size=1)
 
 
 # plotting test stat for testing b2=0
@@ -33,8 +35,8 @@ sigma_hat <- fit.stat$sigma
 obs.residuals <- as.vector(resid(fit)) 
 
 qplot(Xk, obs.residuals,colour=Xk,geom="boxplot", ylab="Residual") + 
-      xlab(expression(X[k]))
-ggsave("../images/stat_category.pdf",height=2,width=2.5)
+      xlab(expression(X[k])) + labs(colour=expression(X[k]))
+ggsave("../images/stat_category.pdf",height=2,width=3)
 
 
 
@@ -42,13 +44,15 @@ ggsave("../images/stat_category.pdf",height=2,width=2.5)
 loc <- sample(1:20, size=1)  # location of observed plot
 simdat <- matrix(rnorm(n=20*n,mean=0, sd=sigma_hat),ncol=20)
 simdat[,loc] <- obs.residuals
-pdat <- data.frame(simdat,x2)
+pdat <- data.frame(simdat,Xk)
 names(pdat) <- c(1:20,"Xk")
 pdat.m <- melt(pdat,id="Xk")
 
 qplot(Xk, value, data=pdat.m ,colour=Xk,,geom="boxplot", ylab="Residual") + 
-     facet_wrap(~variable)+ xlab(expression(X[k]))
-ggsave("../images/test_category.pdf",height=5,width=5)
+     facet_wrap(~variable)+ xlab(expression(X[k])) + 
+     labs(colour=expression(X[k]))
+ggsave("../images/test_category.pdf",height=7,width=7.5)
+ggsave("../images/test_category_small.pdf",height=4.5,width=5) # for comparizon table
 
 
 

### 2012-05-22 paper_jasa.tex
tweaking visual inference explanations

--- 2012-05-22 paper_jasa.tex
@@
 \section{Introduction}
+
+%\checkmark
 
@@
 If it is extreme on this scale then the null hypothesis is rejected.
-In constrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
+In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
 If the data plot is selected as the most different, then this results in rejection of the null hypothesis.
@@
 \hline
-  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
+  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta > 0$& $H_0: \beta=0$ vs $H_1: \beta > 0$\\
  & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
@@
 
-Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
+%Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
 
-Power of a plot and $p$-value have been derived in Buja and Wickham as ...
+%Power of a plot and $p$-value have been derived in Buja and Wickham as ...
 
-for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
+%for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
 
-This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
+%This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
 Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
 
-Let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space. 
+In general, let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space.
 Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-
-\blue{Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.}
-%\subsection{Visual Statistic}
-
+Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.
 For visual inference, the statistic is -- unlike in the classical hypothesis test -- not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
@@
 
-\blue{
+\begin{dfn} \label{dfn:test}
+A \textbf{visual test (statistic)} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot,  consisting of type and specification of aesthetics necessary for complete reproducibility.
+\end{dfn}
 
-\begin{dfn} \label{dfn:test}
-A \textit{visual test (statistic)} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot,  consisting of type and specification of aesthetics necessary for complete reproducibility.
-\end{dfn}
-}
+The visual test statistic is compared with values $ \theta \in \Theta_0$ using a lineup, which is defined as:
 
 \begin{dfn}\label{dfn:lplot}
-A lineup plot is a layout of $m$ visual statistics, consisting of 
+A \textbf{lineup}  is a layout of $m$ visual statistics, consisting of 
 \begin{itemize}\itemsep-3pt
-\item $m-1$ plots simulated from the model specified by $H_0$  (null plots) and 
+\item $m-1$ statistics simulated from the model specified by $H_0$  (null plots) and 
 \item the test statistic produced by plotting the observed data, possibly arising from $H_1$.

### 2012-05-23 paper_jasa.tex
tweaking first two paragraphs of definitions/explanations

@@
 One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?
-*** Make the legend text use a subscript (MM: is it ok now?)}
+Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?}
        \label{fig:test_category}
@@
 \begin{table*}[hbtp]
-\caption{Comparison of visual inference with existing inference technique.
-*** The plot for the traditional test is wrong, for two sided, p-value is area to the right and left of $|T(y|$.
-(MM: the vertical bar is showing the observed $T$ location not p-value.) }
+\caption{Comparison of visual inference with existing inference technique.  }
 \centering 
@@
 
-In general, let $\theta$ be a population parameter of interest, with $\theta \in \Theta$, the parameter space.
+In general, we define $\theta$ to be a population parameter of interest, with $\theta \in \Theta$, the parameter space.
 Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-Similarly, for any test, we have a rejection region $R$ and, complementary to that, $R^c$,  that partition the space of the test statistic, which is usually a subspace of the real numbers.
-For visual inference, the statistic is -- unlike in the classical hypothesis test -- not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
+A test statistic, $T(y)$, is a function that mapss the sample into a numerical summary, that can be used to test the null hypothesis.
+The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
+$T(y)$ is assessed relative to null values of this statistic $T_0$ which would be possible values if $H_0$ was true.
+%Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
+For visual inference, unlike in the classical hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
 When the alternative hypothesis is true, it is expected that the plot of the observed data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
@@
 \begin{dfn} \label{dfn:test}
-A \textbf{visual test (statistic)} $T_\theta$ is defined as the grammar \citep{wilkinson:1999} of the plot,  consisting of type and specification of aesthetics necessary for complete reproducibility.
+A \textbf{visual test (statistic)} $T(y)$ is as a plot of the sample, and $T_0$ are plots of null data.
 \end{dfn}
 
-The visual test statistic is compared with values $ \theta \in \Theta_0$ using a lineup, which is defined as:
+Ideally, the visual test statistic is defined using the grammar of graphics\citep{wilkinson:1999,hadley:2009}, consisting of type and specification of aesthetics necessary for complete reproducibility.
+The visual test statistic is compared with values $T_0$ using a lineup, which is defined as:
 
@@
 \begin{itemize}\itemsep-3pt
-\item $m-1$ statistics simulated from the model specified by $H_0$  (null plots) and 
-\item the test statistic produced by plotting the observed data, possibly arising from $H_1$.
+\item $m-1$ statistics, $T_0$ simulated from the model specified by $H_0$  (null plots) and 
+\item the test statistic, $T(y)$ produced by plotting the observed data, possibly arising from $H_1$.
 \end{itemize}
@@
 
-The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the test statistic assuming that the null hypothesis is true.
-If $H_1$ is true, \blue{we expect this to be reflected as a feature in the test statistic, i.e. the plot of the data, that makes it visually distinguishable from the null plots.}
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}  A careful visual inspection follows, and the observer is asked to point out the plot most different from the lineup.
+\noindent The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the test statistic assuming that the null hypothesis is true.
+If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e.
+the plot of the data, that makes it visually distinguishable from the null plots.
+{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}
 
-Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$. Thus we have type-I error probability of $1/m$.
+A careful visual inspection of the lineup by an independent observer follows, and the observer is asked to point out the plot most different from the lineup.
+Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$.
+Thus we have type-I error probability of $1/m$.
 

### 2012-05-23 paper_jasa.tex
slight clarification of T, T0, theta
comments added in def 2.1.1

@@
 The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
-$T(y)$ is assessed relative to null values of this statistic $T_0$ which would be possible values if $H_0$ was true.
+$T(y)$ is assessed relative to null values of this statistic $T_0$ which would be possible values of $T$ if $\theta \in \Theta$.
 %Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
@@
 A \textbf{visual test (statistic)} $T(y)$ is as a plot of the sample, and $T_0$ are plots of null data.
+\red{MM: I was thinking, visual test statistic $T(.)$ is a function that maps (using the grammar of graphics) data into a statistical plot.
+For observed data ($y$) we have observed test statistic (plot) $T(y)$ and for null data $y_0$ we have $T(y_0)$ which could be denoted by $T_0$.}
+
 \end{dfn}

### 2012-05-23 paper_jasa.tex
slight tweak of second paragraph

--- 2012-05-23 paper_jasa.tex
@@
 \begin{itemize}\itemsep-3pt
-\item $m-1$ statistics, $T_0$ simulated from the model specified by $H_0$  (null plots) and 
-\item the test statistic, $T(y)$ produced by plotting the observed data, possibly arising from $H_1$.
+\item $m-1$ statistics, $T_0$, simulated from the model specified by $H_0$  (null plots) and 
+\item the test statistic, $T(y)$, produced by plotting the observed data, possibly arising from $H_1$.
 \end{itemize}
@@
 Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$.
-Thus we have type-I error probability of $1/m$.
+Thus we have Type I error probability of $1/m$.
 
@@
         \begin{cases} 
-              \text{Type-I error}=\frac1m & \text{if $\theta \in \Theta_0$,} \\
+              \text{Type I error}=\frac1m & \text{if $\theta \in \Theta_0$,} \\
               Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}

### 2012-05-23 paper_jasa.tex
done with definitions up to power, p-value

@@
 Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-A test statistic, $T(y)$, is a function that mapss the sample into a numerical summary, that can be used to test the null hypothesis.
+A test statistic, $T(y)$, is a function that maps the sample into a numerical summary, that can be used to test the null hypothesis.
 The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
-$T(y)$ is assessed relative to null values of this statistic $T_0$ which would be possible values of $T$ if $\theta \in \Theta$.
+$T(y)$ is assessed relative to null values of this statistic $T(y_0)$ which would be possible values of $T$ if $\theta \in \Theta$.
 %Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
@@
 \begin{dfn} \label{dfn:test}
-A \textbf{visual test (statistic)} $T(y)$ is as a plot of the sample, and $T_0$ are plots of null data.
+A \textbf{visual test statistic}, $T(.)$, is a function of a sample that makes a plot.
+$T(y)$ maps the observed data to the plot, and we call this the \textbf{observed data plot}, and $T(y_0)$ maps samples drawn from the null distribution into the same plot form, and these are called \textbf{null plots}.
 \end{dfn}
 
-Ideally, the visual test statistic is defined using the grammar of graphics\citep{wilkinson:1999,hadley:2009}, consisting of type and specification of aesthetics necessary for complete reproducibility.
-The visual test statistic is compared with values $T_0$ using a lineup, which is defined as:
+\noindent Ideally, the visual test statistic is defined using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
+The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
 
@@
 \begin{itemize}\itemsep-3pt
-\item $m-1$ statistics, $T_0$, simulated from the model specified by $H_0$  (null plots) and 
+\item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
 \item the test statistic, $T(y)$, produced by plotting the observed data, possibly arising from $H_1$.
@@
 
-\noindent The $(m-1)$ null plots can be considered to be samples drawn from the sampling distribution of the test statistic assuming that the null hypothesis is true.
+\noindent The $(m-1)$ null plots can be members of the sampling distribution of the test statistic assuming that the null hypothesis is true.
 If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e.
 the plot of the data, that makes it visually distinguishable from the null plots.
-{\em If the test statistic cannot be identified} in the lineup the conclusion is to {\em not reject the null hypothesis.}
+A careful visual inspection of the lineup by an independent observer follows, and the observer is asked to point out the plot most different from the lineup.
+If the test statistic is identified in the lineup, the conclusion is to reject the null hypothesis.
+If it is not identified the conclusion is to not reject the null hypothesis.
 
-A careful visual inspection of the lineup by an independent observer follows, and the observer is asked to point out the plot most different from the lineup.
-Since the lineup plot consists of $m$ plots, the probability of choosing any one of them is $1/m$.
-Thus we have Type I error probability of $1/m$.
+\begin{dfn}\label{dfn:test}
+The \textbf{test} is defined to be 
+\begin{itemize}\itemsep-3pt
+\item \textbf{Reject} $H_0$ if the observer selects the observed data plot, and
+\item \textbf{Fail to reject} $H_0$ if the observer selects a null plot.
+\end{itemize}
+\end{dfn}
+
+The number of plots in the lineup, $m$, controls the Type I error.
+The probability of choosing any one of the plots is $1/m$, so the chance of making a wrong decision by chance is $1/m$.
+Classically, Type II error is the probability of failing to reject $H_0$ when it is not true, which is more difficult to quantify.
+
+\begin{dfn}\label{dfn:error}
+The \textbf{Type I error} is the probability of identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots.
+\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true.
+\end{dfn}
 
@@
 
-For two different visual test statistics of the same observed data, the one  is better, in which a specific pattern is more easily distinguishable to the observer.
-This should be reflected in the power of the test.
+\noindent
+
+Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
+This is akin to different choices of statistics in classical hypothesis testing, for example, mean, median, or trimmed mean.
+One is typically better than another.
+For two different visual test statistics of the same observed data, the one is considered to be better, if $T(y)$ is more easily distinguishable to the observer.
+Power is typically used to measure this characteristic of a test.
+Power is the probability of rejecting $H_0$ when it is not true.
 We can assess power therefore both empirically based on experimental data and through theoretical considerations.
@@
 %Next, we will develop power theoretically and then relate it to the empirical results.
-\blue{We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
+%We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
 we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$.
-In a next step we will then relate the theoretical results to our empirical findings.}
+In a next step we will then relate the theoretical results to our empirical findings.
 
 \begin{dfn} \label{dfn:power}
-For a lineup of $m$ plots the power of \blue{a visual test} $T$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
+For a lineup of $m$ plots the \textbf{power} of a visual test $T$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
     \begin{equation*}

### 2012-05-23 paper_jasa.tex
some comments added

@@
 
-\noindent Ideally, the visual test statistic is defined using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
+\noindent Ideally, the visual test statistic is defined (\red{should it be constructed/plotted/drawn?}) using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
 The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
@@
 \begin{dfn}\label{dfn:test}
-The \textbf{test} is defined to be 
+The \textbf{(\red{Visual}) test} is defined to be 
 \begin{itemize}\itemsep-3pt
@@
 The number of plots in the lineup, $m$, controls the Type I error.
-The probability of choosing any one of the plots is $1/m$, so the chance of making a wrong decision by chance is $1/m$.
+The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$.
 Classically, Type II error is the probability of failing to reject $H_0$ when it is not true, which is more difficult to quantify.

### 2012-05-24 paper_jasa.tex
notation for visual test added

@@
 \usepackage{color}
-\newcommand{\blue}[1]{{\color{blue} #1}}
-\newcommand{\red}[1]{{\color{red} #1}}
+\newcommand{\blue}[1]{{\color{blue} #1}} %MM
+\newcommand{\red}[1]{{\color{red} #1}} %HH
+\newcommand{\green}[1]{{\color{green} #1}} %DC
 
@@
 
-\noindent Ideally, the visual test statistic is defined (\red{should it be constructed/plotted/drawn?}) using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
+\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
 The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
@@
 \begin{dfn}\label{dfn:test}
-The \textbf{(\red{Visual}) test} is defined to be 
+The \textbf{visual test}, $V_{\theta}$, is defined to be 
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if the observer selects the observed data plot, and
+\item \textbf{Reject} $H_0$ if the observer selects the observed data plot in the lineup of $m$ plots, and
 \item \textbf{Fail to reject} $H_0$ if the observer selects a null plot.
@@
 
+\green{Ok, I'm really getting confused now.
+I think that it might be better to define p-value, and significance level next, and then talk about power.
+Power and p-value are confounded in the discussion below.}
+
 \noindent 
-
 Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
@@
 \begin{dfn} \label{dfn:power}
-For a lineup of $m$ plots the \textbf{power} of a visual test $T$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
+The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
     \begin{equation*}
-      \text{Power}_T(\theta)= 
+      \text{Power}_V(\theta)= 
         \begin{cases} 
-              \text{Type I error}=\frac1m & \text{if $\theta \in \Theta_0$,} \\
+              \frac1m & \text{if $\theta \in \Theta_0$,} \\
               Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
@@
 
-For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
-
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$. This leads us to obtain the probability 
+\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
+When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
+This leads us to obtain the probability
 \begin{equation}\label{eqn:pvalue}
@@
 \end{equation}
-where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $T_{\theta}$ as follows:
+where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
 \begin{dfn} \label{dfn:pval}
-The $p$-value associated with the decision of a visual test $T_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
+The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
 where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).

### 2012-05-24 paper_jasa.tex
comments on definition of lineup
pvalue and some type I and II errors

@@
 \newtheorem{con}{Conjecture}[thm]
-%\setlength{\parindent}{0in}   % for no indent
+\setlength{\parindent}{0in}   % for no indent
 
@@
 \begin{dfn}\label{dfn:lplot}
-A \textbf{lineup} is a layout of $m$ visual statistics, consisting of 
+A \textbf{lineup} is a layout of $m$ randomly placed visual statistics, consisting of 
 \begin{itemize}\itemsep-3pt
@@
 the plot of the data, that makes it visually distinguishable from the null plots.
-A careful visual inspection of the lineup by an independent observer follows, and the observer is asked to point out the plot most different from the lineup.
-If the test statistic is identified in the lineup, the conclusion is to reject the null hypothesis.
-If it is not identified the conclusion is to not reject the null hypothesis.
+A careful visual inspection of the lineup by %\strikethrough{an}
+independent observer\red{s} follows;  observers are asked to point out the plot most different from the lineup.
+If the test statistic is identified in the lineup, \red{this is considered as evidence against}
+% conclusion is to reject 
+the null hypothesis. 
+%If it is not identified the conclusion is to not reject the null hypothesis.
+
+\red{This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
+For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
+Under the null hypothesis $X \sim B_{K, 1/m}$, therefore:
+
+\begin{dfn}\label{dfn:pvalue}
+The $p$-value of a lineup of size $m$ is given as 
+\[
+P(X \ge x) = 1 - B_{K, 1/m} (x-1),
+\]
+for $x > 0$, with $X$ defined as above.
+\end{dfn}
+
+}
 
@@
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if the observer selects the observed data plot in the lineup of $m$ plots, and
-\item \textbf{Fail to reject} $H_0$ if the observer selects a null plot.
+\item \textbf{Reject} $H_0$ if \red{`enough'} observers select the observed data plot in the lineup of $m$ plots, and
+\item \textbf{Fail to reject} $H_0$  \red {otherwise}. %if the observer selects a null plot.
 \end{itemize}
+\red{Here, `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.}
 \end{dfn}
 
-The number of plots in the lineup, $m$, controls the Type I error.
-The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$.
+\red{The Type I error is controlled by both $m$, the size of the lineup, and $K$, the number of observers. $(1/m)^K$ is the minimal value, at which we can control the test.}
+
+%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$. 
 Classically, Type II error is the probability of failing to reject $H_0$ when it is not true, which is more difficult to quantify.
@@
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} is the probability of identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots.
-\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true.
+The \textbf{Type I error} is the probability of \red{rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$.
+Note that due to the discrete nature of $X$ we can not always control exactly for $\alpha$.
+}
+% identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
+\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, \red{i.e.
+$P( X \le x_{1-\alpha$}), where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$, the distribution of $X$ under the null hypothesis.}
+\red{Under the alternative hypothesis, we have to account for different visual abilities of  individuals: let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.
+Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.}
 \end{dfn}

### 2012-05-24 paper_jasa.tex
small latex error

@@
 \textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, \red{i.e.
-$P( X \le x_{1-\alpha$}), where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$, the distribution of $X$ under the null hypothesis.}
+$P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$, the distribution of $X$ under the null hypothesis.}
 \red{Under the alternative hypothesis, we have to account for different visual abilities of  individuals: let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.

### 2012-05-24 paper_jasa.tex
removed parindent 0
working out some values for pvalue

@@
 \newtheorem{con}{Conjecture}[thm]
-\setlength{\parindent}{0in}   % for no indent
 
@@
 %\textheight 9.15in  % when making pdf
-
-%\setlength\parindent{0pt}
 
@@
 A careful visual inspection of the lineup by %\strikethrough{an}
-independent observer\red{s} follows;  observers are asked to point out the plot most different from the lineup.
-If the test statistic is identified in the lineup, \red{this is considered as evidence against}
-% conclusion is to reject 
-the null hypothesis. 
+independent observers follows;  observers are asked to point out the plot most different from the lineup.
+If the test statistic is identified in the lineup, this is considered as evidence against the null hypothesis.
 %If it is not identified the conclusion is to not reject the null hypothesis.
 
-\red{This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
+This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
 For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
-Under the null hypothesis $X \sim B_{K, 1/m}$, therefore:
+Under the null hypothesis $X \sim B_{K, 1/m}$, where $B$ is a binomial distribution, therefore:
 
@@
 \[
-P(X \ge x) = 1 - B_{K, 1/m} (x-1),
+P(X \ge x) = 1 - B_{K, 1/m} (x-1) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
 \]
@@
 \end{dfn}
+\noindent When $K=1, m=20$ the $p$-value can only be $1/20=0.05$. When $K=2, m=20$ and if $x=2$ the $p$-value is $0.0025$, but if $x=1$ the $p$-value is$0.905$. When $K=3, m=20$
 
-}
+\begin{table}[htp]
+\begin{center}
+\begin{tabular}{rrrr}
+K & m & x & $p$-value \\\hline
+1 & 20 & 1 & 0.0500 \\\hline
+2 & 20 & 1 & 0.9050 \\
+2 & 20 & 2 & 0.0025 \\\hline
+\end{tabular}
+\end{center}
+\caption{}
+\label{pvalue}
+\end{table}
 
@@
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if \red{`enough'} observers select the observed data plot in the lineup of $m$ plots, and
-\item \textbf{Fail to reject} $H_0$  \red {otherwise}. %if the observer selects a null plot.
+\item \textbf{Reject} $H_0$ if `enough' observers select the observed data plot in the lineup of $m$ plots, and
+\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
 \end{itemize}
-\red{Here, `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.}
+where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
 \end{dfn}
 
-\red{The Type I error is controlled by both $m$, the size of the lineup, and $K$, the number of observers. $(1/m)^K$ is the minimal value, at which we can control the test.}
-
+The resolution of the Type I error is controlled by both $m$, the size of the lineup, and $K$, the number of observers, with $(1/m)^K$ being the minimal value at which we can control the test. 
 %The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$. 

### 2012-05-25 paper_jasa.tex
comments on p-value
massaged the pvalue explanation, added some questions about Type I, II error

@@
 \[
-P(X \ge x) = 1 - B_{K, 1/m} (x-1) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
+P(X \ge x) = 1 - \sum_{i=0}^{x-1} B_{K, 1/m} (i) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
 \]
@@
 \end{dfn}
-\noindent When $K=1, m=20$ the $p$-value can only be $1/20=0.05$. When $K=2, m=20$ and if $x=2$ the $p$-value is $0.0025$, but if $x=1$ the $p$-value is$0.905$. When $K=3, m=20$
+\noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m$.
+It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the observed data plot.
 
@@
 \begin{center}
-\begin{tabular}{rrrr}
-K & m & x & $p$-value \\\hline
-1 & 20 & 1 & 0.0500 \\\hline
-2 & 20 & 1 & 0.9050 \\
-2 & 20 & 2 & 0.0025 \\\hline
+\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
+$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\\hline
+1 &  1 & 0.0500 & 2 &  1 & 0.9050 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
+&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
+&&& &&& 3 & 3 & 0.0049 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
+&&&  &    &        &   &   &        & 4 & 4 & 0.0000 & 5 & 4 & 0.0000 \\\hline
 \end{tabular}
 \end{center}
-\caption{}
+\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m$ lineups.}
 \label{pvalue}
@@
 
-The resolution of the Type I error is controlled by both $m$, the size of the lineup, and $K$, the number of observers, with $(1/m)^K$ being the minimal value at which we can control the test. 
-%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$. 
-Classically, Type II error is the probability of failing to reject $H_0$ when it is not true, which is more difficult to quantify.
+Associated with any test is the risk of making a Type I error, or Type II error, which for visual inference are defined as follows:
 
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} is the probability of \red{rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$.
-Note that due to the discrete nature of $X$ we can not always control exactly for $\alpha$.
-}
+The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$.
 % identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
-\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, \red{i.e.
-$P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$, the distribution of $X$ under the null hypothesis.}
-\red{Under the alternative hypothesis, we have to account for different visual abilities of  individuals: let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.
-Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.}
+\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$.
 \end{dfn}
+
+\green{Why are we including $x_{1-\alpha}$ in both sides?}
+
+\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, finding an exact $x_{1-\alpha}$.
+For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.
+
+\green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text}.
+
+%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$.
+
+\green{Why????
+Under the alternative hypothesis, we have to account for different visual abilities of  individuals:} let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.
+Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.
 
@@
 %\subsection{Power}
-
-\green{Ok, I'm really getting confused now.
-I think that it might be better to define p-value, and significance level next, and then talk about power.
-Power and p-value are confounded in the discussion below.}
 

### 2012-05-25 paper_jasa.tex
type I error

@@
 For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
-Under the null hypothesis $X \sim B_{K, 1/m}$, where $B$ is a binomial distribution, therefore:
+Under the null hypothesis $X \sim \text{Binom}({K, 1/m})$,  therefore:
 
@@
 \[
-P(X \ge x) = 1 - \sum_{i=0}^{x-1} B_{K, 1/m} (i) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
+P(X \ge x) = 1 - \text{Binom}({K, 1/m}) (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
 \]
-for $x > 0$, with $X$ defined as above.
+with $X$ defined as above. 
 \end{dfn}
-\noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m$.
+Note that for $x=0$ the $p$-value becomes --mathematically-- equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
+being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
+
+\noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
 It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the observed data plot.
@@
 \end{center}
-\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m$ lineups.}
+\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
 \label{pvalue}
@@
 % identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
-\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$.
+\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X <  x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$.
 \end{dfn}
@@
 \green{Why are we including $x_{1-\alpha}$ in both sides?}
+\red{fixed  now-- $X < x_{1 - \alpha}$ is the acceptance region, $X \ge x_{1-\alpha}$ is rejection region.}
 

### 2012-05-25 paper_jasa.tex
type II

@@
 
-\green{Why????
-Under the alternative hypothesis, we have to account for different visual abilities of  individuals:} let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.
-Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.
+\green{Why???? Under the alternative hypothesis, we have to account for different visual abilities of  individuals:} 
+\red{I think that's more a matter of writing it - sorry for the confusion - }
+
+let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise, \red{where $p_i$ is the probability that individual $i$ picks out the data plot from the lineup.
+If all individuals have the same probability $p$ for picking out the data plot, $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$, and we can estimate $p$ as $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the data plot from the lineup.
+
+If there is evidence for individual skills influencing the probability $p_i$ for picking out the data plot, 
+ $X = \sum_i X_i$ has  a Poisson-Binomial distribution.}
 

### 2012-05-25 paper_jasa.tex
typo in table

@@
 &&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
-&&& &&& 3 & 3 & 0.0049 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
+&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
 &&&  &    &        &   &   &        & 4 & 4 & 0.0000 & 5 & 4 & 0.0000 \\\hline

### 2012-05-25 paper_jasa.tex
power questions

@@
 \begin{dfn}\label{dfn:test}
+\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
 The \textbf{visual test}, $V_{\theta}$, is defined to be 
@@
 
+\red{How are $\Theta_0$ and $\Theta^c_0$ defined? - why do we need them? wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?}
 \begin{dfn} \label{dfn:power}

### 2012-05-25 paper_jasa.tex
power definitions

@@
 
-\red{How are $\Theta_0$ and $\Theta^c_0$ defined? - why do we need them? wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?}
+\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
+- why do we need them?
+wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?
+- $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
 \begin{dfn} \label{dfn:power}
@@
 
-\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability
-\begin{equation}\label{eqn:pvalue}
-Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
-\end{equation}
-where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
-\begin{dfn} \label{dfn:pval}
-The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
-where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
-Since $u$ is a discrete value with boundaries 0 and $K$,
- we can also re-formulate the $p$-value in terms of upper and lower boundaries:
+\red{Alternative Power definition:}
+\begin{dfn} \label{dfn:power}
+The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
     \begin{equation*}
- \begin{array}{rcll}
- P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
-( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
- \end{array}
- \end{equation*}
- }
-\red{HH: Tried to re-write it to answer the question below.
-Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
-
-    \begin{equation*}
+      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{1-\alpha}) = 
         \begin{cases} 
-              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
-              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
+               1 - \text{Binom}_{K, 1/m}(x_{1-\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
+              1 - F_{X, \theta} (x_{1-\alpha} - 1) & \text{ if } H_1 \text{ is true}.
         \end{cases}
@@
 
+\red{Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.}
+
+%\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
+When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
+This leads us to obtain the probability
+%\begin{equation}\label{eqn:pvalue}
+%Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
+%\end{equation}
+%where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
+%\begin{dfn} \label{dfn:pval}
+%The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
+%where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
+Since $u$ is a discrete value with boundaries 0 and $K$,
+% we can also re-formulate the $p$-value in terms of upper and lower boundaries:
+% \begin{equation*}
+% \begin{array}{rcll}
+% P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
+%( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
+% \end{array}
+% \end{equation*}
+% }
+
+%\red{HH: Tried to re-write it to answer the question below.
+%Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
+%
+%    \begin{equation*}
+%        \begin{cases} 
+%              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
+%              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
+%        \end{cases}
+%    \end{equation*}
+%\end{dfn}
+
 %\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
-\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
+%\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
+
+\red{The following considerations deal with assessing the `difficulty' of  a lineup of size $m$. I think it's very relevant, but I don't see a natural equivalent in the classical testing situation.}
 

### 2012-05-25 paper_jasa.tex
typo p-value table

@@
 \begin{table}[htp]
+\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
 \begin{center}
@@
 $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\\hline
-1 &  1 & 0.0500 & 2 &  1 & 0.9050 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
+1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
 &&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
@@
 \end{center}
-\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
 \label{pvalue}
@@
 
-\green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text}.
+\green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text.} \blue{its jasa guideline to use fraction like $1/m$ whenever possible to save space}.
 

### 2012-05-25 paper_jasa.tex
power and lineup assessment

@@
 &&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
-&&&  &    &        &   &   &        & 4 & 4 & 0.0000 & 5 & 4 & 0.0000 \\\hline
+&&&  &    &        &   &   &        & 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
 \end{tabular}
@@
 
+\noindent
 \red{Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.}
@@
 
-\red{The following considerations deal with assessing the `difficulty' of  a lineup of size $m$. I think it's very relevant, but I don't see a natural equivalent in the classical testing situation.}
+\subsection*{Lineup size $m$}
+\noindent
+One of the big differences between visual inference and the classical testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
+The choice of $m$ has an obvious impact on the test, and we will discuss this in the following paragraph.
 
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
-Suppose the associated test statistic in classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
+The following properties are only derived for the situation of a  fully parameterized  simulation study.
+This allows us to directly compare lineup tests against their classical counterparts and also identify  properties relevant for a quality assessment of  lineups, which are typically used  in non-parametric situations.
+
+We need to make two crucial assumptions:
+\begin{enumerate}
+\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
+\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
+\end{enumerate}
+We will be able to show evidence for the second assumption with the help of the study discussed in section~\ref{sec:simulation}, the first assumption well be reflected by the power of a lineup.
+The better suited a design is for a particular task, the higher its power will be.
+
+We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
+
+suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
+Suppose the associated test statistic in a classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
 By definition we have
@@
 %, with non-centrality parameter $\delta$. 
-Then the distribution function of $p$-value $p_D$ under $H_0$ is uniform, since:  
+Then the distribution function of the $p$-value $p_D$ under $H_0$ is uniform, since:  
 \begin{eqnarray}\label{dist_p}
@@
 
-\blue{Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots.
-Since this data is generated consistently with the null hypothesis,  the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$ (by equation \ref{dist_p}).
-}
+Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots.
+Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
 
@@
 
-With that, we define $X$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.
+With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.
 
-Then $X \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
+Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
 \[
-P(X=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
+P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
 \]
@@
 We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$.
-\red{The following sentence should be moved back} This pattern is also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
+This pattern is confirmed by our findings described in section~ref{sec:simulation}
+%also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
+
+\red{\#\#\# re-done up to here, HH}
 

### 2012-05-25 paper_jasa.tex
definition of visual test

@@
 \end{dfn}
+
+\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
+
+\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.
+
+\begin{dfn}\label{dfn:test}
+The \textbf{visual test}, $V_{\theta}$  of size $\alpha$, is defined to be 
+\begin{itemize}\itemsep-3pt
+\item \textbf{Reject} $H_0$ if out of $K$ observers $x_{\alpha}$ of them correctly identify the observed data plot in the lineup, and
+\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
+\end{itemize}
+where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
+\end{dfn}
+
+Notice that now we can fix $type-I$ error to some values even less than $1/m$ by increasing $K$. $type-I$ error is $1/m$ only when $K=1$.
+}
 

### 2012-05-25 paper_jasa.tex
section 2 done

@@
 \newtheorem{con}{Conjecture}[thm]
+\newtheorem{lemma}[thm]{Lemma}
 
@@
 
-\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
+%\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
 - why do we need them?
@@
 - $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
-\begin{dfn} \label{dfn:power}
-The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-    \begin{equation*}
-      \text{Power}_V(\theta)= 
-        \begin{cases} 
-              \frac1m & \text{if $\theta \in \Theta_0$,} \\
-              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
-        \end{cases}
-    \end{equation*}
-\end{dfn}
-
-\red{Alternative Power definition:}
+%\begin{dfn} \label{dfn:power}
+%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
+%    \begin{equation*}
+%      \text{Power}_V(\theta)= 
+%        \begin{cases} 
+%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
+%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
+%        \end{cases}
+%    \end{equation*}
+%\end{dfn}
+%
+%
+%\red{Alternative Power definition:}
 \begin{dfn} \label{dfn:power}
@@
 \red{Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.}
+
+\paragraph{Subject-specific abilities}
+Calculating the power  is intricately related to the probability to identify the data plot from the lineup.
+
+In the case that observers are exposed to multiple lineups we can estimate the probability $p_i$ while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
+
+Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$ be the random variable, with $X_{j\ell}= 1$, if  subject $j$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
+We can model $\pi_{j\ell}$ with a logistic regression with subject specific random effects as:
+\[
+\text{logit } \pi_{j\ell} = X_\ell \beta + Z u_j,  
+\]
+where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., J$.
 
@@
 
+\begin{lemma}
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
+Suppose the associated test statistic in a classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
+
+The probability of picking the data plot from a lineup depends on the size $m$ of the lineup and the strength of the signal in the data plot. 
+Under the above assumptions, we can express the probability as:
+\[
+P(p_D < p_0) = \left(1 - E[p_D] \right)^{m-1} 
+\]
+where $p_D$ is the $p$-value associated with the data in the test statistic, and $p_0$ is the minimum of all $p$-values in the data going into null plots.
+\end{lemma}
+
+\begin{proof}
 We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
 
-suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
-Suppose the associated test statistic in a classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
  By definition we have
@@
 \]
+\end{proof}
 
@@
 
-\red{\#\#\# re-done up to here, HH}
-
-Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
-Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
+%Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
+%Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
 
@@
 
-\red{We need to comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.
-$p$ value of data plot and $p$ values from null are negatively correlated (if one is low, the other has to be higher.
-We need to address this.}
-\red{HH: should we re-run some of the lineups with different null plots?}
-
-\blue{
-\paragraph{Subject-specific Power}
-Probability to reject null hypothesis is power.
-This is intricately related to the probability to identify the data plot from the lineup.
-We need to address the problematic of individual specific abilities in identifying the data plot
-}. 
-\red{HH: need to cite some literature from cognitive science here}
-
-\blue{Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$ be the random variable, with $X_{j\ell}= 1$, if  subject $j$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $\pi_{j\ell}$ with a logistic regression with subject specific random effects as:
-\[
-\text{logit } \pi_{j\ell} = X_\ell \beta + Z u_j,  
-\]
-where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., J$.
-}
+\red{\#\#\# re-done up to here, HH}
 

### 2012-05-29 paper_jasa.tex
modeling effect size

@@
 
+The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
+combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
+Using equation \ref{mixed}, we then have the power as
+
+\begin{equation}\label{eqn:effect} 
+Power(E) = \pi=Pr(Y=1| E) 
+\end{equation}
+
 \section{Results of Simulation Experiments} \label{sec:simulation}
@@
    \centering
+       \scalebox{0.7}{\includegraphics{effect_power_exp1.pdf}}
+       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1.}
+       \label{fig:power_loess1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
        \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}

### 2012-05-29 R_codes/paper_jasa.R

@@ -14,6 +14,26 @@ qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5) + ylab("Densit
 ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
 
 
+# ===== computation of p-values =============
+
+get_pval <- function(m,K,x){
+	p <- pbinom(size=K,prob=1/m,q=x)
+	px <- dbinom(size=K,prob=1/m,x=x)
+	pval <- 1-p + px
+	return(pval)
+}
+
+pvals <- NULL
+for (K in 1:5){
+	for (x in 0:K){
+		pval <- get_pval(m=20,K=K,x=x)
+		pvals <- rbind(pvals,c(m=20,K=K,x=x,pval=round(pval,4)))
+	}
+}
+
+
+
+
 # === plotting test stat and lineup for categorical variable
 
 # generating data

### 2012-05-29 R_codes/turk1_data_analysis_jasa.R

@@ -1,6 +1,7 @@
 #Turk1 data analysis for jasa paper last modified on Mar 30, 2012 by Mahbub.
 
 library(ggplot2)
+library(plyr)
 
 
 raw.dat <- read.csv("../data/raw_data_turk1.csv")
@@ -1086,8 +1087,19 @@ ggplot(data=effect.d,aes(x=effect,y=power)) +
    
 ggplot(data=effect.d,aes(x=effect,y=power)) + 
    geom_point() + 
-   geom_smooth(method="loess", span=0.8) +
-   geom_line(aes(x=effect,y=ump_power))  
+   geom_smooth(method="loess", span=0.8,aes(colour=factor(replica))) +
+   geom_line(aes(x=effect,y=ump_power)) +
+   scale_colour_discrete(name = "Replication")+
+   xlab(expression(Effect(E)))
+   
+ggsave(filename="../images/effect_power_exp1.pdf", height=7, width=7)   
+   
+ggplot(data=effect.d,aes(x=effect,y=power-ump_power)) + 
+   geom_point() + 
+   geom_smooth(method="loess", span=0.8,aes(colour=factor(replica))) +
+   scale_colour_discrete(name = "Replication")  
+   
+     
 
 
 

### 2012-05-29 R_codes/turk1_data_analysis_jasa.R
subject specific power from model

@@ -2,6 +2,7 @@
 
 library(ggplot2)
 library(plyr)
+library(reshape)
 
 
 raw.dat <- read.csv("../data/raw_data_turk1.csv")
@@ -595,7 +596,7 @@ p + scale_linetype_manual(values =c(4,3,2,1)) + facet_grid(sample_size~sigma)
 
 
 
-# ----------------- estimating subjectwise estimated power curve -------------------------
+# ----------------- estimating subjectwise power curve -------------------------
 library(lme4)
 fit.mixed <- lmer(response ~ factor(sample_size)+beta+factor(sigma)
               + (1|id)
@@ -1105,6 +1106,58 @@ ggplot(data=effect.d,aes(x=effect,y=power-ump_power)) +
 
 # and then fit the following model:
        response ~ effect + (effect | id)
+
+# ------- subjectwise power curve with random slope  for covariate effect 
+library(lme4)
+fit.mixed <- lmer(response ~ effect + (1+effect|id)
+              , family="binomial"
+              , data=dat)
+res <- summary(fit.mixed)
+B <- res@coefs[,1]
+effect <- seq(0.01,16, by=.2)
+#res@frame
+ones <- rep(1,length(effect))
+#conf_level <- rep(1,length(beta))
+X <- cbind(ones,effect)
+dim(X)
+
+Z <- cbind(ones,effect)
+cov <- as.numeric(res@REmat[2,5])*as.numeric(res@REmat[1,4])*as.numeric(res@REmat[2,4])
+vv <- matrix(as.numeric(c(res@REmat[1,3],rep(cov,2),res@REmat[2,3])),ncol=2, byrow=F)
+
+
+# library(MASS) is needed for mvrnorm
+
+power=NULL
+M <- 20
+set.seed(79345)
+tau <- mvrnorm(n = M, mu=c(0,0),Sigma=vv)
+for (i in 1:M){
+xb <- X %*% B + Z %*%tau[i,]
+power <- cbind(power,exp(xb)/(1+exp(xb)))
+}
+colnames(power) <- 1:M
+
+ump_pow <- NULL
+for (i in effect)ump_pow <- c(ump_pow,calculate_ump_power(beta=i/10,n=100,sigma=1))
+
+
+pow.dat <- data.frame(effects = c(-effect, effect)
+                    , rbind(power,power)
+                    , UMP=rep(ump_pow,2))
[…]

### 2012-05-29 paper_jasa.tex
subject specific power from model
cleaning up definitions still, now stuck

@@
 
-\begin{dfn}\label{dfn:test}
-\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
-The \textbf{visual test}, $V_{\theta}$, is defined to be 
-\begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if `enough' observers select the observed data plot in the lineup of $m$ plots, and
-\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
-\end{itemize}
-where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
-\end{dfn}
+%\begin{dfn}\label{dfn:test}
+%\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
+%The \textbf{visual test}, $V_{\theta}$, is defined to be 
+%\begin{itemize}\itemsep-3pt
+%\item \textbf{Reject} $H_0$ if `enough' observers select the observed data plot in the lineup of $m$ plots, and
+%\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
+%\end{itemize}
+%where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
+%\end{dfn}
 
-\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
+%\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
 
-\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.
+%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.
 
@@
 
-Notice that now we can fix $type-I$ error to some values even less than $1/m$ by increasing $K$. $type-I$ error is $1/m$ only when $K=1$.
-}
-
 Associated with any test is the risk of making a Type I error, or Type II error, which for visual inference are defined as follows:
@@
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$.
+The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
 % identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
-\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X <  x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$.
+\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$, where $x_{\alpha}$ is the (approximate) $\alpha$ quantile of  $B_{K, 1/m}$.
+\green{Is this Binom????}
 \end{dfn}
 
-\green{Why are we including $x_{1-\alpha}$ in both sides?}
-\red{fixed  now-- $X < x_{1 - \alpha}$ is the acceptance region, $X \ge x_{1-\alpha}$ is rejection region.}
-
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, finding an exact $x_{1-\alpha}$.
+\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, finding an exact $x_{\alpha}$.
 For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.
+It can be set to be smaller, even arbitrarily small, by increasing $K$.
 
 \green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text.} \blue{its jasa guideline to use fraction like $1/m$ whenever possible to save space}.
+\green{so please make sure that they are all $1/m$.}
 
@@
 
-\green{Why???? Under the alternative hypothesis, we have to account for different visual abilities of  individuals:} 
-\red{I think that's more a matter of writing it - sorry for the confusion - }
-
-let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise, \red{where $p_i$ is the probability that individual $i$ picks out the data plot from the lineup.
+\green{What does this part apply to ????
+power?}let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise, \red{where $p_i$ is the probability that individual $i$ picks out the data plot from the lineup.
 If all individuals have the same probability $p$ for picking out the data plot, $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$, and we can estimate $p$ as $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the data plot from the lineup.
@@
        \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1.}
-       \label{fig:power_loess1}
+       \label{fig:power_effect1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.7}{\includegraphics{effect_power_subject_exp1.pdf}}
+       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1.}
+       \label{fig:power_effect_subject1}
 \end{figure*}

### 2012-05-29 paper_jasa.tex
power with effect for experiment 1and 2

@@
    \centering
-       \scalebox{0.7}{\includegraphics{effect_power_exp1.pdf}}
-       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1.}
+       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
+       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
        \label{fig:power_effect1}
@@
    \centering
-       \scalebox{0.7}{\includegraphics{effect_power_subject_exp1.pdf}}
-       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1.}
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
+       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
        \label{fig:power_effect_subject1}

### 2012-05-29 R_codes/turk2_data_analysis_jasa.R

@@ -2,6 +2,8 @@
 # R script file last modified on Mar 29, 2012 by Mahbub
 
 library(ggplot2)
+library(plyr)
+library(reshape)
 
 
 raw.dat <- read.csv("../data/raw_data_turk2.csv")
@@ -513,13 +515,14 @@ p + xlab(expression(beta)) + ylab("Power")
 
 
 
+
 # ----------- glm with effect as covariate ---------------
 
-dat$effect <-  with(dat,beta/sigma*sqrt(sample_size))
+dat$effect <-  abs(with(dat,beta/sigma*sqrt(sample_size)))
 
 effect.d <- ddply(dat,.(effect, replica), summarize, 
    power = mean(response),
-   ump_power = calculate_ump_power(beta[1],sample_size[1],sigma[1]))
+   ump_power = calculate_ump_power(abs(beta[1]),sample_size[1],sigma[1]))
 
 ggplot(data=effect.d,aes(x=effect,y=power)) + 
    geom_point() + 
@@ -535,13 +538,77 @@ ggplot(data=effect.d,aes(x=effect,y=power)) +
    
 ggplot(data=effect.d,aes(x=effect,y=power)) + 
    geom_point() + 
-   geom_smooth(method="loess", span=0.8) +
-   geom_line(aes(x=effect,y=ump_power))  
+   geom_smooth(method="loess", span=0.8,aes(colour=factor(replica))) +
+   geom_line(aes(x=effect,y=ump_power)) +
+   scale_colour_discrete(name = "Replication")+
+   xlab(expression(Effect(E)))
+   
+ggsave(filename="../images/effect_power_exp2.pdf", height=7, width=7)   
+   
+ggplot(data=effect.d,aes(x=effect,y=power-ump_power)) + 
+   geom_point() + 
+   geom_smooth(method="loess", span=0.8,aes(colour=factor(replica))) +
+   scale_colour_discrete(name = "Replication")  
+   
+     
 
 
 
 # and then fit the following model:
        response ~ effect + (effect | id)
+
+# ------- subjectwise power curve with random slope  for covariate effect 
+library(lme4)
+fit.mixed <- lmer(response ~ effect + (1+effect|id)
+              , family="binomial"
+              , data=dat)
+res <- summary(fit.mixed)
[…]

### 2012-05-29 R_codes/turk1.Rnw
lmer with offset

<<read-data, echo=FALSE, resuls=hide>>=

suppressMessages(library(ggplot2))
setwd("/Users/heike/papers/paper-inference-jasa/R_codes/")
raw.dat <- read.csv("../data/raw_data_turk1.csv")
@

# ========== policy for excluding bad data =========
# Everyone is given at least one easy lineup. If a person can't correctly
# evaluate 50\% of those extremely easy lineups, all his responses are discarded.
# A lineup with p_value < 0.0002 is considered easy

# where are our reference charts?
# we should not draw inference on charts that we use for study inclusion

<<select, dependson=read-data, echo=FALSE, resuls=hide>>=
raw.dat$response <- raw.dat$response_no == raw.dat$plot_location

raw.dat$effect <- with(raw.dat, beta/sigma*sqrt(sample_size))

raw.dat$gender <- factor(raw.dat$gender, exclude=NULL)
levels(raw.dat$gender) <- c("male", "female", "unknown")
raw.dat$mage <- is.na(raw.dat$age)
raw.dat$meduc <- is.na(raw.dat$academic_study)

raw.dat$academic_study <- addNA(factor(raw.dat$academic_study))
raw.dat$age <- addNA(factor(raw.dat$age))

raw.dat$mgender <- raw.dat$gender=="unknown"

#dp <- ddply(subset(raw.dat, p_value < 0.0002),.(id),summarise,
#   percent_correct = mean(response)*100
#  )

#included_id <- dp$id[ dp$percent_correct > 0]
#dat <- subset(raw.dat, id %in% included_id)
dat <- raw.dat

suppressMessages(require(plyr))
turker <- ddply(dat,.(id), summarise,
	n = length(id),
	correct = sum(response),
  gender=gender[1]
)

## if we include all correct percentages above 0.2 we get rid of any differences due to gender, age or schooling
included <- unique(turker$id[with(turker, correct/n >= 0.1)])

## if we exclude everybody with just a single lineup, we have the difference introduced by missing values show up massively
#included <- unique(turker$id[turker$n > 1])

## if we exclude everybody with just a single lineup, we have the difference introduced by missing values show up massively
#included <- unique(turker$id[turker$gender != "unknown"])

dat <- subset(raw.dat, id %in% included)

lineups <- ddply(dat,.(pic_id), summarise,
	n = length(id),
	correct = sum(response),
  p_value = mean(p_value),
  beta=beta[1],
  sigma=sigma[1],
  sample_size=sample_size[1],
  effect = effect[1]               
)
@

<<fig-basic, dependson=select, echo=FALSE>>=
qplot(effect, p_value, data=lineups, colour=correct/n)
qplot(effect, correct/n, data=lineups, colour=p_value) + geom_smooth(method="loess")
@

<<model0, dependson=select>>=
suppressMessages(library(lme4))
dat$setup <- with(dat, sprintf("%d:%d:%d", beta, sigma, sample_size))

m0 <- lmer(response~ 1 + (1|id) + (1|setup), data=dat, family=binomial())
q01 <- qplot(ranef(m0)[[1]][,1], binwidth=0.2)
q02 <- qplot(ranef(m0)[[2]][,1], binwidth=0.2)
summary(m0)
summary(ranef(m0)[[1]][,1])
summary(ranef(m0)[[2]][,1])

dat$offset <- log(0.05/.95)
base <- lmer(response~ offset(offset) - 1 + effect + (effect|id) , data=dat, family=binomial())
anova(m0, base)
q11 <- qplot(ranef(base)[[1]][,1], binwidth=0.2)
q12 <- qplot(ranef(base)[[2]][,1], binwidth=0.2)
summary(base)

base.1 <- lmer(response~ effect + (effect|id) , data=dat, family=binomial())

m1 <- lmer(response~ effect + (1|id), data=dat, family=binomial())
anova(base, m1)

m1a <- lmer(response~ effect + gender + (1|id), data=dat, family=binomial())
m1b <- lmer(response~ effect + academic_study + (1|id), data=dat, family=binomial())
m1c <- lmer(response~ effect + age + (1|id), data=dat, family=binomial())

m2a <- lmer(response~ effect + mgender + (1|id), data=dat, family=binomial())

anova(m0, base,  m2a, m1a)
anova(base, m1,  m2a, m1b)
anova(base, m2a, m1c)

dat$alpha <- 0.05
m3c <- lmer(response~ offset(alpha)+effect -1 + age + (effect|id), data=dat, family=binomial())
anova(base, m2a, m1c, m3c)

# there's something in model m1c that's not explained by missingness alone - but it might be an artefact - there's nothing immediately visible in the estimates
@

<<plots, dependson=model0, echo=FALSE, results='hide', out.width='\\text.width'>>=
qplot(ranef(m3c)[[1]][,1], binwidth=0.1) + xlab("Individuals' Skill")

qqnorm(ranef(m1c)[[1]][,1])
abline(a=0,b=1)

qplot(effect, fitted(base), data=dat, colour=difficulty,  group=factor(sample_size)) + geom_smooth()

qplot(effect, fitted(m1c), data=dat, colour=factor(sample_size))
qplot(group=effect, x=resid(m1c), geom="histogram", data=dat, facets=~effect, binwidth=0.2)
@

levels(dat$age) <- c("below 18", "18 - 25", "26 - 30", "31 - 35", "36 - 40", "41 - 45", "46- 50", "51 - 55", "56 - 60", "above 60")

# academic_study 1 = High school or less
                 2 = Some under graduate courses 
                 3 = Under graduate degree
                 4 = Some graduate courses
                 5 = Graduate degree

# conf_level 1 = most, 5 = least

# choice_reason 1 = Big vertical difference 
                2 = Medians are different
                3 = Outliers
                4 = others

# unit(time_taken) = second

<<predict, dependson=model0>>=
predict.lme <- function(model, newdata, skills=0) {
  design <- ldply(newdata$age, function(x) levels(model@frame$age) %in% x)
  xnew <- cbind(effect=newdata$effect, design)
  preds <- drop(as.matrix(xnew) %*% fixef(model)) + skills
  exp(preds)/(exp(preds)+1)
}

newdata <- data.frame(expand.grid(effect=seq(0,17, by=0.1), age=c(0:10, NA)))
newdata$m1c <- predict.lme(m3c, newdata=newdata)

newdata_all <- data.frame(expand.grid(effect=seq(0,17, by=0.1), age=c(0:10, NA), skills=ranef(m1c)[[1]][,1]))
newdata_all$m1c <- predict.lme(m1c, newdata=newdata_all, skills=ranef(m1c)[[1]][,1])

qplot(effect, m1c, data=newdata, colour=factor(age))
qplot(effect, m1c, data=newdata_all, colour=factor(age))
@

# ------function to calculate UMP power -----------------------

calculate_ump_power <- function (beta, n, sigma){
	alpha <- 0.05/2
	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs for derivation of power
	mu <- beta/se_beta
	t_n <- qt(p=1-alpha,df=n-3)
	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
	return(res)
}

calculate_ump_power(3,100,5)
calculate_ump_power(3,300,5)

calculate_ump_power(3/5,300,1)

# ---------------------- fitting loess smoother  ------------

fit <- loess(as.numeric(response) ~ beta,data=subset(dat,sigma==5 & sample_size==100),span=1.2)
betas <- seq(0,9,by=.1)
with(subset(dat,sigma==5 & sample_size==100),qplot(betas,predict(fit,data.frame(beta=betas))))

fit <- loess(as.numeric(response) ~ beta,data=subset(dat,sigma==5 & sample_size==300),span=.75)
betas <- seq(0,5,by=.1)
with(subset(dat,sigma==5 & sample_size==300),qplot(betas,predict(fit,data.frame(beta=betas))))

get_smooth_power <- function(dat.n,test="Empirical"){
 betas <- seq(0.01,16, by=.2)
 dat_smooth <- NULL
 for (s in c(5,12)){
	for (n in c(100,300)){
		dats <- subset(dat.n,sigma==s & sample_size==n)
		dats$y <- as.numeric(dats$response)
		fit <- loess(y ~ beta,data=dats,span=1+(n==100)*.2)
		dat_smooth <- rbind(dat_smooth,cbind(betas,predict(fit,betas),n,s))
		}
	}
 colnames(dat_smooth) <- c("beta","pow","sample_size","sigma")
 dat_empirical <- data.frame(rbind(dat_smooth,cbind(-dat_smooth[,1],dat_smooth[,-1])))
 dat_empirical$test <- test
 return(dat_empirical)
}

get_ump_power <- function(dat,test="UMP"){
 beta <- seq(0.01,16, by=.2)
 dat_pow <- NULL
 for(n in c(100,300)){
  for(sg in c(5,12)){
     pow <- NULL
     for (i in beta)pow <- c(pow,calculate_ump_power(beta=i,n=n,sigma=sg))
     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))
     dat_ump$sample_size <- n
     dat_ump$sigma <- sg
     #head(dat_ump)
     dat_pow <- rbind(dat_pow,dat_ump)
  }
 }
 dat_pow$test <- test	
 return(dat_pow)
}

dat_emp_pow <- get_smooth_power(dat)
dat_ump_pow <- get_ump_power(dat)

p <- ggplot(dat_emp_pow,aes(beta,pow)) + geom_line(aes(colour=test)) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power")
p

# ---------- bootstrap band for empirical power -----------------------

dat_boot_pow <- NULL
for (i in 1:1000){
dat.b <- ddply(dat,.(beta,sample_size,sigma), summarize,
           response = sample(response,replace=T)
           )
dat_boot_pow <- rbind(dat_boot_pow,get_smooth_power(dat.b,test=paste("smooth",i,sep="")))
}

#write.csv(dat_boot_pow,file="../data/dat_bootstrap_power.txt",row.names=F)
#dat_boot_pow <- read.csv("../data/dat_bootstrap_power.txt")

p <- ggplot(dat_boot_pow, aes(beta,pow,colour=grey)) + geom_line(aes(group=test),alpha=I(0.1))
p + facet_grid(sample_size~sigma)

p <- ggplot() +
     geom_line(aes(beta,pow, group =test), data=dat_boot_pow,alpha=I(0.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_boot_limit <-  ddply(dat_boot_pow,.(beta,sample_size,sigma), summarize,
           limit1 = quantile(pow,.025, na.rm=T),
           limit2 = quantile(pow,.975, na.rm=T)
           )

dat_boot_pow_limit <- melt(dat_boot_limit, id=c("beta","sample_size","sigma"))
colnames(dat_boot_pow_limit) <- c("beta","sample_size","sigma","test","pow")

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,as.numeric(response)), data=dat) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_obs_val <- ddply(dat,.(beta,sample_size,sigma,response), summarize,
           responses = length(response)
           )
dat_obs_val <- rbind(dat_obs_val,cbind(beta= -dat_obs_val[,1],dat_obs_val[,-1]))

p <- ggplot() +
     geom_line(aes(beta,pow, group =test), data=dat_boot_pow,alpha=I(0.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_obs_pow <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,
           pow = sum(response=="TRUE")/sum(response==response)
           )
dat_obs_pow <- rbind(dat_obs_pow,cbind(beta= - dat_obs_pow[,1],dat_obs_pow[,-1]))

p <- ggplot() +
     geom_line(aes(beta,pow, group =test,linetype=5), data=dat_boot_pow_limit,colour=alpha("red",.3)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     geom_point(aes(beta,pow), data=dat_obs_pow, colour=alpha("red",.3)) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

dat_boot_ribbon <-  ddply(dat_boot_pow,.(beta,sample_size,sigma), summarize,
           limit1 = quantile(pow,.025, na.rm=T),
           limit2 = quantile(pow,.975, na.rm=T)
           )
dat_boot_ribbon <- rbind(dat_boot_ribbon,cbind(beta= -dat_boot_ribbon[,1],dat_boot_ribbon[,-1]))

p <- ggplot() +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val, colour=alpha("red",.3)) +
     geom_ribbon(aes(x=beta,ymin=limit1,ymax=limit2), data=dat_boot_ribbon, fill=alpha("red",.2)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

p <- ggplot() +
     geom_point(aes(beta,as.numeric(response), size=responses), data=dat_obs_val, colour=alpha("black",.3)) +
     geom_ribbon(aes(x=beta,ymin=limit1,ymax=limit2), data=dat_boot_ribbon, fill=alpha("grey",.9)) +
     geom_line(aes(beta,pow, colour=test), data=dat_emp_pow) +
     geom_line(aes(beta,pow, colour=test), data=dat_ump_pow) +
     facet_grid(sample_size~sigma) +
     xlab(expression(beta))+ylab("Power") 
p

ggsave(p, filename="../images/power_loess_exp1.pdf", height=5.5, width=8.5)

# --------------- p_value vs %correct ----------------------------------

pdat <- ddply(dat, .(pic_name), summarize,
   p_value = mean(p_value),
	attempted = sum(response==response),
	corrected = sum(response=="TRUE"),
	percent_correct = sum(response=="TRUE")*100/sum(response==response)
	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat) +
     geom_point(aes(p_value,percent_correct)) + 
     xlab(expression(paste("p-value(",p[B],")"))) +
     ylab("Percentage of correct responses") 
p

ggsave("../images/p_val_percent_correct.pdf")

pdat$lbl <-""
indx <- with(pdat,(percent_correct>30 & p_value>.05) | (percent_correct>5 & p_value>.5))
pdat$lbl[indx] <- as.character(pdat$pic_name[indx])
pdat$angle <- 0
pdat$angle[pdat$percent_correct>5 & pdat$p_value>.5] <- 90

p <- ggplot(pdat,aes(p_value,percent_correct,label=lbl)) +
     geom_point() + geom_text(hjust=-.05,vjust=0,aes(angle=angle)) +
     xlab(expression(paste("p-value(",p[B],")"))) +
     ylab("Percentage of correct responses") +
     geom_vline(xintercept=0.05) + geom_hline(yintercept=5) +geom_smooth()
p

subset(dat,pic_name=="plot_turk1_100_1_5_1.png")
subset(dat,pic_name=="plot_turk1_100_0_5_1.png")
subset(dat,pic_name=="plot_turk1_300_1_12_1.png")

subset(dat,pic_name=="plot_turk1_100_1_5_3.png")
subset(dat,pic_name=="plot_turk1_100_1_12_3.png")

# ------------- observed and theorectical power curves ------------------------------

calculate_theoretical_power <- function (beta, n, sigma){

	# n <- 100

	# sigmasq <- 12^2

	alpha <- 0.05

	sigmasq <- sigma^2

	beta_not <- 0

	se_beta <- sqrt(sigmasq/(n*(.5^2)))    # refer to docs derivation of power

	mu <- beta/se_beta

	alpha <- alpha/2

	t_n <- qt(p=1-alpha,df=n-3)

	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1

	return(res)

}

exact_ci <- function(yy, nn, alpha=0.05) {
	pL <- rep(0,length(yy))
	pU <- rep(1,length(yy))

	y <- yy[yy!=0]
	n <- nn[yy!=0]
	pL[yy!=0] <- 1/(1+ (n-y+1)/(y*qf(alpha/2, 2*y, 2*(n-y+1))))

	y <- yy[yy!=nn]
	n <- nn[yy!=nn]
	pU[yy!=nn] <- 1/(1+ (n-y)/((y+1)*qf(1-alpha/2, 2*(y+1), 2*(n-y))))
	pw <- yy/nn
	return(data.frame(pw,pL, pU))
}
exact_ci(yy=c(3,4),nn=c(5,7))

dat_obs <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,
           empirical_power = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,1]
           )
dat_obs1 <- dat_obs

power_obs_ump <- cbind(dat_obs[,1:4],ump=sapply(dat_obs[,1],calculate_theoretical_power ,n=100,sigma=5))
#setwd("U:/phd research/simulation_study/turk")
#write.table(power_obs_ump,file="power_obs_ump_exp1.txt",row.names=F)

dat_obs1$beta <- -dat_obs1$beta
dat_obs <- rbind(dat_obs1,dat_obs)
dat_obs.m <- melt(dat_obs, id=c("beta","sample_size","sigma", "replica"))
colnames(dat_obs.m) <- c("beta","sample_size","sigma", "replica","test","pow")
head(dat_obs.m)

p <- qplot(beta,pow, geom=c("point","line"), data=dat_obs.m, colour=factor(replica)) 
p + facet_grid(sample_size~sigma)

p <- qplot(beta,pow, geom=c("smooth","point"), data=dat_obs.m) 
p + facet_grid(sample_size~sigma)

p <- ggplot(dat_obs.m, aes(beta,pow)) + geom_point()+geom_smooth()
p + facet_grid(sample_size~sigma)

=c("smooth","point"), data=dat_obs.m) 
p + facet_grid(sample_size~sigma)

beta <- seq(0.01,16, by=.2)
dat_pow <- NULL
for(n in c(100,300)){
  for(sg in c(5,12)){
     pow <- NULL
     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))
     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))
     dat_ump$sample_size <- n
     dat_ump$sigma <- sg
     #head(dat_ump)
     dat_pow <- rbind(dat_pow,dat_ump)
}}
dat_pow$test <- "UMP"
head(dat_pow)

plot_dat <- rbind(dat_obs.m,dat_pow[,c(1,3,4,5,2)])
head(plot_dat)
tail(plot_dat)

p <- qplot(beta,pow, geom="line", data=plot_dat, colour=test, linetype=test) 
p <- p + facet_grid(sample_size~sigma)
p <- p + scale_colour_manual(values=c("red","red","red","cyan4"))
p <- p + scale_linetype_manual(values=c(1,2,2,1))
p + xlab(expression(beta)) + ylab("Power")

p <- ggplot(dat_obs.m, aes(beta,pow, colour=test)) + geom_point(alpha=0.5)+geom_smooth()
p <- p + geom_line(aes(beta,pow, colour=test), data=dat_pow)
p <- p + facet_grid(sample_size~sigma)
p <- p + xlab(expression(beta)) + ylab("Power")

ggsave(p, filename="../images/emperical_ump_power_exp1.pdf", height=5.5, width=8.5)

# ----------------- glm power curve ----------------------------------

my_glm_predict <- function(dat.new){

lb <- function(b){

   y <- as.numeric(dat$response)-1

   b0 <- log(0.05/(1-0.05))

   x <- with(dat,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma, x4=beta*sample_size*sigma))

   xb <- b0 + x%*%b

   p <- exp(xb)/(1+exp(xb))

   return(-sum(y*log(p)+(1-y)*log(1-p)))

}

#lb(c(.2,.002,.02,.0002))

bh <- optim(par=c(.2,.002,.02,.0002),lb)$par

xnew <- with(dat.new,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma, x4=beta*sample_size*sigma))

xbh <- b0 + xnew%*%bh

ph <- exp(xbh)/(1+exp(xbh))

return(ph)

}

beta <- rep(seq(.1,16, by=.25),4)

sample.size <- c(rep(100,length(beta)/2),rep(300,length(beta)/2))

sigma <- rep(c(12,5,12,5),each=length(beta)/4)

dat.new <- data.frame(beta, sample_size=sample.size,sigma)

power <- my_glm_predict(dat.new)

vdat <- data.frame(beta=c(beta,-beta),power=c(power,power),sample.size=rep(sample.size,2),sigma=c(sigma,sigma))

vdat$Test="Visual"

qplot(beta,power, geom="line", data=vdat)+facet_grid(sample.size~sigma)

qplot(beta,power, geom="line", data=subset(vdat,(sigma==12)&(sample.size==100)))

udat <- NULL

for(n in c(100,300)){

  for(sg in c(5,12)){

     pow <- NULL

     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))

     dat_ump <- data.frame(beta=c(-beta,beta),power=c(pow,pow))

     dat_ump$sample.size <- n

     dat_ump$sigma <- sg

     #head(dat_ump)

     udat <- rbind(udat,dat_ump)

}}

udat$Test <- "UMP"

dat.pow <- rbind(vdat,udat)

qplot(beta,power, geom="line", data=dat.pow, colour=Test)+facet_grid(sample.size~sigma)

# ------------------ power curve using glm function -----------------------

fit.power <- glm(response ~ sample_size+beta+sigma

             , family=binomial,data=dat)

res <- summary(fit.power)

str(res)

res$coef

beta <- seq(.1,16, by=.25)

sample.size <- rep(100,length(beta))

sigma <- rep(12,length(beta))

newdat <- data.frame(sample_size=sample.size,sigma,beta)

power <- predict(fit.power, newdata = newdat, type="response", se.fit = TRUE)

pr <- predict(fit.power, newdata = newdat, se.fit = TRUE)

family <- family(fit.power)

lower <- family$linkinv(pr$fit - qnorm(0.95) * pr$se.fit)

upper <- family$linkinv(pr$fit + qnorm(0.95) * pr$se.fit)

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , empirical=rep(power$fit,2)

                    , lower_limit=rep(lower,2),upper_limit=rep(upper,2)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

colnames(pow.dat.m) <- c("betas","Test","power")

p <- qplot(betas, power, geom="line", data=pow.dat.m, color=Test, linetype=Test)

p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(1,2,2,1))

# ---------------- estimating power curves for each replicate of linup ----------------

my_glm_predict.rep <- function(dat.new){

  b0 <- log(0.05/(1-0.05))

  lb <- function(b){

     y <- as.numeric(dat$response)-1

     x <- with(dat,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma,

             x4=replica*beta,x5=beta*sample_size*sigma*replica))

     xb <- b0 + x%*%b

     p <- exp(xb)/(1+exp(xb))

     return(-sum(y*log(p)+(1-y)*log(1-p)))

  }

  #lb(c(.2,.007,.002,-.0002,-.002))

  bh <- optim(par=c(.2,.007,.002,-.0002,-.002),lb)$par

  # 0.2162250624  0.0056327718 -0.0174778836  0.0773441682 -0.0001383814

  xnew <- with(dat.new,cbind(x1=beta,x2=beta*sample_size, x3=beta*sigma,

             x4=replica*beta,x5=beta*sample_size*sigma*replica))

  xbh <- b0 + xnew%*%bh

  ph <- exp(xbh)/(1+exp(xbh))

  return(ph)

}

# ----- investigating ------

matrix(my_glm_predict.rep(subset(dat.new, sample_size==100 & sigma==5)),ncol=3)

# ----- end investigation -------

beta <- seq(.1,16, by=.25)

sample.size <- c(100,300)

sigma <- c(12,5)

replica <- 1:3

dat.new <- expand.grid(beta,sample.size,sigma,replica)

colnames(dat.new) <- c("beta","sample_size","sigma","replica")

dat.new$power <- my_glm_predict.rep(dat.new)

vdat <- with(dat.new,data.frame(beta=c(beta,-beta),power=c(power,power),

        sample.size=rep(sample_size,2),sigma=c(sigma,sigma), replica=c(replica,replica)))

vdat$Test=paste("Replicate",vdat$replica,sep="")

qplot(beta,power, geom="line", data=vdat, colour=Test)+facet_grid(sample.size~sigma)

qplot(beta,power, geom="line",colour=Test,data=subset(vdat,(sigma==5)&(sample.size==100)))

dat.X <- expand.grid(beta,sample.size,sigma)

colnames(dat.X) <- c("beta","sample.size","sigma")

upow <- apply(dat.X,1,function(x){return(calculate_theoretical_power(beta=x[1],n=x[2],sigma=x[3]))})

udat <- data.frame(beta=c(dat.X$beta, -dat.X$beta),

                  power=c(upow,upow), rbind(dat.X[,-1],dat.X[,-1]),

                  replica=0, Test="UMP")

dat.pow <- rbind(vdat,udat)

p <- qplot(beta,power, geom="line", data=dat.pow, colour=Test, linetype=Test)

p <- p + facet_grid(sample.size~sigma)

p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

dat_obs <- ddply(dat,.(beta,sample_size,sigma,replica), summarize,

	power=sum(response=="TRUE")/length(response)

   )

dat_obs$sample.size=dat_obs$sample_size

dat_obs$Test <- paste("Replicate",dat_obs$replica, sep="")

ggplot()+

geom_line(data=dat.pow,aes(x=beta,y=power,colour=Test, linetype=Test)) +

geom_point(data=dat_obs,aes(x=beta,y=power,colour=Test)) +

facet_grid(sample.size~sigma) +

xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

dat.pow.diff <- vdat

dat.pow.diff$power <- rep(udat$power,3)- vdat$power

p <- qplot(beta,power, geom="line", data=subset(dat.pow.diff,beta>=0), colour=Test, linetype=Test)

p <- p + facet_grid(sample.size~sigma)

p + xlab(expression(beta)) + ylab("Difference from UMP test Power") + scale_linetype_manual(values =c(2,3,4,1))

# ---------------- estimating power curves for each replicate of linup using glm ----------------

dat$bf <- dat$beta*dat$replica

fit.power.rep <- glm(response ~ sample_size+sigma+bf

             , family=binomial,data=dat)

res <- summary(fit.power.rep)

str(res)

res$coef

dat.pw <- NULL

for ( nn in c(100,300)){

for (sg in c(5,12)){

beta <- seq(.1,16, by=.25)

nb <- length(beta)

sample.size <- rep(nn,nb)

sigma <- rep(sg,nb)

newdat <- data.frame(sample_size=sample.size,sigma)

newdat <- rbind(newdat,newdat,newdat)

newdat$bf  <- rep(beta,3)* rep(1:3, each=nb)

power.rep <- predict(fit.power.rep, newdata = newdat, type="response", se.fit = TRUE)

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=nn,sigma=sg))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , replicate1=rep(power.rep$fit[1:nb],2)

                    , replicate2=rep(power.rep$fit[(nb+1):(nb*2)],2)

                    , replicate3=rep(power.rep$fit[(2*nb+1):(nb*3)],2)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

colnames(pow.dat.m) <- c("betas","Test","power")

#p <- qplot(betas, power, geom="line", data=pow.dat.m, color=Test, linetype=Test)

#p + xlab(expression(beta)) + ylab("Power") + scale_linetype_manual(values =c(2,3,4,1))

replicate1 <- ump_pow-power.rep$fit[1:nb]

replicate2 <- ump_pow-power.rep$fit[(nb+1):(nb*2)]

replicate3 <- ump_pow-power.rep$fit[(2*nb+1):(nb*3)]

overall <- ump_pow-power$fit

pow.diff <- data.frame(beta,replicate1,replicate2,replicate3,overall)

pow.diff.m <- melt(pow.diff,id="beta")

colnames(pow.diff.m) <- c("betas","Replicate","power")

#p <- qplot(betas, power, geom="line", data=pow.diff.m, color=Replicate, linetype=Replicate)

#p + xlab(expression(beta)) + ylab("Difference from UMP Power") + scale_linetype_manual(values =c(4,3,2,1))

dat.pw <- rbind(dat.pw,cbind(pow.diff.m,sample_size=nn,sigma=sg))

}}

p <- qplot(betas, power, geom="line", data=dat.pw, color=Replicate, linetype=Replicate)

p <- p + xlab(expression(beta)) + ylab("Difference from UMP Power")

p + scale_linetype_manual(values =c(4,3,2,1)) + facet_grid(sample_size~sigma)

# ----------------- estimating subjectwise estimated power curve -------------------------

library(lme4)

fit.mixed <- lmer(response ~ factor(sample_size)+beta+factor(sigma)

              + (1|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#res@frame

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

X <- cbind(rep(1,length(beta)),sample.size,beta,sigma.val)

dim(X)

delta <- as.numeric(res@REmat[4])

power=NULL

M <- 30

for (i in 1:M){

tau <- rnorm(mean=0,sd=delta, n=1)

xb <- X %*% B + tau

power <- cbind(power,exp(xb)/(1+exp(xb)))

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=I(alpha("black", 0.1)), data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="hotpink", size=1)

# ----------------- subjectwise power curve with random slope ---------------

library(lme4)

fit.mixed <- lmer(response ~ factor(sample_size)+beta+factor(sigma)

              + (1+beta|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#res@frame

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

ones <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

X <- cbind(ones,sample.size,beta,sigma.val)

dim(X)

Z <- cbind(ones,beta)

cov <- as.numeric(res@REmat[2,5])*as.numeric(res@REmat[1,4])*as.numeric(res@REmat[2,4])

vv <- matrix(as.numeric(c(res@REmat[1,3],rep(cov,2),res@REmat[2,3])),ncol=2, byrow=F)

power=NULL

M <- 20

#library(MASS)

for (i in 1:M){

tau <- mvrnorm(n = 1, mu=c(0,0),Sigma=vv)

xb <- X %*% B + Z %*%tau

power <- cbind(power,exp(xb)/(1+exp(xb)))

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

pow.dat.m <- melt(pow.dat, id="betas")

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=I(alpha("grey", 1)), data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="hotpink", size=1)

# ----------------- estimating subjectwise power curve by gender etc.------------------

library(lme4)

fit.mixed <- lmer(response ~ gender+factor(sample_size)+beta+factor(sigma)

              + (1|id)

              , family="binomial"

              , data=dat)

res <- summary(fit.mixed)

B <- res@coefs[,1]

beta <- seq(0.01,16, by=.2)

#head(res@frame)

sample.size <- rep(0,length(beta))

sigma.val <- rep(1,length(beta))

#conf_level <- rep(1,length(beta))

delta <- as.numeric(res@REmat[4])

power=NULL

gnd <- NULL

tau <- res@ranef    # estimates for existing subject

M <- length(tau)

for (i in 1:M){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,length(beta))

  X <- cbind(rep(1,length(beta)),gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + tau[i]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- 1:M

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1:M, each=length(beta)*2)],rep(1,length(beta)*2)), labels=c("Male","Female"))

head(pow.dat.m)

tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="maroon4", size=1)

# ----------------- estimating subjectwise power curve by gender and random slope------------------

library(lme4)

fit.mixed <- glmer(response ~ gender+factor(sample_size)+beta+factor(sigma)

              + (1 + beta|id)

              , family=binomial

              , data=dat)

res <- summary(fit.mixed)

B <- fixef(fit.mixed)

u <- as.matrix(ranef(fit.mixed)$id)

head(u)

beta <- seq(0.01,16, by=.2)

nb <- length(beta)

sample.size <- rep(1,nb)

sigma.val <- rep(0,nb)

ones <- rep(1,nb)

#Z <- cbind(ones,beta,sample.size,sigma.val)

#Z <- cbind(ones,beta,sigma.val)

Z <- cbind(ones,beta)

power=NULL

gnd <- NULL

M <- length(u[,1])

s <- sample(1:M,size=20)

#s <- c(213,8)

for (i in s){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,nb)

  X <- cbind(ones,gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + Z %*% u[i,]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- s

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

#head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1:length(s), each=nb*2)],rep(1,nb*2)), labels=c("Male","Female"))

#head(pow.dat.m)

#tail(pow.dat.m)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="maroon4", size=1)

p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="darkviolet", size=1)

# ----------------- individual observed and estimated power curve by gender and random slope----------

power=NULL

gnd <- NULL

M <- length(u[,1])

s <- sample(1:M,size=1)

for (i in s){

  #tau <- rnorm(mean=0,sd=delta, n=1) # estimates for new subject

  gender.val <- rep(as.numeric(dat$gender[dat$id==i][1])-1,nb)

  X <- cbind(ones,gender.val,sample.size,beta,sigma.val)

  #dim(X)

  xb <- X %*% B + Z %*% u[i,]

  power <- cbind(power,exp(xb)/(1+exp(xb)))

  gnd <- c(gnd,gender.val[1])

}

colnames(power) <- s

ump_pow <- NULL

for (i in beta)ump_pow <- c(ump_pow,calculate_theoretical_power(beta=i,n=100,sigma=12))

pow.dat <- data.frame(betas = c(-beta, beta)

                    , rbind(power,power)

                    , UMP=rep(ump_pow,2))

#head(pow.dat)

pow.dat.m <- melt(pow.dat, id="betas")

pow.dat.m$gender <- factor(c(gnd[rep(1, each=nb*2)],rep(1,nb*2)), labels=c("Male","Female"))

#head(pow.dat.m)

#tail(pow.dat.m)

dat_s=as.data.frame(subset(dat,id==s))

dat_s$variable=1

dat_pi <- merge(dat_s,ddply(dat_s,.(beta),summarize, obs_pi = mean(as.numeric(response)-1) ), by.beta=by.beta)

p <- qplot(betas,value,group=variable,geom="line",colour=gender, data=pow.dat.m)

p <- p + xlab(expression(beta)) + ylab("Power")

p <- p + geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="darkviolet", size=1)

p <- p + geom_point(aes(beta,as.numeric(response)-1), data=dat_s)

p + geom_point(aes(beta,obs_pi, colour="Observed"), data=dat_pi)

# ------------- observed and theorectical power curves ------------------------------

exact_ci <- function(yy, nn, alpha=0.05) {

	pL <- rep(0,length(yy))

	pU <- rep(1,length(yy))

	y <- yy[yy!=0]

	n <- nn[yy!=0]

	pL[yy!=0] <- 1/(1+ (n-y+1)/(y*qf(alpha/2, 2*y, 2*(n-y+1))))

	y <- yy[yy!=nn]

	n <- nn[yy!=nn]

	pU[yy!=nn] <- 1/(1+ (n-y)/((y+1)*qf(1-alpha/2, 2*(y+1), 2*(n-y))))

	pw <- yy/nn

	return(data.frame(pw,pL, pU))

}

exact_ci(yy=c(3,4),nn=c(5,7))

dat_obs <- ddply(dat,.(beta,sample_size,sigma), summarize,

               empirical = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,1],

               lower_limit = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,2],

               upper_limit = exact_ci(yy=sum(response=="TRUE"),nn= sum(response==response))[,3]

              )

dat_obs1 <- dat_obs

power_obs_ump <- cbind(dat_obs[,1:4],ump=sapply(dat_obs[,1],calculate_theoretical_power ,n=100,sigma=5))

#setwd("U:/phd research/simulation_study/turk")

#write.table(power_obs_ump,file="power_obs_ump_exp1.txt",row.names=F)

dat_obs1$beta <- -dat_obs1$beta

dat_obs <- rbind(dat_obs1,dat_obs)

dat_obs.m <- melt(dat_obs, id=c("beta","sample_size","sigma"))

colnames(dat_obs.m) <- c("beta","sample_size","sigma","test","pow")

head(dat_obs.m)

qplot(beta,pow, geom="line", data=dat_obs.m, colour=test) + facet_grid(sample_size~sigma)

beta <- seq(0.01,16, by=.2)

dat_pow <- NULL

for(n in c(100,300)){

  for(sg in c(5,12)){

     pow <- NULL

     for (i in beta)pow <- c(pow,calculate_theoretical_power(beta=i,n=n,sigma=sg))

     dat_ump <- data.frame(beta=c(-beta,beta),pow=c(pow,pow))

     dat_ump$sample_size <- n

     dat_ump$sigma <- sg

     #head(dat_ump)

     dat_pow <- rbind(dat_pow,dat_ump)

}}

dat_pow$test <- "UMP"

head(dat_pow)

plot_dat <- rbind(dat_obs.m,dat_pow[,c(1,3,4,5,2)])

head(plot_dat)

tail(plot_dat)

p <- qplot(beta,pow, geom="line", data=plot_dat, colour=test, linetype=test)

p <- p + facet_grid(sample_size~sigma)

p <- p + scale_colour_manual(values=c("red","red","red","cyan4"))

p <- p + scale_linetype_manual(values=c(1,2,2,1))

p + xlab(expression(beta)) + ylab("Power")

# --------------- p_value vs time_taken ------------------------------

ggplot(dat,aes((p_value)^.2,(time_taken)^.1))+geom_point()+geom_smooth()

p <- ggplot(dat, aes(factor(p_value), time_taken))

p <- p + geom_boxplot() +ylim(0,450) + facet_grid(gender~.)

p

pdat1 <- ddply(dat, .(p_value, sample_size), summarize,

	avg_time_taken = median(time_taken),

	num_responded = sum(p_value==p_value),

      nn = length(p_value)

    )

p <- ggplot(subset(pdat1, p_value<= 0.2), aes(p_value,avg_time_taken))

p <- p+ geom_point(aes(size=nn))

p <- p + stat_smooth(se=F)

p + facet_grid(sample_size~.)

# --------------- Comparison of p_values from theory and actual ----------------------

pval <- function(x){

	N <- x[1]

	u <- x[2]

	k <- u:N

	binom <- choose(N,k)*((0.05)^k)*((0.95)^(N-k))

	return(sum(binom))

}

pval(c(13,3))

pdat <- ddply(dat, .(p_value, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

    )

pval_calculated <- apply(pdat[,3:4],1,pval)

qplot(pdat$p_value,pval_calculated, xlab="p_value of Mathematical Test", ylab="p_value of Visual Test")

ddply(dat, .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(subset(dat, pic_id==92), .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(subset(dat, substr(pic_name,1,12)=='plot_100_3_5'), .(gender), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

ddply(dat, .(pic), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# --------------- p_value vs %correct ----------------------------------

pdat <- ddply(dat, .(p_value, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat, aes(p_value,percent_correct))

p+ geom_point(size=2)

ggsave("../images/p_val_percent_correct.pdf")

p <- ggplot(pdat, aes(p_value,percent_correct))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(se=F)

p + facet_grid(sample_size~.)

p <- ggplot(subset(pdat, p_value <= 0.099), aes(p_value,percent_correct))

p <- p+ geom_point(size=3)

p <- p + stat_smooth( se=F)

p + facet_grid(sample_size~.)

# --------------- distribution of time taken ----------------------------------

p <- ggplot(dat, aes(time_taken))+geom_histogram(binwidth=2)+xlim(0,350)

p <- p + facet_grid(gender~.)

p

# --------------- p_value vs %correct by male female ----------------------------------

pdat <- ddply(dat, .(p_value, gender, sample_size), summarize,

	attempted = sum(response==response),

	corrected = sum(response=="TRUE"),

	percent_correct = sum(response=="TRUE")*100/sum(response==response)

	)

# pdat$percent_correct <- pdat$corrected*100/pdat$attempted

p <- ggplot(pdat, aes(p_value,percent_correct, colour=gender, shape=gender))

p+ geom_point(size=3)

p <- ggplot(pdat, aes(factor(p_value),percent_correct, colour=gender, shape=gender))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p + facet_grid(sample_size~.)

p <- ggplot(subset(pdat, p_value <= 0.099), aes(factor(p_value),percent_correct, colour=gender, shape=gender))

p <- p+ geom_point(size=3)

p <- p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p + facet_grid(sample_size~.)

# ------------- number of attempts and % correct -----------

dat$attempt_no <- unlist(tapply(dat$start_time, dat$nick_name,

      function(x){return(as.integer(factor(x)))}), use.names=F)

adat <- ddply(dat, .(attempt_no, gender), summarize,

	attempted = sum(response_no==response_no),

	corrected = sum(response_no==obs_no),

      percent_correct = sum(response_no==obs_no)*100/sum(response_no==response_no)

    )

p <- ggplot(subset(adat, attempt_no <= 12), aes(attempt_no,percent_correct, color=factor(gender), shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)

# --------------- number of attempts vs time_taken ------------------------

adat1 <- ddply(dat, .(attempt_no, gender), summarize,

	avg_time_taken = mean(time_taken),

	avg_p_value = mean(p_value)

    )

p <- ggplot(subset(adat1, attempt_no <= 12), aes(attempt_no,avg_time_taken, color=gender, shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)

p <- ggplot(subset(adat1, attempt_no <= 12), aes(attempt_no,avg_p_value, color=gender, shape=gender))

p <- p + geom_point(size=3)

p + stat_smooth(aes(group=gender, linetype=gender), se=F)

### 2012-05-31 paper_jasa.tex
re-worked the explanation of errors and power

@@
  & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
- Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
+ Reject $H_0$ if & actual $T$ is extreme & actual plot is identifiable \\
 \hline 
@@
 For visual inference, unlike in the classical hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
-When the alternative hypothesis is true, it is expected that the plot of the observed data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
+When the alternative hypothesis is true, it is expected that the plot of the actual data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
 We will call a plot with this property a {\it visual statistic} for $\theta$.
@@
 A \textbf{visual test statistic}, $T(.)$, is a function of a sample that makes a plot.
-$T(y)$ maps the observed data to the plot, and we call this the \textbf{observed data plot}, and $T(y_0)$ maps samples drawn from the null distribution into the same plot form, and these are called \textbf{null plots}.
+$T(y)$ maps the actual data to the plot, and we call this the \textbf{actual data plot}, and $T(y_0)$ maps samples drawn from the null distribution into the same plot form, and these are called \textbf{null plots}.
 \end{dfn}
@@
 \item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
-\item the test statistic, $T(y)$, produced by plotting the observed data, possibly arising from $H_1$.
+\item the test statistic, $T(y)$, produced by plotting the actual data, possibly arising from $H_1$.
 \end{itemize}
@@
 \noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
-It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the observed data plot.
+It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
 
@@
 %\begin{itemize}\itemsep-3pt
-%\item \textbf{Reject} $H_0$ if `enough' observers select the observed data plot in the lineup of $m$ plots, and
+%\item \textbf{Reject} $H_0$ if `enough' observers select the actual data plot in the lineup of $m$ plots, and
 %\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
@@
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if out of $K$ observers $x_{\alpha}$ of them correctly identify the observed data plot in the lineup, and
+\item \textbf{Reject} $H_0$ if out of $K$ observers $x_{\alpha}$ of them correctly identify the actual data plot in the lineup, and
 \item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
@@
 The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
-% identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
-\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$, where $x_{\alpha}$ is the (approximate) $\alpha$ quantile of  $B_{K, 1/m}$.
-\green{Is this Binom????}
+% identifying the actual data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
+\textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
 \end{dfn}
 
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, finding an exact $x_{\alpha}$.
+\green{Oh, we are using observed and observer, which might be confusing. May need to change observed data plot to something else. Will think about this. Will use ``actual'' for now.}
+
+\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$.
 For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.
 It can be set to be smaller, even arbitrarily small, by increasing $K$.
-
-\green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text.} \blue{its jasa guideline to use fraction like $1/m$ whenever possible to save space}.
-\green{so please make sure that they are all $1/m$.}
-
-%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$.
-
-\green{What does this part apply to ????
-power?}let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise, \red{where $p_i$ is the probability that individual $i$ picks out the data plot from the lineup.
-If all individuals have the same probability $p$ for picking out the data plot, $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$, and we can estimate $p$ as $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the data plot from the lineup.
-
-If there is evidence for individual skills influencing the probability $p_i$ for picking out the data plot, 
- $X = \sum_i X_i$ has  a Poisson-Binomial distribution.}
+Type II error is harder to calculate, as is usually the case.
+In visual inference, individual abilities need to be incorporated to calculate Type II error.
+This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
+For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different.
+This involves understanding the individual's visual skills.
+Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup.
+If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$.
+Ee can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
+If there is evidence for individual skills influencing the probability $p_i$, then $X = \sum_i X_i$ will have a Poisson-Binomial distribution.
+This leads to the definition of the power of the visual test, as follows:
 
@@
 %\subsection{Power}
-
-\noindent 
-Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
-This is akin to different choices of statistics in classical hypothesis testing, for example, mean, median, or trimmed mean.
-One is typically better than another.
-For two different visual test statistics of the same observed data, the one is considered to be better, if $T(y)$ is more easily distinguishable to the observer.
-Power is typically used to measure this characteristic of a test.
-Power is the probability of rejecting $H_0$ when it is not true.
-We can assess power therefore both empirically based on experimental data and through theoretical considerations.
 
@@
 
-\noindent
-\red{Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.}
+\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
+
+Power is an important consideration in deciding which test to use for solving a problem. %In classical testing, new test procedures are examined for their power under a variety of scenarios. 
+For visual inference, this corresponds to the choice of plot to use.
+Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
+This is akin to different choices of statistics in classical hypothesis testing, for example, mean, median, or trimmed mean.
+One is typically better than another.
+For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer.
+Power is typically used to measure this characteristic of a test.
+%Power is the probability of rejecting $H_0$ when it is not true.
+We can assess power therefore both empirically based on experimental data and through theoretical considerations.
+
+\green{Done to here - the next parts are pretty long!}
 
@@
 Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the observed data, the resulting residual plot shows two groups of residuals.
+If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the actual data, the resulting residual plot shows two groups of residuals.
 To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic.
@@
 The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
-The test statistic, the plot containing the observed data, is randomly placed among these null plots.
+The test statistic, the plot containing the actual data, is randomly placed among these null plots.
 If the test statistic is identifiable the null hypothesis is rejected.
@@
 We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $N$ independent observers with $u$ identifications of the observed plot.
+Suppose, we have responses from $N$ independent observers with $u$ identifications of the actual plot.
 This gives an estimated power of
@@
 Three replicates of each level were generated.
-These produced 60 different ``observed data sets''.
+These produced 60 different ``actual data sets''.
 
@@
 
-For added control, to ensure a signal in the simulated observed data a blocking structure was used to filter data sets.
+For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
 A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
@@
 The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}.
-We arbitrarily fix $\beta_0 = 6$ while generating observed data.
-For each combination of parameter, we generate at least three different observed data sets constituting a total of 65 lineups.
+We arbitrarily fix $\beta_0 = 6$ while generating actual data.
+For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.
 
@@
 
-For this experiment we generate five replications of observed data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.
+For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.
 
@@
        \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 1.5 it is very unlikely to identify the observed plot.
+After p-value exceeds 1.5 it is very unlikely to identify the actual plot.
 The theoretical justification of this is shown in figure \ref{fig:pval_power}}

### 2012-05-31 paper_jasa.tex
Working on the lineup size and number of observers section

@@
 If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$.
-Ee can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
+We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
 If there is evidence for individual skills influencing the probability $p_i$, then $X = \sum_i X_i$ will have a Poisson-Binomial distribution.
-This leads to the definition of the power of the visual test, as follows:
+\green{WHY???
+And how does this relate to the logistic model below?} This leads to the definition of the power of the visual test, as follows:
 
@@
     \begin{equation*}
-      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{1-\alpha}) = 
+      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{\alpha}) = 
         \begin{cases} 
-               1 - \text{Binom}_{K, 1/m}(x_{1-\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
-              1 - F_{X, \theta} (x_{1-\alpha} - 1) & \text{ if } H_1 \text{ is true}.
+               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
+              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
         \end{cases}
@@
 
-\green{Done to here - the next parts are pretty long!}
+\section{Effect of Lineup Size and Number of Observers}
+
+\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
+Can someone please decide which to use, and make the appropriate changes?
+How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above?
+It looks like this is a repeat.
+One needs to go once notation is clarified.}
 
@@
 
-Let us make the assumption that an onlooker is able to identify  the chart corresponding to the data with the smallest $p$-value.
-Further we will assume that all onlookers  have the same ability in identifying the data plot.
+Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
+Further we will assume that all observers  have the same ability in identifying the data plot.
 
@@
 
-Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
+Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
 \[
@@
 \blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
-While this allows us to compute this value independently of the parametric setting, it also means that onlookers might pick out the data plot for a reason not related to the value of $\theta$.
+While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
 

### 2012-05-31 R_codes/turk1.Rnw
updated version with offsets

@@
 	n = length(id),
-	correct = sum(response),
+	correct = sum(response[p_value < 0.0002]),
+  gotchance = sum(p_value < 0.0002),  
+  pvalue = mean(p_value),  
+  pcorrect = sum(response)/length(id),
   gender=gender[1]
@@
 ## if we include all correct percentages above 0.2 we get rid of any differences due to gender, age or schooling
-included <- unique(turker$id[with(turker, correct/n >= 0.1)])
+
+## Mahbub
+# included <- unique(turker$id[with(turker, correct/gotchance > 0.5)])
 
@@
 dat$offset <- log(0.05/.95)
-base <- lmer(response~ offset(offset) - 1 + effect + (effect|id) , data=dat, family=binomial())
-anova(m0, base)
-q11 <- qplot(ranef(base)[[1]][,1], binwidth=0.2)
-q12 <- qplot(ranef(base)[[2]][,1], binwidth=0.2)
-summary(base)
-
-base.1 <- lmer(response~ effect + (effect|id) , data=dat, family=binomial())
-
-m1 <- lmer(response~ effect + (1|id), data=dat, family=binomial())
-anova(base, m1)
-
-m1a <- lmer(response~ effect + gender + (1|id), data=dat, family=binomial())
-m1b <- lmer(response~ effect + academic_study + (1|id), data=dat, family=binomial())
-m1c <- lmer(response~ effect + age + (1|id), data=dat, family=binomial())
-
-m2a <- lmer(response~ effect + mgender + (1|id), data=dat, family=binomial())
-
-anova(m0, base,  m2a, m1a)
-anova(base, m1,  m2a, m1b)
-anova(base, m2a, m1c)
-
-dat$alpha <- 0.05
-m3c <- lmer(response~ offset(alpha)+effect -1 + age + (effect|id), data=dat, family=binomial())
-anova(base, m2a, m1c, m3c)
-
+base <- lmer(response~ offset(offset) - 1 + effect + (effect-1|id) , data=dat, family=binomial())
+
+qqnorm(ranef(base)[[1]][,1])
+q11 <- qplot(ranef(base)[[1]][,1], geom="histogram", binwidth=0.002)
+q11
+
+base.2 <- lmer(response~ offset(offset) - 1 + effect + (effect-1|id) , data=dat, family=binomial(link="probit"))
+base.3 <- lmer(response~ offset(offset) - 1 + effect + (effect-1|id) , data=dat, family=binomial(link="cloglog"))
+# 
+anova(base.2, base.3, base)
+
+m1a <- lmer(response~  offset(offset) - 1 + effect:mgender + (effect-1|id), data=dat, family=binomial())
+m1b <- lmer(response~ offset(offset) - 1 + effect:academic_study + (effect-1|id), data=dat, family=binomial())
+m1c <- lmer(response~ offset(offset) - 1 + effect:age + (effect-1|id), data=dat, family=binomial())
+
+anova(base, m1a)
+anova(base, m1b)
+anova(base, m1c)
+
+qplot(effect, fitted(m1a), data=dat, colour=factor(academic_study))
 # there's something in model m1c that's not explained by missingness alone - but it might be an artefact - there's nothing immediately visible in the estimates

### 2012-05-31 paper_jasa.tex
model notation changed

@@
 
+\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $Y$ to indicate binomial random variable since it is a kind of response.
+\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-vaue.}
+
 \paragraph{Subject-specific abilities}
@@
 
-In the case that observers are exposed to multiple lineups we can estimate the probability $p_i$ while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
+In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{i \ell}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
 
-Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$ be the random variable, with $X_{j\ell}= 1$, if  subject $j$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $\pi_{j\ell}$ with a logistic regression with subject specific random effects as:
+Let $X_{i\ell} \sim Binom_{1, p_{i\ell}}$ be the random variable, with $X_{i\ell}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
+We can model $p_{i\ell}$ with a logistic regression with subject specific random effects as:
 \[
-\text{logit } \pi_{j\ell} = X_\ell \beta + Z u_j,  
+\text{logit } p_{i\ell} = X_\ell B + Z u_i,  
 \]
-where $X_\ell$ is a matrix of fixed effects with parameter vector $\beta$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $j = 1, ..., J$.
+where $X_\ell$ is a matrix of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $i = 1, ..., K$.
 

### 2012-05-31 paper_jasa.tex
model notation updated
notation suggestion

@@
 
-\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $Y$ to indicate binomial random variable since it is a kind of response.
-\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-vaue.}
+\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data.
+\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-value.
+\\ Should I change everything accordingly?
+}
 

### 2012-06-01 paper_jasa.tex
comment on power definition
model notation updated

@@
 
+\red{there may be a flow in this definition.
+Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power.
+I feel like we should not define power using binomial.}
+
 Power is an important consideration in deciding which test to use for solving a problem. %In classical testing, new test procedures are examined for their power under a variety of scenarios. 
@@
 
-In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{i \ell}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
+In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{ \ell i}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
 
-Let $X_{i\ell} \sim Binom_{1, p_{i\ell}}$ be the random variable, with $X_{i\ell}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $p_{i\ell}$ with a logistic regression with subject specific random effects as:
+Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
+We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
 \[
-\text{logit } p_{i\ell} = X_\ell B + Z u_i,  
+\text{logit } p_{\ell i} = X_{\ell i} B +  Z_{\ell i} \tau_i,  
 \]
-where $X_\ell$ is a matrix of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $i = 1, ..., K$.
+where $X_{\ell i}$ is a vector of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $\tau \sim MVN(0, \sigma_{\tau} I)$  is a vector of random variables $i = 1, ..., K$.
+
+\red{This model is elaborately described in equation \ref{mixed}.}
 
@@
 We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $N$ independent observers with $u$ identifications of the actual plot.
+Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
 This gives an estimated power of
-\begin{equation}\label{power_est} Power(\beta)=\frac{u}{N} \hspace{1cm} 0 \le u \le N \end{equation}
+\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
 
-Suppose each of $N$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{ij}$.
-Let $Y_{ij} = 1$, if subject $j$ correctly identifies the test statistic on lineup $i$, and 0 otherwise.
-We model $\pi_{ij} = E(Y_{ij})$ as  a mixed effects logistic regression
+Suppose each of $K$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{\ell i}$.
+Let $Y_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$, and 0 otherwise.
+We model $\pi_{\ell i} = E(Y_{\ell i})$ as  a mixed effects logistic regression
 
 \begin{equation} 
-g(\pi_{ij}) = X_{ij}B  + Z_{ij} \tau_j 
+g(\pi_{\ell i}) = X_{\ell i}B  + Z_{\ell i} \tau_i 
 \label{mixed} 
@@
 %
-where $\tau_j$ is random effects coefficient vector of length $q$ for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+where $\tau_i$ is random effects coefficient vector of length $q$ for subject $i$,  $\tau_i  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{\ell i}$ is the $\ell$th row vector of random effects covariates for subject $i$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{\ell i}$ is the $\ell$th row vector of the fixed effects covariates for subject $i$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
 The covariates could be demographic information of people such as age, gender, education level etc.

### 2012-06-01 paper_jasa.tex
Cleaning up experimental results section

@@
 
-\subsection{Experiment with a discrete covariate}
+Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
+There are three experiments.
+The first two are classical settings where we would not expect the lineup protocol to do better than the traditional test, but we hope it to perform favorably.
+The third experiment is a scenario where assumptions required for the traditional test are violated, and we would expect the lineup protocol to outperform the traditional test.
+
+\subsection{Discrete covariate}
 
@@
 
-\subsection{Experiment with a continuous covariate}
+\subsection{Continuous covariate}
 
@@
 
-\subsection{Experiment with contaminated data}
+\subsection{Contaminated data}
 

### 2012-06-01 paper_jasa.tex
Organized section 5, Simulation experiments. A lot more needs to be added to this section.

@@
 Which plot do you think is the most different from the others?
-Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
+(The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
 The null hypothesis typically derives from the task at hand, or the type of plot being made.
@@
 One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?}
+Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?
+(The position of the actual data plot is provided in Section \ref{sec:category}.)}
        \label{fig:test_category}
@@
         \begin{cases} 
-               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
+               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
               1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
@@
 
-\section{Application of Visual Inference in Linear Models} \label{sec:regression}
+\section{Application to Linear Models} \label{sec:regression}
 
@@
 
-\section{Results of Simulation Experiments} \label{sec:simulation}
+\section{Simulation Experiments} \label{sec:simulation}
 
@@
 The third experiment is a scenario where assumptions required for the traditional test are violated, and we would expect the lineup protocol to outperform the traditional test.
+(Data and lineups used in the experiments are available in the supplementary material.)
 
-\subsection{Discrete covariate}
+\subsection{Discrete covariate}\label{sec:category}
 
-The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \eqref{multi}).
-Data is simulated for different values of $\beta_2 (=0, 1, 3, 5, 7, 8, 10, 16)$, with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
-The set of $\beta_2$ values was chosen so that estimates of the power should produce reasonable power curves, comparable to the theoretical UMP test.
-We fixed the values of $\beta_0 = 5$,  $\beta_1=15$ and values for $X_1$ were generated as a random sample from a Poisson(30) distribution.
-Data sets with different combinations of $\beta_2$,  $n$ and $\sigma$ were generated with frequencies shown in Table \ref{tbl:experiment_params}.
-Three replicates of each level were generated.
+The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
+Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
+The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test.
+Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson(30) distribution.
+\green{Explain why this was used.} Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
+Three replicates at each level were generated.
 These produced 60 different ``actual data sets''.
 
+Each subject viewed 10 lineups.
+Selection of lineups shown to each subject followed these rules ....
+\green{fill this in}.
+Figure \ref{fig:test_category} is an example of one of these lineups.
+The visual test statistic is a side-by-side boxplot.
+
+\green{Did you in fact bin like this???} For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
+A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
+The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
+Additional control was applied to the 19 null plots.
+Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
+
 \begin{table}[hbtp]
-\caption{Values of parameters considered for three survey experiments} % title name of the table
+\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
 \centering
 % centering table
-\begin{tabular}{c c l l l}
+\begin{tabular}{c c c c c}
 % creating 10 columns
@@
 % inserting double-line
-& & \multicolumn{3}{c}{values for slope ($\beta$)} \\
+& & \multicolumn{3}{c}{Slope ($\beta$)} \\
  \cline{3-5}
 
-\raisebox{1.5ex}  {Sample size ($n$)} &   \raisebox{1.5ex} {$\sigma$} &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3}
+Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
+ ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
 \\ [0.5ex]
@@
 
-For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
-A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
-The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
-Additional control was applied to the 19 null plots.
-Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
-
 \subsection{Continuous covariate}
 
-This experiment is designed for the same regression model (Equation \eqref{multi}) with $p=1$ and covariate $X_1$ being continuous.
+This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous.
 Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
@@
 For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.
+
+\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
 
@@
 For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.
+
+\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
+Show a sample lineup, probably the most difficult, instead of the single plot.}
 
@@
 
+\subsection{Results}
+
 Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
@@
 \end{figure}
+
+\green{A lot more needs to be done here!!!}
 

### 2012-06-01 paper_jasa.tex
cleaned up section 4, needs some additions as indicated

@@
 Theoretical power for the regression parameters is derived in the next section.
+\subsection{Estimating Power from Lineups}
+
+We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
+Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
+This gives an estimated power of
+\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
+
+Suppose each of $K$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{\ell i}$.
+Let $Y_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$, and 0 otherwise.
+We model $\pi_{\ell i} = E(Y_{\ell i})$ as  a mixed effects logistic regression
+
+\begin{equation} 
+g(\pi_{\ell i}) = X_{\ell i}B  + Z_{\ell i} \tau_i 
+\label{mixed} 
+\end{equation}
+%
+where $\tau_i$ is random effects coefficient vector of length $q$ for subject $i$,  $\tau_i  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{\ell i}$ is the $\ell$th row vector of random effects covariates for subject $i$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{\ell i}$ is the $\ell$th row vector of the fixed effects covariates for subject $i$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+The covariates could be demographic information of people such as age, gender, education level etc.
+as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.
+
+From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
+the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
+%Thus we have the overall probability of success ($\pi$ or power) 
+for specified sample size ($n$) and  variance ($\sigma$)  as
+
+\begin{equation}\label{eqn:power} 
+Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
+\end{equation}
+
+The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
+combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
+Using equation \ref{mixed}, we then have the power as
+
+\begin{equation}\label{eqn:effect} 
+Power(E) = \pi=Pr(Y=1| E) 
+\end{equation}
 
@@
 
-Consider a linear regression model 
+To make these concepts more concrete consider how this would operate in the linear models setting. Consider a linear regression model 
 \begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
@@
 
-Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the actual data, the resulting residual plot shows two groups of residuals.
-To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic.
-If $\beta_k\ne 0$ the boxplots show a vertical displacement.
+In this setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model \citep{cook:99}.
+Table \ref{tbl:stat_multiple} shows an overview of possible hypotheses that would be tested in the regression setting, and what visual test statistic might be used in these tests.
+These plots shoul be familiar, because they are all commonly used for evaluating and diagnosing regression models.
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
+To assess whether the linear assumption is appropriate we would make a plot of residuals against fitted values (case 5).
 
-In a regression setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model \citep{cook:99}.
-Table \ref{tbl:stat_multiple} shows an overview of some commonly used charts and their corresponding hypothesis tests.
-Note that some charts can be associated with several different testing situations.
-In our paper, we pick the examples for situation 2 and 3 as candidates.
-
-A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
-The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
-The test statistic, the plot containing the actual data, is randomly placed among these null plots.
-If the test statistic is identifiable the null hypothesis is rejected.
-% with a $p$-value of at most 0.05.
+%Note that some charts can be associated with several different testing situations. In our paper, we pick the examples for situation 2 and 3 as candidates.
 
@@
 \end{table*}
+
+For the purposes of validating visual inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
+
+Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
+If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the actual data, the resulting residual plot shows two groups of residuals.
+To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic.
+If $\beta_k\ne 0$ the boxplots show a vertical displacement.
+
+%A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
+The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
+The test statistic, the plot containing the actual data, is randomly placed among these null plots.
+If the test statistic is identifiable the null hypothesis is rejected.
+% with a $p$-value of at most 0.05.
 
@@
 Figure \ref{fig:power_expected} shows the power of the uniformly most powerful test (UMP) in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
-Notice that the expected power of the visual test is almost as good as the power of UMP test.
+\green{Give more details on what is the UMP test in this setting.} Notice that the expected power of the visual test is almost as good as the power of UMP test.
 In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
@@
 %
-
-\subsection{Estimating Power from Lineups}
-
-We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
-This gives an estimated power of
-\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
-
-Suppose each of $K$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{\ell i}$.
-Let $Y_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$, and 0 otherwise.
-We model $\pi_{\ell i} = E(Y_{\ell i})$ as  a mixed effects logistic regression
-
-\begin{equation} 
-g(\pi_{\ell i}) = X_{\ell i}B  + Z_{\ell i} \tau_i 
-\label{mixed} 
-\end{equation}
-%
-where $\tau_i$ is random effects coefficient vector of length $q$ for subject $i$,  $\tau_i  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{\ell i}$ is the $\ell$th row vector of random effects covariates for subject $i$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{\ell i}$ is the $\ell$th row vector of the fixed effects covariates for subject $i$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-The covariates could be demographic information of people such as age, gender, education level etc.
-as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.
-
-From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
-the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
-%Thus we have the overall probability of success ($\pi$ or power) 
-for specified sample size ($n$) and  variance ($\sigma$)  as
-
-\begin{equation}\label{eqn:power} 
-Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
-\end{equation}
-
-The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
-combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
-Using equation \ref{mixed}, we then have the power as
-
-\begin{equation}\label{eqn:effect} 
-Power(E) = \pi=Pr(Y=1| E) 
-\end{equation}
 

### 2012-06-02 paper_jasa.tex
type II, power consideration

@@
 \newcommand{\blue}[1]{{\color{blue} #1}} %MM
-\newcommand{\red}[1]{{\color{red} #1}} %HH
+\newcommand{\red}[1]{{\color{red} #1}} 
 \newcommand{\green}[1]{{\color{green} #1}} %DC
+
+\definecolor{orange}{rgb}{1,0.5,0}
+\newcommand{\hh}[1]{{\color{orange} #1}} %HH
 
@@
 
-%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.
+%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.}
 
@@
 We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
-If there is evidence for individual skills influencing the probability $p_i$, then $X = \sum_i X_i$ will have a Poisson-Binomial distribution.
-\green{WHY???
-And how does this relate to the logistic model below?} This leads to the definition of the power of the visual test, as follows:
+
+\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
+This makes the distribution of $X$ a Poisson-Binomial by definition, see XXX for details.
+estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
+
+%\green{WHY??? And how does this relate to the logistic model below?} 
+%This leads to the definition of the power of the visual test, as follows:
 
@@
     \begin{equation*}
-      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{\alpha}) = 
-        \begin{cases} 
-               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
-              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
-        \end{cases}
+      \text{Power}_V(\theta)= Pr(\text{Reject } H_0 \mid \theta) 
+%      = P(X \ge x_{\alpha}) =
+%        \begin{cases} 
+%               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
+%              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
+%        \end{cases}
     \end{equation*}
@@
 
-\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
+\hh{\noindent An important difference between classical testing and the lineup scenarios is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
+we are assuming, again, that $X$ counts the number of observers (out of $K$) who identify the data plot from the lineup.
+Then power is calculated as
+\[
+\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
+\] 
+where $F_{X, \theta}$ is the distribution of $X$ -- this depends on which hypothesis is true:
 
-\red{there may be a flow in this definition.
+Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a power of 
+\[
+\text{Power}_V(\theta)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
+\]
+where $x_\alpha$ is defined as in (\ref{dfn:test}), as  $P(X \ge x_{\alpha}) < \alpha$.
+
+If the alternative hypothesis is true with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
+Again, assessing an individual's probability to identify the data plot  is central to the actual calculation .
+}
+%\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
+
+%\red{there may be a flow in this definition.
 Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power.
@@
 
-In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{ \ell i}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:
+Suppose each of $K$ independent observers gives evaluations on multiple lineups and responses are associated with binary random variable $X_{\ell i}$, where  $X_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$,  $\ell = 1, ..., L$, and 0 otherwise.
+
+\hh{
+We model $p_{\ell i} = E(X_{\ell i})$ as  a mixed effects logistic regression
+ accommodating both for different abilities of individuals as well as differences in the difficulty of lineups:
 
@@
 \[
-\text{logit } p_{\ell i} = X_{\ell i} B +  Z_{\ell i} \tau_i,  
+g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{i\ell},  
 \]
-where $X_{\ell i}$ is a vector of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $\tau \sim MVN(0, \sigma_{\tau} I)$  is a vector of random variables $i = 1, ..., K$.
 
-\red{This model is elaborately described in equation \ref{mixed}.}
+where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and
+
+\begin{tabular}{lp{5.5in}}
+$X$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$.
+Covariates could include  demographic information of individuals, such as age, gender, education level etc.
+as well lineup-specific elements, e.g.
+difficulty level.
+\\
+$B$ & is a vector of parameters for fixed effects.\\
+$Z_{i\ell}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
+$\tau$ & is a vector of independent real-valued random variables $\tau_{i\ell}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
+$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.\\
+\end{tabular}
+
+}
+
+%\red{This model is  described in more detail in equation \ref{mixed}.}
 
@@
 
-\subsection*{Lineup size $m$}
 \noindent
+\paragraph{Lineup size $m$}
 One of the big differences between visual inference and the classical testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
@@
 
-\blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
@@
 %In a parametric setting, signal strength of a test coincides with the power of the test.
-}
 
-\red{\#\#\# re-done up to here, HH}
-
-Theoretical power for the regression parameters is derived in the next section.
-\subsection{Estimating Power from Lineups}
-
-We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
+%Theoretical power for the regression parameters is derived in the next section.
+%\subsection{Estimating Power from Lineups}
+%
+%We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
 Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
 This gives an estimated power of
-\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
+%\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
 
-Suppose each of $K$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{\ell i}$.
-Let $Y_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$, and 0 otherwise.
-We model $\pi_{\ell i} = E(Y_{\ell i})$ as  a mixed effects logistic regression
-
-\begin{equation} 
-g(\pi_{\ell i}) = X_{\ell i}B  + Z_{\ell i} \tau_i 
-\label{mixed} 
-\end{equation}
+%From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
+%the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
+%%Thus we have the overall probability of success ($\pi$ or power) 
+%for specified sample size ($n$) and  variance ($\sigma$)  as 
 %
-where $\tau_i$ is random effects coefficient vector of length $q$ for subject $i$,  $\tau_i  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{\ell i}$ is the $\ell$th row vector of random effects covariates for subject $i$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{\ell i}$ is the $\ell$th row vector of the fixed effects covariates for subject $i$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-The covariates could be demographic information of people such as age, gender, education level etc.
-as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.
-
-From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
-the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
-%Thus we have the overall probability of success ($\pi$ or power) 
-for specified sample size ($n$) and  variance ($\sigma$)  as
-
-\begin{equation}\label{eqn:power} 
-Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
-\end{equation}
-
-The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
+%\begin{equation}\label{eqn:power} 
+%Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
+%\end{equation}
+%
+%The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
 combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
 Using equation \ref{mixed}, we then have the power as
-
-\begin{equation}\label{eqn:effect} 
-Power(E) = \pi=Pr(Y=1| E) 
-\end{equation}
+%
+%\begin{equation}\label{eqn:effect} 
+%Power(E) = \pi=Pr(Y=1| E) 
+%\end{equation}
 

### 2012-06-02 paper_jasa.tex
experiment explanation

@@
 We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
-\[
+\begin{equation} \label{mixed}
 g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{i\ell},  
-\]
+\end{equation}
 
@@
 
-Figure \ref{fig:power_expected} shows the power of the uniformly most powerful test (UMP) in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
-\green{Give more details on what is the UMP test in this setting.} Notice that the expected power of the visual test is almost as good as the power of UMP test.
+The classical uniformly most powerful (UMP) test rejects $H_0$ if the observed test statistic $T= \hat{\beta_k}/ se(\hat{\beta_k})$ takes an extreme value in the distribution of theoretical $t$ distribution with $n-p$ degrees of freedom.
+Figure \ref{fig:power_expected} shows the power of the UMP test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
+Notice that the expected power of the visual test is almost as good as the power of UMP test.
 In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
@@
 Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson(30) distribution.
-\green{Explain why this was used.} Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
+\green{Explain why this was used.} \red{There was not any real reason for this.
+It was just arbitrarily chosen to generate some data.
+When I first started to think of applying visual inference in regression setting, I was thinking of a dataset (I found in a book) of weight of babies by their gestational age (in week) where author wanted to test the difference of mean weight for gender.
+Thus for our experiment I generated values similar to a variable like age.
+the discrete variable was thought similar to gender.} Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
 Three replicates at each level were generated.

### 2012-06-03 paper_jasa.tex
experiment1 details

@@
 
-Each subject viewed 10 lineups.
-Selection of lineups shown to each subject followed these rules ....
-\green{fill this in}.
-Figure \ref{fig:test_category} is an example of one of these lineups.
-The visual test statistic is a side-by-side boxplot.
+{\bf Number of evaluations} required for each lineup is determined based on the precision level of the estimated power.
+Suppose $\gamma$ denotes the classical UMP test power for each parameter combination shown in table \ref{tbl:experiment_params}.
+Since expected power of visual inference is very close to the power of classical UMP test as in figure \ref{fig:power_expected} the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
+For a given proportion $\gamma$ we want to have margin of error (ME) to be 0.1.
+Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
+Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted.
+Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed as shown in table \ref{tbl:dist_lineup1}.
+For each difficulty level a specific number of lineups were randomly picked to show for evaluation.
+This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$.
+To satisfy this plan we needed to recruit at least 300 subjects.
+
+%For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
+Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
+We collect data such that each of the parameter combinations in the table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
+  Figure \ref{fig:test_category} is an example of one of these lineups. The visual test statistic is a side-by-side boxplot.
+
+\begin{table}[hbtp]
+\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
+\centering
+\begin{tabular}{c c c c  c c c}
+\hline\hline
+Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
+level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
+\hline
+easy&100& 5&8 & 1& 12 & 1\\
+&100&12&16 &1&&\\
+&300& 5&5 &1&&\\
+&300&12&10 &1 &&\\
+\hline
+medium&100& 5&3 &203 & 9 &2\\
+&300& 5&2,3 & 97, 1&&\\
+\hline
+hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
+&300& 5&1 & 371 &&\\
+&300&12&3,5& 375,74 &&\\
+\hline
+mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
+&100&12&1& 100& &\\
+&300& 5&0 & 73&&\\
+&300&12&7,1& 2, 152&&\\
+\hline
+Total &&&&&60&10\\
+\hline
+\end{tabular}
+\label{tbl:dist_lineup1}
+\end{table}
 

### 2012-06-03 R_codes/turk1_data_analysis_jasa.R

@@ -53,6 +53,8 @@ dat <- subset(raw.dat, id %in% included_id)
 #
 ## unit(time_taken) = second 
 
+      
+
 
 # ------function to calculate UMP power -----------------------
 
@@ -70,6 +72,21 @@ calculate_ump_power(3,300,5)
 
 calculate_ump_power(3/5,300,1)
 
+
+# ---- distribution of beta as the effect of algorithm for showing lineups ---
+
+params <- ddply(dat, .(difficulty, sample_size, sigma), summarise,
+   params_beta <- paste(unique(beta), collapse=",") )
+   
+sample_size <- function(g){
+	res <- g*(1-g)/((.05/1.96)^2)
+	return(round(res,0))
+} 
+
+params_comb <- ddply(dat, .(difficulty, sample_size, sigma), summarise,
+      params_beta <- unique(beta),
+      g <- calculate_ump_power(beta=unique(beta),n=sample_size[1],sigma=sigma[1]),
+      s.size <- sample_size(calculate_ump_power(beta=unique(beta),n=sample_size[1],sigma=sigma[1])) ) 
 
 
 # ---------------------- fitting loess smoother  ------------

### 2012-06-03 paper_jasa.tex
generation of lineup

@@
    \centering
-       \includegraphics[width=0.95\textwidth]{test_category.pdf}
+       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
        \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
@@
 
- Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{test_category_small.pdf}} \end{center} \end{minipage} \\
+ Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{lineup_category_small.pdf}} \end{center} \end{minipage} \\
  & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
@@
 \\
-2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+2 & $H_0: \beta_k=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_beta_k.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
+For lineup plot, we simulate data from fitted null model.
 \\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
@@
 
+{\bf Plotting visual statistic:} For a given data set we fit the null model and obtain the residuals.
+Using these residuals the visual statistics for each of the experiments were generated as described in table \ref{tbl:stat_multiple}.
+Actual data plot and null plots in a lineup are obtained from actual data set and null data set respectively.
+The procedure for obtaining actual data sets are described in the following sections.
+
 \subsection{Discrete covariate}\label{sec:category}
@@
 
+{\bf Steps to obtain lineup}
+\begin{enumerate}
+	\item For a given parameter setting obtain actual data set
+	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
+	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
+	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+\end{enumerate}
+
+ Figure \ref{fig:test_category} is an example of one of these lineups.
+The visual test statistic is a side-by-side boxplot.
+The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
+
 {\bf Number of evaluations} required for each lineup is determined based on the precision level of the estimated power.
@@
 We collect data such that each of the parameter combinations in the table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
-  Figure \ref{fig:test_category} is an example of one of these lineups. The visual test statistic is a side-by-side boxplot.
 
@@
 
-\green{Did you in fact bin like this???} For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
+\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
+This produces an empirical distribution of p-value under alternative hypothesis.
+Finally, three data sets that have 16th, 50th and 84th quantile of the p-values were taken.
+Similar procedure was used to generate 19 null plots based on the empirical distribution of p-value under null hypothesis}.
+For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
 A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
@@
 This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous.
+$X_1$ is simulated from standard normal(0,1).
 Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
@@
 
-\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
+{\bf Steps to obtain lineup}
+\begin{enumerate}
+	\item For a given parameter setting obtain actual data set
+	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
+	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
+	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+\end{enumerate}
+
+%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
+
+\begin{figure}[htp]
+%\begin{figurehere}
+   \centering
+       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
+       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous.
+One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
+Which plot is the most different from the others, in the sense that there is the steepest slope?}
+       \label{fig:test_continous}
+\end{figure}
 

### 2012-06-03 R_codes/paper_jasa.R

@@ -38,13 +38,14 @@ for (K in 1:5){
 
 # generating data
 set.seed(29)
+
 n <- 300
-beta <- 15
+beta <- 10
 sigma <- 12	
 x1 <- rpois(n,lambda=20)
 Xk <- factor(sample(c("A","B"), size=n, replace=T))
 cnd <- sample(c("A","B"), size=1)
-y <- round(5 + 1.5 * x1 + beta * (Xk==cnd) + rnorm(n=n, mean=0, sd=sigma))
+y <- round(5 + 15 * x1 + beta * (Xk==cnd) + rnorm(n=n, mean=0, sd=sigma))
 #qplot(x1,y, colour=factor(Xk)) + geom_smooth(method="lm", se=F, size=1)
 
 
@@ -56,7 +57,7 @@ obs.residuals <- as.vector(resid(fit))
 
 qplot(Xk, obs.residuals,colour=Xk,geom="boxplot", ylab="Residual") + 
       xlab(expression(X[k])) + labs(colour=expression(X[k]))
-ggsave("../images/stat_category.pdf",height=2,width=3)
+ggsave("../images/stat_category.pdf",height=2,width=2.75)
 
 
 
@@ -71,8 +72,53 @@ pdat.m <- melt(pdat,id="Xk")
 qplot(Xk, value, data=pdat.m ,colour=Xk,,geom="boxplot", ylab="Residual") + 
      facet_wrap(~variable)+ xlab(expression(X[k])) + 
      labs(colour=expression(X[k]))
-ggsave("../images/test_category.pdf",height=7,width=7.5)
-ggsave("../images/test_category_small.pdf",height=4.5,width=5) # for comparizon table
+ggsave("../images/lineup_category.pdf",height=7,width=7.5)
+ggsave("../images/lineup_category_small.pdf",height=4.5,width=5) # for comparizon table
+
+# === plotting test stat and lineup for continuous variable variable (turk2)
+
+generate_turk2_stat <- function(n,beta,sigma){
+	#n <- 100; a <- 6; b <- 2; sigma <- 12
+	a <- 6
+	b <- beta*sign(rnorm(1,0,1))
+	x <- subset(read.csv("../data/Xdata.csv"),N==n)[,1]
+	y <- a + b*x + rnorm(n=n, mean=0, sd=sigma)
[…]

### 2012-06-03 paper_jasa.tex
experiment2 setup

@@
 \centering 
-\begin{tabular}{m{0.5cm}m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
+\begin{tabular}{m{0.5cm}m{3cm}m{2cm}m{3cm}m{5.5cm}} 
 \hline\hline 
@@
 \hline 
-1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
+1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
 For lineup plot, we simulate data from fitted null model.
 \\
-2 & $H_0: \beta_k=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_beta_k.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
-For lineup plot, we simulate data from fitted null model.
-\\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+\\
+4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid.
 For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
-5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
+5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
 For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
+6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
 For lineup plot, we simulate data from standard normal.
 \\
-7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
+7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
 For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
 \\
-8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data.
+8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data.
 For lineup plot, we simulate data from fitted model.
 \\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
+9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
 \hline 
@@
 For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.
+Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.
 
@@
 
+A lineup for this experiment is shown in figure \ref{fig:test_continuous}.
+This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
+The actual data plot location is 11.
+
 %\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
@@
 Which plot is the most different from the others, in the sense that there is the steepest slope?}
-       \label{fig:test_continous}
+       \label{fig:test_continuous}
 \end{figure}

### 2012-06-03 R_codes/paper_jasa.R

@@ -75,21 +75,7 @@ qplot(Xk, value, data=pdat.m ,colour=Xk,,geom="boxplot", ylab="Residual") +
 ggsave("../images/lineup_category.pdf",height=7,width=7.5)
 ggsave("../images/lineup_category_small.pdf",height=4.5,width=5) # for comparizon table
 
-# === plotting test stat and lineup for continuous variable variable (turk2)
-
-generate_turk2_stat <- function(n,beta,sigma){
-	#n <- 100; a <- 6; b <- 2; sigma <- 12
-	a <- 6
-	b <- beta*sign(rnorm(1,0,1))
-	x <- subset(read.csv("../data/Xdata.csv"),N==n)[,1]
-	y <- a + b*x + rnorm(n=n, mean=0, sd=sigma)
-	p <- qplot(x, y, xlab=expression(X[k]),geom="point", ylab="Y", alpha=I(.2))
-	p <- p + geom_smooth(method="lm", se=F, size=1)
-	return(p)
-}
-p <- generate_turk2_stat(n=100,beta=1.25,sigma=5)
-ggsave(plot=p, file="../images/stat_beta_k.pdf",height=2,width=2)
-
+# === plotting lineup for continuous variable variable (turk2)
 
 generate_turk2_lineup <- function(n,beta,sigma){
 	#n <- 100; a <- 6; b <- 2; sigma <- 12
@@ -119,6 +105,21 @@ generate_turk2_lineup <- function(n,beta,sigma){
 
 l2 <- generate_turk2_lineup(n=100,beta=1.25,sigma=5)
 ggsave(plot=l2$p, file="../images/lineup_continuous.pdf",height=7,width=7.5)
+
+# ==== plotting test statistics --------
+
+generate_stat_slope <- function(n,beta,sigma){
+	#n <- 100; a <- 6; b <- 2; sigma <- 12
+	a <- 6
+	x <- rnorm(n=n, mean=0, sd=1)
+	y <- a + beta*x + rnorm(n=n, mean=0, sd=sigma)
+	fit <- lm(y~1)
+	p <- qplot(x, resid(fit), xlab=expression(X[k]),geom="point", ylab="Residual", alpha=I(.3))
+	p <- p + geom_smooth(method="lm", se=F, size=1)
+	return(p)
+}
+p <- generate_stat_slope(n=100,beta=2.5,sigma=5)
+ggsave(plot=p, file="../images/stat_beta_k.pdf",height=2,width=2.05)
 
 
 

### 2012-06-03 R_codes/turk2_data_analysis_jasa.R

@@ -541,7 +541,7 @@ ggplot(data=effect.d,aes(x=effect,y=power)) +
    geom_smooth(method="loess", span=0.8,aes(colour=factor(replica))) +
    geom_line(aes(x=effect,y=ump_power)) +
    scale_colour_discrete(name = "Replication")+
-   xlab(expression(Effect(E)))
+   xlab(expression(Effect(E))) 
    
 ggsave(filename="../images/effect_power_exp2.pdf", height=7, width=7)   
    
@@ -602,7 +602,7 @@ tail(pow.dat.m)
 qplot(effects,value,group=variable,geom="line",alpha=I(.2), data=pow.dat.m) + 
  xlab(expression(beta)) + ylab("Power") + 
  geom_line(data=subset(pow.dat.m, variable=="UMP"),colour="hotpink", size=1) + 
- xlab("Effect(E)")
+ xlab("Effect(E)") 
  
 ggsave(filename="../images/effect_power_subject_exp2.pdf", height=5, width=7)   
 

### 2012-06-04 paper_jasa.tex
experiment 3 details

@@
 For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.
+For some parameter combinations we have more than three replications.
 Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.
@@
 
+{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
+For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}.
+For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
+The 10 lineups are choosen as follows;
+
+\begin{enumerate}
+	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
+	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
+	\item Finally select another lineup with difficulty level 0
+\end{enumerate}
+
 \begin{figure}[htp]
@@
   \begin{array}{l l}
-    5+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad \text{ $i$ =1,...,85}\\
-    15+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=86,...,n\\
+    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
+    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
   \end{array} \right.
 \]
-where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$.
-One such plot of the contaminated data for $\sigma$=5 and $\beta$=5 is shown in figure \ref{fig:cont_dat}.
-For this experiment we only consider sample size $n=100$ and error standard deviation $\sigma=5$.
-
-For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.
-
-\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
+where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data.
+For the experiment we consider $n=100$, $c=0.15$ which produced actual data size 115, $\alpha=0$ and $\lambda=10$ .
+$\sigma$ is chosen such that overall error standard deviation of the data is 5 which is approximately 3.5.
+One such plot of the contaminated data for  $\beta$=5 is shown in figure \ref{fig:cont_dat}.
+We fit model (Equation \eqref{multi}) with $p=1$  and intercept $\beta_0=0$ to this contaminated data to test the hypothesis whether slope is zero or not.
+
+{\bf Steps to obtain lineup}
+\begin{enumerate}
+	\item For a given parameter setting obtain actual data set
+	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
+	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
+	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+\end{enumerate}
+
+One lineup for slope $\beta=0.4$ is shown in figure \ref{fig:test_contaminated}.
+For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}.
+This produces 30 lineups.
+each subject got at least 10 lineups to evaluate.
+Based on the slope parameter we group lineups into six groups with difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.
+Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
+
+\begin{figure}[htp]
+%\begin{figurehere}
+   \centering
+       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
+       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure.
+One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
+Which plot is the most different from the others, in the sense that there is the steepest slope?}
+       \label{fig:test_contaminated}
+\end{figure}
+
+%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
 Show a sample lineup, probably the most difficult, instead of the single plot.}

### 2012-06-04 R/codes_for_heike.R

@@ -1,0 +1,41 @@
+lineups <- ddply(raw.dat, .(pic_name, academic_study), summarise,
+	ni = sum(response),
+	K = length(response),
+	sample_size=sample_size[1],
+	p_value=p_value[1],
+	beta=beta[1],
+	sigma=sigma[1],
+	age=mean(age)
+)
+lineups$effect <- with(lineups, beta/sigma)
+
+
+lineups.glm <- glm(cbind(ni,K-ni)~effect+ sample_size, data=lineups, family=binomial(link="logit"))
+preds <- predict(lineups.glm, type="response", se.fit=TRUE)
+
+lineups$preds<- preds$fit
+lineups$se <- preds$se.fit
+
+
+calculate_ump_power <- function (beta, n, sigma){
+	alpha <- 0.05/2
+	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs for derivation of power
+	mu <- beta/se_beta
+	t_n <- qt(p=1-alpha,df=n-3)
+	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
+	return(res)
+}
+
+ump100 <- data.frame(effect=seq(0,1.6,by=.01),
+          power = sapply(seq(0,1.6,by=.01),calculate_ump_power,n=100,sigma=1),
+          sample_size=100)
+ump300 <- data.frame(effect=seq(0,1.6,by=.01),
+          power = sapply(seq(0,1.6,by=.01),calculate_ump_power,n=300,sigma=1),
+          sample_size=300)
+ump <- rbind(ump100,ump300)   
+ 
+ggplot() + 
+geom_point(aes(effect, preds,colour=factor(sample_size)), data=lineups) +
+geom_errorbar(aes(effect,preds, colour=factor(sample_size),ymax = preds + 1.95*se, ymin=preds - 1.95*se), data=lineups)+ 
+geom_line(data=ump,aes(effect,power, colour=factor(sample_size)))   # how do I get that line here?
+

### 2012-06-04 R/heike-data-turk1.R

@@ -1,0 +1,77 @@
+raw.dat$beta <- as.numeric(gsub("([0-9]+)_[0-9]+","\\1", as.character(raw.dat$param_value)))
+raw.dat$sigma <- as.numeric(gsub("[0-9]+_([0-9]+)","\\1", as.character(raw.dat$param_value)))
+
+lineups <- ddply(raw.dat, .(pic_name, academic_study), summarise,
+	ni = sum(response),
+	K = length(response),
+	sample_size=sample_size[1],
+	p_value=p_value[1],
+	beta=beta[1],
+	sigma=sigma[1],
+	age=mean(age)
+)
+lineups$effect <- with(lineups, beta/sigma)
+
+lineups.glm <- glm(cbind(ni,K-ni)~effect+ sample_size, data=lineups, family=binomial(link="logit"))
+preds <- predict(lineups.glm, type="response", se.fit=TRUE)
+
+lineups$preds<- preds$fit
+lineups$se <- preds$se.fit
+
+#lineups$preds[! is.na(lineups$age)] <- preds$fit
+#lineups$se[! is.na(lineups$age)] <- preds$se.fit
+
+ggplot(aes(x=effect, y=preds, ymax = preds + 1.95*se, ymin=preds - 1.95*se), data=lineups) +
+facet_grid(sample_size~.)+  geom_point()  + geom_errorbar()
+#geom_line(data=ump,aes(effect,power))   # how do I get that line here?
+
+
+calculate_ump_power <- function (beta, n, sigma){
+	alpha <- 0.05/2
+	se_beta <- sigma/(0.5 * sqrt(n))    # refer to docs for derivation of power
+	mu <- beta/se_beta
+	t_n <- qt(p=1-alpha,df=n-3)
+	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
+	return(res)
+}
+
+ump100 <- data.frame(effect=seq(0,1.6,by=.01),
+          power = sapply(seq(0,1.6,by=.01),calculate_ump_power,n=100,sigma=1),
+          sample_size=100)
+ump300 <- data.frame(effect=seq(0,1.6,by=.01),
+          power = sapply(seq(0,1.6,by=.01),calculate_ump_power,n=300,sigma=1),
+          sample_size=300)
+ump <- rbind(ump100,ump300)   
+ 
+ggplot() + 
+geom_point(aes(effect, preds,colour=factor(sample_size)), data=lineups) +
[…]

### 2012-06-04 R/paper_jasa.R

@@ -106,6 +106,44 @@ generate_turk2_lineup <- function(n,beta,sigma){
 l2 <- generate_turk2_lineup(n=100,beta=1.25,sigma=5)
 ggsave(plot=l2$p, file="../images/lineup_continuous.pdf",height=7,width=7.5)
 
+# plotting turk3 lineups --------------
+
+generate_turk3_lineup <- function(n,sigma,beta){
+	#n <- 100;nc <- 15; beta <- .1; sigma <- 4; 
+	alpha <- 0
+	lambda <- 10 
+	beta <- sign(rnorm(1))*beta
+	nc <- n*15/100
+	x1 <- rnorm(n,0,1)
+	x2 <- rnorm(n=nc,mean=-1.75*sign(beta), sd=1/3)
+	x <- c(x1,x2)
+	sig_hat <- 0
+	while(sig_hat<4.99 || sig_hat>5.01){
+		y1 <- alpha + beta*x1 + rnorm(n,0,sigma)
+		y2 <- rnorm(nc,lambda,sigma/3)
+		y <- c(y1,y2)
+		fit <- lm(y~x-1)
+		sig_hat <- summary(fit)$sigma
+	}
+	pval <- summary(fit)$coefficients[,4]
+
+	sim <- matrix(rnorm(n=20*length(x),mean=0,sd=sig_hat),ncol=20)
+	loc <- sample(1:20,size=1)
+	sim[,loc] <- y
+
+	dat <- data.frame(x,sim)
+	colnames(dat) <- c("X",1:20)
+	m.dat <- melt(dat, id="X")
+
+	p <- qplot(X,value,data=m.dat,alpha=I(.2), ylab="Y")+facet_wrap(~variable)
+	result = data.frame(n,beta,sigma=5,pval,loc)
+	return(list(result=result,p=p, dat_used=data.frame(x,y)))
+}
+
+pl <- generate_turk3_lineup(n=100,sigma=3.5,beta=0.4)
+pl$result
+ggsave(plot=pl$p, file="../images/lineup_contaminated.pdf",height=7,width=7.5)
+
 # ==== plotting test statistics --------
 
 generate_stat_slope <- function(n,beta,sigma){

### 2012-06-04 paper_jasa.tex
edited experiment section

@@
 
-{\bf Plotting visual statistic:} For a given data set we fit the null model and obtain the residuals.
-Using these residuals the visual statistics for each of the experiments were generated as described in table \ref{tbl:stat_multiple}.
-Actual data plot and null plots in a lineup are obtained from actual data set and null data set respectively.
-The procedure for obtaining actual data sets are described in the following sections.
-
 \subsection{Discrete covariate}\label{sec:category}

### 2012-06-04 paper_jasa.tex
working on experimental explanation section

@@
 
+After many small pilot studies with local personnel, it was clear that some care was needed to set up the human subjects experiments.
+It was best for a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
+The explanations about each experiment (below) includes and explanation of how the lineups were sampled and provided to the observers.
+
 \subsection{Discrete covariate}\label{sec:category}
@@
 The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test.
-Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson(30) distribution.
-\green{Explain why this was used.} \red{There was not any real reason for this.
+Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
+%\green{Explain why this was used.} \red{There was not any real reason for this.
 It was just arbitrarily chosen to generate some data.
@@
 Thus for our experiment I generated values similar to a variable like age.
-the discrete variable was thought similar to gender.} Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
+the discrete variable was thought similar to gender.}
+Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
 Three replicates at each level were generated.
-These produced 60 different ``actual data sets''.
-
-{\bf Steps to obtain lineup}
-\begin{enumerate}
-	\item For a given parameter setting obtain actual data set
-	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
-	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-\end{enumerate}
-
+These produced 60 different ``actual data sets'', and thus, 60 different lineups.
+To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
+The actual ydata plot was generated by making a side-by-side boxplot of the residuals (Table\ref{tbl:stat_multiple} case 3).
+The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2$, and these were plotted in the same way as side-by-side boxplots.
+The actual data plot was randomly placed among these null data plots to produce the lineup.
 Figure \ref{fig:test_category} is an example of one of these lineups.
+The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
+\green{I think it is plot11, not 13!!!!!!!!
+Please check!}
+
+%{\bf Steps to obtain lineup}
+%\begin{enumerate}
+%	\item For a given parameter setting obtain actual data set
+%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+%	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
+%	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
+%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+%\end{enumerate}
+
+ %Figure \ref{fig:test_category} is an example of one of these lineups.
 The visual test statistic is a side-by-side boxplot.
@@
 
-{\bf Number of evaluations} required for each lineup is determined based on the precision level of the estimated power.
+The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
 Suppose $\gamma$ denotes the classical UMP test power for each parameter combination shown in table \ref{tbl:experiment_params}.
-Since expected power of visual inference is very close to the power of classical UMP test as in figure \ref{fig:power_expected} the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
-For a given proportion $\gamma$ we want to have margin of error (ME) to be 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
+Since the expected power of visual inference is very close to the power of classical UMP test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
+For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.1.
+Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ \green{0.05 or 0.1???} which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
+
 Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted.
-Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed as shown in table \ref{tbl:dist_lineup1}.
+Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the Appendix).
 For each difficulty level a specific number of lineups were randomly picked to show for evaluation.
@@
 
+\begin{table}[hbtp]
+\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
+\centering
+% centering table
+\begin{tabular}{c c c c c}
+% creating 10 columns
+\hline\hline
+
+% inserting double-line
+& & \multicolumn{3}{c}{Slope ($\beta$)} \\
+ \cline{3-5}
+
+Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
+ ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
+\\ [0.5ex]
+\hline
+% inserts single-line
+
+% Entering 1st row
+&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
+\raisebox{1.5ex}{100} &12
+& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]
+
+% Entering 2rd row
+&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
+\raisebox{1.5ex}{300} & 12
+& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
+% [1ex] adds vertical space
+\hline
+% inserts single-line
+\end{tabular}
+\label{tbl:experiment_params}
+\end{table}
+
 %For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
@@
 
-\begin{table}[hbtp]
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
-\centering
-\begin{tabular}{c c c c  c c c}
-\hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
-\hline
-easy&100& 5&8 & 1& 12 & 1\\
-&100&12&16 &1&&\\
-&300& 5&5 &1&&\\
-&300&12&10 &1 &&\\
-\hline
-medium&100& 5&3 &203 & 9 &2\\
-&300& 5&2,3 & 97, 1&&\\
-\hline
-hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
-&300& 5&1 & 371 &&\\
-&300&12&3,5& 375,74 &&\\
-\hline
-mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
-&100&12&1& 100& &\\
-&300& 5&0 & 73&&\\
-&300&12&7,1& 2, 152&&\\
-\hline
-Total &&&&&60&10\\
-\hline
-\end{tabular}
-\label{tbl:dist_lineup1}
-\end{table}
-
-\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
+%\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
 This produces an empirical distribution of p-value under alternative hypothesis.
@@
 Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
-
-\begin{table}[hbtp]
-\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
-\centering
-% centering table
-\begin{tabular}{c c c c c}
-% creating 10 columns
-\hline\hline
-
-% inserting double-line
-& & \multicolumn{3}{c}{Slope ($\beta$)} \\
- \cline{3-5}
-
-Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
- ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
-\\ [0.5ex]
-\hline
-% inserts single-line
-
-% Entering 1st row
-&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
-\raisebox{1.5ex}{100} &12
-& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]
-
-% Entering 2rd row
-&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
-\raisebox{1.5ex}{300} & 12
-& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
-% [1ex] adds vertical space
-\hline
-% inserts single-line
-\end{tabular}
-\label{tbl:experiment_params}
-\end{table}
 
@@
 %\end{multicols}
+\section*{Appendix}
+
+\begin{table}[hbtp]
+\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
+\centering
+\begin{tabular}{c c c c  c c c}
+\hline\hline
+Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
+level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
+\hline
+easy&100& 5&8 & 1& 12 & 1\\
+&100&12&16 &1&&\\
+&300& 5&5 &1&&\\
+&300&12&10 &1 &&\\
+\hline
+medium&100& 5&3 &203 & 9 &2\\
+&300& 5&2,3 & 97, 1&&\\
+\hline
+hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
+&300& 5&1 & 371 &&\\
+&300&12&3,5& 375,74 &&\\
+\hline
+mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
+&100&12&1& 100& &\\
+&300& 5&0 & 73&&\\
+&300&12&7,1& 2, 152&&\\
+\hline
+Total &&&&&60&10\\
+\hline
+\end{tabular}
+\label{tbl:dist_lineup1}
+\end{table}
 

### 2012-06-04 paper_jasa.tex
updated comments

@@
 Figure \ref{fig:test_category} is an example of one of these lineups.
-The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
-\green{I think it is plot11, not 13!!!!!!!!
-Please check!}
+The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
+
+%\green{I think it is plot11, not 13!!!!!!!! Please check!}
 
@@
 Since the expected power of visual inference is very close to the power of classical UMP test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
-For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ \green{0.05 or 0.1???} which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
+For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
+Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
 

### 2012-06-04 paper_jasa.tex
data summary

@@
 
+{\bf Cleaning data:} There are five suggested approach to clean the data. They are
+
+\begin{enumerate}
+\item not throw out anything
+\item throw out people where we don't have demographic information
+\item throw out all people with 0 percent correct
+\item only include people with at least 20 percent correct (current suggestion)
+\item include people with 50\% on all easy lineups (as suggested before)
+\end{enumerate}
+
+\begin{table}[hbtp]
+\caption{Summary of unique subjects and their total feedbacks after applying five screening process on all three experimental data}
+\begin{center}
+\begin{tabular}{rrrrr|rrrr|rrrr}
+  \hline
+Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
+ \cline{2-5}  \cline{6-9}   \cline{10-13}
+criteria & total & male & female & feedbacks & total & male & female & feedbacks & total & male & female & feedbacks \\ 
+  \hline
+1 & 520 & 264 & 201 & 4914 & 520 & 264 & 201 & 4914 & 520 & 264 & 201 & 4914 \\ 
+  2 & 465 & 264 & 201 & 4624 & 465 & 264 & 201 & 4624 & 465 & 264 & 201 & 4624 \\ 
+  3 & 384 & 200 & 165 & 4187 & 384 & 200 & 165 & 4187 & 384 & 200 & 165 & 4187 \\ 
+  4 & 318 & 168 & 135 & 3395 & 318 & 168 & 135 & 3395 & 318 & 168 & 135 & 3395 \\ 
+  5 & 277 & 139 & 124 & 3008 & 277 & 139 & 124 & 3008 & 277 & 139 & 124 & 3008 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:summary}
+\end{table}
+
 Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.

### 2012-06-04 R/paper_jasa.R

@@ -160,11 +160,46 @@ p <- generate_stat_slope(n=100,beta=2.5,sigma=5)
 ggsave(plot=p, file="../images/stat_beta_k.pdf",height=2,width=2.05)
 
 
+# ==== summary of the data =====
+
+get_summary <- function(dat){
+  g <- ddply(dat,.(gender),summarise, length(unique(id)))
+  male <- g[1,2]
+  female <- g[2,2]
+  total <- length(unique(dat$id))
+  evaluations <- nrow(dat)
+  return(data.frame(total,male,female, evaluations))
+}
+
+get_sceering_index <- function(dat){
+  indx1 <- dat$id==dat$id
+  indx2 <- !is.na(dat$gender)
+  d34 <- ddply(dat,.(id), summarise, percent_correct=mean(response)*100)
+  included_id3 <- d34$id[ d34$percent_correct > 0]
+  indx3 <- dat$id %in% included_id3
+  included_id4 <- d34$id[ d34$percent_correct >= 20 ]
+  indx4 <- dat$id %in% included_id4
+  d4 <- ddply(subset(dat, p_value < 0.0002),.(id),summarise,
+              easy_cnt = length(response),
+              percent_correct = mean(response)*100
+              )
+  included_id5 <- d4$id[ d4$percent_correct > 49]
+  indx5 <- dat$id %in% included_id5
+  return(data.frame(indx1,indx2,indx3,indx4,indx5))
+}
+
+get_screened_summary <- function(dat){
+  indx <- get_sceering_index(dat)
+  s <- NULL
+  for (i in 1:5) s <-  rbind(s,get_summary(subset(dat,indx[,i])))
+  return(s)
+}
+
 
 
 # =================== Turk1 data analysis  ================================
 
-dat <- read.csv("../data/feedback_data_turk1_50p.txt")
+dat <- read.csv("../data/raw_data_turk1.csv")
 
 # ------function to calculate UMP power -----------------------
 

### 2012-06-04 R/paper_jasa.R
number of subjects after applying different screening criteria

@@ -167,8 +167,8 @@ get_summary <- function(dat){
   male <- g[1,2]
   female <- g[2,2]
   total <- length(unique(dat$id))
-  evaluations <- nrow(dat)
-  return(data.frame(total,male,female, evaluations))
+  feedbacks <- nrow(dat)
+  return(data.frame(total,male,female, feedbacks))
 }
 
 get_sceering_index <- function(dat){
@@ -195,7 +195,15 @@ get_screened_summary <- function(dat){
   return(s)
 }
 
+dat1 <- read.csv("../data/raw_data_turk1.csv")
+dat2 <- read.csv("../data/raw_data_turk1.csv")
+dat3 <- read.csv("../data/raw_data_turk1.csv")
 
+sm <- cbind(get_screened_summary(dat1),
+      get_screened_summary(dat2),
+      get_screened_summary(dat3))
+library(xtable)
+xtable(sm)
 
 # =================== Turk1 data analysis  ================================
 

### 2012-06-05 paper_jasa.tex
number of subjects after applying different screening criteria
teensy typos

@@
 \[
-\text{Power}_V(\theta)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
+\text{Power}_V(\theta, K)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
 \]
@@
 \begin{equation} \label{mixed}
-g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{i\ell},  
+g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{\ell i},  
 \end{equation}
@@
 $B$ & is a vector of parameters for fixed effects.\\
-$Z_{i\ell}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
-$\tau$ & is a vector of independent real-valued random variables $\tau_{i\ell}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
+$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
+$\tau$ & is a vector of independent real-valued random variables $\tau_{\ell i}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.\\
@@
 \begin{table}[hbtp]
-\caption{Summary of unique subjects and their total feedbacks after applying five screening process on all three experimental data}
+\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
 \begin{center}
@@
   \hline
-1 & 520 & 264 & 201 & 4914 & 520 & 264 & 201 & 4914 & 520 & 264 & 201 & 4914 \\ 
-  2 & 465 & 264 & 201 & 4624 & 465 & 264 & 201 & 4624 & 465 & 264 & 201 & 4624 \\ 
-  3 & 384 & 200 & 165 & 4187 & 384 & 200 & 165 & 4187 & 384 & 200 & 165 & 4187 \\ 
-  4 & 318 & 168 & 135 & 3395 & 318 & 168 & 135 & 3395 & 318 & 168 & 135 & 3395 \\ 
-  5 & 277 & 139 & 124 & 3008 & 277 & 139 & 124 & 3008 & 277 & 139 & 124 & 3008 \\ 
+1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
+  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
+  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
+  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
+  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
    \hline

### 2012-06-05 paper_jasa.tex
Mosteller reference to type III error (pdf of paper is on dropbox)

@@
 
+\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size.
+This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}
+
 \green{A lot more needs to be done here!!!}

### 2012-06-06 paper_jasa.tex
Poisson Binomial reference
lead into lemma 3.1

@@
 \hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
-This makes the distribution of $X$ a Poisson-Binomial by definition, see XXX for details.
+This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details.
 estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
@@
 \end{enumerate}
-We will be able to show evidence for the second assumption with the help of the study discussed in section~\ref{sec:simulation}, the first assumption well be reflected by the power of a lineup.
+We will be able to show evidence for the second assumption with the help of the study discussed in section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup.
 The better suited a design is for a particular task, the higher its power will be.
+
+\hh{In order to compare power of classical tests and lineups side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
+}
 

### 2012-06-06 paper_jasa.tex
visual versus classical curves
results

--- 2012-06-05 paper_jasa.tex
@@
 \[
-\text{Power}_V(\theta, K)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
+\text{Power}_V(\theta)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
 \]
@@
 \begin{equation} \label{mixed}
-g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{\ell i},  
+g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{i\ell},  
 \end{equation}
@@
 $B$ & is a vector of parameters for fixed effects.\\
-$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
-$\tau$ & is a vector of independent real-valued random variables $\tau_{\ell i}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
+$Z_{i\ell}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
+$\tau$ & is a vector of independent real-valued random variables $\tau_{i\ell}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.\\
@@
 For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
-The 10 lineups are choosen as follows;
+The 10 lineups are chosen as follows;
 
@@
 \begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
-\begin{center}
+\caption{Number of participants and their total feedbacks after applying five screening process on all three experimental data sets.}
+\begin{center} {\small
 \begin{tabular}{rrrrr|rrrr|rrrr}
@@
    \hline
-\end{tabular}
+\end{tabular} }
 \end{center}

### 2012-06-06 paper_jasa.tex
not sure what i was chagning

@@
 Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
-%\green{Explain why this was used.} \red{There was not any real reason for this.
-It was just arbitrarily chosen to generate some data.
-When I first started to think of applying visual inference in regression setting, I was thinking of a dataset (I found in a book) of weight of babies by their gestational age (in week) where author wanted to test the difference of mean weight for gender.
-Thus for our experiment I generated values similar to a variable like age.
-the discrete variable was thought similar to gender.}
 Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
@@
 To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
-The actual ydata plot was generated by making a side-by-side boxplot of the residuals (Table\ref{tbl:stat_multiple} case 3).
-The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2$, and these were plotted in the same way as side-by-side boxplots.
+The actual data plot was generated by making a side-by-side boxplot of the residuals (Table \ref{tbl:stat_multiple}, case 3).
+The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and these were plotted in the same way as side-by-side boxplots.
 The actual data plot was randomly placed among these null data plots to produce the lineup.
@@
 The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
-
-%\green{I think it is plot11, not 13!!!!!!!! Please check!}
 
@@
 %\end{enumerate}
-
- %Figure \ref{fig:test_category} is an example of one of these lineups.
-The visual test statistic is a side-by-side boxplot.
-The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
 
@@
 For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
-The 10 lineups are chosen as follows;
+The 10 lineups are choosen as follows;
 
@@
 \begin{table}[hbtp]
-\caption{Number of participants and their total feedbacks after applying five screening process on all three experimental data sets.}
-\begin{center} {\small
+\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
+\begin{center}
 \begin{tabular}{rrrrr|rrrr|rrrr}
@@
    \hline
-\end{tabular} }
+\end{tabular}
 \end{center}

### 2012-06-06 paper_jasa.tex
fixed conflicts

--- 2012-06-06 paper_jasa.tex
@@
        \scalebox{.65}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.}
+       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.
+\hh{what is $m$?
+- would it make sense to show a curves for varying values of $m$?}}
        \label{fig:power_expected}
@@
 \begin{table}[hbtp]
+
 \caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
@@
   4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\
+
   5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 

### 2012-06-06 paper_jasa.tex
re-worked the section on continuous covariate

@@
 It was best for a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
-The explanations about each experiment (below) includes and explanation of how the lineups were sampled and provided to the observers.
+The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the observers, for each of the three experiments.
+
+\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.}
 
@@
 
-The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
+The experiment is designed to study the ability of human observers to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
 Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
@@
 
-This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous.
-$X_1$ is simulated from standard normal(0,1).
-Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
-The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}.
-We arbitrarily fix $\beta_0 = 6$ while generating actual data.
-For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups.
-For some parameter combinations we have more than three replications.
+This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2.
+Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}.
+We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
+For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
+\green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications.
 Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.
 
-{\bf Steps to obtain lineup}
-\begin{enumerate}
-	\item For a given parameter setting obtain actual data set
-	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
-	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
-	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-\end{enumerate}
-
-A lineup for this experiment is shown in figure \ref{fig:test_continuous}.
+The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid.
+To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data.
+An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
 The actual data plot location is 11.
+\green{Check this} This lineup, we think, is a difficult one.
+To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
+For the smallest slopes the difficulty is 0 (hardest) and for the largest slopes the difficulty is 4 (easiest).
+For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
+Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
+The order that the lineups are shown to participants is randomized.
+\green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups.
+The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants.
+It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones.
+\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
+
+%{\bf Steps to obtain lineup}
+%\begin{enumerate}
+%	\item For a given parameter setting obtain actual data set
+%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+%	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
+%	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
+%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+%\end{enumerate}
 
@@
 
-{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
+%{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
 For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}.
@@
 
-\begin{enumerate}
-	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
-	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
-	\item Finally select another lineup with difficulty level 0
-\end{enumerate}
+%\begin{enumerate}
+%	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
+%	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
+%	\item Finally select another lineup with difficulty level 0
+%\end{enumerate}
 

### 2012-06-06 paper_jasa.tex
Finished re-working the contaminated data section

@@
 
+\green{Discuss the UMP test, and give the reference for it here.}
+
 \subsection{Continuous covariate}
@@
 
+\green{Discuss the UMP test, and give the reference for it here.}
+
 \subsection{Contaminated data}
 
-The first two simulation experiments are based on normal error model which satisfies the conditions for uniformly most powerful (UMP) test.
-To study how the proposed visual inference work for a scenario where the data are not really coming from a normal model we design a contaminated data experiment.
-To generate the contaminated data set of size $n=100$ we used the following model
+The first two simulation experiments use data generated with a normal error model, satisfying the conditions for uniformly most powerful (UMP) test procedures.
+In these situations there exists a test, and there would, in general, be no need to use visual inference.
+The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the UMP, will compare favorably with it.
+This third simulation is closer to the mark for the purpose of visual inference.
+The assumptions for the UMP are violated, by contaminating the data.
+The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.
+The data is generated from the following model:
 
@@
 where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data.
-For the experiment we consider $n=100$, $c=0.15$ which produced actual data size 115, $\alpha=0$ and $\lambda=10$ .
-$\sigma$ is chosen such that overall error standard deviation of the data is 5 which is approximately 3.5.
-One such plot of the contaminated data for  $\beta$=5 is shown in figure \ref{fig:cont_dat}.
-We fit model (Equation \eqref{multi}) with $p=1$  and intercept $\beta_0=0$ to this contaminated data to test the hypothesis whether slope is zero or not.
-
-{\bf Steps to obtain lineup}
-\begin{enumerate}
-	\item For a given parameter setting obtain actual data set
-	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
-	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-\end{enumerate}
-
-One lineup for slope $\beta=0.4$ is shown in figure \ref{fig:test_contaminated}.
-For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}.
-This produces 30 lineups.
-each subject got at least 10 lineups to evaluate.
-Based on the slope parameter we group lineups into six groups with difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.
-Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
+For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points.
+Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????}.
+Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
+A linear model (Equation \eqref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
+This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2.
+The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.
+
+Five replications of actual data sets for each slope value shown in Table \ref{tbl:experiment_params}, producing 30 lineups.
+Six groups of difficulty level based on slope are defined, difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.
+\green{This is the reverse of the previous experiment}  Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
+
+%{\bf Steps to obtain lineup}
+%\begin{enumerate}
+%	\item For a given parameter setting obtain actual data set
+%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
+%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
+%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
+%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
+%\end{enumerate}
+
+An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? Plot 7 is the actual data plot.
 

### 2012-06-06 power.R
edited figure 2

@@ -7,7 +7,8 @@ pb=c(seq(0.05,0.5, by=0.05), seq(0.25, 0.5, by=0.05))))
 powerdf$power <- with(powerdf, (1-pb)^(m-1))
 
 K <- 10
-qplot(m, power, geom="line", group=pb, data=powerdf, colour=pb) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B])) + geom_ribbon(aes(x=m, ymin=1/m-1.96*sqrt(1/m*(m-1)/m)/K, ymax=1/m+1.96*sqrt(1/m*(m-1)/m)/K), fill="grey80", lty=0, alpha=0.1)
+ggplot(aes(x=m, power,  group=pb, colour=pb), data=powerdf) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f", pb), show_guide=FALSE), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B])) + geom_ribbon(aes(x=m, ymin=1/m-1.96*sqrt(1/m*(m-1)/m)/K, ymax=1/m+1.96*sqrt(1/m*(m-1)/m)/K), fill="grey80", lty=0, alpha=0.1) +
+geom_line()
 ggsave(file="powerplot.pdf")
 
 

### 2012-06-06 paper_jasa.tex
small updates

@@
 \begin{center}
-\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
-$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\\hline
-1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
-&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
-&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
-&&&  &    &        &   &   &        & 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
+\begin{tabular}{rrrrrrrrrrrrrrrrrrr}
+$K$ &  $x$ & $p$-value && $K$ &  $x$ & $p$-value && $K$ &  $x$ & $p$-value && $K$  & $x$ & $p$-value && $K$  & $x$ & $p$-value\\ \cline{1-3} \cline{5-7}\cline{9-11}\cline{13-15}\cline{17-19}
+1 &  1 & 0.0500 && 2 &  1 & 0.0975 && 3 & 1 & 0.1426 && 4 & 1 & 0.1855 && 5 & 1 & 0.2262 \\%\cline{1-3}
+&&&& 2 &  2 & 0.0025 && 3 & 2 & 0.0073 && 4 & 2 & 0.0140 && 5 & 2 & 0.0226 \\
+&&&& &&&& 3 & 3 & 0.0001 && 4 & 3 & 0.0005 && 5 & 3 & 0.0012 \\%\cline{1-6}
+&&&&  &    &        &   &   &        &&& 4 & 4 & $< 0.0001$ && 5 & 4 & $< 0.0001$ 
 \end{tabular}

### 2012-06-06 paper_jasa.tex
updating comments

@@
 \hh{what is $m$?
-- would it make sense to show a curves for varying values of $m$?}}
+- would it make sense to show a curves for varying values of $m$?} \blue{$m$=20 here.
+When $m$ changes the type-I error also changes.
+Thus I can produce the difference between power curves for each $m$.
+Do you suggest that?}}
        \label{fig:power_expected}
@@
 
-\green{Discuss the UMP test, and give the reference for it here.}
+\green{Discuss the UMP test \blue{(discussed right after equation \ref{power_exp}) }, and give the reference for it here \blue { I am looking for a good reference. Do you have any suggestion?}.}
 
@@
 For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
-\green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications.
-Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.
+\green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications \blue{(since we run an extension of this experiment with more lineups for some parameter combinations}.
+Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots \blue{(we can ignore this now as I see that blocking discussion in section \ref{sec:category} is removed.) }.
 
@@
 Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
-The order that the lineups are shown to participants is randomized.
-\green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups.
+The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after.
+The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups.
 The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants.

### 2012-06-06 paper_jasa.tex
comments updated

@@
 To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
-For the smallest slopes the difficulty is 0 (hardest) and for the largest slopes the difficulty is 4 (easiest).
+For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest).
 For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
@@
 Six groups of difficulty level based on slope are defined, difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.
-\green{This is the reverse of the previous experiment}  Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
+Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
 

### 2012-06-06 paper_jasa.tex
comments added

@@
 The actual data plot location is 11.
-\green{Check this} This lineup, we think, is a difficult one.
+\green{Check this} \blue{(Yes, it is 11 !)} This lineup, we think, is a difficult one.
 To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
@@
 For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points.
-Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????}.
+Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????} \blue{(Yes, this is how we did it.
+This value (3.5) makes the overall sd to be 5 because of the inclusion of the contaminated part of the data.)}.
 Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.

### 2012-06-07 paper_jasa.tex
power curves with different screening criteria

@@
 
+Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
+Summary of the data we obtained for all three experiments are shown in table \ref{tbl:summary}.
+Participants are asked to select the plot they think best match the question asked, give a reason for their choice, and determine a level of  confidence for their decision.
+Gender, age, education and geographic location of each participant are also collected.
+In this process of collecting data some participants did not provide demographic information as we see in table \ref{tbl:summary} that male and female participants do not add up to total participants.
+It is not clear why they did not provide the information.
+Now the concern arises whether we should keep their responses in the study.
+It is also possible that participants were not serious enough to provide feedback.
+For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
+This could be a criteria to determine their attentiveness in the study.
+For this we consider five different data cleaning criteria as shown below;
+
 {\bf Cleaning data:} There are five suggested approach to clean the data. They are 
-
 \begin{enumerate}
@@
 \item only include people with at least 20 percent correct (current suggestion)
-\item include people with 50\% on all easy lineups (as suggested before)
+\item include people with 50\% correct on all very easy (could be evaluated correctly if the participant is not ignorant) lineups (as suggested before)
 \end{enumerate}
@@
 \begin{table}[hbtp]
-
 \caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
@@
   4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-
   5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
@@
 
-Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
+We fit model \ref{mixed} with effect $E=\beta*n/\sqrt(n)$ as the only fixed effect covariate without intercept.
+We also set $offset = log(0.05/95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
+The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
+This shows how the screening criteria may affect the results.
+The corresponding power curves of UMP test are also shown.
+Notice that no mater what criteria we apply the result does not change much for experiment 2.
+Also, estimated power curve with criteria 5 for experiment 3 is above the UMP test power curve.
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
+       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
+The corresponding power curve for Uniformly Most powerful (UMP) test is shown in black solid lines for comparison.}
+       \label{fig:power_screening}
+\end{figure}
+
 Each participant was shown a sequence of 10 lineup plots.
-Participants are asked to select the plot with the biggest vertical difference, give a reason for their choice, and determine a level of  confidence for their decision.
-Gender, age, education and geographic location of each participant are also collected.
 In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
@@
 
-We fit model \eqref{mixed} to the survey data obtained from the simulation experiment.
+%We fit model \eqref{mixed} to the survey data obtained from the simulation experiment.
 The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}.

### 2012-06-07 R/paper_jasa.R

@@ -4,6 +4,13 @@
 library(ggplot2)
 library(plyr)
 library(reshape)
+library(xtable)
+library(lme4)
+
+
+dat1 <- read.csv("../data/raw_data_turk1.csv")
+dat2 <- read.csv("../data/raw_data_turk2.csv")
+dat3 <- read.csv("../data/raw_data_turk3.csv")
 
 
 # =================== mathematical test t distribution  ===================
@@ -144,7 +151,7 @@ pl <- generate_turk3_lineup(n=100,sigma=3.5,beta=0.4)
 pl$result
 ggsave(plot=pl$p, file="../images/lineup_contaminated.pdf",height=7,width=7.5)
 
-# ==== plotting test statistics --------
+# ==== plotting visual test statistics --------
 
 generate_stat_slope <- function(n,beta,sigma){
 	#n <- 100; a <- 6; b <- 2; sigma <- 12
@@ -195,15 +202,78 @@ get_screened_summary <- function(dat){
   return(s)
 }
 
-dat1 <- read.csv("../data/raw_data_turk1.csv")
-dat2 <- read.csv("../data/raw_data_turk2.csv")
-dat3 <- read.csv("../data/raw_data_turk3.csv")
-
 sm <- cbind(get_screened_summary(dat1),
       get_screened_summary(dat2),
       get_screened_summary(dat3))
-library(xtable)
+#library(xtable)
 xtable(sm)
+
+# --- fit model with each screening criteria and plot power
+#library(lme4)
+
+pred.mixed <- function(X, intercept=0, fit) {
+  alpha <- log(.05/0.95)
+  eta <- alpha + X * fixef(fit) + intercept
+  g.eta <- exp(eta)/(1+exp(eta))
+  return(g.eta)
+}
+
+get_predict_mixed <- function(dat, newdat){
+  dat$effect <-  with(dat,abs(beta)/sigma*sqrt(sample_size))
+  dat$alpha <- log(.05/0.95)
+  fit.mixed <- lmer(response ~ offset(alpha) + effect -1 + (effect-1|id),
+                    family="binomial",
+                    data=dat) 
+  res <- data.frame(effect=newdat$effect,pred=pred.mixed(X=newdat$effect, fit=fit.mixed))
+  return(res)
+}
+effect <- seq(0.01,16, by=.2)
+get_predict_mixed(dat1, newdat=data.frame(effect))
+
+
[…]

### 2012-06-07 R/turk1_data_analysis_jasa.R

@@ -1129,7 +1129,7 @@ ggplot(data=effect.d,aes(x=effect,y=power-ump_power)) +
 library(lme4)
 
 dat$alpha <- log(.05/0.95)
-fit.mixed <- lmer(response ~ offset(alpha) + effect -1 + (effect|id),
+fit.mixed <- lmer(response ~ offset(alpha) + effect -1 + (effect-1|id),
                   family="binomial",
                   data=dat)
 qplot(group=effect, x=resid(fit.mixed), geom="histogram", data=dat, facets=~effect, binwidth=0.2)
@@ -1149,6 +1149,23 @@ Z <- cbind(ones,effect)
 # vv <- matrix(as.numeric(c(res@REmat[1,3],rep(cov,2),res@REmat[2,3])),ncol=2, byrow=F)
 vv <- VarCorr(fit.mixed)$id
 
+
+#################
+pred.mixed <- function(X, intercept=0) {
+  alpha <- log(.05/0.95)
+  fit <- X * fixef(fit.mixed) + intercept + alpha
+  return(exp(fit)/(1+exp(fit)))
+}
+
+qplot(effect, pred(X=effect)) + ylim(c(0,1))
+
+newdata <- data.frame(expand.grid(effect=effect, subject=ranef(fit.mixed)[[1]][,1]))
+newdata$pred <- pred(newdata$effect, intercept=newdata$subject)
+
+
+qplot(effect, pred, group=subject, data=subset(newdata, subject %in% sample(size=2, x=unique(newdata$subject))), geom="line")
+
+#################
 # library(MASS) is needed for mvrnorm
 
 power=NULL

### 2012-06-07 paper_jasa.tex
worked "Appliation to Linear Models" over

@@
 
-\begin{lemma}
+\begin{lemma}~\label{lemma}
 Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
@@
 
-Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the right hand side of the equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
@@
 
-In this setting, there is a multitude of established graphics that help with the evaluation and diagnostics of a model \citep{cook:99}.
-Table \ref{tbl:stat_multiple} shows an overview of possible hypotheses that would be tested in the regression setting, and what visual test statistic might be used in these tests.
-These plots shoul be familiar, because they are all commonly used for evaluating and diagnosing regression models.
+In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g.
+\citet{cook:99}).
+Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and the visual test statistic that might be used for testing these hypotheses.
+These plots should be familiar, because they are all in common use.
 For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
 To assess whether the linear assumption is appropriate we would make a plot of residuals against fitted values (case 5).
+For the purposes of comparing visual against classical inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
 
@@
 \begin{table*}[hbtp] 
-\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
+\caption{Visual test statistics for testing hypotheses related to the model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
 \centering 
@@
 
-For the purposes of validating visual inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
-
 Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are significantly different and we fit the null model to the actual data, the resulting residual plot shows two groups of residuals.
-To test this we generate side-by-side boxplots of the residuals conditioned on the two levels of $X_k$, as the test statistic.
-If $\beta_k\ne 0$ the boxplots show a vertical displacement.
+If the responses for the two levels of the categorical variable $X_k$ in the model are different and we fit the null model to the actual data, the resulting residuals should show a significant difference between the two groups.
+To test this boxplots of the residuals conditioned on the two levels of $X_k$ are made, as the observed test statistic.
+If $\beta_k\ne 0$ the boxplots should show a vertical displacement.
 
@@
 
-Under the assumption that an observer is able to pick the plot with the smallest $p$-value from a lineup plot,  this leads to the decision to reject $H_0$ when $p_{B} < p_0$.
-Thus we have the expected power as
+The classical test in this scenario uses $T= \hat{\beta_k}/
+se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme
+on the scale of a $t$ distribution with $n-p$ degrees of freedom. This
+is the uniformly most powerful (UMP) test \green{reference needed
+  here} and it forms the benchmark upon which we evaluate the visual
+test. To calculate what we might expect for the power of the visual
+test, under perfect conditions, first assume that the observer is able
+to pick the plot with the smallest $p$-value from a lineup plot.  This
+leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}. Thus 
+the expected power of the visual test in this scenario is
 
@@
 \end{equation}
-
-The classical uniformly most powerful (UMP) test rejects $H_0$ if the observed test statistic $T= \hat{\beta_k}/ se(\hat{\beta_k})$ takes an extreme value in the distribution of theoretical $t$ distribution with $n-p$ degrees of freedom.
-Figure \ref{fig:power_expected} shows the power of the UMP test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
+Figure \ref{fig:power_expected} shows the power of the UMP test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
 Notice that the expected power of the visual test is almost as good as the power of UMP test.
-In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
+\green{What do you mean by this?
+Where does this ``fact'' come from?} In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
 
@@
        \scalebox{.65}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.
-\hh{what is $m$?
-- would it make sense to show a curves for varying values of $m$?} \blue{$m$=20 here.
-When $m$ changes the type-I error also changes.
-Thus I can produce the difference between power curves for each $m$.
-Do you suggest that?}}
+       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$, $\sigma = 12$ and $m$=20. \green{Produce these curves for $m=12$, $m=20$, $m=30$.}}
        \label{fig:power_expected}
@@
 Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
-A linear model (Equation \eqref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
+A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
 This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2.
@@
        \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
-       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \eqref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
+       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
 }
@@
        \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
        \label{fig:power_observed}
@@
        \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-       \caption{Observed power of visual test from equation \eqref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
 }
@@
        \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \eqref{plot_signal}}
+       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}}
        \label{fig:pval_plot_signal}
@@
        \scalebox{0.40}{\includegraphics{power_model.pdf}}
-       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
+       \caption{Estimated power curve from Equation \ref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}

### 2012-06-07 paper_jasa.tex
subject specific power curves added

@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-       \label{fig:power_observed}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-       \label{fig:power_loess1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
-       \label{fig:power_loess2}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect_subject1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
-       \label{fig:power_loess3}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
-       \label{fig:power_diff_exp}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-}
-       \label{fig:power_observed_exp3}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
-       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 1.5 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in figure \ref{fig:pval_power}}
-       \label{fig:pval_pcorrect}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}}
-       \label{fig:pval_plot_signal}
-\end{figure*}
-
 \subsection{Results}
@@
 
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
+       \caption{The subject wise power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
+The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}
+       \label{fig:power_screening_subject}
+\end{figure}
+
 Each participant was shown a sequence of 10 lineup plots.
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.40}{\includegraphics{power_model.pdf}}
-       \caption{Estimated power curve from Equation \ref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.40}{\includegraphics{power_model.pdf}}
+%       \caption{Estimated power curve from Equation \ref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-       \label{fig:power_model}
-\end{figure}
+%       \label{fig:power_model}
+%\end{figure}
 
@@
 Notice that the power curve estimated for one subject is above the UMP test power curve.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
-       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.
+%
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
+%       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-       \label{fig:power_subject}
-\end{figure}
+%       \label{fig:power_subject}
+%\end{figure}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
+       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
+After p-value exceeds 1.5 it is very unlikely to identify the actual plot.
+The theoretical justification of this is shown in figure \ref{fig:pval_power}}
+       \label{fig:pval_pcorrect}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
+       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}}
+       \label{fig:pval_plot_signal}
+\end{figure*}
 
@@
 the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
+       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+       \label{fig:power_observed}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+       \label{fig:power_loess1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
+       \label{fig:power_loess2}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
+       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \label{fig:power_effect1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
+       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \label{fig:power_effect_subject1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
+Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
+Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
+       \label{fig:power_loess3}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
+       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
+       \label{fig:power_diff_exp}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
+       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+}
+       \label{fig:power_observed_exp3}
+\end{figure*}
 

### 2012-06-07 R/paper_jasa.R

@@ -211,31 +211,37 @@ xtable(sm)
 # --- fit model with each screening criteria and plot power
 #library(lme4)
 
-pred.mixed <- function(X, intercept=0, fit) {
+pred.mixed <- function(X, subject=0,fit) {
   alpha <- log(.05/0.95)
-  eta <- alpha + X * fixef(fit) + intercept
+  eta <- alpha + X * fixef(fit) + subject
   g.eta <- exp(eta)/(1+exp(eta))
   return(g.eta)
 }
 
-get_predict_mixed <- function(dat, newdat){
+get_predict_mixed <- function(dat, newdat, intercept=F){
   dat$effect <-  with(dat,abs(beta)/sigma*sqrt(sample_size))
   dat$alpha <- log(.05/0.95)
   fit.mixed <- lmer(response ~ offset(alpha) + effect -1 + (effect-1|id),
                     family="binomial",
                     data=dat) 
-  res <- data.frame(effect=newdat$effect,pred=pred.mixed(X=newdat$effect, fit=fit.mixed))
+  X <- newdat$effect
+  if(intercept){
+    subject <- ranef(fit.mixed)[[1]][,1]
+    d <- data.frame(expand.grid(effect=X, subject=subject))
+    pred <- pred.mixed(d$effect, subject=d$subject, fit=fit.mixed)
+    res <- data.frame(effect=d$effect, subject=d$subject, pred)
+  } else res <- data.frame(effect=X,pred=pred.mixed(X, fit=fit.mixed))
   return(res)
 }
 effect <- seq(0.01,16, by=.2)
 get_predict_mixed(dat1, newdat=data.frame(effect))
+#get_predict_mixed(dat1, newdat=data.frame(effect), intercept=T)
 
-
-get_screened_predict <- function(dat, newdat){
+get_screened_predict <- function(dat, newdat,intercept=F){
   indx <- get_sceering_index(dat)
   res <- NULL
   for (i in 1:5){
-    pred <- get_predict_mixed(subset(dat,indx[,i]), newdat)
+    pred <- get_predict_mixed(subset(dat,indx[,i]), newdat, intercept)
     res <- rbind(res, data.frame(screening=i,pred)) 
   }
   return(res)
@@ -263,6 +269,19 @@ ggsave( file="../images/power_screening.pdf",height=4.25,width=10)
 
 
 # ----- subject specific power from mixed model
+
+
[…]

### 2012-06-07 paper_jasa.tex
lots of cleanup everywhere, division of results section

@@
 This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
-One of the protocols, the lineup, places the data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
+One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
 Figure \ref{fig:test_category} shows an example lineup.
@@
 The null hypothesis typically derives from the task at hand, or the type of plot being made.
-The alternative encompasses all possible antitheses, all types of patterns that might be detected in the data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
+The alternative encompasses all possible antitheses, all types of patterns that might be detected in the actual data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
 The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.
@@
 In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
-If the data plot is selected as the most different, then this results in rejection of the null hypothesis.
+If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.
 
@@
 \\ Should I change everything accordingly?
+I see that we have been careful not to use $\beta$ for power!
 }
@@
 \end{enumerate}
-We will be able to show evidence for the second assumption with the help of the study discussed in section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup.
+We will be able to show evidence for the second assumption with the help of the study discussed in Section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup.
 The better suited a design is for a particular task, the higher its power will be.
@@
 
-Figure \ref{fig:pval_power} gives an overview of the probability of picking the data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
-We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$.
-This pattern is confirmed by our findings described in section~ref{sec:simulation}
+Figure \ref{fig:pval_power} gives an overview of the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
+We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the actual data plot from null plots in a lineup of size $m=20$.
+This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
 %also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
@@
    \centering
-   \includegraphics[width=3.5in]{images/powerplot.pdf} 
-   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_D$.
-The grey line shows probability of picking the data plot under $H_0$.
+   \includegraphics[width=3.8in]{images/powerplot.pdf} 
+   \caption{Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
+The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true.
 }
@@
 Now assume that the same lineup plot is evaluated by $K$ independent observers.
-W.l.o.g.
-let the last plot be the data plot and plots 1 through $m-1$ the null plots.
+W.l.o.g.\green{Can we get rid of cliches like WLOG?} let the last plot be the data plot and plots 1 through $m-1$ the null plots.
 Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
@@
 
-Note that the right hand side of the equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the right hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
@@
 
-The classical test in this scenario uses $T= \hat{\beta_k}/
-se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme
-on the scale of a $t$ distribution with $n-p$ degrees of freedom. This
-is the uniformly most powerful (UMP) test \green{reference needed
-  here} and it forms the benchmark upon which we evaluate the visual
-test. To calculate what we might expect for the power of the visual
-test, under perfect conditions, first assume that the observer is able
-to pick the plot with the smallest $p$-value from a lineup plot.  This
-leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}. Thus 
-the expected power of the visual test in this scenario is
+The classical test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
+This is the uniformly most powerful (UMP) test \green{reference needed here} and it forms the benchmark upon which we evaluate the visual test.
+(This is also the UMP test for the continuous covariate, case 2.) To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
+This leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}.
+Thus the expected power of the visual test in this scenario is
 
@@
 
-\section{Simulation Experiments} \label{sec:simulation}
+\section{Human Subjects Experiments with Simulated Data} \label{sec:simulation}
 
@@
 The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the observers, for each of the three experiments.
+
+Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
+A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}.
+Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice.
+Gender, age, education and geographic location of each participant are also collected.
+The data was collected through the web.
+For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.
 
@@
 
-\green{Discuss the UMP test \blue{(discussed right after equation \ref{power_exp}) }, and give the reference for it here \blue { I am looking for a good reference. Do you have any suggestion?}.}
-
 \subsection{Continuous covariate}
@@
 For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
-\green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications \blue{(since we run an extension of this experiment with more lineups for some parameter combinations}.
-Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots \blue{(we can ignore this now as I see that blocking discussion in section \ref{sec:category} is removed.) }.
+For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
 
@@
 The actual data plot location is 11.
-\green{Check this} \blue{(Yes, it is 11 !)} This lineup, we think, is a difficult one.
+This lineup, we think, is a difficult one.
 To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
@@
 Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
-The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after.
-The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups.
+The order that the lineups are shown to participants is randomized.
+Each subject saw a substantial number of difficult lineups.
 The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants.
@@
 \end{figure}
-
-\green{Discuss the UMP test, and give the reference for it here.}
 
@@
 For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points.
-Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????} \blue{(Yes, this is how we did it.
-This value (3.5) makes the overall sd to be 5 because of the inclusion of the contaminated part of the data.)}.
+Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
 Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
@@
 
-An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? Plot 7 is the actual data plot.
+An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? %Plot 7 is the actual data plot. 
+\begin{figure}[hbt]
+%\begin{figurehere}
+   \centering
+       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
+       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
+}
+       \label{fig:cont_dat}
+\end{figure}
 
@@
 
-\begin{figure}[hbt]
-%\begin{figurehere}
-   \centering
-       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
-       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
+%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+%       \label{fig:power_observed}
+%\end{figure*}
+
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
+%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+%       \label{fig:power_loess1}
+%\end{figure*}
+
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
+%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
+%       \label{fig:power_loess2}
+%\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
+       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \label{fig:power_effect1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
+       \caption{Subject specific power obtained from Equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \label{fig:power_effect_subject1}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
+Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
+Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
+       \label{fig:power_loess3}
+\end{figure*}
+
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
+%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
+%       \label{fig:power_diff_exp}
+%\end{figure*}
+
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
+%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
 }
-       \label{fig:cont_dat}
-\end{figure}
-
-\subsection{Results}
-
-Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
-Summary of the data we obtained for all three experiments are shown in table \ref{tbl:summary}.
-Participants are asked to select the plot they think best match the question asked, give a reason for their choice, and determine a level of  confidence for their decision.
-Gender, age, education and geographic location of each participant are also collected.
-In this process of collecting data some participants did not provide demographic information as we see in table \ref{tbl:summary} that male and female participants do not add up to total participants.
+%       \label{fig:power_observed_exp3}
+%\end{figure*}
+
+\section{Results}
+
+\subsection{Data Cleaning}
+
+% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants.
 It is not clear why they did not provide the information.
@@
  \cline{2-5}  \cline{6-9}   \cline{10-13}
-criteria & total & male & female & feedbacks & total & male & female & feedbacks & total & male & female & feedbacks \\ 
+criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
   \hline
@@
 \end{table}
+
+\subsection{Power comparison}
 
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-       \caption{The subject wise power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}
-       \label{fig:power_screening_subject}
-\end{figure}
-
 Each participant was shown a sequence of 10 lineup plots.
 In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
-The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{power_est} along with 95\% confidence interval calculated using Fisher's exact method.
+The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using Equation \ref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
 
@@
 Notice that the power curve estimated for one subject is above the UMP test power curve.
-%
+
 %\begin{figure}[hbtp]
@@
 
+\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size.
+This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}
+
+\subsection{Subject-specific variation}
+
+\subsection{Estimating $p$-value in on-Classical Settings}
+
 \begin{figure*}[hbtp]
@@
 \end{figure*}
-
-\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size.
-This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-       \label{fig:power_observed}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-       \label{fig:power_loess1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
-       \label{fig:power_loess2}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect_subject1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
-       \label{fig:power_loess3}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
-       \label{fig:power_diff_exp}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-}
-       \label{fig:power_observed_exp3}
-\end{figure*}
-
-\green{A lot more needs to be done here!!!}
 

### 2012-06-07 paper_jasa.tex
subject specific power curves added again

@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-       \caption{Subject specific power obtained from Equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect_subject1}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
-       \label{fig:power_loess3}
-\end{figure*}
-
 %\begin{figure*}[hbtp]
@@
 
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
+       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
+       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \label{fig:power_effect1}
+\end{figure*}
+
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
+%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
+%       \caption{Subject specific power obtained from Equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+%       \label{fig:power_effect_subject1}
+%\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
+Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
+Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
+       \label{fig:power_loess3}
+\end{figure*}
+
 %\begin{figure}[hbtp]
@@
 
-\subsection{Estimating $p$-value in on-Classical Settings}
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
+      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
+The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}
+       \label{fig:power_screening_subject}	
+\end{figure}
+
+\subsection{Estimating $p$-value in Non-Classical Settings}
 

### 2012-06-07 R/paper_jasa.R

@@ -167,7 +167,7 @@ p <- generate_stat_slope(n=100,beta=2.5,sigma=5)
 ggsave(plot=p, file="../images/stat_beta_k.pdf",height=2,width=2.05)
 
 
-# ==== summary of the data =====
+# ==== summary of the data applying different screening criteria =====
 
 get_summary <- function(dat){
   g <- ddply(dat,.(gender),summarise, length(unique(id)))
@@ -208,7 +208,7 @@ sm <- cbind(get_screened_summary(dat1),
 #library(xtable)
 xtable(sm)
 
-# --- fit model with each screening criteria and plot power
+# --- fit mixed model with each screening criteria and plot power
 #library(lme4)
 
 pred.mixed <- function(X, subject=0,fit) {
@@ -268,7 +268,7 @@ ggplot()+
 ggsave( file="../images/power_screening.pdf",height=4.25,width=10)
 
 
-# ----- subject specific power from mixed model
+# ----- subject specific power for each screening criteria from mixed model
 
 
 pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(dat1, newdat=data.frame(effect=seq(0.01,18, by=.2)),intercept=T)),
@@ -282,15 +282,13 @@ ggplot()+
   ylab("power") + xlab(expression(Effect(E)))
 ggsave( file="../images/power_screening_subject.pdf",height=7,width=10 )  
 
-
-
-ump1 <- calculate_ump_power1(3, 100, 5)
-qplot(effect, pred.mixed(X=effect, fit=fit.mixed)) + ylim(c(0,1))
-
-newdata <- data.frame(expand.grid(effect=effect, subject=ranef(fit.mixed)[[1]][,1]))
-newdata$pred <- pred.mixed(newdata$effect, intercept=newdata$subject, fit=fit.mixed)
-
-qplot(effect, pred, group=subject, data=subset(newdata, subject %in% sample(size=2, x=unique(newdata$subject))), geom="line")
+# ump1 <- calculate_ump_power1(3, 100, 5)
+# qplot(effect, pred.mixed(X=effect, fit=fit.mixed)) + ylim(c(0,1))
+# 
+# newdata <- data.frame(expand.grid(effect=effect, subject=ranef(fit.mixed)[[1]][,1]))
+# newdata$pred <- pred.mixed(newdata$effect, intercept=newdata$subject, fit=fit.mixed)
+# 
[…]

### 2012-06-08 paper_jasa.tex
added criteria 6 for screening data

@@
 This could be a criteria to determine their attentiveness in the study.
-For this we consider five different data cleaning criteria as shown below;
-
-{\bf Cleaning data:} There are five suggested approach to clean the data. They are 
+For this we consider six different data cleaning criteria as shown below;
+
+{\bf Cleaning data:} There are six suggested approach to clean the data. They are 
 \begin{enumerate}
@@
 \item include people with 50\% correct on all very easy (could be evaluated correctly if the participant is not ignorant) lineups (as suggested before)
+\item each person saw at least one easy lineup.
+we select one of those easy lineups as the inclusion criteria.
+If that lineup is evaluated correctly we include only the rest of the evaluations of that subject.
+Otherwise, we exclude all the feedbacks of that subject.
 \end{enumerate}
@@
 \begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
+\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets.}
 \begin{center}
@@
   5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
+  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 &  98 &  64 &  34 & 1095 \\ 
    \hline
@@
        \scalebox{0.70}{\includegraphics{power_screening.pdf}}
-       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
+       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown in black solid lines for comparison.}
@@
        \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.
+      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}

### 2012-06-08 R/paper_jasa.R

@@ -186,19 +186,25 @@ get_sceering_index <- function(dat){
   indx3 <- dat$id %in% included_id3
   included_id4 <- d34$id[ d34$percent_correct >= 20 ]
   indx4 <- dat$id %in% included_id4
-  d4 <- ddply(subset(dat, p_value < 0.0002),.(id),summarise,
+  d56 <- ddply(subset(dat, p_value < 0.0002),.(id),summarise,
               easy_cnt = length(response),
-              percent_correct = mean(response)*100
+              percent_correct = mean(response)*100,
+              include_criteria_6 = response[1],
+              exclude_lineup_criteria_6 = paste(id[1],"_",pic_id[1],sep=""),
+              exclude_lineup_p_value = p_value[1]
               )
-  included_id5 <- d4$id[ d4$percent_correct > 49]
+  included_id5 <- d56$id[ d56$percent_correct > 49]
   indx5 <- dat$id %in% included_id5
-  return(data.frame(indx1,indx2,indx3,indx4,indx5))
+  included_id6 <- d56$id[d56$include_criteria_6]
+  indx6 <- dat$id %in% included_id6
+  indx6[paste(dat$id,"_",dat$pic_id,sep="") %in% d56$exclude_lineup_criteria_6] <- FALSE
+  return(data.frame(indx1,indx2,indx3,indx4,indx5,indx6))
 }
 
 get_screened_summary <- function(dat){
   indx <- get_sceering_index(dat)
   s <- NULL
-  for (i in 1:5) s <-  rbind(s,get_summary(subset(dat,indx[,i])))
+  for (i in 1:6) s <-  rbind(s,get_summary(subset(dat,indx[,i])))
   return(s)
 }
 

### 2012-06-08 paper_jasa.tex
added subsection which follows up checking lowest p-value pick
minimum p-value assumption holds

@@
 
+\subsection{Do people tend to pick the lowest $p$-value?}
+
+This section takes the data and looks at the assumption made for the early expected power.
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_rank_counts.pdf}}
+       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 with raw data (nothing is dropped).
+This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
+There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error.
+Can we?}
+       \label{fig:P-val_rank}
+\end{figure*}
+
 \section{Conclusions}

### 2012-06-08 R/paper_jasa.R

@@ -13,6 +13,7 @@ dat2 <- read.csv("../data/raw_data_turk2.csv")
 dat3 <- read.csv("../data/raw_data_turk3.csv")
 
 
+
 # =================== mathematical test t distribution  ===================
 
 t <- seq(-4, 4, by = 0.1)
@@ -246,23 +247,23 @@ get_predict_mixed(dat1, newdat=data.frame(effect))
 get_screened_predict <- function(dat, newdat,intercept=F){
   indx <- get_sceering_index(dat)
   res <- NULL
-  for (i in 1:5){
+  for (i in 1:6){
     pred <- get_predict_mixed(subset(dat,indx[,i]), newdat, intercept)
     res <- rbind(res, data.frame(screening=i,pred)) 
   }
   return(res)
 }
 get_screened_predict(dat1, newdat=data.frame(effect))
-pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(dat1, newdat=data.frame(effect=seq(0.01,18, by=.2)))),
-            data.frame(experiment="Experiment 2",get_screened_predict(dat2, newdat=data.frame(effect=seq(0.01,8, by=.2)))),
-            data.frame(experiment="Experiment 3",get_screened_predict(dat3, newdat=data.frame(effect=seq(0.01,8, by=.2)))))
+pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(dat1, newdat=data.frame(effect=seq(0,18, by=.2)))),
+            data.frame(experiment="Experiment 2",get_screened_predict(dat2, newdat=data.frame(effect=seq(0,7, by=.2)))),
+            data.frame(experiment="Experiment 3",get_screened_predict(dat3, newdat=data.frame(effect=seq(0,6, by=.2)))))
 set.seed(2035)
-ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0.01,18, by=.2),
-                        ump= calculate_ump_power1(beta=seq(0.01,18, by=.2)/10, n=100, sigma=1)),
-             data.frame(experiment="Experiment 2",effect=seq(0.01,8, by=.2),
-                        ump= calculate_ump_power2(beta=seq(0.01,8, by=.2)/10, n=100, sigma=1)),
-             data.frame(experiment="Experiment 3",effect=seq(0.01,8, by=.2),
-                        ump= calculate_ump_power3(beta=seq(0.01,8, by=.2)/sqrt(115), n=115, sigma=1, x=getx(100))))
[…]

### 2012-06-09 R/generate_expected_visual_power_simulation-hh.R
changed axis label to use greek letter, size of figure 3

@@ -1,58 +1,52 @@
 # This file computes the expected power for visual test.
 # Last modified Jan 19, 2012
-
-
 library(ggplot2)
 library(plyr)
 library(MASS)
 
 generate_visual_power <- function(n,beta,sigma, m=20){
-	#This program generates expected visual power by simulation.
-	age <- rpois(n,lambda=30)
-	grp <- factor(sample(c("A","B"), size=n, replace=T))
+  #This program generates expected visual power by simulation.
+  age <- rpois(n,lambda=30)
+  grp <- factor(sample(c("A","B"), size=n, replace=T))
 
-	X <- model.matrix(~age + factor(grp))
+  X <- model.matrix(~age + factor(grp))
 
+  sdbeta <- sigma *sqrt(ginv(t(X) %*% X)[3,3])
+  ts <- abs(rt(5000, ncp=beta/sdbeta, df=n-3))
+  pvals <- 2*pt(ts, n-3, lower.tail=FALSE)
 
-	sdbeta <- sigma *sqrt(ginv(t(X) %*% X)[3,3])
-	ts <- abs(rt(5000, ncp=beta/sdbeta, df=n-3))
-	pvals <- 2*pt(ts, n-3, lower.tail=FALSE)
-
-	
+  
 # minimum of (m-1) uniform random variables has a Beta (1, m-1) distribution
-	pow <- mean(pbeta(pvals, shape1=1, shape2=m-1, lower.tail=FALSE))
+  pow <- mean(pbeta(pvals, shape1=1, shape2=m-1, lower.tail=FALSE))
 # equivalent:
-#	pow <- mean(vpower(pvals, m=20))
+#  pow <- mean(vpower(pvals, m=20))
 
-	return(pow)
+  return(pow)
 }
-
-
 
 calculate_theoretical_power <- function (beta, n, sigma, alpha=0.05){
-	# standard deviation of beta depends on modelling situation
-	# only valid for experiment (1)
-	se_beta <- sigma/sqrt(n*(.5^2))  
-	mu <- beta/se_beta
-	# get quantile for rejection at alpha for H0: beta=0
-	t_n <- qt(p=1-alpha/2,df=n-3)
-	
-	# prob to reject for given beta
-	res <- pt(q=-t_n, df=n-3, ncp=mu)-pt(q=t_n, df=n-3, ncp=mu)+1
-	
-	return(res)
+  # standard deviation of beta depends on modelling situation
[…]

### 2012-06-09 paper_jasa.tex
changed axis label to use greek letter, size of figure 3
lemma 3.1

@@
 \[
-P(p_D < p_0) = \left(1 - E[p_D] \right)^{m-1} 
+P(p_D < p_0) = 1 - E\left[ (1 - p_D)^{m-1}\right]
 \]
@@
 
-Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
-%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
+\hh{Their minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function 
+\[
+F_{p_0} (x) =  (1- x) ^{m-1} \text{ for } x \in [0,1].
+\]
+
 \begin{eqnarray*}
-P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
+P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
+&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right],
 \end{eqnarray*}
-which, in particular, is independent of $p_{0,i}$ for all $i$.
-
-Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
+
+}
+%
+%
+%
+%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
+%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
+%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
+%\begin{eqnarray*}
+%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
+%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
+%\end{eqnarray*}
+%which, in particular, is independent of $p_{0,i}$ for all $i$.
+%
+%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
 Further we will assume that all observers  have the same ability in identifying the data plot.
-
-With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.
-
-Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
-\[
-P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
-\]
+%
+%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
+%
+%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
+%\[
+%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
+%\]
 \end{proof}
@@
    \centering
-       \scalebox{.65}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$, $\sigma = 12$ and $m$=20. \green{Produce these curves for $m=12$, $m=20$, $m=30$.}}
+       \scalebox{.95}{\includegraphics{power_expected.pdf}}
+       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
        \label{fig:power_expected}

### 2012-06-09 R/generate_expected_visual_power_simulation-hh.R
fixed merge

@@ -4,6 +4,12 @@ library(ggplot2)
 library(plyr)
 library(MASS)
 
+vpower <- function(pd, m=20) {
+# same as pbeta(pd, shape1=1, shape2=m-1, lower.tail=FALSE)	
+
+	(1 - pd)^(m-1)
+}
+
 generate_visual_power <- function(n,beta,sigma, m=20){
   #This program generates expected visual power by simulation.
   age <- rpois(n,lambda=30)
@@ -12,9 +18,10 @@ generate_visual_power <- function(n,beta,sigma, m=20){
   X <- model.matrix(~age + factor(grp))
 
   sdbeta <- sigma *sqrt(ginv(t(X) %*% X)[3,3])
+# simulate from non central t distribution
   ts <- abs(rt(5000, ncp=beta/sdbeta, df=n-3))
+# compute p-values 	
   pvals <- 2*pt(ts, n-3, lower.tail=FALSE)
-
   
 # minimum of (m-1) uniform random variables has a Beta (1, m-1) distribution
   pow <- mean(pbeta(pvals, shape1=1, shape2=m-1, lower.tail=FALSE))

### 2012-06-09 paper_jasa.tex
still lemma 3.1

@@
 P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right],
+&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right] = \\
+&=& 1 - \sum_{k=0}^{m-1} {m-1 \choose k} (-E[p_D])^{k} = (m-1)E[p_D] + O(E[p_D^2]),
 \end{eqnarray*}
 
+This can't be right - it's completely counter-intuitive: as $p_D$, the $p$-value is smaller, the probability that we pick the data plot should go up not down. 
+Conversely, as $m$, the number of choices given, goes up the probability to pick the data plot should go down, not up. 
 }
@@
 
-Now assume that the same lineup plot is evaluated by $K$ independent observers.
-W.l.o.g.\green{Can we get rid of cliches like WLOG?} let the last plot be the data plot and plots 1 through $m-1$ the null plots.
+Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
+Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
 Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
@@
 \begin{array}{ll}
-B_{K, (1-E[p_D])^{m-1}} & \text { under } H_1\\
+B_{K, 1-E[1 - p_D])^{m-1}} & \text { under } H_1\\
 B_{K, 1/m} & \text { under } H_0

### 2012-06-09 paper_jasa.tex
lemma 3.1 final? - figure 3 fixed

--- 2012-06-09 paper_jasa.tex
@@
 \[
-P(p_D < p_0) = 1 - E\left[ (1 - p_D)^{m-1}\right]
+P(p_D < p_0) =  E\left[ (1 - p_D)^{m-1}\right]
 \]
@@
 \[
-F_{p_0} (x) =  (1- x) ^{m-1} \text{ for } x \in [0,1].
+F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
 \]
@@
 P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt = 1 - E\left[ (1 - p_D)^{m-1}\right],
+&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt =  E\left[ (1 - p_D)^{m-1}\right].
 \end{eqnarray*}
 
+This proves the statement above. We further see:
+\[
+E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-E[p_D])^{k} = 1-(m-1)E[p_D] + O(E[p_D^2]).
+\]
+This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and,
+conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
 }
@@
 \begin{array}{ll}
-B_{K, (1-E[p_D])^{m-1}} & \text { under } H_1\\
+B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
 B_{K, 1/m} & \text { under } H_0
@@
 \begin{equation}\label{eqn:power_estimate}
-{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
+%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
+\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
 \end{equation}
 
-Note that the right hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the \hh{left}  hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.

### 2012-06-09 power.R

@@ -1,18 +1,26 @@
 
 library(ggplot2)
 
-powerdf <- data.frame(expand.grid(m = 5:50,
-pb=c(seq(0.05,0.5, by=0.05), seq(0.25, 0.5, by=0.05))))
+powerdf <- data.frame(expand.grid(pd = seq(0.0001, 0.20, by=0.001), m = seq(10, 30, by=5)))
+powerdf$power <- with(powerdf, pbeta(pd, 1, m-1, lower.tail=FALSE))
+powerdf$size <- with(powerdf, I(.75+(m==20)))
+powerdf$labels <- paste("m =",powerdf$m)
+powerdf$hjust <- c(-0.15,1.15)[(powerdf$m > 20) + 1]
+powerdf$power <- with(powerdf, pmax(power, 1/m))
 
-powerdf$power <- with(powerdf, (1-pb)^(m-1))
+qplot(pd, power, data=powerdf, geom="line", group=m, size=size) + ylab("Prob that data plot has lowest p-value") + xlab(expression(p[D])) + scale_size_identity() +geom_text(aes(x=pd, y=power, label=labels, hjust=hjust), size=3, data=subset(powerdf, abs(4*pd - power) < 0.005 )) + geom_hline(y=1/seq(15,30, by=5), linetype=2, alpha=0.3) 
 
-K <- 10
-ggplot(aes(x=m, power,  group=pb, colour=pb), data=powerdf) + geom_line(aes(m, 1/m), colour="grey30", lty=2) + geom_text(aes(label=sprintf("%.2f", pb), show_guide=FALSE), data=subset(powerdf, m==5), hjust=1.1) + xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B])) + geom_ribbon(aes(x=m, ymin=1/m-1.96*sqrt(1/m*(m-1)/m)/K, ymax=1/m+1.96*sqrt(1/m*(m-1)/m)/K), fill="grey80", lty=0, alpha=0.1) +
-geom_line()
 ggsave(file="powerplot.pdf")
 
 
+
+
+
 powerdf$EX <- with(powerdf, (m-1)*(1-pb))
+
+
+ggplot(aes(x=pb, m,  group = power, colour=power), data=powerdf) + geom_point()
+
 
 
[…]

### 2012-06-09 R/paper_jasa.R
new figure to verify pvalue pick

@@ -318,11 +318,24 @@ p <- qplot(factor(rank_pval), counts, geom="boxplot", data=prr) +
   xlab("rank of p-value") + ylab("Number of subjects")
 ggsave(file="../images/p_val_rank_counts.pdf", height=4, width=7)
 
-prr$pic_name2 <- str_sub(prr$pic_name, 12, 20)
-p <- qplot(pvalue, counts, geom="line", data=prr) +
+prr$pic_name2 <- str_sub(prr$pic_name, 12, -5L)
+x <- str_split(prr$pic_name2, "_")
+prr$n <-  unlist(lapply(x, function(x) x[1]))
+prr$beta <-  unlist(lapply(x, function(x) x[2]))
+prr$beta <-  factor(prr$beta, levels=c("0", "1", "2", "3", "5", "7", "8", "10", "16"))
+prr$sd <-  unlist(lapply(x, function(x) x[3]))
+prr$rep <-  unlist(lapply(x, function(x) x[4]))
+prr$pic_name3 <- factor(paste(prr$beta, prr$n, prr$sd, prr$rep, sep="."))
+p <- qplot(pvalue, counts, geom="point", data=prr) +
   xlab("rank of p-value") + ylab("Number of subjects") +
-  facet_wrap(~pic_name2, ncol=10, scales="free_y") + scale_x_log10()
+  facet_wrap(~pic_name3, ncol=10, scales="free_y") + scale_x_log10()
+prr$lpvalue<-log10(prr$pvalue+0.01)
+p <- qplot(lpvalue, counts, geom="point", data=prr) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
+  xlab(expression(paste(log[10], " p-value"))) + ylab("Number of subjects") +
+  facet_grid(beta~n+sd+rep, scales="free_y", labeller="label_both") 
 p
+ggsave(file="../images/p_val_log_counts.pdf", height=10, width=12)
 
 # =================== Turk1 data analysis  ================================
 

### 2012-06-09 paper_jasa.tex
new figure to verify pvalue pick
experiment2 p-value data and plot added

@@
        \scalebox{0.6}{\includegraphics{p_val_rank_counts.pdf}}
-       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 with raw data (nothing is dropped).
+       \scalebox{0.6}{\includegraphics{p_val_rank_counts2.pdf}}
+       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 (bottom) with raw data (nothing is dropped).
 This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
@@
        \label{fig:P-val_rank}
+\end{figure*}
+
+\green{I think it is very interesting to look at the individual lineups by beta, and other parameters, and we can see the tendency to pick the lowest $p$-value. Figure \ref{fig:P-val_log}.}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
+       \caption{Frequency of plot picks by log$_10$ $p$-value in individual lineups}
+       \label{fig:P-val_log}
 \end{figure*}

### 2012-06-09 R/paper_jasa.R

@@ -295,6 +295,7 @@ ggsave( file="../images/power_screening_subject.pdf",height=8,width=10 )
 # ======= checking minimum p-value assumption for expected power ======
 
 pval1 <- read.csv("../data/pvalue_turk1.csv")
+#pval2 <- read.csv("../data/pvalue_turk2.csv")
 
 get_response_count <- function(response_no){
   res <- rep(0,20)

### 2012-06-09 R/paper_jasa.R
colours added on p-value-log-count plot based on the observed data plot p-value

@@ -306,15 +306,17 @@ get_response_count <- function(response_no){
 
 response_count <- ddply(dat1, .(pic_name), summarise,
                   counts = get_response_count(response_no)$counts,
-                  variable = paste("X",1:20,sep="") )
+                  variable = paste("X",1:20,sep=""),
+                  obs_pvalue = rep(p_value[1],20))
 p_values <- melt(pval1, id="pic_name")
 
 pr <- merge(p_values,response_count, by=c("pic_name","variable"))
 
 prr <- ddply(pr, .(pic_name),summarise,
              counts=counts[order(value)],
-             pvalue=round(value,5)[order(value)],
-             rank_pval = 1:20)
+             pvalue=round(value,4)[order(value)],
+             rank_pval = 1:20,
+             obs_pvalue = obs_pvalue)
 
 p <- qplot(factor(rank_pval), counts, geom="boxplot", data=prr) +
   xlab("rank of p-value") + ylab("Number of subjects")
@@ -333,10 +335,14 @@ p <- qplot(pvalue, counts, geom="point", data=prr) +
   xlab("rank of p-value") + ylab("Number of subjects") +
   facet_wrap(~pic_name3, ncol=10, scales="free_y") + scale_x_log10()
 prr$lpvalue<-log10(prr$pvalue+0.01)
-p <- qplot(lpvalue, counts, geom="point", data=prr) +
+prr$lobs_pvalue<-log10(prr$obs_pvalue+0.01)
+p <- qplot(lpvalue, counts, geom="point", data=subset(prr,lpvalue != lobs_pvalue)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
+  geom_segment(mapping=aes(xend=lobs_pvalue, yend=0, colour="hotpink"), data= subset(prr,lpvalue==lobs_pvalue))+ 
+  geom_point(mapping=aes(xend=lobs_pvalue, yend=0, colour="hotpink"), data= subset(prr,lpvalue==lobs_pvalue))+ 
   xlab(expression(paste(log[10], " p-value"))) + ylab("Number of subjects") +
-  facet_grid(beta~n+sd+rep, scales="free_y", labeller="label_both") 
+  facet_grid(beta~n+sd+rep, scales="free_y", labeller="label_both") +
+  opts(legend.position="none")
 p
 ggsave(file="../images/p_val_log_counts.pdf", height=10, width=12)
 

### 2012-06-10 paper_jasa.tex
colours added on p-value-log-count plot based on the observed data plot p-value
typo in proof
exclusion criteria

@@
 \[
-E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-E[p_D])^{k} = 1-(m-1)E[p_D] + O(E[p_D^2]).
+E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
 \]
@@
 
-\subsection{Data Cleaning}
+\subsection{Data Screening}
 
@@
 It is also possible that participants were not serious enough to provide feedback.
-For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
+%For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
 This could be a criteria to determine their attentiveness in the study.
@@
 
-{\bf Cleaning data:} There are six suggested approach to clean the data. They are 
+{\bf Screening scenarios:} \hh{Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
+In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.}
+%There are six suggested approach to clean the data. 
+They are 
 \begin{enumerate}
-\item not throw out anything
-\item throw out people where we don't have demographic information
-\item throw out all people with 0 percent correct
-\item only include people with at least 20 percent correct (current suggestion)
-\item include people with 50\% correct on all very easy (could be evaluated correctly if the participant is not ignorant) lineups (as suggested before)
-\item each person saw at least one easy lineup.
-we select one of those easy lineups as the inclusion criteria.
-If that lineup is evaluated correctly we include only the rest of the evaluations of that subject.
-Otherwise, we exclude all the feedbacks of that subject.
+\item {\bf include all participants} and their evaluations
+\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
+\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
+\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
+\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
+\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
+If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
 \end{enumerate}
@@
 \begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets.}
+\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
+\hh{Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information.}}
 \begin{center}
@@
        \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
-       \caption{Frequency of plot picks by log$_10$ $p$-value in individual lineups}
+       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of observed data plot.}
        \label{fig:P-val_log}

### 2012-06-10 paper_jasa.tex
figure 2 caption

@@
    \includegraphics[width=3.8in]{images/powerplot.pdf} 
-   \caption{Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
+   \caption{Probability that the data plot has the smallest probability in a lineup of size $m$.
+With increasing $p$-value the probability drops -- when it reaches $1/m$ a horizontal line is drawn to emphasize insufficient sensitivity of the test due to the lineup size.
+   %Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
 The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true.
+}
 }

### 2012-06-10 paper_jasa.tex
small things

@@
 Each subject saw a substantial number of difficult lineups.
-The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants.
-It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones.
+\hh{While this allows us a better estimation for the power in small slopes, }
+%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
+but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
 \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
@@
        \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{p-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}}
+       \caption{$p$-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}
+       \hh{Mahbub, could you exchange this plot for one with  $p_D$ estimated as $(1 - \#  correct/ \# shown)/(m-1)$ as given in lemma 3.1 now?}}
        \label{fig:pval_plot_signal}

### 2012-06-10 paper_jasa.tex
figure plot-signal-strength modified

@@
 (This is also the UMP test for the continuous covariate, case 2.) To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
-This leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}.
+This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the classical $p$-value as details given in Lemma \ref{lemma}.
 Thus the expected power of the visual test in this scenario is
@@
 \begin{equation}\label{power_exp} 
-   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for}  \quad \beta \ne 0
+   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
 \end{equation}
@@
        \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{$p$-value vs plot signal strength. The solid line represents 45 degree line which represents the equity shown in equation \ref{plot_signal}
-       \hh{Mahbub, could you exchange this plot for one with  $p_D$ estimated as $(1 - \#  correct/ \# shown)/(m-1)$ as given in lemma 3.1 now?}}
+       \caption{Classical test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
+The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.}
+       %\hh{Mahbub, could you exchange this plot for one with  $p_D$ estimated as $(1 - \#  correct/ \# shown)/(m-1)$ as given in lemma 3.1 now?}}
        \label{fig:pval_plot_signal}
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_rank_counts.pdf}}
-       \scalebox{0.6}{\includegraphics{p_val_rank_counts2.pdf}}
-       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 (bottom) with raw data (nothing is dropped).
+       \scalebox{0.6}{\includegraphics{p_val_rank_counts12.pdf}}
+       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
 This indicates that most of the subjects pick the plot with minimum p-value (rank 1).

### 2012-06-10 R/paper_jasa.R

@@ -295,7 +295,7 @@ ggsave( file="../images/power_screening_subject.pdf",height=8,width=10 )
 # ======= checking minimum p-value assumption for expected power ======
 
 pval1 <- read.csv("../data/pvalue_turk1.csv")
-#pval2 <- read.csv("../data/pvalue_turk2.csv")
+pval2 <- read.csv("../data/pvalue_turk2.csv")
 
 get_response_count <- function(response_no){
   res <- rep(0,20)
@@ -304,23 +304,33 @@ get_response_count <- function(response_no){
   return(data.frame(counts=res))
 }
 
-response_count <- ddply(dat1, .(pic_name), summarise,
-                  counts = get_response_count(response_no)$counts,
-                  variable = paste("X",1:20,sep=""),
-                  obs_pvalue = rep(p_value[1],20))
-p_values <- melt(pval1, id="pic_name")
+get_merged_pvalue <- function(dat,pval){
+  response_count <- ddply(dat, .(pic_name), summarise,
+                          counts = get_response_count(response_no)$counts,
+                          variable = paste("X",1:20,sep=""),
+                          obs_pvalue = rep(p_value[1],20))
+  p_values <- melt(pval, id="pic_name")
+  pr <- merge(p_values,response_count, by=c("pic_name","variable"))
+  prr <- ddply(pr, .(pic_name),summarise,
+               counts=counts[order(value)],
+               pvalue=round(value,4)[order(value)],
+               rank_pval = 1:20,
+               obs_pvalue = obs_pvalue)  
+  return(prr)
+}
 
-pr <- merge(p_values,response_count, by=c("pic_name","variable"))
 
-prr <- ddply(pr, .(pic_name),summarise,
-             counts=counts[order(value)],
-             pvalue=round(value,4)[order(value)],
-             rank_pval = 1:20,
-             obs_pvalue = obs_pvalue)
-
+prr <- get_merged_pvalue(dat1,pval1)
 p <- qplot(factor(rank_pval), counts, geom="boxplot", data=prr) +
   xlab("rank of p-value") + ylab("Number of subjects")
 ggsave(file="../images/p_val_rank_counts.pdf", height=4, width=7)
+
+prr12 <- rbind(data.frame(Experiment="Experiment1",get_merged_pvalue(dat1,pval1)),
[…]

### 2012-06-10 paper_jasa.tex
discussion: plot signal strength

@@
 
-Each participant was shown a sequence of 10 lineup plots.
+%Each participant was shown a sequence of 10 lineup plots.
 In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
@@
 This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}
 
@@
 
+Lemma \ref{lemma} helps us estimate the plot signal strength based on which the visual test decision could be made based on multiple evaluations on a lineup.
+From the equation \ref{fig:pval_plot_signal} following from Lemma \ref{lemma} we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+This gives an estimate as  $$\hat p_D= (1- u/K)/(m-1)$$ where $K$ is the total number of independent evaluations on a lineup with $u$ being number of successful evaluations.
+Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with classical test $p$-values.
+Notice that plot signal strength is equivalent to classical test $p$-value when $p$-value is less than or equal to 0.05 and beyond that the plot signal strength levels off on a point bigger than 0.05.
+Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
+
+Figure \ref{fig:pval_pcorrect} shows evidence of direct association of percentage of correct evaluations with classical test $p$-values.
+When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
+We observed this for both experiment 1 and 2.
+For experiment 3, the scenario is different as in this case people might have observed something other than the effect $E$.
+
 \begin{figure*}[hbtp]
@@
        \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 1.5 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in figure \ref{fig:pval_power}}
+After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
+The theoretical justification of this is shown in figure \ref{fig:pval_power}.
+\blue{The scenario is different for experiment 3.
+$p$-value does not seem to have any effect on percent correct.
+This may be due to Type-III error.} }
        \label{fig:pval_pcorrect}

### 2012-06-10 R/paper_jasa.R

@@ -601,9 +601,9 @@ pdat <- NULL
 for (i in 1:3){
    dati <- eval(parse(text=paste("dat",i,sep="")))
    pdati <- ddply(dati, .(p_value, sample_size), summarize,
-	  attempted = sum(response==response),
-	  corrected = sum(response=="TRUE"),
-	  percent_correct = sum(response=="TRUE")*100/sum(response==response)
+	  attempted = length(response),
+	  corrected = sum(response),
+	  percent_correct = mean(response)*100
 	)
    pdati$experiment=paste("experiment",i)
    pdat <- rbind(pdat,pdati)
@@ -614,11 +614,11 @@ for (i in 1:3){
 p <- ggplot() +
      geom_point(aes(p_value,percent_correct),data=pdat,size=2) + 
      facet_grid(.~experiment) +
-     xlab(expression(paste("p-value(",p[B],")"))) +
+     xlab(expression(paste("p-value(",p[D],")"))) +
      ylab("Percentage of correct responses") 
 p 
 
-ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4.25, width = 10)
+ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4, width = 10)
 
 # ----- p-value vs plot signal strength --------------------
 

### 2012-06-10 paper_jasa.tex
summary table for minimum p-value assumption

@@
 
+\begin{table}[hbtp]
+\caption{Number of lineups for which most people pick the plot with minimum $p$-value.
+The data used is from screening criteria 1.
+\blue{Is this table useful?
+I can do the same for all the screening criteria.}}
+\begin{center}
+\begin{tabular}{rrr}
+  \hline
+  Experiment & Total lineups & most  pick minimum $p$-value\\ 
+  \hline
+1 &   60 &  37 \\ 
+  2 &   70 &  60 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{figure*}[hbtp]
@@
        \label{fig:P-val_log}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.5}{\includegraphics{p_val_log_counts2.pdf}}
+       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of observed data plot.}
+       \label{fig:P-val_log2}
 \end{figure*}

### 2012-06-10 R/paper_jasa.R

@@ -356,6 +356,48 @@ p <- qplot(lpvalue, counts, geom="point", data=subset(prr,lpvalue != lobs_pvalue
 p
 ggsave(file="../images/p_val_log_counts.pdf", height=10, width=12)
 
+# now same plot for turk2 data
+prr2 <- get_merged_pvalue(dat2,pval2)
+prr2$pic_name2 <- str_sub(prr2$pic_name, 12, -5L)
+x <- str_split(prr2$pic_name2, "_")
+prr2$n <-  unlist(lapply(x, function(x) x[1]))
+prr2$B <-  unlist(lapply(x, function(x) as.numeric(x[2])/100))
+#prr$beta <-  factor(prr$beta, levels=c("0", "1", "2", "3", "5", "7", "8", "10", "16"))
+prr2$sd <-  unlist(lapply(x, function(x) x[3]))
+prr2$rep <-  unlist(lapply(x, function(x) x[4]))
+prr2$pic_name3 <- factor(paste(prr2$beta, prr2$n, prr2$sd, prr2$rep, sep="."))
+p <- qplot(pvalue, counts, geom="point", data=prr2) +
+  xlab("rank of p-value") + ylab("Number of subjects") +
+  facet_wrap(~pic_name3, ncol=10, scales="free_y") + scale_x_log10()
+prr2$lpvalue<-log10(prr2$pvalue+0.01)
+prr2$lobs_pvalue<-log10(prr2$obs_pvalue+0.01)
+p <- qplot(lpvalue, counts, geom="point", data=subset(prr2,lpvalue != lobs_pvalue)) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
+  geom_segment(mapping=aes(xend=lobs_pvalue, yend=0, colour="hotpink"), data= subset(prr2,lpvalue==lobs_pvalue))+ 
+  geom_point(mapping=aes(xend=lobs_pvalue, yend=0, colour="hotpink"), data= subset(prr2,lpvalue==lobs_pvalue))+ 
+  xlab(expression(paste(log[10], " p-value"))) + ylab("Number of subjects") +
+  facet_grid(B~n+sd+rep, scales="free_y", labeller="label_both") +
+  opts(legend.position="none")
+p
+ggsave(file="../images/p_val_log_counts2.pdf", height=10, width=14)
+
+
+# --- summary of minimum p-value data -----
+
+prr1 <- get_merged_pvalue(dat1,pval1)
+min_pval1 <- ddply(prr1,.(pic_name), summarize,
[…]

### 2012-06-10 R/paper_jasa.R
minimum p-value summary for all screening criteria

@@ -394,7 +394,33 @@ min_pval2 <- ddply(prr2,.(pic_name), summarize,
 pval_sm <- rbind(data.frame(Experiment=1,Total = nrow(min_pval1), Most_pick_minimum = sum(min_pval1$most_pick_minimum_pval)),
                  data.frame(Experiment=2,Total = nrow(min_pval2), Most_pick_minimum = sum(min_pval2$most_pick_minimum_pval))
                  )
-xtable(pval_sm)
+
+# ---- getting minimum p-value summary for all screening criteria
+get_all_pval_sam <- function(dat1,dat2,pval1,pval2){
+  prr1 <- get_merged_pvalue(dat1,pval1)
+  min_pval1 <- ddply(prr1,.(pic_name), summarize,
+                     most_pick_minimum_pval = max(counts)==counts[rank_pval==1])
+  prr2 <- get_merged_pvalue(dat2,pval2)
+  min_pval2 <- ddply(prr2,.(pic_name), summarize,
+                     most_pick_minimum_pval = max(counts)==counts[rank_pval==1])
+  
+  pval_sm <- rbind(data.frame(Experiment="1",Total = nrow(min_pval1), Most_pick_minimum = sum(min_pval1$most_pick_minimum_pval)),
+                   data.frame(Experiment="2",Total = nrow(min_pval2), Most_pick_minimum = sum(min_pval2$most_pick_minimum_pval))
+                   )
+  return(pval_sm)
+}
+
+
+indx1 <- get_sceering_index(dat1)
+indx2 <- get_sceering_index(dat2)
+pval_sm <- NULL
+for (i in 1:6) {
+  d1 <- subset(dat1,indx1[,i])
+  d2 <- subset(dat2,indx2[,i])
+  pval_sm <- rbind(pval_sm,cbind(screening=i,get_all_pval_sam(d1,d2,pval1,pval2)))
+}
+
+print(xtable(pval_sm), include.rownames=FALSE)
 
 
 

### 2012-06-10 R/paper_jasa.R
minimum p-value summary table modified

@@ -404,8 +404,8 @@ get_all_pval_sam <- function(dat1,dat2,pval1,pval2){
   min_pval2 <- ddply(prr2,.(pic_name), summarize,
                      most_pick_minimum_pval = max(counts)==counts[rank_pval==1])
   
-  pval_sm <- rbind(data.frame(Experiment="1",Total = nrow(min_pval1), Most_pick_minimum = sum(min_pval1$most_pick_minimum_pval)),
-                   data.frame(Experiment="2",Total = nrow(min_pval2), Most_pick_minimum = sum(min_pval2$most_pick_minimum_pval))
+  pval_sm <- cbind(data.frame(Experiment1=sum(min_pval1$most_pick_minimum_pval)),
+                   data.frame(Experiment2=sum(min_pval2$most_pick_minimum_pval))
                    )
   return(pval_sm)
 }
@@ -419,7 +419,7 @@ for (i in 1:6) {
   d2 <- subset(dat2,indx2[,i])
   pval_sm <- rbind(pval_sm,cbind(screening=i,get_all_pval_sam(d1,d2,pval1,pval2)))
 }
-
+pval_sm <- rbind(pval_sm,c("Total lineups",60,70))
 print(xtable(pval_sm), include.rownames=FALSE)
 
 

### 2012-06-10 paper_jasa.tex
minimum p-value summary for all screening criteria
minimum p-value summary table modified
lowest p-value discussion

@@
 
+Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2.
+The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
+We see the same pattern in figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
+
+%\begin{table}[hbtp]
+%\caption{Number of lineups for which most people pick the plot with minimum $p$-value. The data used is from screening criteria 1. } 
+%\begin{center}
+%\begin{tabular}{cccc}
+%  \hline
+%screening & Experiment & Total &  minimum $p$-value\\ 
+%criteria & & lineups & picked by the most \\
+%  \hline
+%   & 1 &  60 &  37 \\ [-1ex]
+%    \raisebox{1.5ex}{1} & 2 &  70 &  60 \\ 
+%\hline
+%     & 1 &  60 &  38 \\  [-1ex]
+%    \raisebox{1.5ex}{2} & 2 &  70 &  61 \\ 
+%\hline
+%     & 1 &  60 &  38 \\  [-1ex]
+%   \raisebox{1.5ex} {3} & 2 &  70 &  60 \\ 
+%\hline
+%     & 1 &  60 &  41 \\  [-1ex]
+%    \raisebox{1.5ex}{4} & 2 &  70 &  60 \\ 
+%\hline
+%     & 1 &  60 &  39 \\  [-1ex]
+%    \raisebox{1.5ex}{5} & 2 &  70 &  60 \\ 
+%\hline
+%     & 1 &  60 &  40 \\  [-1ex]
+%    \raisebox{1.5ex}{6} & 2 &  70 &  60 \\ 
+%   \hline
+%\end{tabular}
+%\end{center}
+%\end{table}
+
 \begin{table}[hbtp]
-\caption{Number of lineups for which most people pick the plot with minimum $p$-value.
-The data used is from screening criteria 1.
-\blue{Is this table useful?
-I can do the same for all the screening criteria.}}
+\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
 \begin{center}
-\begin{tabular}{rrr}
-  \hline
-  Experiment & Total lineups & most  pick minimum $p$-value\\ 
-  \hline
-1 &   60 &  37 \\ 
-  2 &   70 &  60 \\ 
+\begin{tabular}{ccc}
+  \hline
+screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
+ \cline{2-3}
+criteria & Experiment 1 & Experiment 2 \\ 
+  \hline
+1 & 37 & 60 \\ 
+  2 & 38 & 61 \\ 
+  3 & 38 & 60 \\ 
+  4 & 41 & 60 \\ 
+  5 & 39 & 60 \\ 
+  6 & 40 & 60 \\ 
+\hline
+  Total lineups & 60 & 70 \\ 
    \hline
@@
 \end{center}
+\label{tbl:pval_assumption} 
 \end{table}

### 2012-06-11 paper_jasa.tex
plots of visual test statistics regenerated

@@
 \hline 
-1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid.
+1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid.
 For lineup plot, we simulate data from fitted null model.
@@
 \\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
 For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid.
+4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
 For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
-5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
+5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
 For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
+6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
 For lineup plot, we simulate data from standard normal.
 \\
-7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
+7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
 For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
 \\
-8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data.
+8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data.
 For lineup plot, we simulate data from fitted model.
 \\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
+9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
 \hline 

### 2012-06-11 R/paper_jasa.R
codes for visual stats plot regeneration

@@ -150,6 +150,8 @@ ggsave(plot=pl$p, file="../images/lineup_contaminated.pdf",height=7,width=7.5)
 
 # ==== plotting visual test statistics --------
 
+set.seed(200)
+
 generate_stat_slope <- function(n,beta,sigma){
 	#n <- 100; a <- 6; b <- 2; sigma <- 12
 	a <- 6
@@ -158,10 +160,45 @@ generate_stat_slope <- function(n,beta,sigma){
 	fit <- lm(y~1)
 	p <- qplot(x, resid(fit), xlab=expression(X[k]),geom="point", ylab="Residual", alpha=I(.3))
 	p <- p + geom_smooth(method="lm", se=F, size=1)
-	return(p)
+	return(list(dat=data.frame(x,y),plot=p))
 }
-p <- generate_stat_slope(n=100,beta=2.5,sigma=5)
-ggsave(plot=p, file="../images/stat_beta_k.pdf",height=2,width=2.05)
+rs <- generate_stat_slope(n=100,beta=2.5,sigma=5)
+ggsave(plot=rs$plot, file="../images/stat_beta_k.pdf",height=2,width=2.05)
+
+p <- qplot(x,y, geom="point", data=rs$dat, alpha=I(0.3)) + 
+     xlab(expression(X[k]))+ ylab("Y") +
+     geom_smooth(method="lm", se=F, size=1)
+ggsave(plot=p, file="../images/stat_intercept.pdf",height=2,width=2.05)
+
+p <- qplot(y, geom="histogram", binwidth=1, data=rs$dat)
+ggsave(plot=p, file="../images/stat_goodness_simple.pdf",height=2,width=2.05)
+
+p <- qplot(0,x, geom="boxplot", data=rs$dat, colour=factor(1),fill=factor(1), alpha=.3) +
+     opts(legend.position="none",axis.text.x = theme_blank()) + 
+     xlim(-1,1) + ylab("Residual")
+ggsave(plot=p, file="../images/stat_sigma_box.pdf",height=2,width=2.05)
+
+generate_stat_interection <- function(n,beta, sigma){  
+  x1 <- rpois(n,lambda=20)
+  x2 <- factor(sample(c("A","B"), size=n, replace=T))
+  cnd <- sample(c("A","B"), size=1)
+  y <- round(5+2*x1+3*(x2==cnd)+beta*x1*(x2==cnd)+rnorm(n=n, mean=0, sd=sigma))
[…]

### 2012-06-14 paper_jasa.tex
tweakign language, working on section 6.5

@@
 
-\section{Effect of Lineup Size and Number of Observers}
+\section{Effect of Lineup Size and Number of Observers}~\label{sec:size}
 
@@
 We need to make two crucial assumptions:
-\begin{enumerate}
+\begin{enumerate} \itemsep 0in
 \item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
@@
 
-This section takes the data and looks at the assumption made for the early expected power.
-
-Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2.
+In Section \ref{sec:size} in order to make comparisons with classical tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
+In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
+Figure \ref{fig:P-val_rank} shows side-by-side boxplots of rank of plot (in its lineup) and how many subjects selected the plot.
+\green{Maybe we don't need this plot.}
+
+Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
+In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizntally and the count of the number of subjects who selected this plot is displayed vertically.
+The dot represents this point, and the lines connect vertically with 0 count.
+Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
+Columns correspond to the other levels of the exeriment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
+(Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value.
+As $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
+
+%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2.
 The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
@@
 \begin{center}
-\begin{tabular}{ccc}
-  \hline
+\begin{tabular}{ccc} \hline
 screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
@@
 There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error.
-Can we?}
+Can we?
+\green{Maybe don't need this plot}}
        \label{fig:P-val_rank}
@@
 
-\green{I think it is very interesting to look at the individual lineups by beta, and other parameters, and we can see the tendency to pick the lowest $p$-value. Figure \ref{fig:P-val_log}.}
-
 \begin{figure*}[hbtp]
@@
        \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
-       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of observed data plot.}
+       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of actual data plot.}
        \label{fig:P-val_log}
@@
        \scalebox{0.5}{\includegraphics{p_val_log_counts2.pdf}}
-       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of observed data plot.}
+       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of actual data plot.}
        \label{fig:P-val_log2}

### 2012-06-14 paper_jasa.tex
made a few place marker comments on which to make changes later

@@
 Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data.
+The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}.
 Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
@@
 
+\green{Need to emphasize that this is important for the third experiment.}
+
 Lemma \ref{lemma} helps us estimate the plot signal strength based on which the visual test decision could be made based on multiple evaluations on a lineup.
@@
 As $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
+
+Experiment 3 does not have sensible $p$-values associated with each plot, so this assumption can only be tested with the first two experiments.
 

### 2012-06-14 paper_jasa.tex
working on picking pvalue assumption section

--- 2012-06-14 paper_jasa.tex
@@
 
-Lemma \ref{lemma} helps us estimate the plot signal strength based on which the visual test decision could be made based on multiple evaluations on a lineup.
-From the equation \ref{fig:pval_plot_signal} following from Lemma \ref{lemma} we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
-This gives an estimate as  $$\hat p_D= (1- u/K)/(m-1)$$ where $K$ is the total number of independent evaluations on a lineup with $u$ being number of successful evaluations.
+\green{third experiment doesn't really have a p-value, does it? Need to explain this. Why are there still so many small p-values in experiment 3?}
+
+In the real setting, where visual inference is to be useful, there will be no classical tests or traditional $p$-values.
+Yet, assessing the strength of the structure is a critical component of inference.
+For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
+For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} displays scatterplots of these two variables for each experiment.
+As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and classical test $p$-values.
+When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
+We observed this for both experiment 1 and 2.
+For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
+
+\green{Should we use proportion of correct, rather than percentage? Wouldn't that work into the equations better?}
+
+Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
+From Equation \ref{fig:pval_plot_signal}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+This gives the signal strength estimate as
+\[
+\hat p_D= \frac{1- \frac{u}{K}}{m-1}
+\]
+where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
 Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with classical test $p$-values.
@@
 Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-Figure \ref{fig:pval_pcorrect} shows evidence of direct association of percentage of correct evaluations with classical test $p$-values.
-When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
-We observed this for both experiment 1 and 2.
-For experiment 3, the scenario is different as in this case people might have observed something other than the effect $E$.
 
@@
 $p$-value does not seem to have any effect on percent correct.
-This may be due to Type-III error.} }
+This may be due to Type-III error.} \green{p-value on log scale?} }
        \label{fig:pval_pcorrect}
@@
        \caption{Classical test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
-The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.}
+The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.
+\green{p-value on log scale?}}
        %\hh{Mahbub, could you exchange this plot for one with  $p_D$ estimated as $(1 - \#  correct/ \# shown)/(m-1)$ as given in lemma 3.1 now?}}
@@
 (Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value.
-As $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
+In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
+This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small.
+The results are clearer for experiment 2, that used a continuous covariate.
+So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
+
+Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
+Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
+Rather it is assumed that the observer has the ability to select the plot with the most extreme structure.
+A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, sd=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
+Some of these anomalous lineups were selcted for follow-up studies using an eyetracker to investigate what people were looking at.
 
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_rank_counts12.pdf}}
+       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
        \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
-       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of actual data plot.}
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
+The colored lines indicate the plot with the lowest $p$-value.
+Rows correspond to $\beta$, smallest to largest values from top to bottom.
+Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
+Empty cells indicate no lineup for that combination of parameters.
+Highest counts tend to be the plot in the lineup having the lowest $p$-value.}
        \label{fig:P-val_log}
@@
    \centering
-       \scalebox{0.5}{\includegraphics{p_val_log_counts2.pdf}}
-       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of actual data plot.}
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
+The colored lines indicate the plot with the lowest $p$-value.
+Rows correspond to $\beta$, smallest to largest values from top to bottom.
+Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
+Empty cells indicate no lineup for that combination of parameters.
+Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.}
        \label{fig:P-val_log2}

### 2012-06-18 paper_jasa.tex
UMP replaced by conventional

@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the uniformly most powerful (UMP) test, out-performs the UMP when data is contaminated, and, there may be some super-individuals who yield better power than the UMP test even in the most difficult trials.
+Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
 \end {abstract}
@@
 Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
-Section \ref{sec:simulation} describes the results of the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent uniformly most powerful (UMP) and an analysis of the resulting data.
+Section \ref{sec:simulation} describes the results of the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and an analysis of the resulting data.
 The last section describes the next steps in this research.
@@
 
-The classical test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
-This is the uniformly most powerful (UMP) test \green{reference needed here} and it forms the benchmark upon which we evaluate the visual test.
-(This is also the UMP test for the continuous covariate, case 2.) To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
+The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
+It forms the benchmark upon which we evaluate the visual test.
+To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
 This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the classical $p$-value as details given in Lemma \ref{lemma}.
@@
 \end{equation}
-Figure \ref{fig:power_expected} shows the power of the UMP test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Notice that the expected power of the visual test is almost as good as the power of UMP test.
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
+Notice that the expected power of the visual test is almost as good as the power of conventional test.
 \green{What do you mean by this?
@@
        \scalebox{.95}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
+       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
        \label{fig:power_expected}
@@
 Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
-The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test.
+The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test.
 Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
@@
 The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
-Suppose $\gamma$ denotes the classical UMP test power for each parameter combination shown in table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of classical UMP test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
+Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in table \ref{tbl:experiment_params}.
+Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
 For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
@@
 
-The first two simulation experiments use data generated with a normal error model, satisfying the conditions for uniformly most powerful (UMP) test procedures.
+The first two simulation experiments use data generated with a normal error model, satisfying the conditions for conventional test procedures.
 In these situations there exists a test, and there would, in general, be no need to use visual inference.
-The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the UMP, will compare favorably with it.
+The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the conventional test, will compare favorably with it.
 This third simulation is closer to the mark for the purpose of visual inference.
-The assumptions for the UMP are violated, by contaminating the data.
+The assumptions for the conventional test are violated, by contaminating the data.
 The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.
@@
 This shows how the screening criteria may affect the results.
-The corresponding power curves of UMP test are also shown.
+The corresponding power curves of conventional test are also shown.
 Notice that no mater what criteria we apply the result does not change much for experiment 2.
-Also, estimated power curve with criteria 5 for experiment 3 is above the UMP test power curve.
+Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
 
@@
        \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown in black solid lines for comparison.}
+The corresponding power curve for conventional test is shown in black solid lines for comparison.}
        \label{fig:power_screening}
@@
        \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-       \caption{Observed power shown by loess smoother and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
        \label{fig:power_effect1}
@@
        \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
 Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
@@
       \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}
+The corresponding power curve for conventional test is shown in pink lines for comparison.}
        \label{fig:power_screening_subject}	
@@
 
-\green{third experiment doesn't really have a p-value, does it? Need to explain this. Why are there still so many small p-values in experiment 3?}
+\green{third experiment doesn't really have a p-value, does it?
+Need to explain this.
+Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.}
 
@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results equivalent to the uniformly most powerful (UMP) test and for some scenarios yields better power than the UMP test.
+Results suggest that the lineup protocol provides results equivalent to the conventional test and for some scenarios yields better power than the conventional test.
 \end {abstract}

### 2012-06-18 paper_jasa.tex
tiny comments

@@
 \caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-\hh{Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information.}}
+Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
+This is due to missing demographic information.
+\hh{Do we need this table or could we make this a visual comparison?
+For me, there's too many numbers with too many digits.}}
 \begin{center}
@@
 \[
-\hat p_D= \frac{1- \frac{u}{K}}{m-1}
+\hat p_D= (1- {u}/{K})/({m-1})
 \]

### 2012-06-18 paper_jasa.tex
notational changes

@@
 \begin{equation} \label{mixed}
-g( p_{\ell i} )= X_{\ell i} B +  Z_{\ell i}  \tau_{\ell i},  
+g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
 \end{equation}
@@
 \begin{tabular}{lp{5.5in}}
-$X$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$.
+$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$.
 Covariates could include  demographic information of individuals, such as age, gender, education level etc.
@@
 \\
-$B$ & is a vector of parameters for fixed effects.\\
+$\delta$ & is a vector of parameters for fixed effects.\\
 $Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\

### 2012-06-18 R/paper_jasa.R
updated screening criteria turk3 applied

@@ -9,9 +9,9 @@ library(lme4)
 library(stringr)
 
 
-dat1 <- read.csv("../data/raw_data_turk1.csv")
-dat2 <- read.csv("../data/raw_data_turk2.csv")
-dat3 <- read.csv("../data/raw_data_turk3.csv")
+raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
+raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
+raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
 
 
 
@@ -201,7 +201,7 @@ p <- qplot(X,resid(fit), ylab="Residual", alpha=I(0.3)) +
 ggsave(plot=p, file="../images/stat_nonlinear.pdf",height=2,width=2.05)
 
 
-# ==== summary of the data applying different screening criteria =====
+# ==== summary of the data applying different screening criteria to the raw data 
 
 get_summary <- function(dat){
   g <- ddply(dat,.(gender),summarise, length(unique(id)))
@@ -220,7 +220,9 @@ get_sceering_index <- function(dat){
   indx3 <- dat$id %in% included_id3
   included_id4 <- d34$id[ d34$percent_correct >= 20 ]
   indx4 <- dat$id %in% included_id4
-  d56 <- ddply(subset(dat, p_value < 0.0002),.(id),summarise,
+  easy_dat <- subset(dat, p_value < 0.0002)
+  if (dat$experiment[1]=='turk3') easy_dat <- subset(dat, difficulty==0)
+  d56 <- ddply(easy_dat,.(id),summarise,
               easy_cnt = length(response),
               percent_correct = mean(response)*100,
               include_criteria_6 = response[1],
@@ -242,9 +244,9 @@ get_screened_summary <- function(dat){
   return(s)
 }
 
-sm <- cbind(get_screened_summary(dat1),
-      get_screened_summary(dat2),
-      get_screened_summary(dat3))
+sm <- cbind(get_screened_summary(raw.dat1),
+      get_screened_summary(raw.dat2),
+      get_screened_summary(raw.dat3))
 #library(xtable)
 xtable(sm)
 
@@ -286,10 +288,10 @@ get_screened_predict <- function(dat, newdat,intercept=F){
   }
   return(res)
 }
-get_screened_predict(dat1, newdat=data.frame(effect))
[…]

### 2012-06-19 paper_jasa.tex
updated screening criteria turk3 applied
comments on section 6.5 and 6.6 (new)

@@
   4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
-  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 &  98 &  64 &  34 & 1095 \\ 
+  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
+  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
    \hline
@@
 Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with classical test $p$-values. 
-Notice that plot signal strength is equivalent to classical test $p$-value when $p$-value is less than or equal to 0.05 and beyond that the plot signal strength levels off on a point bigger than 0.05.
-Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
+\hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$.
+This estimate has to be interpreted carefully, since the  choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense.
+Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
+\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: XXX, XXX, and XXX based on $n_1 = YYY, n_2 = YYY$, and $n_3=YYY$ lineups, respectively.
+Mahbub, could you please compute correlation coefficients for $p$ values below $1/(m-1)$ and fill in the numbers.
+}
+%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
+%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
 
@@
 \green{p-value on log scale?}}
-       %\hh{Mahbub, could you exchange this plot for one with  $p_D$ estimated as $(1 - \#  correct/ \# shown)/(m-1)$ as given in lemma 3.1 now?}}
+       \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
        \label{fig:pval_plot_signal}
@@
 Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
-In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizntally and the count of the number of subjects who selected this plot is displayed vertically.
+In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically.
 The dot represents this point, and the lines connect vertically with 0 count.
 Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
-Columns correspond to the other levels of the exeriment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
+Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
 (Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value.
@@
 A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, sd=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
-Some of these anomalous lineups were selcted for follow-up studies using an eyetracker to investigate what people were looking at.
+Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.
 
@@
 
+\hh{
+\subsection{How much do null plots affect the choice?}
+Another difference between classical tests and visual inference is the potential influence that null plots have on an observer's choice of plot.
+If one or more null plots represent data with -- by chance -- a smaller $p$-value than the data plot, there's a strong possibility that an observer will pick the null plot instead of the data plot, independently of the actual value of the $p$-value.
+
+See figure \ref{still-to-do} of evidence for this phenomenon.
+Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
+Thanks!
+
+This problem is further investigated in the paper \cite{waldo}.
+
+}
+
 \section{Conclusions}

### 2012-06-20 R/paper_jasa.R
power by effet with bootstrap confidence band

@@ -13,6 +13,29 @@ raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
 raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
 
+# cleaning the data based on criteria 6
+clean_data <- function(raw_dat){
+  easy_dat <- subset(raw_dat, p_value < 0.0002)
+  if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
+  d <- ddply(subset(easy_dat),.(id),summarise,
+             easy_cnt = length(response),
+             percent_correct = mean(response)*100,
+             include_id = response[1],
+             excluded_lineup = paste(id[1],"_",pic_id[1],sep=""),
+             excluded_lineup_p_value = round(p_value[1],4)
+             )
+  included_id <- d$id[d$include_id]
+  indx <- raw_dat$id %in% included_id
+  excluded_lineup <- paste(raw_dat$id,"_",raw_dat$pic_id,sep="") %in% d$excluded_lineup
+  indx[excluded_lineup] <- FALSE
+  cleaned_dat <- subset(raw_dat,indx)
+  return(cleaned_dat)
+}
+
+
+dat1 <- clean_data(raw.dat1)
+dat2 <- clean_data(raw.dat2)
+dat3 <- clean_data(raw.dat3)
 
 
 # =================== mathematical test t distribution  ===================
@@ -23,7 +46,7 @@ qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5) + ylab("Densit
 ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
 
 
-# ===== computation of p-values =============
+# ===== computation of visual p-values for diffrent combination of K and x 
 
 get_pval <- function(m,K,x){
 	p <- pbinom(size=K,prob=1/m,q=x)
@@ -65,7 +88,7 @@ qplot(Xk, obs.residuals,colour=Xk,geom="boxplot", ylab="Residual") +
       xlab(expression(X[k])) + labs(colour=expression(X[k]))
 ggsave("../images/stat_category.pdf",height=2,width=2.75)
 
-# plotting lineup for testing b2=0
+# plotting lineup (turk1) for testing b2=0
 loc <- sample(1:20, size=1)  # location of observed plot
[…]

### 2012-06-20 paper_jasa.tex
power by effet with bootstrap confidence band
correlation between visual and conventional p-values

@@
 The protocol has only been informally tested until now.
-In the work described in this paper, the lineup protocol is compared head to head with the equivalent classical test.
+In the work described in this paper, the lineup protocol is compared head to head with the equivalent conventional test.
 Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
@@
 Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
-The results are compared with those of the classical test.
+The results are compared with those of the conventional test.
 
@@
 Both start from the same place, the same null hypothesis.
-The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
+The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
 In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
@@
 
-\hh{\noindent An important difference between classical testing and the lineup scenarios is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
+\hh{\noindent An important difference between conventional testing and the lineup scenarios is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
 we are assuming, again, that $X$ counts the number of observers (out of $K$) who identify the data plot from the lineup.
@@
 \paragraph{Lineup size $m$}
-One of the big differences between visual inference and the classical testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
+One of the big differences between visual inference and the conventional testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
 The choice of $m$ has an obvious impact on the test, and we will discuss this in the following paragraph.
@@
 
-\hh{In order to compare power of classical tests and lineups side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
+\hh{In order to compare power of conventional tests and lineups side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
 }
@@
 \begin{lemma}~\label{lemma}
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the classical test statistic.
-Suppose the associated test statistic in a classical test is observed as  $t_{obs}$ with $p$-value $p_D$.
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic.
+Suppose the associated test statistic in a conventional test is observed as  $t_{obs}$ with $p$-value $p_D$.
 
@@
 
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.70}{\includegraphics{power_loess_effect.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test. }
+       \label{fig:power_screening}
+\end{figure}
+
 %Each participant was shown a sequence of 10 lineup plots.
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for conventional test is shown in pink lines for comparison.}
-       \label{fig:power_screening_subject}	
-\end{figure}
-
 \subsection{Estimating $p$-value in Non-Classical Settings}
@@
 
-In the real setting, where visual inference is to be useful, there will be no classical tests or traditional $p$-values.
+In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values.
 Yet, assessing the strength of the structure is a critical component of inference.
@@
 For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} displays scatterplots of these two variables for each experiment.
-As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and classical test $p$-values.
+As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
 When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
@@
 where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations. 
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with classical test $p$-values. 
+Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values. 
 \hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$.
@@
 Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
-\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: XXX, XXX, and XXX based on $n_1 = YYY, n_2 = YYY$, and $n_3=YYY$ lineups, respectively.
-Mahbub, could you please compute correlation coefficients for $p$ values below $1/(m-1)$ and fill in the numbers.
+\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
+}  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.
+But it is true for experiment 2, right?.
 }
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
-       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
+%       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
 After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
@@
 This may be due to Type-III error.} \green{p-value on log scale?} }
-       \label{fig:pval_pcorrect}
-\end{figure*}
+%       \label{fig:pval_pcorrect}
+%\end{figure*}
 
@@
        \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{Classical test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
+       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
 The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.
-\green{p-value on log scale?}}
-       \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
+p-values are shown on log scale.}
+    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
        \label{fig:pval_plot_signal}
@@
 
-In Section \ref{sec:size} in order to make comparisons with classical tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
+In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
 In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
@@
 \subsection{How much do null plots affect the choice?}
-Another difference between classical tests and visual inference is the potential influence that null plots have on an observer's choice of plot.
+Another difference between conventional tests and visual inference is the potential influence that null plots have on an observer's choice of plot.
 If one or more null plots represent data with -- by chance -- a smaller $p$-value than the data plot, there's a strong possibility that an observer will pick the null plot instead of the data plot, independently of the actual value of the $p$-value.
@@
 \end{table}
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
+      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+The corresponding power curve for conventional test is shown in pink lines for comparison.}
+       \label{fig:power_screening_subject}	
+\end{figure}
 

### 2012-06-20 R/calculate_ump_power.R

@@ -1,0 +1,59 @@
+calculate_ump_power1 <- function(beta, n, sigma) {
+  # n <- 100
+  # sigmasq <- 12^2
+  alpha <- 0.05
+  sigmasq <- sigma^2
+  beta_not <- 0
+  se_beta <- sqrt(sigmasq/(n * (0.5^2)))  # refer to docs derivation of power
+  mu <- beta/se_beta
+  alpha <- alpha/2
+  t_n <- qt(p = 1 - alpha, df = n - 3)
+  res <- pt(q = -t_n, df = n - 3, ncp = mu) - pt(q = t_n, df = n - 3, ncp = mu) + 1
+  return(res)
+}
+
+calculate_ump_power2 <- function (beta, n, sigma){
+  alpha <- 0.05
+  beta_not <- 0
+  df <- n-2 # two parameters
+  x <- subset(read.csv("../data/Xdata.csv"),N==n)[,1]
+  ssx <- sum((x-mean(x))^2)
+  se_beta <- sigma/sqrt(ssx) 
+  mu <- beta/se_beta
+  t_n <- qt(p=1-alpha/2,df=df)
+  res <- pt(q=-t_n, df=df, ncp=mu)-pt(q=t_n, df=df, ncp=mu)+1
+  return(res)
+}
+
+getx <- function(n){
+  x1 <- rnorm(n,0,1)
+  nc <- 15*n/100
+  x2 <- rnorm(n=nc,mean=-1.75, sd=1/3)
+  return(c(x2,x1))
+}
+
+calculate_ump_power3 <- function (beta, n, sigma,x){
+  alpha <- 0.05
+  sigmasq <- sigma^2
+  beta_not <- 0
+  df <- n-1 # one parameter(s)
+  se_beta <- sqrt(sigmasq/sum((x-mean(x))^2))
+  mu <- beta/se_beta
+  alpha <- alpha/2
+  t_n <- qt(p=1-alpha,df=df)
+  res <- pt(q=-t_n, df=df, ncp=mu)-pt(q=t_n, df=df, ncp=mu)+1
+  return(res)
+}
+
+get_ump_power_by_effect <- function(){
+  set.seed(2035)
+  ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0,18, by=.2),
+                pow= calculate_ump_power1(beta=seq(0,18, by=.2)/10, n=100, sigma=1)),
+               data.frame(experiment="Experiment 2",effect=seq(0,6, by=.2),
[…]

### 2012-06-20 R/paper_jasa.R

@@ -8,7 +8,6 @@ library(xtable)
 library(lme4)
 library(stringr)
 
-
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
 raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
@@ -315,17 +314,13 @@ get_screened_predict(raw.dat1, newdat=data.frame(effect))
 pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(raw.dat1, newdat=data.frame(effect=seq(0,18, by=.2)))),
             data.frame(experiment="Experiment 2",get_screened_predict(raw.dat2, newdat=data.frame(effect=seq(0,7, by=.2)))),
             data.frame(experiment="Experiment 3",get_screened_predict(raw.dat3, newdat=data.frame(effect=seq(0,6, by=.2)))))
-set.seed(2035)
-ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0,18, by=.2),
-                        ump= calculate_ump_power1(beta=seq(0,18, by=.2)/10, n=100, sigma=1)),
-             data.frame(experiment="Experiment 2",effect=seq(0,7, by=.2),
-                        ump= calculate_ump_power2(beta=seq(0,7, by=.2)/10, n=100, sigma=1)),
-             data.frame(experiment="Experiment 3",effect=seq(0,6, by=.2),
-                        ump= calculate_ump_power3(beta=seq(0,6, by=.2)/sqrt(115), n=115, sigma=1, x=getx(100))))
+
+source("calculate_ump_power.R")
+ump <- get_ump_power_by_effect()
 
 ggplot()+
   geom_line(aes(effect,pred, colour=factor(screening)), data=pi) +
-  geom_line(aes(effect,ump), data=ump) +
+  geom_line(aes(effect,pow), data=ump) +
   facet_grid(.~experiment, scales="free") +
   ylab("power") + xlab(expression(Effect(E))) +
   scale_colour_discrete(name = "Screening \ncriteria") 
@@ -524,20 +519,7 @@ print(xtable(pval_sm), include.rownames=FALSE)
 
 # ------function to calculate UMP power -----------------------
 
-calculate_ump_power1 <- function(beta, n, sigma) {
[…]

### 2012-06-20 paper_jasa.tex
difference of p-value to min.p-value figure added

@@
    \centering
-       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
-       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for conventional test is shown in black solid lines for comparison.}
-       \label{fig:power_screening}
+       \scalebox{0.70}{\includegraphics{power_loess_effect.pdf}}
+       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test.
+Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
+Size of the dots represents the number of responses for each lineup.}
+       \label{fig:power_loess_effect}
 \end{figure}
@@
    \centering
-       \scalebox{0.70}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test. }
-       \label{fig:power_screening}
+       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
+       \caption{The overall power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
+       \label{fig:power_mixed}
 \end{figure}
@@
 \subsection{Subject-specific variation}
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
+       \caption{Subject specific  power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
+       \label{fig:power_mixed_subject}
+\end{figure}
 
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
-       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
+%\begin{figure*}[hbtp]
+%   \centering
+%       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
+%       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
 This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
@@
 \green{Maybe don't need this plot}}
-       \label{fig:P-val_rank}
-\end{figure*}
+%       \label{fig:P-val_rank}
+%\end{figure*}
 
@@
 
-This problem is further investigated in the paper \cite{waldo}.
+This problem is further investigated in the paper \cite{waldo}. \blue{Here is the plot in figure \ref{fig:pval_difference} as proposed.}
 
 }
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
+       \caption{Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.}
+       \label{fig:pval_difference}
+\end{figure}
 
@@
 \end{table}
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
+       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+The corresponding power curve for conventional test is shown in black solid lines for comparison.}
+       \label{fig:power_screening}
+\end{figure}
 

### 2012-06-20 R/paper_jasa.R

@@ -1,5 +1,4 @@
 # The complete R codes for JASA Paper
-# Last modified by Di on Jun 7, 2012.
 
 library(ggplot2)
 library(plyr)
@@ -8,9 +7,12 @@ library(xtable)
 library(lme4)
 library(stringr)
 
+# getting the data and some common functions
+
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
 raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
+source("calculate_ump_power.R") # functions to compute power
 
 # cleaning the data based on criteria 6
 clean_data <- function(raw_dat){
@@ -30,7 +32,6 @@ clean_data <- function(raw_dat){
   cleaned_dat <- subset(raw_dat,indx)
   return(cleaned_dat)
 }
-
 
 dat1 <- clean_data(raw.dat1)
 dat2 <- clean_data(raw.dat2)
@@ -272,7 +273,7 @@ sm <- cbind(get_screened_summary(raw.dat1),
 #library(xtable)
 xtable(sm)
 
-# --- fit mixed model with each screening criteria and plot power
+# --- fit mixed model with cleaned data
 #library(lme4)
 
 pred.mixed <- function(X, subject=0,fit) {
@@ -301,6 +302,24 @@ effect <- seq(0.01,16, by=.2)
 get_predict_mixed(dat1, newdat=data.frame(effect))
 #get_predict_mixed(dat1, newdat=data.frame(effect), intercept=T)
 
+pi_effect <- rbind(data.frame(experiment="Experiment 1",get_predict_mixed(dat1, newdat=data.frame(effect=seq(0,18, by=.2)))),
+            data.frame(experiment="Experiment 2",get_predict_mixed(dat2, newdat=data.frame(effect=seq(0,6, by=.2)))),
+            data.frame(experiment="Experiment 3",get_predict_mixed(dat3, newdat=data.frame(effect=seq(0,5, by=.2)))))
+
+# source("calculate_ump_power.R")
+ump <- get_ump_power_by_effect()
+
+ggplot()+
+  geom_line(aes(effect,pred, colour="Visual"), data=pi_effect) +
+  geom_line(aes(effect,pow, colour="Conventional"), data=ump) +
+  facet_grid(.~experiment, scales="free") +
+  ylab("power") + xlab(expression(Effect(E))) +
[…]

### 2012-06-20 R/paper_jasa.R
p-value plot updated

@@ -667,92 +667,17 @@ p <- ggplot() +
 p
 ggsave(p, filename = "../images/power_loess_exp1.pdf", height = 5.5,width = 8.5)
 
-# ------------------- Expected power calculations old --------
-n <- 300
-beta <- 3
-sigma <- 12
-ncp <- beta * sqrt(n)/(2 * sigma)  # refer to documentation for derivation of power (xbar=1/2 here)
-#Density of the absolute value of a noncentral
-#t random variable with df degrees of freedom
-#and noncentrality parameter ncp.
-#refer to Dr. nettleton power point slide in gmail
-f.abs.t = function(t, df, ncp = 0) {
-    (t > 0) * dt(t, df = df, ncp = ncp) + dt(-t, df = df, ncp = ncp)
-}
-#Density of the p-value from a t-test
-#when the noncentrality parameter is ncp
-#and the degrees of freedom are df.
-f = function(p, df, ncp = 0) {
-    tt = qt(1 - p/2, df = df)
-    f.abs.t(tt, df = df, ncp = ncp)/f.abs.t(tt, df = df)
-}
-#Example: Density of p-value for t-test with
-#df=8 and ncp=2.
-df <- n - 3  # since 3 parameters in the regression model
-p = seq(0.01, 1, by = 0.01)
-y = f(p, df = df, ncp = ncp)
-plot(p, y, type = "l", ylim = c(0, max(y)))
-library(ggplot2)
-qplot(p, y, geom = "line", ylim = c(0, max(y))) + xlab("p-value") + 
-    ylab("density")
-n <- 100
-sigma <- 12
-expected_power <- function(beta) {
-    #beta <- 3
-    m <- 19
-    ncp <- beta * sqrt(n)/(2 * sigma)
-    df <- n - 3
-    p <- seq(1e-05, 1, by = 1e-04)
-    N <- length(p)
-    y <- f(p, df = df, ncp = ncp)
-    x <- sample(p, prob = y, replace = T)
-    z <- rbeta(n = N, shape1 = 1, shape2 = m - 1)
-    return(sum(x < z)/N)
-}
-expected_power(6)
-n <- 100
-sigma <- 12
-expected_power_thm <- function(beta) {
-    #beta <- 3
-    m <- 20
-    ncp <- beta * sqrt(n)/(2 * sigma)
-    df <- n - 3
-    p <- runif(n = 1e+06)
[…]

### 2012-06-20 paper_jasa.tex
p-value plot updated
started discussion of subject specific variation

@@
 We fit model \ref{mixed} with effect $E=\beta*n/\sqrt(n)$ as the only fixed effect covariate without intercept.
-We also set $offset = log(0.05/95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
+We also set $offset = log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
 The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
-       \label{fig:power_effect1}
-\end{figure*}
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
+%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
+%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
+%       \label{fig:power_effect1}
+%\end{figure*}
 
@@
 Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
+\blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
+Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
        \label{fig:power_loess3}
@@
 \subsection{Subject-specific variation}
+
+Model \ref{mixed} with effect as fixed covariate and subject as random effect is fitted to the data to estimate the subject specific power curves shown in figure \ref{fig:power_mixed_subject}.
+Subject specific power is quite different for three experiments.
+It appears that scatter plot serves better job compared to boxplot as a visual test statistic.
+But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
 

### 2012-06-20 R/paper_jasa.R
plot 7 updated

@@ -820,16 +820,16 @@ loess.limts <- rbind(data.frame(experiment="Experiment 1", get_bootstrap_limit_l
 source("calculate_ump_power.R")
 ump.power <- get_ump_power_by_effect()
 power.dat <- rbind(data.frame(Test="Visual",loess.power),
-                   data.frame(Test="Conventional", ump.power))
-#power.dat$emphasize <- 0.5 +  (power.dat$Test=="Conventional")                             
+                   data.frame(Test="Conventional", ump.power))                          
 
 ggplot()+
   geom_point(aes(effect,as.numeric(response), size=responses), data=effect.dat, alpha=.3) +
-  geom_line(aes(effect,pow,colour=Test), data=power.dat) +
+  geom_line(aes(effect,pow,linetype=Test, colour=Test), data=power.dat, size=1.2) +
   geom_ribbon(aes(x = effect, ymin = limit1, ymax = limit2), 
               data = loess.limts, alpha=.3) +
   facet_grid(.~experiment, scales="free") + 
-  ylab("Power")
+  scale_colour_manual(values=c("Blue","Black")) +
+  ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses")
 ggsave(filename = "../images/power_loess_effect.pdf", height = 4,width = 10)
 
 

### 2012-06-20 paper_jasa.tex
plot 7 updated
plot 3 updated

@@
    \centering
-       \scalebox{.95}{\includegraphics{power_expected.pdf}}
+       \scalebox{.6}{\includegraphics{power_expected.pdf}}
        \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
@@
 
-Model \ref{mixed} with effect as fixed covariate and subject as random effect is fitted to the data to estimate the subject specific power curves shown in figure \ref{fig:power_mixed_subject}.
+In all the three experiments, each subject saw a lineup only once.
+Assuming that there is no random effect for lineup, we set  $\tau_{\ell  i}=\tau_i $ in model \ref{mixed} and $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
+This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}.
+Parameter estimates of the model is shown in table \ref{tbl:model_par}.
+The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}.
+The offset is set to $log(0.05/0.95)$ so that Type-I error is fixed to be 0.05 when effect = 0.
 Subject specific power is quite different for three experiments.
@@
 But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
+
+\begin{table}[hbtp]
+\caption{Parameter estimates of model \ref{mixed}.}
+\begin{center}
+\begin{tabular}{ccc}
+  \hline
+ &  \multicolumn{2}{c} {Estimates}  \\
+ \cline{2-3}
+ Experiment & Effect & inter subject correlation\\
+  \hline
+  1 &  &  \\ 
+  2 &  &  \\ 
+  3 &  & \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:model_par}
+\end{table}
 

### 2012-06-20 R/generate_expected_visual_power_simulation-hh.R

@@ -45,25 +45,25 @@ calculate_theoretical_power <- function (beta, n, sigma, alpha=0.05){
   return(res)
 }
 
-powerdf <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=seq(10, 30, by=5), test="visual"))
+powerdf <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=seq(10, 30, by=5), test="Visual"))
 powerdf$power <- ldply(1:nrow(powerdf), function(x) {
   mean(replicate(10, with(powerdf, generate_visual_power(n[x],
     beta=beta[x], sigma=sigma[x], m=m[x]))))
 })$V1
 
-power2 <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=NA, test="theoretical"))
+power2 <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=NA, test="Conventional"))
 power2$power <- ldply(1:nrow(power2), function(x) {
   with(powerdf, calculate_theoretical_power(n[x], beta=beta[x], sigma=sigma[x]))
 })$V1
 
 powerdf <- rbind(powerdf, power2)
 levels(powerdf$test)
-powerdf$emphasize <- 0.5 + ((powerdf$m==20) | (powerdf$test=="theoretical"))
+powerdf$emphasize <- 0.5 + ((powerdf$m==20) | (powerdf$test=="Conventional"))
 
 ggplot(aes(beta, power, group=m, linetype=test, size=emphasize), data=powerdf) + 
   geom_smooth(aes(colour=m), alpha=0.5, fill=NA, method="loess", span=0.1) + 
   scale_colour_gradient("Lineup size m", breaks=c(10,20,30)) + 
-  scale_size_identity() + 
+  scale_size_identity() + ylab("Power") +
   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
 
-ggsave("images/power_expected.pdf", width=5, height=3, units="in")
+ggsave("../images/power_expected.pdf", width=7, height=5, units="in")

### 2012-06-20 R/paper_jasa.R

@@ -310,11 +310,12 @@ pi_effect <- rbind(data.frame(experiment="Experiment 1",get_predict_mixed(dat1, 
 ump <- get_ump_power_by_effect()
 
 ggplot()+
-  geom_line(aes(effect,pred, colour="Visual"), data=pi_effect) +
-  geom_line(aes(effect,pow, colour="Conventional"), data=ump) +
+  geom_line(aes(effect,pred, colour="Visual"), data=pi_effect, size=1.2) +
+  geom_line(aes(effect,pow,colour="Conventional"), data=ump, size=1.2) +
   facet_grid(.~experiment, scales="free") +
-  ylab("power") + xlab(expression(Effect(E))) +
-  scale_colour_discrete(name = "Test") 
+  ylab("Power") + xlab(expression(Effect(E))) +
+  scale_colour_discrete(name = "Test") +
+  scale_colour_manual( values=c("Black","Blue"))
 
 ggsave( file="../images/power_mixed.pdf",height=4,width=10)
 
@@ -341,7 +342,7 @@ ggplot()+
   geom_line(aes(effect,pred, colour=factor(screening)), data=pi) +
   geom_line(aes(effect,pow), data=ump) +
   facet_grid(.~experiment, scales="free") +
-  ylab("power") + xlab(expression(Effect(E))) +
+  ylab("Power") + xlab(expression(Effect(E))) +
   scale_colour_discrete(name = "Screening \ncriteria") 
 
 ggsave( file="../images/power_screening.pdf",height=4,width=10)
@@ -356,7 +357,7 @@ ggplot()+
   geom_line(aes(effect,pred, group=subject), data=pi_subject, alpha=.1) +
   geom_line(aes(effect,pow), data=ump, colour="hotpink", size=1) +
   facet_grid(.~experiment, scales="free") +
-  ylab("power") + xlab(expression(Effect(E)))
+  ylab("Power") + xlab(expression(Effect(E)))
 
 ggsave( file="../images/power_mixed_subject.pdf",height=4,width=10)
 
@@ -371,7 +372,7 @@ ggplot()+
   geom_line(aes(effect,pred, group=subject), data=pi, alpha=.1) +
   geom_line(aes(effect,pow), data=ump, colour="hotpink", size=1) +
   facet_grid(screening~experiment, scales="free") +
-  ylab("power") + xlab(expression(Effect(E)))
+  ylab("Power") + xlab(expression(Effect(E)))
 ggsave( file="../images/power_screening_subject.pdf",height=8,width=10 )  
 
[…]

### 2012-06-21 R/paper_jasa.R
difference between data plot p-value and min(null plot p-values)

@@ -171,6 +171,21 @@ pl <- generate_turk3_lineup(n=100,sigma=3.5,beta=0.4)
 pl$result
 ggsave(plot=pl$p, file="../images/lineup_contaminated.pdf",height=7,width=7.5)
 
+# == plotting contaminated data example
+set.seed(509)
+n <- 100 ; nc <- 15; beta <- 3; sigma <- 5; gamma <- 15
+x1 <- rnorm(n,0,1)
+y1 <- 5 + beta*x1 + rnorm(n,0,sigma)
+x2 <- rnorm(n=nc,mean=-1.75, sd=1/3)
+y2 <- rnorm(nc,gamma,sigma/3)
+X <- c(x2,x1)
+Y <- c(y2,y1)
+summary(lm(Y~X)) # p-value for slope is large
+qplot(X,Y)
+
+ggsave("../images/contaminated_data.pdf", width=5,height=5)
+
+
 # ==== plotting visual test statistics --------
 
 set.seed(200)
@@ -549,31 +564,31 @@ pval_sm <- rbind(pval_sm,c("Total lineups",60,70))
 print(xtable(pval_sm), include.rownames=FALSE)
 
 
-# Difference between data plot p-value and minimum p-value 
+# Difference between data plot p-value and minimum null plot p-value 
 
-min.pval1 <- ddply(melt(pval1, id="pic_name"),.(pic_name), summarize, 
-                   min.pval = round(min(value),4))
-min.pval2 <- ddply(melt(pval2, id="pic_name"),.(pic_name), summarize, 
-                   min.pval = round(min(value)))
+get_pval_success <- function (dat){
+  res <- ddply(dat, .(pic_name), summarize,      
+        prop_correct=mean(response),
+        p_value=p_value[1],
+        actual_plot = plot_location[1])
+  res$experiment=paste("Experiment", substr(dat$experiment,5,5)[1])
+  return(res)
+}
+pval.success <- rbind(get_pval_success(dat1),get_pval_success(dat2))
+pval.dat <- merge(pval.success,rbind(pval1,pval2), by="pic_name")
+pval.dat.m <- melt(pval.dat, id=c("pic_name","prop_correct",
+                                  "p_value","actual_plot","experiment"))
+pval.diff <- ddply(pval.dat.m, .(pic_name), summarise,
+                   experiment=experiment[1],
+                   prop_correct=prop_correct[1],
[…]

### 2012-06-21 paper_jasa.tex
difference between data plot p-value and min(null plot p-values)
comments on figure 7

@@
    \centering
-       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
+       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
        \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
+The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope.
 }
@@
 Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-Size of the dots represents the number of responses for each lineup.}
+Size of the dots represents the number of responses for each lineup.
+\blue{Response to the intuition of Di: We are overestimating the effect for experiment 3 here because to compute effect I used the beta values of the linear part of the contaminated data.
+True effect should be much smaller since overall contaminated data should have a smaller slope.
+Thus the effect of 5 that the plot is showing now would actually be effect 2 (may be).
+That means the  power that we see in the plot for effect 5 is actually the power for effect 2.
+How can we get true effect for this data?
+}}
        \label{fig:power_loess_effect}

### 2012-06-21 R/paper_jasa.R

@@ -172,7 +172,7 @@ pl$result
 ggsave(plot=pl$p, file="../images/lineup_contaminated.pdf",height=7,width=7.5)
 
 # == plotting contaminated data example
-set.seed(509)
+set.seed(59)
 n <- 100 ; nc <- 15; beta <- 3; sigma <- 5; gamma <- 15
 x1 <- rnorm(n,0,1)
 y1 <- 5 + beta*x1 + rnorm(n,0,sigma)
@@ -853,6 +853,17 @@ ggplot()+
   ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses") + 
   opts(asp.ratio=1)
 ggsave(filename = "../images/power_loess_effect.pdf", height = 4,width = 10)
+
+
+# examining dat3
+dat3$effect= with(dat3, sqrt(sample_size)*abs(beta)/sigma)
+d3 <- ddply(dat3, .(pic_name), summarize,
+            prop_correct=mean(response),
+            effect=effect[1],
+            num_responses = length(effect))
+qplot(effect, prop_correct, data=d3, size=num_responses) + geom_smooth(method="loess")
+
+qplot(effect,as.numeric(response), data=dat3)+ geom_smooth(method="loess")
 
 
 

### 2012-06-21 paper_jasa.tex
model only plot and experiment3 plot removed

@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
-       \caption{The overall power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
-       \label{fig:power_mixed}
-\end{figure}
+%
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
+%       \caption{The overall power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
+%       \label{fig:power_mixed}
+%\end{figure}
 
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
+%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
 Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
@@
 Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
-       \label{fig:power_loess3}
-\end{figure*}
+%       \label{fig:power_loess3}
+%\end{figure*}
 

### 2012-06-21 paper_jasa.tex
figure related references updated

@@
 
-\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size.
+\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
 This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
@@
 For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
-For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} displays scatterplots of these two variables for each experiment.
+For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
 As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
@@
 In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
-Figure \ref{fig:P-val_rank} shows side-by-side boxplots of rank of plot (in its lineup) and how many subjects selected the plot.
-\green{Maybe we don't need this plot.}
 
@@
 
-See figure \ref{still-to-do} of evidence for this phenomenon.
-Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
+See figure \ref{fig:pval_difference} of evidence for this phenomenon.
+%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
 Thanks!
 
-This problem is further investigated in the paper \cite{waldo}. \blue{Here is the plot in figure \ref{fig:pval_difference} as proposed.}
+This problem is further investigated in the paper \cite{waldo}.
 

### 2012-06-21 paper_jasa.tex
experiment 3 choice reasons

@@
 %\end{figure}
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
+       \caption{Reasons for choosing the plot in a lineup for experiment 3.
+Reasons 1 = Most different plot,  2 = Visible trend,  3 = Clustering visible,  4= Other.
+Reason 12 means both reasons 1 and 2 are chosen.
+Notice that 345 people choose only reason 3 out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
+Those who chose reason 123 got the actual plot about 100\% times.}
+       \label{fig:choice_reason}
+\end{figure}
 

### 2012-06-21 R/paper_jasa.R

@@ -698,6 +698,24 @@ calculate_ump_power2(beta=.1,n=100,sigma=5)
 # source("calculate_ump_power.R")
 calculate_ump_power3(beta=.1,n=100,sigma=5,x=getx(100))
 
+# experiment 3 reasoning (reasons for choice)
+
+# 1. Most different plot
+# 2. Visible trend
+# 3. Clustering visible
+# 4. Other
+
+reason.exp3 <- c("Most different plot", "Visible trend","Clustering visible","Other")
+
+reason.dat3 <- ddply(dat3,.(choice_reason), summarise,
+      prop_correct = mean(response),
+      counts = length(response))
+reason.dat3$choice_reason <- with(reason.dat3, factor(choice_reason, 
+                             levels=choice_reason[order(counts, decreasing=T)]))
+qplot(choice_reason, counts, geom="bar", data=reason.dat3, fill=prop_correct) +
+  xlab("Reasons for choice") + ylab("Number of Subjects") +
+  scale_fill_continuous(name="Proportion \ncorrect")
+ggsave("../images/choice_reason.pdf", height = 4, width = 8)
 
 
 # ======================= p_value vs %correct =====================
@@ -864,6 +882,3 @@ d3 <- ddply(dat3, .(pic_name), summarize,
 qplot(effect, prop_correct, data=d3, size=num_responses) + geom_smooth(method="loess")
 
 qplot(effect,as.numeric(response), data=dat3)+ geom_smooth(method="loess")
-
-
-

### 2012-06-21 paper_jasa.tex
discussion: model fitting

@@
 
-We fit model \ref{mixed} with effect $E=\beta*n/\sqrt(n)$ as the only fixed effect covariate without intercept.
-We also set $offset = log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
+For each parameter combination, the effect of the lineup is computed as $$E=\beta*\sqrt {n}/\sigma.$$ The experimental data is displayed in figure \ref{fig:power_loess_effect} for different effect sizes.
+The loess curves fitted to the response data provide the estimated observed power for each test.
+The corresponding power curves of conventional tests are shown for comparison.
+The observed power for experiment 2 (scatter plot) is better than what we observed for experiment 1 (box plot).
+For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
+
+Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2.
+For experiment 3 the scenario is different since visually we can easily notice this type of contamination in the data.
+Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
+While fitting model \ref{mixed} we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
+Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
+For experiment 3 we fit the intercept two allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.
+
 The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
@@
    \centering
-       \scalebox{0.70}{\includegraphics{power_loess_effect.pdf}}
+       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
        \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test.
@@
 
+\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
+This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.} \blue{We also see this feature in figure \ref{fig:choice_reason}}.
+
 \begin{figure}[hbtp]
@@
 
-\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
-This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}
-
 \subsection{Subject-specific variation}
@@
 In all the three experiments, each subject saw a lineup only once.
-Assuming that there is no random effect for lineup, we set  $\tau_{\ell  i}=\tau_i $ in model \ref{mixed} and $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
+This allows no random effect for lineup and this simplifies model \ref{mixed} as $\tau_{\ell  i}=\tau_i$.
 This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}.
@@
 The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}.
-The offset is set to $log(0.05/0.95)$ so that Type-I error is fixed to be 0.05 when effect = 0.
 Subject specific power is quite different for three experiments.
-It appears that scatter plot serves better job compared to boxplot as a visual test statistic.
+It appears that scatter plot serves better job compared to box plot as a visual test statistic.
 But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject specific  power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
+       \caption{Subject specific  power estimated from model \ref{mixed} are shown.
+The corresponding power curve for conventional test (dashed line) is shown for comparison.
+The overall average power curve estimated from the model is shown (light blue).}
        \label{fig:power_mixed_subject}

### 2012-06-21 R/paper_jasa.R

@@ -369,10 +369,11 @@ pi_subject <- rbind(data.frame(experiment="Experiment 1",get_predict_mixed(dat1,
                    data.frame(experiment="Experiment 3",get_predict_mixed(dat3, newdat=data.frame(effect=seq(0,5, by=.2)), intercept=T)))
 
 ggplot()+
-  geom_line(aes(effect,pred, group=subject), data=pi_subject, alpha=.1) +
-  geom_line(aes(effect,pow), data=ump, colour="hotpink", size=1) +
+  geom_line(aes(effect,pred, group=subject), data=pi_subject, colour="blue",alpha=.1) +
+  geom_line(aes(effect,pow), data=ump, size=1, linetype=2) +
+  geom_line(aes(effect,pred), data=pi_effect, colour="skyblue", size=1) +
   facet_grid(.~experiment, scales="free") +
-  ylab("Power") + xlab(expression(Effect(E)))
+  ylab("Power") + xlab("Effect")
 
 ggsave( file="../images/power_mixed_subject.pdf",height=4,width=10)
 
@@ -870,7 +871,7 @@ ggplot()+
   scale_colour_manual(values=c("Blue","Black")) +
   ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses") + 
   opts(asp.ratio=1)
-ggsave(filename = "../images/power_loess_effect.pdf", height = 4,width = 10)
+ggsave(filename = "../images/power_loess_effect.pdf", height = 4.5,width = 12)
 
 
 # examining dat3

### 2012-06-21 R/paper_jasa.R
p-value plot changed to square root scale

@@ -761,10 +761,10 @@ p
 p <- ggplot(pdat) +
   geom_point(aes(p_value,strength),size=2) + 
   facet_grid(.~experiment) +
-  xlab(expression(paste("Conventional test p-value (",p[D],") on ",log[10]," scale"))) +
-  ylab(expression(paste("Estimate of visual p-value (", hat(p)[D],") on ",log[10]," scale"))) + 
+  xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
+  ylab(expression(paste("Visual p-value (", hat(p)[D],") on square root scale"))) + 
   geom_abline(aes(intercept=0,slope=1)) +
-  scale_x_log10() +scale_y_log10()
+  scale_x_sqrt() +scale_y_sqrt()
 p
 
 ggsave(p,file="../images/p_val_plot_signal.pdf", height = 4, width = 10)

### 2012-06-21 paper_jasa.tex
p-value plot changed to square root scale
New section added

@@
 
-\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
-This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.} \blue{We also see this feature in figure \ref{fig:choice_reason}}.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
-       \caption{Reasons for choosing the plot in a lineup for experiment 3.
-Reasons 1 = Most different plot,  2 = Visible trend,  3 = Clustering visible,  4= Other.
-Reason 12 means both reasons 1 and 2 are chosen.
-Notice that 345 people choose only reason 3 out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
-Those who chose reason 123 got the actual plot about 100\% times.}
-       \label{fig:choice_reason}
-\end{figure}
-
 \subsection{Subject-specific variation}
@@
 The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.
-p-values are shown on log scale.}
+p-values are shown on square root scale scale.
+\blue{Is it better now?}}
     %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
@@
 
+\subsection{Alternative - Type III error}
+
+\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
+This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}
+
+%\green{Not sure that this goes here, but something needs to be said about Type III.}
+
+\blue{We also see this feature in figure \ref{fig:choice_reason}}.
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
+       \caption{Reasons for choosing the plot in a lineup for experiment 3.
+Reasons 1 = Most different plot,  2 = Visible trend,  3 = Clustering visible,  4= Other.
+Reason 12 means both reasons 1 and 2 are chosen.
+Notice that 345 people choose only reason 3 out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
+Those who chose reason 123 got the actual plot about 100\% times.}
+       \label{fig:choice_reason}
+\end{figure}
+
 \section{Conclusions}

### 2012-06-22 R/calculate_ump_power.R
loess picture updated

@@ -47,12 +47,12 @@ calculate_ump_power3 <- function (beta, n, sigma,x){
 
 get_ump_power_by_effect <- function(){
   set.seed(2035)
-  ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0,18, by=.2),
-                pow= calculate_ump_power1(beta=seq(0,18, by=.2)/10, n=100, sigma=1)),
-               data.frame(experiment="Experiment 2",effect=seq(0,6, by=.2),
-                pow= calculate_ump_power2(beta=seq(0,6, by=.2)/10, n=100, sigma=1)),
-               data.frame(experiment="Experiment 3",effect=seq(0,5, by=.2),
-                pow= calculate_ump_power3(beta=seq(0,5, by=.2)/sqrt(115), n=115, sigma=1, x=getx(100))))
+  ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0,16, by=.2),
+                pow= calculate_ump_power1(beta=seq(0,16, by=.2)/10, n=100, sigma=1)),
+               data.frame(experiment="Experiment 2",effect=seq(0,5.5, by=.2),
+                pow= calculate_ump_power2(beta=seq(0,5.5, by=.2)/10, n=100, sigma=1)),
+               data.frame(experiment="Experiment 3",effect=seq(0,4.5, by=.2),
+                pow= calculate_ump_power3(beta=seq(0,4.5, by=.2)/sqrt(115), n=115, sigma=1, x=getx(100))))
   return(ump)
 }
 

### 2012-06-22 R/paper_jasa.R

@@ -12,6 +12,7 @@ library(stringr)
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
 raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
+
 source("calculate_ump_power.R") # functions to compute power
 
 # cleaning the data based on criteria 6
@@ -848,8 +849,8 @@ loess.power <- rbind(data.frame(experiment="Experiment 1", get_power_loess(dat1)
                     data.frame(experiment="Experiment 2", get_power_loess(dat2)),
                     data.frame(experiment="Experiment 3", get_power_loess(dat3)))
 
-# the following codes generates bootstrap limits and takes 10 minutes to run
-# The limits are saved so without running this we can get the saved limit data
+#the following codes generates bootstrap limits and takes 10 minutes to run
+#The limits are saved so without running this we can get the saved limit data
 # loess.limts <- rbind(data.frame(experiment="Experiment 1", get_bootstrap_limit_loess(dat1)),
 #                      data.frame(experiment="Experiment 2", get_bootstrap_limit_loess(dat2)),
 #                      data.frame(experiment="Experiment 3", get_bootstrap_limit_loess(dat3)))
@@ -861,18 +862,19 @@ source("calculate_ump_power.R")
 ump.power <- get_ump_power_by_effect()
 power.dat <- rbind(data.frame(Test="Visual",loess.power),
                    data.frame(Test="Conventional", ump.power))                          
-
+power.dat$m <- NA
+power.dat$m[power.dat$Test=="Visual"] <- 20
 ggplot()+
-  geom_point(aes(effect,as.numeric(response), size=responses), data=effect.dat, alpha=.3) +
-  geom_line(aes(effect,pow,linetype=Test, colour=Test), data=power.dat, size=1.2) +
   geom_ribbon(aes(x = effect, ymin = limit1, ymax = limit2), 
               data = loess.limts, alpha=.3) +
+  geom_point(aes(effect,as.numeric(response), size=responses), data=effect.dat, alpha=.3) +
+  geom_line(aes(effect,pow,linetype=Test, colour=m), data=power.dat, size=1.2) +
   facet_grid(.~experiment, scales="free") + 
-  scale_colour_manual(values=c("Blue","Black")) +
[…]

### 2012-06-25 R/paper_jasa.R
figure for choice reason updated

@@ -713,11 +713,13 @@ reason.dat3 <- ddply(dat3,.(choice_reason), summarise,
       prop_correct = mean(response),
       counts = length(response))
 reason.dat3$choice_reason <- with(reason.dat3, factor(choice_reason, 
-                             levels=choice_reason[order(counts, decreasing=T)]))
-qplot(choice_reason, counts, geom="bar", data=reason.dat3, fill=prop_correct) +
-  xlab("Reasons for choice") + ylab("Number of Subjects") +
-  scale_fill_continuous(name="Proportion \ncorrect")
-ggsave("../images/choice_reason.pdf", height = 4, width = 8)
+                             levels=choice_reason[order(prop_correct, decreasing=T)]))
+reason.dat3$n <- cut(reason.dat3$counts, breaks=c(0,50, 100, 150, 300, 500))
+qplot(choice_reason, prop_correct, geom="point", data=reason.dat3) +
+  xlab("Reasons for choice") + ylab("Proportion correct") + facet_wrap(~n) +
+ # opts(legend.position="none")
+ # scale_size_continuous(name="Number of \nsubjects")
+ggsave("../images/choice_reason.pdf", height = 4, width = 6)
 
 
 # ======================= p_value vs %correct =====================

### 2012-06-26 paper_jasa.tex
loess picture updated
figure for choice reason updated
experiment 3 reasoning plot updated

@@
 This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}.
+Different subject performs differently depending on effect as well.
+So, we include a random interaction term with effect.
 Parameter estimates of the model is shown in table \ref{tbl:model_par}.
@@
        \caption{Reasons for choosing the plot in a lineup for experiment 3.
-Reasons 1 = Most different plot,  2 = Visible trend,  3 = Clustering visible,  4= Other.
 Reason 12 means both reasons 1 and 2 are chosen.
-Notice that 345 people choose only reason 3 out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
-Those who chose reason 123 got the actual plot about 100\% times.}
+Number of subjects choosing that reason are shown on the right hand side.
+Notice that 345 people choose only reason (3) "Clustering visible", out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
+Those who chose reason 123 got the actual plot about 97\% times.}
        \label{fig:choice_reason}

### 2012-06-26 R/paper_jasa.R

@@ -715,8 +715,9 @@ reason.dat3 <- ddply(dat3,.(choice_reason), summarise,
 reason.dat3$choice_reason <- with(reason.dat3, factor(choice_reason, 
                              levels=choice_reason[order(prop_correct, decreasing=T)]))
 reason.dat3$n <- cut(reason.dat3$counts, breaks=c(0,50, 100, 150, 300, 500))
-qplot(choice_reason, prop_correct, geom="point", data=reason.dat3) +
-  xlab("Reasons for choice") + ylab("Proportion correct") + facet_wrap(~n) +
+qplot(prop_correct,choice_reason,  geom="point", data=reason.dat3) +
+  ylab("Reasons for choice") + xlab("Proportion correct") +
+  facet_grid(.~n, scales="free", space="free_y")  + opts(strip.text.y = theme_text())
  # opts(legend.position="none")
  # scale_size_continuous(name="Number of \nsubjects")
 ggsave("../images/choice_reason.pdf", height = 4, width = 6)

### 2012-06-26 R/paper_jasa.R
p-value from definition

@@ -708,18 +708,21 @@ calculate_ump_power3(beta=.1,n=100,sigma=5,x=getx(100))
 # 4. Other
 
 reason.exp3 <- c("Most different plot", "Visible trend","Clustering visible","Other")
-
 reason.dat3 <- ddply(dat3,.(choice_reason), summarise,
-      prop_correct = mean(response),
-      counts = length(response))
-reason.dat3$choice_reason <- with(reason.dat3, factor(choice_reason, 
-                             levels=choice_reason[order(prop_correct, decreasing=T)]))
+                     prop_correct = mean(response),
+                     counts = length(response))
+reason.dat3$choice_reason <- with(reason.dat3, factor(choice_reason,
+                                  levels=choice_reason[order(prop_correct,decreasing=T)]))
 reason.dat3$n <- cut(reason.dat3$counts, breaks=c(0,50, 100, 150, 300, 500))
-qplot(prop_correct,choice_reason,  geom="point", data=reason.dat3) +
-  ylab("Reasons for choice") + xlab("Proportion correct") +
-  facet_grid(.~n, scales="free", space="free_y")  + opts(strip.text.y = theme_text())
- # opts(legend.position="none")
- # scale_size_continuous(name="Number of \nsubjects")
+reason.dat3$n <- reorder(reason.dat3$n, reason.dat3$count, function(x) -max(x))
+reason.dat3$choice_reason <- factor(reason.dat3$choice_reason, levels=rev(levels(reason.dat3$choice_reason)))
+levels(reason.dat3$choice_reason)[c(7,3,5,1)] <- paste("(",1:4,") ",reason.exp3, sep="")
+levels(reason.dat3$choice_reason)[3] <- "Visible Gap"
+
+qplot(prop_correct,choice_reason, geom="point", data=reason.dat3, size=I(3)) +
+  ylab("Reasons for choice") + xlab("Proportion correct") + xlim(c(0,1)) +
+  facet_grid(n~., scales="free", space="free_y") + opts(strip.text.y = theme_text())
+
 ggsave("../images/choice_reason.pdf", height = 4, width = 6)
 
 
@@ -767,12 +770,31 @@ p <- ggplot(pdat) +
   geom_point(aes(p_value,strength),size=2) + 
   facet_grid(.~experiment) +
   xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
[…]

### 2012-06-26 paper_jasa.tex
p-value from definition
subject specific power corve updated

@@
 \red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data.
-\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-value.
+\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
 \\ Should I change everything accordingly?
@@
 
-{\bf Screening scenarios:} \hh{Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
-In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.}
+{\bf Screening scenarios:} Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
+In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.
 %There are six suggested approach to clean the data. 
@@
 
+Table \ref{tbl:summary} shows  the summary results after applying different screening on the data.
+For our analysis we picked criteria six (6) and present results based on this in the following sections.
+
 \subsection{Power comparison}
@@
 Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2.
-For experiment 3 the scenario is different since visually we can easily notice this type of contamination in the data.
+For experiment 3 the scenario is different since contamination in the data is visible and this sets the lower limit of proportion correct at a higher value.
 Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
@@
 Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
-For experiment 3 we fit the intercept two allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.
+For experiment 3 we fit the intercept to allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.
 
@@
 Need to explain this.
-Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.}
+Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.
+So many small p-values are coming from the fact that contamination makes the slope negatively significant.
+For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant.
+It is like the effect of influential outlier.}
 
@@
 Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
-From Equation \ref{fig:pval_plot_signal}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
 This gives the signal strength estimate as
@@
 %\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
+       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . p-values are shown on square root scale scale. }
+    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
+       \label{fig:pval_definition}
+\end{figure*}
 

### 2012-06-26 R/paper_jasa.R

@@ -369,11 +369,14 @@ pi_subject <- rbind(data.frame(experiment="Experiment 1",get_predict_mixed(dat1,
                    data.frame(experiment="Experiment 2",get_predict_mixed(dat2, newdat=data.frame(effect=seq(0,6, by=.2)), intercept=T)),
                    data.frame(experiment="Experiment 3",get_predict_mixed(dat3, newdat=data.frame(effect=seq(0,5, by=.2)), intercept=T)))
 
+pi_effect$col <- 30
+pi_effect$col[1] <- NA
 ggplot()+
-  geom_line(aes(effect,pred, group=subject), data=pi_subject, colour="blue",alpha=.1) +
+  geom_line(aes(effect,pred, group=subject), data=pi_subject,alpha=.08) +
   geom_line(aes(effect,pow), data=ump, size=1, linetype=2) +
-  geom_line(aes(effect,pred), data=pi_effect, colour="skyblue", size=1) +
+  geom_line(aes(effect,pred, colour=col), data=pi_effect, size=1) +
   facet_grid(.~experiment, scales="free") +
+  scale_colour_gradient("Subject",limits=c(10,30), guide="none") +
   ylab("Power") + xlab("Effect")
 
 ggsave( file="../images/power_mixed_subject.pdf",height=4,width=10)

### 2012-06-27 paper_jasa.tex
cleaned up the data cleaning section

@@
 
-\subsection{Data Screening}
+\subsection{Data Cleaning}
 
@@
 
-{\bf Screening scenarios:} Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
+%{\bf Screening scenarios:}
+Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA.
+Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data.
+For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
+For the purposes of identifying these participants and cleaning the data, each block of 10 lineups that a subject received included several very easy ones.
+One of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
+The data of participants who did not answer correctly was removed.
+For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
+That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.
+
+%Amazon Mechanical Turk data suffers from the problem that some participants are trying to 'game' the system, because they are being paid for their efforts, not much by the standards of the US minimum wage, but enough to encourage honest efforts.
+in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
 In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.
 %There are six suggested approach to clean the data. 
-They are
-\begin{enumerate}
-\item {\bf include all participants} and their evaluations
-\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
-\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
-\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
+%They are
+%\begin{enumerate}
+%\item {\bf include all participants} and their evaluations
+%\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
+%\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
+%\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
+%\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
+%\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
 If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-\end{enumerate}
-
-\begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
+%\end{enumerate}
+
+%\begin{table}[hbtp]
+%\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
+%Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
 This is due to missing demographic information.
@@
 For me, there's too many numbers with too many digits.}}
+%\begin{center}
+%\begin{tabular}{rrrrr|rrrr|rrrr}
+%  \hline
+%Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
+% \cline{2-5}  \cline{6-9}   \cline{10-13}
+%criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
+%  \hline
+%1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
+%  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
+%  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
+%  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
+%  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
+%  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+%   \hline
+%\end{tabular}
+%\end{center}
+%\label{tbl:summary}
+%\end{table}
+
+\green{Would it make sense to expand this table now to show number of lineups for different experimental conditions?? Maybe not}
+\begin{table}[hbtp]
+\caption{Number of subjects, gender and total lineups for all three experimental data sets.
+Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
+}
 \begin{center}
 \begin{tabular}{rrrrr|rrrr|rrrr}
-  \hline
-Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
- \cline{2-5}  \cline{6-9}   \cline{10-13}
-criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-  \hline
-1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+ & \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\ \cline{2-5}  \cline{6-9}   \cline{10-13}
+ & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
+   & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
    \hline
@@
 
-Table \ref{tbl:summary} shows  the summary results after applying different screening on the data.
-For our analysis we picked criteria six (6) and present results based on this in the following sections.
+Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
 

### 2012-06-27 paper_jasa.tex
small comment

@@
 
-\paragraph{Subject-specific abilities}
+\subsection{Subject-specific abilities}
 Calculating the power  is intricately related to the probability to identify the data plot from the lineup.
@@
 Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
-\begin{equation} \label{mixed}
+We can model $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
+\begin{equation} \label{eqn:mixed}
 g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
@@
 \end{tabular}
-
 }
 
-%\red{This model is  described in more detail in equation \ref{mixed}.}
+The model in equation \eqref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
+\begin{equation} \label{eqn:mixed_power}
+\hat p_{\ell i} = \frac{exp(\hat g_{\ell i})}{1+ exp(\hat g_{\ell i})}
+\end{equation}
+where $\hat g_{\ell i} = W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}$.
+
+%\red{This model is  described in more detail in equation \eqref{eqn:mixed}.}
 
@@
 \noindent
-\paragraph{Lineup size $m$}
+\subsection{Lineup size, $m$}
+
 One of the big differences between visual inference and the conventional testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
@@
 %Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
 
@@
 We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the actual data plot from null plots in a lineup of size $m=20$.
-This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
+% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
 %also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
@@
 
-Note that the \hh{left}  hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the \hh{left}  hand side in equation \eqref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
@@
 
-%From model \ref{mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
+%From model \ref{eqn:mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
 %the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
@@
 combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
-Using equation \ref{mixed}, we then have the power as
+Using equation \eqref{eqn:mixed}, we then have the power as
 %
@@
 %%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.
+As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
 %In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
@@
 \end{equation}
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}. 
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Notice that the expected power of the visual test is almost as good as the power of conventional test.
-\green{What do you mean by this?
-Where does this ``fact'' come from?} In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
+Notice that the expected power of the visual test is %almost as good as
+\hh{close to} the power of the conventional test.
 
@@
 %\begin{thm}\label{thm:power}
-% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
+% equation \eqref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
 %\end{thm}
@@
 %
-%Thus expected power in equation \ref{power_exp} be
+%Thus expected power in equation \eqref{power_exp} be
 %\begin{eqnarray*}
@@
 
-\begin{figure}[htp]
+\begin{figure}[p]
 %\begin{figurehere}
@@
 Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
-Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
+%Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
 A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
@@
 An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? %Plot 7 is the actual data plot.
-\begin{figure}[hbt]
+
+%\begin{figure}[hbt]
 %\begin{figurehere}
-   \centering
-       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
-       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
+%   \centering
+%       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
+%       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
 The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope.
-}
-       \label{fig:cont_dat}
-\end{figure}
-
-\begin{figure}[htp]
+%}
+%       \label{fig:cont_dat}
+%\end{figure}
+
+\begin{figure}[p]
 %\begin{figurehere}
@@
 %       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
+%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
 %       \label{fig:power_observed}
@@
 %       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
+%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
 }
@@
 
-\green{Would it make sense to expand this table now to show number of lineups for different experimental conditions?? Maybe not}
 \begin{table}[hbtp]
-\caption{Number of subjects, gender and total lineups for all three experimental data sets.
+\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets.
 Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
@@
 \begin{center}
-\begin{tabular}{rrrrr|rrrr|rrrr}
- & \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\ \cline{2-5}  \cline{6-9}   \cline{10-13}
- & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-   & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+\begin{tabular}{ccrrcc}
+   \hline 
+ Experiment & Subject & Male & Female & Responses &Lineup\\ 
+    \hline
+  1 & 239 & 121 & 107 & 2317 & 60 \\ 
+  2 & 351 & 185 & 164 & 3858 & 70 \\ 
+  3 & 156 & 106 &  50 & 1681 & 30 \\ 
    \hline
@@
 
+\subsection{Model fitting}
+For each parameter combination, the effect is computed as
+
+\[
+E=\beta (\sqrt {n}/\sigma).
+\]
+
+The model in equation \eqref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept.
+we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
+Different subjects are expected to have different skill levels in evaluating lineups.
+To account for subject specific variation we use subject as random effect.
+We also use an interaction between subject and effect $E$ to allow for subject specific random slope.
+This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in equation \eqref{eqn:mixed}.
+While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$.
+This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (figure \ref{fig:power_loess_effect}).
+
+For all the three experiments, each subject saw a lineup just once.
+This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
+The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using equation \eqref{eqn:mixed_power}.
+For computation we use R \citep{R} package lme4.
+
+Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
+
+\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
+
+\begin{table}[hbtp]
+\caption{Parameter estimates of model \ref{eqn:mixed}.}
+\begin{center}
+\begin{tabular}{ccc}
+  \hline
+ &  \multicolumn{2}{c} {Estimates}  \\
+ \cline{2-3}
+ Experiment & Effect & inter subject correlation\\
+  \hline
+  1 & 0.39 &  \\ 
+  2 & 1.22 &  \\ 
+  3 &  & \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:model_par}
+\end{table}
+
 \subsection{Power comparison}
 
-For each parameter combination, the effect of the lineup is computed as $$E=\beta*\sqrt {n}/\sigma.$$ The experimental data is displayed in figure \ref{fig:power_loess_effect} for different effect sizes.
-The loess curves fitted to the response data provide the estimated observed power for each test.
-The corresponding power curves of conventional tests are shown for comparison.
-The observed power for experiment 2 (scatter plot) is better than what we observed for experiment 1 (box plot).
+\noindent Figure \ref{fig:power_loess_effect} shows the power against effect for the three experiments.
+Data values are represented as dots with size indicating count at this value.
+The loess curves fit the data, estimating the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands.
+For comparison, the dashed lines show the corresponding power curves of the conventional tests.
+The observed power for experiments 1 and 2 are lower than the power of the conventional test.
+This is not a surprise -- it is not expected that visual inference will beat conventional tests, in the scenario when assumptions are satisfied.
+It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
+       \caption{Power in comparison to effect for the three experiments.
+Points indicate subject responses, with size indicating count.
+Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
+The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.
+Conventional test power is drawn as a dashed line.
+For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
+Power for contaminated data is shown by cross mark.
+}
+       \label{fig:power_loess_effect}
+\end{figure}
+
+Results for experiment 3, are quite different.
+This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
+There is no actual conventional power in this situation, because assumptions are violated.
+It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
+The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
+For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
+Notice that the power structure is completely broken for contaminated data.
+For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
+
+It is really curious to see in the results of experiment 3 that the power of the visual test is really high when effect is small.
+These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
+Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data.
+This will be explored further in Section \ref{sec:TypeIII}.
+
+%(scatter plot) is better than what we observed for experiment 1 (box plot).
 For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
 
-Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2.
-For experiment 3 the scenario is different since contamination in the data is visible and this sets the lower limit of proportion correct at a higher value.
-Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
-While fitting model \ref{mixed} we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
-Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
-For experiment 3 we fit the intercept to allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.
-
-The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
+It can also be seen that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2.
+For experiment 3 the scenario is different.
+The contamination in the data is visible which sets the lower limit of proportion correct at a higher value.
+
+%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
 This shows how the screening criteria may affect the results.
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test.
-Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-Size of the dots represents the number of responses for each lineup.
-\blue{Response to the intuition of Di: We are overestimating the effect for experiment 3 here because to compute effect I used the beta values of the linear part of the contaminated data.
-True effect should be much smaller since overall contaminated data should have a smaller slope.
-Thus the effect of 5 that the plot is showing now would actually be effect 2 (may be).
-That means the  power that we see in the plot for effect 5 is actually the power for effect 2.
-How can we get true effect for this data?
-}}
-       \label{fig:power_loess_effect}
-\end{figure}
-
 %
@@
 %       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
-%       \caption{The overall power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
+%       \caption{The overall power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
 %       \label{fig:power_mixed}
@@
 In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
-The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using Equation \ref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
+The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
 
@@
 %       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-%       \caption{Subject specific power obtained from Equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
+%       \caption{Subject specific power obtained from equation \eqref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
 %       \label{fig:power_effect_subject1}
@@
 %       \scalebox{0.40}{\includegraphics{power_model.pdf}}
-%       \caption{Estimated power curve from Equation \ref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
+%       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
@@
 
-%We fit model \eqref{mixed} to the survey data obtained from the simulation experiment.
+%We fit model \eqref{eqn:mixed} to the survey data obtained from the simulation experiment.
 The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}.
-Model \ref{mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}.
+Model \ref{eqn:mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}.
 The plot includes 20 randomly selected subject-specific power curves.
@@
 %       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
-%       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.
+%       \caption{Estimated subject specific power curve from model \ref{eqn:mixed} for sample size $n$ = 100 and $\sigma$ = 12.
 The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
@@
 
-In all the three experiments, each subject saw a lineup only once.
-This allows no random effect for lineup and this simplifies model \ref{mixed} as $\tau_{\ell  i}=\tau_i$.
-This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}.
-Different subject performs differently depending on effect as well.
-So, we include a random interaction term with effect.
-Parameter estimates of the model is shown in table \ref{tbl:model_par}.
-The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}.
-Subject specific power is quite different for three experiments.
-It appears that scatter plot serves better job compared to box plot as a visual test statistic.
-But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
-
-\begin{table}[hbtp]
-\caption{Parameter estimates of model \ref{mixed}.}
-\begin{center}
-\begin{tabular}{ccc}
-  \hline
- &  \multicolumn{2}{c} {Estimates}  \\
- \cline{2-3}
- Experiment & Effect & inter subject correlation\\
-  \hline
-  1 &  &  \\ 
-  2 &  &  \\ 
-  3 &  & \\ 
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:model_par}
-\end{table}
+%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
+ The estimated subject-specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
+The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
+Subject-specific power is quite different for three experiments.
+In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
+In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects.
+In experiment 3 there is the most subject-specific variation.
+Some subjects performed substantially better than the conventional test, and on average the visual test was better.
+
+\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?}
 
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject specific  power estimated from model \ref{mixed} are shown.
+       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
 The corresponding power curve for conventional test (dashed line) is shown for comparison.
@@
 
-\subsection{Estimating $p$-value in Non-Classical Settings}
-
-\green{third experiment doesn't really have a p-value, does it?
+\subsection{Estimating the $p$-value in the real world}
+
+%\green{third experiment doesn't really have a p-value, does it?
 Need to explain this.
@@
 In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values.
-Yet, assessing the strength of the structure is a critical component of inference.
-For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
+Assessing the strength of perceived structure is a critical component of visual inference.
+In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup.
+As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values.
+When $p$-value is larger than 0.15, it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
+
+\green{Add figure showing proportion correct vs p-value of actual data plot for experiments 1 and 2.}
+
+\green{You need to explain visual $p$-value - how is this calculated. How is this different from the signal strength. And please say why you introduce a term signal strength.}
+
+With multiple observers it is possible to estimate the strength of the signal, using the proportion of correct picks.
+From equation \eqref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+This gives the signal strength estimate as
+\[
+\hat p_D= (1- {u}/{K})/({m-1})
+\]
+where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
+Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
+For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions.
+The blue line is a loess fit to the data, and the grey line indicates equality of the two axes.
+The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
+As the conventional $p$-value increases, the estimated signal strength increases -- observers find it harder to detect the actual data plot at larger $p$-values.
+The rise is very sharp for experiment 1, which indicates that observers have a harder time detecting the actual data plot for even small $p$-values.
+
+\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
+
+\green{Be consistent with capitalization of plot labels - check this everywhere - you have lower case signal and upper case Strength}
+
+%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
 For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
@@
 
-\green{Should we use proportion of correct, rather than percentage? Wouldn't that work into the equations better?}
-
-Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
-From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
-This gives the signal strength estimate as
-\[
-\hat p_D= (1- {u}/{K})/({m-1})
-\]
-where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations. 
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values. 
-\hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$.
+%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
+
+Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
 This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense.
-Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
-\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
+Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
+%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
 }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.
@@
        \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
-The solid line is a 45 degree line which represents the equity shown in equation \ref{eqn:power_estimate}.
+The solid line is a 45 degree line which represents the equity shown in equation \eqref{eqn:power_estimate}.
 p-values are shown on square root scale scale.
@@
 
-\begin{table}[hbtp]
-\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
-\begin{center}
-\begin{tabular}{ccc} \hline
-screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
- \cline{2-3}
-criteria & Experiment 1 & Experiment 2 \\ 
-  \hline
-1 & 37 & 60 \\ 
-2 & 38 & 61 \\  
-3 & 38 & 60 \\ 
-4 & 41 & 60 \\ 
-5 & 39 & 60 \\
-6 & 40 & 60 \\ 
-\hline
-Total lineups & 60 & 70 \\  
-\hline
-\end{tabular}
-\end{center}
-\label{tbl:pval_assumption} 
-\end{table}
+%\begin{table}[hbtp]
+%\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
+%\begin{center}
+%\begin{tabular}{ccc} \hline
+%screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
+% \cline{2-3}
+%criteria & Experiment 1 & Experiment 2 \\ 
+%  \hline
+%1 & 37 & 60 \\ 
+%2 & 38 & 61 \\  
+%3 & 38 & 60 \\ 
+%4 & 41 & 60 \\ 
+%5 & 39 & 60 \\
+%6 & 40 & 60 \\ 
+%\hline
+%Total lineups & 60 & 70 \\  
+%\hline
+%\end{tabular}
+%\end{center}
+%\label{tbl:pval_assumption} 
+%\end{table}
 
@@
 
-\subsection{Alternative - Type III error}
+\subsection{Alternative - Type III error}~\label{sec:TypeIII}
 
@@
 \blue{We also see this feature in figure \ref{fig:choice_reason}}.
+\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
 
@@
        \scalebox{0.70}{\includegraphics{power_screening.pdf}}
-       \caption{The overall power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+       \caption{The overall power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for conventional test is shown in black solid lines for comparison.}
@@
        \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+      \caption{The subject-specific power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for conventional test is shown in pink lines for comparison.}

### 2012-06-27 paper_jasa.tex
power of true effect in the contaminated data

--- 2012-06-26 paper_jasa.tex
@@
 Size of the dots represents the number of responses for each lineup.
-\blue{Response to the intuition of Di: We are overestimating the effect for experiment 3 here because to compute effect I used the beta values of the linear part of the contaminated data.
-True effect should be much smaller since overall contaminated data should have a smaller slope.
-Thus the effect of 5 that the plot is showing now would actually be effect 2 (may be).
-That means the  power that we see in the plot for effect 5 is actually the power for effect 2.
-How can we get true effect for this data?
+\blue{For experiment 3, the dotted line represents the conventional power based on the slopes of the linear part of the contaminated data.
+The cross marked points are the conventional power of true effect of the contaminated data.
+We obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated and plotted as cross marked points.
+Notice that the power structure is completely broken for contaminated data.
+For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3.
 }}
@@
   \hline
-  1 &  &  \\ 
-  2 &  &  \\ 
+  1 & 0.39 &  \\ 
+  2 & 1.22 &  \\ 
   3 &  & \\ 

### 2012-06-27 R/paper_jasa.R

@@ -291,6 +291,17 @@ xtable(sm)
 
 # --- fit mixed model with cleaned data
 #library(lme4)
+fit_model <- function(dat){
+  dat$alpha <- log(.05/0.95)
+  dat$effect <- with(dat,abs(beta)/sigma*sqrt(sample_size))
+  model <- as.formula(response ~ offset(alpha) + effect -1 + (effect-1|id))
+  fit <- lmer(model,family="binomial",data=dat)
+  return(summary(fit))
+}
+
+fit_model(dat1)
+fit_model(dat2)
+
 
 pred.mixed <- function(X, subject=0,fit) {
   alpha <- log(.05/0.95)
@@ -873,6 +884,13 @@ get_bootstrap_limit_loess <- function(dat){
 effect.dat <- rbind(data.frame(experiment="Experiment 1", get_response_by_effect(dat1)),
                     data.frame(experiment="Experiment 2", get_response_by_effect(dat2)),
                     data.frame(experiment="Experiment 3", get_response_by_effect(dat3)))
+dat3$effect_hat <- with(dat3,sqrt(sample_size)*abs(beta_hat)/sigma)
+dat3$effect <- with(dat3,sqrt(sample_size)*abs(beta)/sigma)
+effect_hat.dat <- ddply(dat3,.(effect_hat), summarize,
+                        experiment = "Experiment 3",
+                        effect = effect[1],
+                        pow = calculate_ump_power3(beta=effect_hat[1]/sqrt(115),n=115,sigma=1,x=getx(100))  
+                        )
 
 loess.power <- rbind(data.frame(experiment="Experiment 1", get_power_loess(dat1)),
                     data.frame(experiment="Experiment 2", get_power_loess(dat2)),
@@ -890,7 +908,8 @@ loess.limts <- read.csv('../data/loess_bootstrap_limits.txt')
 source("calculate_ump_power.R")
 ump.power <- get_ump_power_by_effect()
 power.dat <- rbind(data.frame(Test="Visual",loess.power),
-                   data.frame(Test="Conventional", ump.power))                          
+                   data.frame(Test="Conventional", ump.power)) 
+
 power.dat$m <- NA
 power.dat$m[power.dat$Test=="Visual"] <- 20
 ggplot()+
@@ -898,9 +917,9 @@ ggplot()+
               data = loess.limts, alpha=.3) +
[…]

### 2012-06-27 paper_jasa.tex
data summary table modified

@@
 \begin{center}
-\begin{tabular}{rrrrr|rrrr|rrrr}
- & \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\ \cline{2-5}  \cline{6-9}   \cline{10-13}
- & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-   & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+\begin{tabular}{ccrrcc}
+   \hline 
+ Experiment & Subject & Male & Female & Total &Lineup\\ 
+    \hline
+  1 & 239 & 121 & 107 & 2317 & 60 \\ 
+  2 & 351 & 185 & 164 & 3858 & 70 \\ 
+  3 & 156 & 106 &  50 & 1681 & 30 \\ 
    \hline

### 2012-06-27 paper_jasa.tex
finished cleaning up power comparison sub-section

@@
 
-For each parameter combination, the effect of the lineup is computed as $$E=\beta*\sqrt {n}/\sigma.$$ The experimental data is displayed in figure \ref{fig:power_loess_effect} for different effect sizes.
-The loess curves fitted to the response data provide the estimated observed power for each test.
-The corresponding power curves of conventional tests are shown for comparison.
-The observed power for experiment 2 (scatter plot) is better than what we observed for experiment 1 (box plot).
+For each parameter combination, the effect is computed as
+
+\[
+E=\beta*\sqrt {n}/\sigma.
+\]
+
+\noindent Figure \ref{fig:power_loess_effect} shows the power against effect for the three experiments.
+Data values are represented as dots with size indicating count at this value.
+The loess curves fit the data, estimating the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands.
+For comparison, the dashed lines show the corresponding power curves of the conventional tests.
+The observed power for experiments 1 and 2 are lower than the power of the conventional test.
+This is not a surprise -- it is not expected that visual inference will beat conventional tests, in the scenario when assumptions are satisfied.
+It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
+Results for experiment 3, are quite different.
+This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
+
+%(scatter plot) is better than what we observed for experiment 1 (box plot).
 For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
 
-Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2.
-For experiment 3 the scenario is different since contamination in the data is visible and this sets the lower limit of proportion correct at a higher value.
-Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$.
-While fitting model \ref{mixed} we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
+It can also be seen that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2.
+For experiment 3 the scenario is different.
+The contamination in the data is visible which sets the lower limit of proportion correct at a higher value.
+For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
+While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
 Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
-For experiment 3 we fit the intercept to allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.
-
-The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
+For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.
+
+%The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
 This shows how the screening criteria may affect the results.
@@
 
-\subsection{Estimating $p$-value in Non-Classical Settings}
+\subsection{Estimating $p$-value in non-classical settings}
 

### 2012-06-27 paper_jasa.tex
tweaked the power comparison subsection a little more

@@
 It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
+
 Results for experiment 3, are quite different.
 This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
+There is no actual conventional power in this situation, because assumptions are violated.
+It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
+The crosses represent \green{...
+please fill this in}.
+It is really curious to see that the power of the visual test is really high when effect is small.
+These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
+Although they were asked to select the plot that exhibited thr highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data.
+This will be explored further in Section \ref{TypeIII}.
 
@@
        \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test.
+       \caption{Power in comparison to effect for the three experiments.
+Points indicate subject responses, with size indicating count.
 Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-Size of the dots represents the number of responses for each lineup.
-\blue{For experiment 3, the dotted line represents the conventional power based on the slopes of the linear part of the contaminated data.
-The cross marked points are the conventional power of true effect of the contaminated data.
+The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.
+Conventional test power is drawn as a dashed line.
+For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
+\green{These next two sentences are confusing.
+Maybe too much for caption, and should be moved to text.} Crosses mark points that are the conventional power of true effect of the contaminated data.
 We obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated and plotted as cross marked points.
@@
 For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3.
-}}
+}
        \label{fig:power_loess_effect}
@@
 
-\subsection{Alternative - Type III error}
+\subsection{Alternative - Type III error}~\label{TypeIII}
 

### 2012-06-27 paper_jasa.tex
summary table heading
power comparison discussion

@@
    \hline 
- Experiment & Subject & Male & Female & Total &Lineup\\ 
+ Experiment & Subject & Male & Female & Responses &Lineup\\ 
     \hline
@@
 It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
-The crosses represent \green{...
-please fill this in}.
+The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
+For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
+Notice that the power structure is completely broken for contaminated data.
+For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
+
 It is really curious to see that the power of the visual test is really high when effect is small.
 These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
-Although they were asked to select the plot that exhibited thr highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data.
-This will be explored further in Section \ref{TypeIII}.
+Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data.
+This will be explored further in Section \ref{sec:TypeIII}.
 
@@
 For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
-\green{These next two sentences are confusing.
-Maybe too much for caption, and should be moved to text.} Crosses mark points that are the conventional power of true effect of the contaminated data.
-We obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated and plotted as cross marked points.
-Notice that the power structure is completely broken for contaminated data.
-For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3.
 }
@@
 
-\subsection{Alternative - Type III error}~\label{TypeIII}
+\subsection{Alternative - Type III error}~\label{sec:TypeIII}
 

### 2012-06-28 paper_jasa.tex
caption changed for loess figure
subecjt-specific variation

@@
 
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
+       \caption{Power in comparison to effect for the three experiments.
+Points indicate subject responses, with size indicating count.
+Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
+The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.
+Conventional test power is drawn as a dashed line.
+For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
+Power for contaminated data is shown by cross mark.
+}
+       \label{fig:power_loess_effect}
+\end{figure}
+
 Results for experiment 3, are quite different.
@@
 
-It is really curious to see that the power of the visual test is really high when effect is small.
+It is really curious to see in the results of experiment 3 that the power of the visual test is really high when effect is small.
 These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
@@
 Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Power in comparison to effect for the three experiments.
-Points indicate subject responses, with size indicating count.
-Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.
-Conventional test power is drawn as a dashed line.
-For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
-}
-       \label{fig:power_loess_effect}
-\end{figure}
 
@@
 
-In all the three experiments, each subject saw a lineup only once.
-This allows no random effect for lineup and this simplifies model \ref{mixed} as $\tau_{\ell  i}=\tau_i$.
-This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}.
-Different subject performs differently depending on effect as well.
+\green{Do we have more than one model??
+If so then it is appropriate to refer to them as model 1, model 2, ...., but if there is only one model then we need to refer readers back to the equation, occasionally.}
+
+\green{Actually I think this paragraph should have already been covered when the model was introduced.
+At this point we just need to discuss the results, so I'm going to cut the paragraph and start again.}
+
+%Different subjects are expected to have different skill levels in evaluating lineups.
+This is accounted for in the mixed effects model (Equation \ref{mixed}).
+For all the three experiments, each subject saw a lineup just once.
+This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i}=\tau_i$.
+The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$.
 So, we include a random interaction term with effect.
-Parameter estimates of the model is shown in table \ref{tbl:model_par}.
-The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}.
-Subject specific power is quite different for three experiments.
-It appears that scatter plot serves better job compared to box plot as a visual test statistic.
-But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
+Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
+The estimated subject specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
+The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
+Subject-specific power is quite different for three experiments.
+In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
+In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects.
+In experiment 3 there is the most subject-specific variation.
+Some subjects performed substantially better than the conventional test, and on average the visual test was better.
+
+\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?}
 

### 2012-06-28 paper_jasa.tex
added a model fit section

@@
 Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
+We can model $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
 \begin{equation} \label{mixed}
@@
 
+\subsection{Model fitting}
+
+The model in Equation \ref{mixed}is fit using \green{....
+fill this in} Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
+For all the three experiments, each subject saw a lineup just once.
+This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell
+  i}=\tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$. So, we include a random interaction term with effect.
+
+Different subjects are expected to have different skill levels in evaluating lineups.
+For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
+While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
+Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
+For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.
+
+\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
+
+\begin{table}[hbtp]
+\caption{Parameter estimates of model \ref{mixed}.}
+\begin{center}
+\begin{tabular}{ccc}
+  \hline
+ &  \multicolumn{2}{c} {Estimates}  \\
+ \cline{2-3}
+ Experiment & Effect & inter subject correlation\\
+  \hline
+  1 & 0.39 &  \\ 
+  2 & 1.22 &  \\ 
+  3 &  & \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:model_par}
+\end{table}
+
 \subsection{Power comparison}
@@
 The contamination in the data is visible which sets the lower limit of proportion correct at a higher value.
-For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
-While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
-Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
-For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.
 
@@
 
-\green{Do we have more than one model??
-If so then it is appropriate to refer to them as model 1, model 2, ...., but if there is only one model then we need to refer readers back to the equation, occasionally.}
-
-\green{Actually I think this paragraph should have already been covered when the model was introduced.
-At this point we just need to discuss the results, so I'm going to cut the paragraph and start again.}
-
-%Different subjects are expected to have different skill levels in evaluating lineups.
-This is accounted for in the mixed effects model (Equation \ref{mixed}).
-For all the three experiments, each subject saw a lineup just once.
-This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i}=\tau_i$.
-The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$.
-So, we include a random interaction term with effect.
-Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-The estimated subject specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
+%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{mixed}).   
+ The estimated subject-specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
 The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
@@
 
-\begin{table}[hbtp]
-\caption{Parameter estimates of model \ref{mixed}.}
-\begin{center}
-\begin{tabular}{ccc}
-  \hline
- &  \multicolumn{2}{c} {Estimates}  \\
- \cline{2-3}
- Experiment & Effect & inter subject correlation\\
-  \hline
-  1 & 0.39 &  \\ 
-  2 & 1.22 &  \\ 
-  3 &  & \\ 
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:model_par}
-\end{table}
-
 \begin{figure}[hbtp]
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject specific  power estimated from model \ref{mixed} are shown.
+       \caption{Subject-specific  power estimated from model \ref{mixed} are shown.
 The corresponding power curve for conventional test (dashed line) is shown for comparison.
@@
 
-\subsection{Estimating $p$-value in non-classical settings}
+\subsection{Estimating the $p$-value in the real world}
 
@@
        \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+      \caption{The subject-specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for conventional test is shown in pink lines for comparison.}

### 2012-06-28 paper_jasa.tex
pvalues in real world updated

@@
 
-\green{third experiment doesn't really have a p-value, does it?
+%\green{third experiment doesn't really have a p-value, does it?
 Need to explain this.
@@
 In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values.
-Yet, assessing the strength of the structure is a critical component of inference.
-For experiments 1 and 2, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
-For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
-As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
-When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
-We observed this for both experiment 1 and 2.
-For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
-
-\green{Should we use proportion of correct, rather than percentage? Wouldn't that work into the equations better?}
-
-Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
+Assessing the strength of perceived structure is a critical component of visual inference.
+In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup.
+As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values.
+When $p$-value is larger than 0.15, it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
+
+\green{Add figure showing proportion correct vs p-value of actual data plot for experiments 1 and 2.}
+
+\green{You need to explain visual $p$-value - how is this calculated. How is this different from the signal strength. And please say why you introduce a term signal strength.}
+
+With multiple observers it is possible to estimate the strength of the signal, using the proportion of correct picks.
 From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
@@
 where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values. 
-\hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$.
+Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
+For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions.
+The blue line is a loess fit to the data, and the grey line indicates equality of the two axes.
+The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
+As the conventional $p$-value increases, the estimated signal strength increases -- observers find it harder to detect the actual data plot at larger $p$-values.
+The rise is very sharp for experiment 1, which indicates that observers have a harder time detecting the actual data plot for even small $p$-values.
+
+\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
+
+\green{Be consistent with capitalization of plot labels - check this everywhere - you have lower case signal and upper case Strength}
+
+%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
+For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
+As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
+When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
+We observed this for both experiment 1 and 2.
+For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
+
+%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
+
+Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
 This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense.
-Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
-\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
+Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
+%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
 }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.

### 2012-06-28 paper_jasa.tex
model fitting details

@@
 We can model $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
-\begin{equation} \label{mixed}
+\begin{equation} \label{eqn:mixed}
 g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
@@
 \end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}.
 Notice that the expected power of the visual test is almost as good as the power of conventional test.
@@
 \subsection{Model fitting}
-
-The model in Equation \ref{mixed}is fit using \green{....
-fill this in} Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
+For each parameter combination, the effect is computed as
+
+\[
+E=\beta (\sqrt {n}/\sigma).
+\]
+
+The model in equation \eqref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept.
+we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
+Different subjects are expected to have different skill levels in evaluating lineups.
+To account for subject specific variation we use subject as random effect.
+We also use an interaction between subject and effect $E$ to allow for subject specific random slope.
+This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in equation \eqref{eqn:mixed}.
+While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$.
+This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (figure \ref{fig:power_loess_effect}).
+
 For all the three experiments, each subject saw a lineup just once.
-This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell
-  i}=\tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$. So, we include a random interaction term with effect.
-
-Different subjects are expected to have different skill levels in evaluating lineups.
-For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
-While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated.
-Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2.
-For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.
+This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
+The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using equation \eqref{eqn:mixed_power}.
+For computation we use R \citep{R} package lme4.
+
+Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
 
@@
 \begin{table}[hbtp]
-\caption{Parameter estimates of model \ref{mixed}.}
+\caption{Parameter estimates of model \ref{eqn:mixed}.}
 \begin{center}
@@
 \subsection{Power comparison}
-
-For each parameter combination, the effect is computed as
-
-\[
-E=\beta*\sqrt {n}/\sigma.
-\]
 
@@
 
-%The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
+%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
 This shows how the screening criteria may affect the results.
@@
 
-%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{mixed}).   
+%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
  The estimated subject-specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject-specific  power estimated from model \ref{mixed} are shown.
+       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
 The corresponding power curve for conventional test (dashed line) is shown for comparison.
@@
 With multiple observers it is possible to estimate the strength of the signal, using the proportion of correct picks.
-From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+From equation \eqref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
 This gives the signal strength estimate as
@@
        \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject-specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.
+      \caption{The subject-specific power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.
 The corresponding power curve for conventional test is shown in pink lines for comparison.}

### 2012-06-28 R/paper_jasa.R

@@ -294,7 +294,7 @@ xtable(sm)
 fit_model <- function(dat){
   dat$alpha <- log(.05/0.95)
   dat$effect <- with(dat,abs(beta)/sigma*sqrt(sample_size))
-  model <- as.formula(response ~ offset(alpha) + effect -1 + (effect-1|id))
+  model <- as.formula(response ~ offset(alpha) + effect -1 + (effect -1|id))
   fit <- lmer(model,family="binomial",data=dat)
   return(summary(fit))
 }
@@ -384,8 +384,8 @@ pi_effect$col <- 30
 pi_effect$col[1] <- NA
 ggplot()+
   geom_line(aes(effect,pred, group=subject), data=pi_subject,alpha=.08) +
-  geom_line(aes(effect,pow), data=ump, size=1, linetype=2) +
   geom_line(aes(effect,pred, colour=col), data=pi_effect, size=1) +
+  geom_line(aes(effect,pow), data=ump, size=1, linetype=2) +
   facet_grid(.~experiment, scales="free") +
   scale_colour_gradient("Subject",limits=c(10,30), guide="none") +
   ylab("Power") + xlab("Effect")
@@ -803,7 +803,6 @@ qplot(p_value,visual_pval, data=pdat) +
   facet_grid(.~experiment)+
   scale_x_sqrt() +scale_y_sqrt()
 ggsave(file="../images/p_val_definition.pdf", height = 4, width = 10)
-
 
 
 qplot(strength,visual_pval, data=pdat) +

### 2012-06-28 R/paper_jasa.R
linear smoother

@@ -1,4 +1,4 @@
-# The complete R codes for JASA Paper
+# The complete R code for JASA Paper
 
 library(ggplot2)
 library(plyr)
@@ -783,7 +783,7 @@ p
 p <- ggplot(pdat) +
   geom_abline(aes(intercept=0,slope=1), color="grey") +
   geom_point(aes(p_value,strength),size=2) + 
-  stat_smooth(aes(p_value,strength),method="loess", se=F, span=1)+
+  stat_smooth(aes(p_value,strength),method="loess", se=F, span=1, degree=1)+
   facet_grid(.~experiment) +
   xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
   ylab(expression(paste("Plot signal Strength (", hat(p)[D],") on square root scale"))) + 
@@ -797,8 +797,10 @@ ggsave(p,file="../images/p_val_plot_signal.pdf", height = 4, width = 10)
 
 pdat$visual_pval <- apply(pdat[,2:3],1,function(x)return(get_pval(m=20,K=x[1],x=x[2])))
 qplot(p_value,visual_pval, data=pdat) +
-  stat_smooth(method="loess", se=F, span=.4)+
-  xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
+  geom_abline(colour="grey") +
+#  stat_smooth(method="loess", se=F, span=.4) + #5, degree=1)+
+  stat_smooth(method="loess", se=F, span=.45, degree=1)+
+    xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
   ylab(expression(paste("Visual p-value on square root scale"))) +
   facet_grid(.~experiment)+
   scale_x_sqrt() +scale_y_sqrt()

### 2012-06-28 paper_jasa.tex
small change from merge

@@
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}. 
-Notice that the expected power of the visual test is almost as good as the power of conventional test.
-\green{What do you mean by this?
-Where does this ``fact'' come from?} In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
+Notice that the expected power of the visual test is %almost as good as
+\hh{close to} the power of the conventional test.
 

### 2012-06-28 paper_jasa.tex
defined visual p-value, tweaked equations and reference

@@
 For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
-Under the null hypothesis $X \sim \text{Binom}({K, 1/m})$,  therefore:
+Under the null hypothesis $X \sim \text{Binom}_{K, 1/m}$,  therefore:
 
@@
 The $p$-value of a lineup of size $m$ is given as
-\[
-P(X \ge x) = 1 - \text{Binom}({K, 1/m}) (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
-\]
-with $X$ defined as above. 
+
+\begin{equation}\label{binom}
+P(X \ge x) = 1 - \text{Binom}_{{K, 1/m}} (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i}
+\end{equation}
+
+\noindent with $X$ defined as above, and $x$ is the number of observers selecting the actual data plot. 
 \end{dfn}
@@
 Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup.
-If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$.
+If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$.
 We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
@@
 
-\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data.
+\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
 \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
@@
 
-The model in equation \eqref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
+The model in Equation \ref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
 \begin{equation} \label{eqn:mixed_power}
@@
 
-Note that the \hh{left}  hand side in equation \eqref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
@@
 \end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}. 
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}. 
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
 Notice that the expected power of the visual test is %almost as good as
-\hh{close to} the power of the conventional test.
+close to the power of the conventional test.
 
@@
 
-With multiple observers it is possible to estimate the strength of the signal, using the proportion of correct picks.
-From equation \eqref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+With multiple observers it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
+The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
+Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more observers select that plot.
+Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.
+
+Similary, the strength of the signal, is calculated using the proportion of observers picking actual data plot.
+It is closely related to the power of the test.
+From Equation \eqref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
 This gives the signal strength estimate as
@@
        \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$) estimated using Lemma \ref{lemma}.
-The solid line is a 45 degree line which represents the equity shown in equation \eqref{eqn:power_estimate}.
+       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
+The solid line is a 45 degree line which represents the equity shown in Equation \eqref{eqn:power_estimate}.
 p-values are shown on square root scale scale.

### 2012-06-28 paper_jasa.tex
equation to Equation, eqref to ref

@@
 
-The model in equation \eqref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept.
+The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept.
 we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
@@
 We also use an interaction between subject and effect $E$ to allow for subject specific random slope.
-This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in equation \eqref{eqn:mixed}.
+This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.
 While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$.
-This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (figure \ref{fig:power_loess_effect}).
+This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (Figure \ref{fig:power_loess_effect}).
 
@@
 This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
-The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using equation \eqref{eqn:mixed_power}.
+The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}.
 For computation we use R \citep{R} package lme4.
@@
 It is closely related to the power of the test.
-From Equation \eqref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
+From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
 This gives the signal strength estimate as
@@
        \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
-The solid line is a 45 degree line which represents the equity shown in Equation \eqref{eqn:power_estimate}.
+The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
 p-values are shown on square root scale scale.

### 2012-06-28 paper_jasa.tex
working on lwest pvalue picks

@@
 The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
-p-values are shown on square root scale scale.
-\blue{Is it better now?}}
+p-values are shown on square root scale scale.}
     %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
@@
 
-In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
+One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
+In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
+This can be examined for the data collected, to see if this assumption is, indeed, reasonable.
+%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
 In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
@@
 Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
-(Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value.
+(Empty cells indicate no lineup shown with this combination.) Red indicates the plot with the lowest $p$-value in the lineup.
 In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
@@
 The results are clearer for experiment 2, that used a continuous covariate.
+But even when they didn't pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
 So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
+
+There are some noticeable exceptions to this rule.
+In experiment 1, when $\beta=5, n=300, sd=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.
+Same experiment, when $\beta=0, n=100, sd=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices.
+\cite{*** yifan's CC needs to be made into a tech report, and referenced here.}
 
@@
 
+\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
+Along the lines of the Army tests that Susan brought to the working group.}
+
 \paragraph{Acknowledgement:}

### 2012-06-28 paper_jasa.tex
model fitting discussion

--- 2012-06-28 paper_jasa.tex
@@
 \begin{equation} \label{eqn:mixed_power}
-\hat p_{\ell i} = \frac{exp(\hat g_{\ell i})}{1+ exp(\hat g_{\ell i})}
+\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
 \end{equation}
-where $\hat g_{\ell i} = W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}$.
 
@@
 To account for subject specific variation we use subject as random effect.
-We also use an interaction between subject and effect $E$ to allow for subject specific random slope.
+We do not include random intercept, instead, we use an interaction between subject and effect $E$ to allow for subject specific random slope.
 This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.
@@
 \begin{table}[hbtp]
-\caption{Parameter estimates of model \ref{eqn:mixed}.}
+\caption{Parameter estimates of model in Equation \ref{eqn:mixed}.}
 \begin{center}
-\begin{tabular}{ccc}
-  \hline
- &  \multicolumn{2}{c} {Estimates}  \\
+\begin{tabular}{cccc}
+  \hline
+ &  \multicolumn{2}{c} {Fixed effect}  & Random effect\\
  \cline{2-3}
- Experiment & Effect & inter subject correlation\\
-  \hline
-  1 & 0.39 &  \\ 
-  2 & 1.22 &  \\ 
-  3 &  & \\ 
+ Experiment & Estimate & Std. error & Variance\\
+  \hline
+  1 & 0.39 & 0.0093 & 0.0077 \\ 
+  2 & 1.22 &  0.0196 &  0.0455 \\ 
+  3 &  & & \\ 
    \hline

### 2012-06-28 R/paper_jasa.R

@@ -358,9 +358,9 @@ get_screened_predict <- function(dat, newdat,intercept=F){
   return(res)
 }
 get_screened_predict(raw.dat1, newdat=data.frame(effect))
-pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(raw.dat1, newdat=data.frame(effect=seq(0,18, by=.2)))),
-            data.frame(experiment="Experiment 2",get_screened_predict(raw.dat2, newdat=data.frame(effect=seq(0,7, by=.2)))),
-            data.frame(experiment="Experiment 3",get_screened_predict(raw.dat3, newdat=data.frame(effect=seq(0,6, by=.2)))))
+pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(raw.dat1, newdat=data.frame(effect=seq(0,16, by=.2)))),
+            data.frame(experiment="Experiment 2",get_screened_predict(raw.dat2, newdat=data.frame(effect=seq(0,5.5, by=.2)))),
+            data.frame(experiment="Experiment 3",get_screened_predict(raw.dat3, newdat=data.frame(effect=seq(0,4.5, by=.2)))))
 
 # source("calculate_ump_power.R")
 ump <- get_ump_power_by_effect()
@@ -395,9 +395,9 @@ ggsave( file="../images/power_mixed_subject.pdf",height=4,width=10)
 
 # ----- subject specific power for each screening criteria from mixed model
 
-pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(raw.dat1, newdat=data.frame(effect=seq(0,18, by=.2)),intercept=T)),
-            data.frame(experiment="Experiment 2",get_screened_predict(raw.dat2, newdat=data.frame(effect=seq(0,7, by=.2)),intercept=T)),
-            data.frame(experiment="Experiment 3",get_screened_predict(raw.dat3, newdat=data.frame(effect=seq(0,6, by=.2)),intercept=T)))
+pi <- rbind(data.frame(experiment="Experiment 1",get_screened_predict(raw.dat1, newdat=data.frame(effect=seq(0,16, by=.2)),intercept=T)),
+            data.frame(experiment="Experiment 2",get_screened_predict(raw.dat2, newdat=data.frame(effect=seq(0,5.5, by=.2)),intercept=T)),
[…]

### 2012-06-29 R/paper_jasa.R
model estimates

@@ -392,6 +392,17 @@ ggplot()+
 
 ggsave( file="../images/power_mixed_subject.pdf",height=4,width=10)
 
+# subjects having more power than ump
+
+subject_ump <- merge(pi_subject,ump, by=c("experiment","effect"))
+dd <- ddply(subject_ump, .(experiment, effect), summarize,
+      prop = mean(pred>pow))
+qplot(effect,prop, data=subset(dd, prop>0), geom="line", colour=experiment) +
+  ylab("Proportion of subject having  power \n above conventional test") +
+  scale_colour_discrete(name="Data") +xlab("Effect")
+ggsave( file="../images/subject_having_more_power.pdf",height=4,width=6)
+
+
 
 # ----- subject specific power for each screening criteria from mixed model
 
@@ -765,6 +776,33 @@ p <- ggplot() +
 p 
 
 ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4, width = 10)
+
+# uncontaminated p-value for experiment 3
+qplot(uncontaminated_pvalue,p_value, data=dat3) +
+  scale_x_sqrt()+scale_y_sqrt()+
+  ylab("Contaminated p-value") +
+  xlab("Uncontaminated p-value")
+
+pval.exp3 <- ddply(dat3, .(pic_name), summarize,
+                   prop_correct = mean(response),
+                   strength = (1 - mean(response))/19,
+                   p_value_contaminated = p_value[1],
+                   p_value_uncontaminated = uncontaminated_pvalue[1])
+
+qplot( p_value_uncontaminated, prop_correct, data=pval.exp3)
+qplot( p_value_contaminated, prop_correct, data=pval.exp3)
+
+pval.exp3.m <- melt(pval.exp3[,-c(1,3)], id="prop_correct")
+qplot(value, prop_correct, data=pval.exp3.m) + 
+  facet_grid(.~variable) + xlab("p-value") +ylab("Proportion correct")+
+  scale_x_sqrt()
+
+pval.exp3.m <- melt(pval.exp3[,-c(1,2)], id="strength")
+qplot(value, strength, data=pval.exp3.m) + 
+  facet_grid(.~variable) + xlab("p-value") +ylab("Plot signal strength")+
+  scale_x_sqrt()+ scale_y_sqrt()
+
+ggsave(file="../images/p_val_plot_signal_exp3.pdf", height = 4, width = 8)
 
 # ----- p-value vs plot signal strength (visual p-value) 
 

### 2012-06-29 paper_jasa.tex
model extimates table caption

@@
 
-The model in Equation \ref{eqn:mixed} provides the subject and lineup specific probability of successful evaluation as 
+Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} we obtain the the estimate of subject and lineup specific probability of successful evaluation as 
 \begin{equation} \label{eqn:mixed_power}
@@
 \begin{table}[hbtp]
-\caption{Parameter estimates of model in Equation \ref{eqn:mixed}.}
+\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
 \begin{center}

### 2012-06-29 paper_jasa.tex
p-value vs prop-correct figure added

@@
 
-\green{Add figure showing proportion correct vs p-value of actual data plot for experiments 1 and 2.}
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
+       \caption{Proportion of correct responses decreases rapidly with the increase of p-value.
+After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
+The theoretical justification of this is shown in figure \ref{fig:pval_power}.
+}
+       \label{fig:pval_pcorrect}
+\end{figure*}
+
+%\green{Add figure showing proportion correct vs p-value of actual data plot for experiments 1 and 2.}
 
@@
 
-\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
+\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?} \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
 
@@
 %and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
-%       \caption{Percentage of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in figure \ref{fig:pval_power}.
-\blue{The scenario is different for experiment 3.
-$p$-value does not seem to have any effect on percent correct.
-This may be due to Type-III error.} \green{p-value on log scale?} }
-%       \label{fig:pval_pcorrect}
-%\end{figure*}
 

### 2012-06-29 R/paper_jasa.R

@@ -769,13 +769,14 @@ for (i in 1:3){
 # pdat$percent_correct <- pdat$corrected*100/pdat$attempted
 
 p <- ggplot() +
-     geom_point(aes(log(p_value),percent_correct),data=pdat,size=2) + 
+     geom_point(aes(p_value,percent_correct/100),data=subset(pdat, experiment !="Experiment 3"),size=2) + 
      facet_grid(.~experiment) +
-     xlab(expression(paste(log[10], "(p-value(",p[D],"))"))) +
-     ylab("Percentage of correct responses") 
+     xlab(expression(paste("p-value(",p[D],") on squqre root scale"))) +
+     ylab("Proportion correct on square root scale") +
+     scale_x_sqrt()+ scale_y_sqrt()
 p 
 
-ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4, width = 10)
+ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4, width = 7.5)
 
 # uncontaminated p-value for experiment 3
 qplot(uncontaminated_pvalue,p_value, data=dat3) +

### 2012-06-29 paper_jasa.tex
effect of null plots

@@
 
-%\green{Add figure showing proportion correct vs p-value of actual data plot for experiments 1 and 2.}
-
-\green{You need to explain visual $p$-value - how is this calculated. How is this different from the signal strength. And please say why you introduce a term signal strength.}
-
 With multiple observers it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
@@
 
-\green{Be consistent with capitalization of plot labels - check this everywhere - you have lower case signal and upper case Strength}
+\green{Explain why the term ``plot signal'' is introduced}
 
@@
 
-\hh{
 \subsection{How much do null plots affect the choice?}
-Another difference between conventional tests and visual inference is the potential influence that null plots have on an observer's choice of plot.
-If one or more null plots represent data with -- by chance -- a smaller $p$-value than the data plot, there's a strong possibility that an observer will pick the null plot instead of the data plot, independently of the actual value of the $p$-value.
-
-See figure \ref{fig:pval_difference} of evidence for this phenomenon.
+
+Another difference between conventional tests and visual inference is that the observer's choice of plot may be influenced by the $m-1$ null plots in the lineup.
+We have already seen that observers tend to pick the plot with the lowest $p$-value.
+If one or more null plots represent data having -- by chance -- a smaller $p$-value than the actual data plot, there's a strong possibility that an observer will pick this null plot instead of the actual data plot.
+
+Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
+The horizontal axis shows the difference between the smalles $p$-value of the null plots, and that of the actual data plot.
+Large values indicate that the actual data plot has a much smaller $p$-value than any of the null plots.
+The vertical axis shows percent subjects choosing the actual data plot.
+\green{Is this really pct correct, or 1-pct correct?
+When there is a big diff in p-value almost no-one gets it right!
+Maybe it is just a scale issue?} We see that when there is a big difference between the $p$-values observers choose the actual data plot more often.
+
 %Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
@@
 
-This problem is further investigated in the paper \cite{waldo}.
-
-}
+This problem is further investigated in the paper \cite{waldo}. Here numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
 
@@
 \green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
-Along the lines of the Army tests that Susan brought to the working group.}
+Along the lines of the Army tests that Susan brought to the working group.
+
+Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
 

### 2012-06-29 R/paper_jasa.R
uncontaminated p-value plot added

@@ -751,7 +751,7 @@ qplot(prop_correct,choice_reason, geom="point", data=reason.dat3, size=I(3)) +
 ggsave("../images/choice_reason.pdf", height = 4, width = 6)
 
 
-# ======================= p_value vs %correct =====================
+# ======================= p_value vs proportion correct =====================
 
 pdat <- NULL
 for (i in 1:3){
@@ -759,56 +759,51 @@ for (i in 1:3){
    pdati <- ddply(dati, .(pic_name), summarize,
 	  attempted = length(response),
 	  corrected = sum(response),
-	  percent_correct = mean(response)*100,
+	  prop_correct = mean(response),
+    strength = (1 - mean(response))/19,              
     p_value = p_value[1]
 	)
    pdati$experiment=paste("Experiment",i)
    pdat <- rbind(pdat,pdati)
 }
 
-# pdat$percent_correct <- pdat$corrected*100/pdat$attempted
+# uncontaminated p-value for experiment 3
+pdati <- ddply(dat3, .(pic_name), summarize,
+               attempted = length(response),
+               corrected = sum(response),
+               prop_correct = mean(response),
+               strength = (1 - mean(response))/19,              
+               p_value = uncontaminated_pvalue[1]
+               )
+pdati$experiment= "Experiment 3 uncontaminated"
+pdat <- rbind(pdat,pdati)
 
-p <- ggplot() +
-     geom_point(aes(p_value,percent_correct/100),data=subset(pdat, experiment !="Experiment 3"),size=2) + 
+
+p <- ggplot(subset(pdat, substr(experiment,1,12) !="Experiment 3")) +
+     geom_point(aes(p_value,prop_correct),size=2) + 
      facet_grid(.~experiment) +
      xlab(expression(paste("p-value(",p[D],") on squqre root scale"))) +
      ylab("Proportion correct on square root scale") +
      scale_x_sqrt()+ scale_y_sqrt()
 p 
+ggsave(p,file="../images/p_val_prop_correct.pdf", height = 4, width = 7.5)
 
-ggsave(p,file="../images/p_val_percent_correct.pdf", height = 4, width = 7.5)
 
-# uncontaminated p-value for experiment 3
 qplot(uncontaminated_pvalue,p_value, data=dat3) +
   scale_x_sqrt()+scale_y_sqrt()+
[…]

### 2012-07-01 paper_jasa.tex
comment on choices
uncontaminated p-value plot added
worked on last two sections of paper - some questions to be answered

@@
 One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots?
+Which plot is the most different from the others, in the sense that there is the largest shift or location difference between the boxplots?
 (The position of the actual data plot is provided in Section \ref{sec:category}.)}
@@
 
-Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} we obtain the the estimate of subject and lineup specific probability of successful evaluation as 
+Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} we obtain the estimate of subject and lineup specific probability of successful evaluation as 
 \begin{equation} \label{eqn:mixed_power}
@@
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
 Notice that the expected power of the visual test is %almost as good as
@@
 Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-There are three experiments.
 The first two are classical settings where we would not expect the lineup protocol to do better than the traditional test, but we hope it to perform favorably.
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_percent_correct.pdf}}
+       \scalebox{0.6}{\includegraphics{p_val_prop_correct.pdf}}
        \caption{Proportion of correct responses decreases rapidly with the increase of p-value.
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
+       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
        \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
@@
 So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
+
+\begin{figure*}[hbtp]
+   \centering
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
+The colored lines indicate the plot with the lowest $p$-value.
+Rows correspond to $\beta$, smallest to largest values from top to bottom.
+Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
+Empty cells indicate no lineup for that combination of parameters.
+Highest counts tend to be the plot in the lineup having the lowest $p$-value.}
+       \label{fig:P-val_log}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+   \centering
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
+The colored lines indicate the plot with the lowest $p$-value.
+Rows correspond to $\beta$, smallest to largest values from top to bottom.
+Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
+Empty cells indicate no lineup for that combination of parameters.
+Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.}
+       \label{fig:P-val_log2}
+\end{figure*}
 
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-The colored lines indicate the plot with the lowest $p$-value.
-Rows correspond to $\beta$, smallest to largest values from top to bottom.
-Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
-Empty cells indicate no lineup for that combination of parameters.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value.}
-       \label{fig:P-val_log}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
-The colored lines indicate the plot with the lowest $p$-value.
-Rows correspond to $\beta$, smallest to largest values from top to bottom.
-Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
-Empty cells indicate no lineup for that combination of parameters.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.}
-       \label{fig:P-val_log2}
-\end{figure*}
-
 \subsection{How much do null plots affect the choice?}
 
-Another difference between conventional tests and visual inference is that the observer's choice of plot may be influenced by the $m-1$ null plots in the lineup.
-We have already seen that observers tend to pick the plot with the lowest $p$-value.
-If one or more null plots represent data having -- by chance -- a smaller $p$-value than the actual data plot, there's a strong possibility that an observer will pick this null plot instead of the actual data plot.
-
+Visual inference is like other types of computationally intensive test procedures, such as permutation tests, where the statistics from the data are compared with those from null data.
+But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
+How this small set of null plots influences the observers choice is important for understanding the reliability of visual inference.
+If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
+Measuring the difference between the plots is almost impossible!
+But in the controlled setting here we can again use $p$-value as a proxy for similarity and difference.
+
+In the previous section we noticed that observers tend to pick the plot with the lowest $p$-value.
+If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an observer will pick this null plot instead of the actual data plot.
 Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
-The horizontal axis shows the difference between the smalles $p$-value of the null plots, and that of the actual data plot.
-Large values indicate that the actual data plot has a much smaller $p$-value than any of the null plots.
-The vertical axis shows percent subjects choosing the actual data plot.
+The plots display the proportion of observers who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
+A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
+If there is a big negative difference we'd expect to see We would expect to see that the observers would more often make the correct choice of plot.
+This can be seen.
+When there is a large positive difference If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
+
+\green{Something is definitely not right! The code looks like it does actual data p-value - min(null pvalue). But how can you get a difference of 0.8 from this??}
+
  \green{Is this really pct correct, or 1-pct correct?
@@
 
-This problem is further investigated in the paper \cite{waldo}. Here numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
+In practice, $p$-value is not going to be a valid way to compare plots.
+Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
+This is investigated in the paper \cite{waldo}.
+Here numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
 
@@
 
-\subsection{Alternative - Type III error}~\label{sec:TypeIII}
-
-\hh{From figure \ref{fig:power_loess_effect} we see that identification of the data plot is high independently of the effect size for experiment 3.
-This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}
+\subsection{Type III error}~\label{sec:TypeIII}
+
+A little known error amongst statisticians is what was coined as Type III in \citet{mosteller:48}.
+This occurs when the null hypothesis is rejected for thw rong reason.
+Experiment 3 is prone to this type of error.
+Observers were asked to identify the plot that had the largest slope, either positive or negative.
+But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see the association.
+For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
+This would be considered to be a Type III error because it is not what was requested, but it still gives the correct answer.
+This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
+Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
+
+Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
+Observers were provided with four possible choices
+
+\begin{enumerate} \itemsep 0in
+\item Most different plot
+\item Visible trend
+\item Clustering visible
+\item Other
+\end{enumerate}
+
+\noindent It was possible to choose more than one reason.
+The question begged for ``visible trend'' (2) to be the reasons that observers gave for correctly choosing the actual data plot.
+Either reasons 1 or 3 would be indicative of Type III error.
+It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
+Figure \ref{fig:choice_reason} examines this.
+For observers who chose the actual data plot, the proportions for the three reasons are shown.
+It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
+\green{What about other??}
+
+%This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
+the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}
 
@@
 
-\blue{We also see this feature in figure \ref{fig:choice_reason}}.
-\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
+%\blue{We also see this feature in figure \ref{fig:choice_reason}}.
+%\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
+
+\green{I find this plot really confusing! Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?}
 
@@
 
+In practice there is no correct answer.
+In these experiments we controlled the effets that observers saw, so there was a correct choice in each case.
+In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
+And observers would provide open-ended answers for their choice of plot.
+
+What are your suggestions for future experiments, and next steps of the research?
+
 \paragraph{Acknowledgement:}

### 2012-07-01 R/paper_jasa.R
model for experiment 3 updated

@@ -251,7 +251,7 @@ get_summary <- function(dat){
   return(data.frame(total,male,female, feedbacks))
 }
 
-get_sceering_index <- function(dat){
+get_screening_index <- function(dat){
   indx1 <- dat$id==dat$id
   indx2 <- !is.na(dat$gender)
   d34 <- ddply(dat,.(id), summarise, percent_correct=mean(response)*100)
@@ -277,7 +277,7 @@ get_sceering_index <- function(dat){
 }
 
 get_screened_summary <- function(dat){
-  indx <- get_sceering_index(dat)
+  indx <- get_screening_index(dat)
   s <- NULL
   for (i in 1:6) s <-  rbind(s,get_summary(subset(dat,indx[,i])))
   return(s)
@@ -291,31 +291,32 @@ xtable(sm)
 
 # --- fit mixed model with cleaned data
 #library(lme4)
+
 fit_model <- function(dat){
   dat$alpha <- log(.05/0.95)
-  dat$effect <- with(dat,abs(beta)/sigma*sqrt(sample_size))
+  dat$effect <- with(dat,abs(beta)*sqrt(sample_size)/sigma)
   model <- as.formula(response ~ offset(alpha) + effect -1 + (effect -1|id))
+  if (dat$experiment[1]=='turk3')model <- as.formula(response ~  effect + (effect -1|id))
   fit <- lmer(model,family="binomial",data=dat)
   return(summary(fit))
 }
 
 fit_model(dat1)
 fit_model(dat2)
+fit_model(dat3)
 
 
 pred.mixed <- function(X, subject=0,fit) {
   alpha <- log(.05/0.95)
-  eta <- alpha + X * fixef(fit) + subject*X
+  if (length(fixef(fit))>1) { # true for experiment 3
+    eta <- fixef(fit)[1] + X * fixef(fit)[2] + subject*X
+  } else eta <- alpha + X * fixef(fit) + subject*X
   g.eta <- exp(eta)/(1+exp(eta))
   return(g.eta)
 }
 
 get_predict_mixed <- function(dat, newdat, intercept=F){
-  dat$effect <-  with(dat,abs(beta)/sigma*sqrt(sample_size))
-  dat$alpha <- log(.05/0.95)
-  fit.mixed <- lmer(response ~ offset(alpha) + effect -1 + (effect-1|id),
-                    family="binomial",
-                    data=dat) 
+  fit.mixed <- fit_model(dat)
   X <- newdat$effect
   if(intercept){
[…]

### 2012-07-01 paper_jasa.tex
model for experiment 3 updated
all figures and tables updated after removing duplicate reponse on a lineup

@@
 
-\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.}
+\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
+This happens when subject accidentally use browser's ``back'' button and same data is saved twice.
+We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject.
+For this we only keep the very first response.}
 
@@
 Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
-}
+\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
 \begin{center}
@@
     \hline
-  1 & 239 & 121 & 107 & 2317 & 60 \\ 
-  2 & 351 & 185 & 164 & 3858 & 70 \\ 
-  3 & 156 & 106 &  50 & 1681 & 30 \\ 
+1 & 239 & 121 & 107 & 2249 &  60 \\ 
+  2 & 351 & 185 & 164 & 3636 &  70 \\ 
+  3 & 155 & 103 &  52 & 1511 &  29 \\  
    \hline
@@
 \begin{center}
-\begin{tabular}{cccc}
-  \hline
- &  \multicolumn{2}{c} {Fixed effect}  & Random effect\\
- \cline{2-3}
- Experiment & Estimate & Std. error & Variance\\
-  \hline
-  1 & 0.39 & 0.0093 & 0.0077 \\ 
-  2 & 1.22 &  0.0196 &  0.0455 \\ 
-  3 &  & & \\ 
+\begin{tabular}{cr@{.}lcc}
+  \hline
+ &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
+ \cline{2-4}
+ Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
+  \hline
+  1 & 0&39 & 0.0094 & 0.0080 \\ 
+  2 & 1&21 &  0.0197 &  0.0443 \\ 
+  3 & 0&31 (Intercept)  &   0.1110 & 0.1407\\ 
+     & 0&34 (Slope)    &  0.0581     & \\ 
+     &-0&72 (correlation) & & \\
    \hline
@@
 
-\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?}
+\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point.
+What can you do to make the dashed line visible?} \blue{The dashed line layer is placed on top of all other layers.
+Still there is trouble to see it.
+Do you think it will work?}
 
@@
 Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
-%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.
+%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.
 }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.

### 2012-07-01 R/paper_jasa.R

@@ -15,7 +15,8 @@ raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
 
 source("calculate_ump_power.R") # functions to compute power
 
-# cleaning the data based on criteria 6
+# cleaning the data based on criteria 6 and removing the duplicated data
+# (id,pic_id) should be unique since no person gets a lineup more than once.
 clean_data <- function(raw_dat){
   easy_dat <- subset(raw_dat, p_value < 0.0002)
   if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
@@ -31,7 +32,9 @@ clean_data <- function(raw_dat){
   excluded_lineup <- paste(raw_dat$id,"_",raw_dat$pic_id,sep="") %in% d$excluded_lineup
   indx[excluded_lineup] <- FALSE
   cleaned_dat <- subset(raw_dat,indx)
-  return(cleaned_dat)
+  indx_dup <- with(cleaned_dat, !duplicated(data.frame(id,pic_id))) #duplication index
+  # now returning cleaned data removing duplication
+  return(subset(cleaned_dat,indx_dup))
 }
 
 dat1 <- clean_data(raw.dat1)
@@ -248,7 +251,8 @@ get_summary <- function(dat){
   female <- g[2,2]
   total <- length(unique(dat$id))
   feedbacks <- nrow(dat)
-  return(data.frame(total,male,female, feedbacks))
+  lineups <- length(unique(dat$pic_id))
+  return(data.frame(total,male,female, feedbacks,lineups))
 }
 
 get_screening_index <- function(dat){
@@ -288,6 +292,13 @@ sm <- cbind(get_screened_summary(raw.dat1),
       get_screened_summary(raw.dat3))
 #library(xtable)
 xtable(sm)
+
+# summary of the cleaned data
+
+sm_clean <- rbind(get_summary(dat1),
+                  get_summary(dat2),
+                  get_summary(dat3))
+xtable(sm_clean)
 
 # --- fit mixed model with cleaned data
 #library(lme4)
@@ -580,8 +591,8 @@ get_all_pval_sam <- function(dat1,dat2,pval1,pval2){
 }
 
 
-indx1 <- get_sceering_index(raw.dat1)
-indx2 <- get_sceering_index(raw.dat2)
+indx1 <- get_screening_index(raw.dat1)
[…]

### 2012-07-02 R/paper_jasa.R
experiment 3 model results

@@ -307,7 +307,7 @@ fit_model <- function(dat){
   dat$alpha <- log(.05/0.95)
   dat$effect <- with(dat,abs(beta)*sqrt(sample_size)/sigma)
   model <- as.formula(response ~ offset(alpha) + effect -1 + (effect -1|id))
-  if (dat$experiment[1]=='turk3')model <- as.formula(response ~  effect + (effect -1|id))
+  if (dat$experiment[1]=='turk3')model <- as.formula(response ~  effect + (effect |id))
   fit <- lmer(model,family="binomial",data=dat)
   return(summary(fit))
 }
@@ -320,7 +320,8 @@ fit_model(dat3)
 pred.mixed <- function(X, subject=0,fit) {
   alpha <- log(.05/0.95)
   if (length(fixef(fit))>1) { # true for experiment 3
-    eta <- fixef(fit)[1] + X * fixef(fit)[2] + subject*X
+    subject <- matrix(unlist(sapply(as.character(subject), strsplit, "_")), ncol=2, byrow=T)
+    eta <- fixef(fit)[1] + X * fixef(fit)[2] + as.numeric(subject[,1]) + as.numeric(subject[,2])*X
   } else eta <- alpha + X * fixef(fit) + subject*X
   g.eta <- exp(eta)/(1+exp(eta))
   return(g.eta)
@@ -330,15 +331,17 @@ get_predict_mixed <- function(dat, newdat, intercept=F){
   fit.mixed <- fit_model(dat)
   X <- newdat$effect
   if(intercept){
-    subject <- ranef(fit.mixed)[[1]][,1]
+    if (dat$experiment=="turk3"){
+      subject <- apply(ranef(fit.mixed)[[1]],1,paste,collapse="_")
+      } else subject <- ranef(fit.mixed)[[1]][,1]
     d <- data.frame(expand.grid(effect=X, subject=subject))
-    pred <- pred.mixed(d$effect, subject=d$subject, fit=fit.mixed)
+    pred <- pred.mixed(X=d$effect, subject=d$subject, fit=fit.mixed)
     res <- data.frame(effect=d$effect, subject=d$subject, pred)
   } else res <- data.frame(effect=X,pred=pred.mixed(X, fit=fit.mixed))
   return(res)
 }
 effect <- seq(0.01,16, by=.2)
-get_predict_mixed(dat1, newdat=data.frame(effect))
[…]

### 2012-07-02 paper_jasa.tex
experiment 3 model results
comment on p-value difference

@@
   2 & 1&21 &  0.0197 &  0.0443 \\ 
-  3 & 0&31 (Intercept)  &   0.1110 & 0.1407\\ 
-     & 0&34 (Slope)    &  0.0581     & \\ 
-     &-0&72 (correlation) & & \\
+  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
+     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
+     &-0&78 (correlation) & & \\
    \hline
@@
 
-\green{Something is definitely not right! The code looks like it does actual data p-value - min(null pvalue). But how can you get a difference of 0.8 from this??}
+\green{Something is definitely not right!
+The code looks like it does actual data p-value - min(null $p$-value).
+But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically!
+Suppose data plot $p$-value is larger than 0.8.
+Under null hypothesis, $p$-value $\sim U(0,1)$.
+Thus min(null $p$-value) could close to zero by chance producing this large difference.
+And this has happened in both experiment 1 and 2.}
 
  \green{Is this really pct correct, or 1-pct correct?
+\blue{This is proportion correct}.
 When there is a big diff in p-value almost no-one gets it right!
-Maybe it is just a scale issue?} We see that when there is a big difference between the $p$-values observers choose the actual data plot more often.
+Maybe it is just a scale issue?} \blue{In fact this is what we expect.
+If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small.
+This should lead people to pick the wrong plot.
+Our minimum $p$-value assumption basically says this fact, right?}.
+We see that when there is a big difference between the $p$-values observers choose the actual data plot more often.
 

### 2012-07-02 R/paper_jasa.R

@@ -632,6 +632,11 @@ ggplot(pval.diff) +
   ylab("Proportion correct")
 ggsave(file="../images/pval_difference.pdf", height=4, width=8)
 
+# investigating large differences between p-values and minimum p-value of null data
+
+dd <- subset(pval.dat, pic_name %in% subset(pval.diff,pval_diff >.8)[,1])
+
+
 # =================== Turk1 data analysis  ================================
 
 # ------function to calculate UMP power -----------------------

### 2012-07-02 paper_jasa.tex
reasoning plot added

@@
 
-\green{I find this plot really confusing! Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?}
+\green{I find this plot really confusing!
+Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in figure \ref{fig:choice_reason} ok?
+Should we remove the left plot?}
 
@@
        \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
+       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
        \caption{Reasons for choosing the plot in a lineup for experiment 3.

### 2012-07-02 R/paper_jasa.R

@@ -770,6 +770,15 @@ qplot(prop_correct,choice_reason, geom="point", data=reason.dat3, size=I(3)) +
 
 ggsave("../images/choice_reason.pdf", height = 4, width = 6)
 
+reason.clustring <- ddply(dat3,.(response), summarize,
+                          cnt = sum(choice_reason==3))
+reason.clustring$response <- factor(as.character(reason.clustring$response), labels=c("Data plot not picked", "Data plot picked"))
+qplot(response,cnt, data=reason.clustring, geom="bar", alpha=I(.7)) +
+  xlab("Responses of subjects choosing Clustering Visible") +
+  ylab("Number of subjects") + opts(legend.position="none")
+
+ggsave("../images/choice_reason_count.pdf", height = 4, width = 4.5)
+
 
 # ======================= p_value vs proportion correct =====================
 

### 2012-07-02 R/paper_jasa.R
clustering visible plot added

@@ -774,9 +774,8 @@ reason.clustring <- ddply(dat3,.(response), summarize,
                           cnt = sum(choice_reason==3))
 reason.clustring$response <- factor(as.character(reason.clustring$response), labels=c("Data plot not picked", "Data plot picked"))
 qplot(response,cnt, data=reason.clustring, geom="bar", alpha=I(.7)) +
-  xlab("Responses of subjects choosing Clustering Visible") +
+  xlab("Responses of those choosing Clustering visible") +
   ylab("Number of subjects") + opts(legend.position="none")
-
 ggsave("../images/choice_reason_count.pdf", height = 4, width = 4.5)
 
 

### 2012-07-02 R/paper_jasa.R
figure 2 caption changed

@@ -896,17 +896,21 @@ get_corr(dat3)
 
 # -------- Hieke's code for power vs m and p-value------------------
 
-powerdf <- data.frame(expand.grid(m = 5:50,
-pb=c(seq(0.05,0.5, by=0.05), seq(0.25, 0.5, by=0.05))))
+powerdf <- data.frame(expand.grid(pd = seq(0.0001, 0.20, by=0.001), m = seq(10, 30, by=5)))
+powerdf$power <- with(powerdf, pbeta(pd, 1, m-1, lower.tail=FALSE))
+powerdf$size <- with(powerdf, I(.75+(m==20)))
+powerdf$labels <- paste("m =",powerdf$m)
+powerdf$hjust <- c(-0.15,1.15)[(powerdf$m > 20) + 1]
+powerdf$power <- with(powerdf, pmax(power, 1/m))
 
-powerdf$power <- with(powerdf, (1-pb)^(m-1))
+qplot(pd, power, data=powerdf, geom="line", group=m, size=size) + 
+  ylab("Probability that data plot has lowest p-value") + 
+  xlab(expression(p[D])) + scale_size_identity() +
+  geom_text(aes(x=pd, y=power, label=labels, hjust=hjust), size=3, 
+            data=subset(powerdf, abs(4*pd - power) < 0.005 )) + 
+  geom_hline(y=1/seq(15,30, by=5), linetype=2, alpha=0.3) 
 
-p <- qplot(m, power, geom="line", group=pb, data=powerdf, colour=pb) + 
-     geom_line(aes(m, 1/m), colour="grey30", lty=2) + 
-     geom_text(aes(label=sprintf("%.2f",pb)), data=subset(powerdf, m==5), hjust=1.1) + 
-     xlim(c(0,50)) + scale_colour_gradient(name=expression(p[B]))
-p     
-ggsave(p,file="../images/p_val_power_m.pdf", height = 4.25, width = 6)
+ggsave(file="../images/powerplot.pdf", height = 5, width = 5)
 
 
 # empirical power by effect with bootstrap confidence band

### 2012-07-03 paper_jasa.tex
clustering visible plot added
figure 2 caption changed
added references

@@
 Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years.
-Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{heer:2010} with findings supporting the original results.
+Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results.
 This body of work provides a framework for evaluating new statistical graphics.
@@
    \centering
-   \includegraphics[width=3.8in]{images/powerplot.pdf} 
+   \includegraphics[width=3in]{images/powerplot.pdf} 
    \caption{Probability that the data plot has the smallest probability in a lineup of size $m$.
@@
 
-By matching the expected values, we therefore get an expression for the $p$-value in the data under $H_1$  as:
+By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
+%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
+as:
 \begin{equation}\label{eqn:power_estimate}
@@
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
+\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
+We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact.
+Both of these topics are further highlighted in the application section.}
 
@@
 It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
-\green{What about other??}
+\green{What about other??} \blue{Not many people chose other as we see in figure \ref{fig:choice_reason}.}
 

### 2012-07-03 paper_jasa.tex
small editorial changes

@@
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if out of $K$ observers $x_{\alpha}$ of them correctly identify the actual data plot in the lineup, and
-\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
+\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
+\item \textbf{Fail to reject} $H_0$  otherwise. 
 \end{itemize}
@@
 
-Associated with any test is the risk of making a Type I error, or Type II error, which for visual inference are defined as follows:
+\noindent
+Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:
 
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
-% identifying the actual data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
+The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
 \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
@@
 
-\green{Oh, we are using observed and observer, which might be confusing. May need to change observed data plot to something else. Will think about this. Will use ``actual'' for now.}
+%\green{Oh, we are using observed and observer, which might be confusing.
+May need to change observed data plot to something else.
+Will think about this.
+Will use ``actual'' for now.} \hh{I think that change is done}
 
@@
 For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.
-It can be set to be smaller, even arbitrarily small, by increasing $K$.
+It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers.
 Type II error is harder to calculate, as is usually the case.
@@
 Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup.
-If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$.
-We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
+If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
 
@@
 This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details.
-estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
+Estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
 

### 2012-07-03 paper_jasa.tex
cut out some of the plot signal strength discussion

@@
 
-This proves the statement above. We further see:
-\[
-E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
-\]
+This proves the statement above. 
+%We further see:
+%\[
+%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
+%\]
 This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and,
@@
 
-Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
-Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
+%Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
+%Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
 Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
@@
 We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
-
-For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
-we get
-\[
-N_m\sim \left \{ 
-\begin{array}{ll}
-B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
-B_{K, 1/m} & \text { under } H_0
-\end{array}
-\right .
-\]
-
-By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
-%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
-as:
-\begin{equation}\label{eqn:power_estimate}
-%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
-\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
-\end{equation}
-
-Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
+%
+%For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
+%we get
+%\[
+%N_m\sim \left \{ 
+%\begin{array}{ll}
+%B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
+%B_{K, 1/m} & \text { under } H_0
+%\end{array}
+%\right .
+%\]
+%
+%By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
+%%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
+%as:
+%\begin{equation}\label{eqn:power_estimate}
+%%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
+%\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
+%\end{equation}
+%
+%Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
 While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
-\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
+%\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
 We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact.
-Both of these topics are further highlighted in the application section.}
+%Both of these topics are further highlighted in the application section.}
 
@@
 
-\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?} \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
-
-\green{Explain why the term ``plot signal'' is introduced}
+\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
+% \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
+%
+%\green{Explain why the term ``plot signal'' is introduced}
 
@@
 
-Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
+%Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
 This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense.
@@
 
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
+%\begin{figure*}[hbtp]
+%   \centering
+%       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
+%       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
 The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
 p-values are shown on square root scale scale.}
-    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-       \label{fig:pval_plot_signal}
-\end{figure*}
+%    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
+%       \label{fig:pval_plot_signal}
+%\end{figure*}
 

### 2012-07-04 paper_jasa.tex
removed signal strength and added visual p-value disccusion

@@
 
-With multiple observers it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
+From the experimental data the visual $p$-values are computed using Definition \ref{dfn:pvalue}.
+Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$-value.
+The pattern of visual $p$-value is interesting.
+If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) giving a clear indication to reject $H_0$ or not.
+This is why we do not see lot of visual $p$-values in between 0.05 and 0.8 especially for experiment 2.
+This guide the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$.
+For visual test this is not common.
+
+\begin{figure*}[hbtp]
+   \centering
+       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
+       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . p-values are shown on square root scale scale. }
+    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
+       \label{fig:pval_definition}
+\end{figure*}
+
+We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}.
+The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
+For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2.
+This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.
+For experiment 3, we see all the visual $p$ values are almost zero no matter what the conventional $p$-values are.
+This is expected since conventional test fails to reject $H_0$ where we should reject and visual test does the job nicely.
+
+%With multiple observers it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
 The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
@@
 
-Similary, the strength of the signal, is calculated using the proportion of observers picking actual data plot.
+%Similary, the strength of the signal, is calculated using the proportion of observers picking actual data plot.
 It is closely related to the power of the test.
@@
 This gives the signal strength estimate as
-\[
-\hat p_D= (1- {u}/{K})/({m-1})
-\]
-where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
+%\[
+%\hat p_D= (1- {u}/{K})/({m-1})
+%\]
+%where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
 Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
@@
 
-\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
+%\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
 % \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
@@
 %and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . p-values are shown on square root scale scale. }
-    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-       \label{fig:pval_definition}
-\end{figure*}
 

### 2012-07-04 paper_jasa.tex
supplimentary plot updated

@@
 For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
-Power for contaminated data is shown by cross mark.
-}
+Power for contaminated data is shown by cross mark.}
        \label{fig:power_loess_effect}
@@
 A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
-If there is a big negative difference we'd expect to see We would expect to see that the observers would more often make the correct choice of plot.
-This can be seen.
-When there is a large positive difference If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
+If there is a big negative difference we'd expect to see that the observers would more often make the correct choice of plot.
+When there is a large positive difference, observers fail to pick the correct plot.
+If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
 
@@
 Under null hypothesis, $p$-value $\sim U(0,1)$.
-Thus min(null $p$-value) could close to zero by chance producing this large difference.
+Thus min(null $p$-value) could be close to zero by chance producing this large difference.
 And this has happened in both experiment 1 and 2.}

### 2012-07-04 paper_jasa.tex
comments on the lemma

@@
 %\]
-This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and,
+This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and \blue{(How can we say that from this?
+The term  $E\left[ (1 - p_D)^{m-1}\right]$ is definitely not a function of $p_D$ since $p_D$ should be eliminated after the expectation.
+Instead, its a function of $m$ only !
+Is this interesting that under the minimum $p$-value assumption the power is a function of lineup size $m$ only?)},
 conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 

### 2012-07-04 R/paper_jasa.R
uncontaminated p-value added

@@ -865,7 +865,17 @@ qplot(p_value,visual_pval, data=subset(pdat,nchar(experiment) < 13)) +
   ylab(expression(paste("Visual p-value on square root scale"))) +
   facet_grid(.~experiment)+
   scale_x_sqrt() +scale_y_sqrt()
-ggsave(file="../images/p_val_definition.pdf", height = 4, width = 10)
+
+pdat$visual_pval <- apply(pdat[,2:3],1,function(x)return(get_pval(m=20,K=x[1],x=x[2])))
+qplot(p_value,visual_pval, data=pdat) +
+  geom_abline(colour="grey") +
+  stat_smooth(method="loess", se=F, span=.45, degree=1)+
+  xlab(expression(paste("Conventional test p-value (",p[D],") on square root scale"))) +
+  ylab(expression(paste("Visual p-value on square root scale"))) +
+  facet_grid(.~experiment)+
+  scale_x_sqrt() +scale_y_sqrt()
+
+ggsave(file="../images/p_val_definition.pdf", height = 4, width = 11.5)
 
 
 qplot(strength,visual_pval, data=pdat) +

### 2012-07-04 paper_jasa.tex
uncontaminated p-value added
some typo fixed

@@
 
-%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the figure on $1/m$ versus power.
+%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the Figure on $1/m$ versus power.
 These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
@@
 % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
-%also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.
+%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.
 
@@
 The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
-Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in table \ref{tbl:experiment_params}.
+Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
 Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
@@
 %For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
-We collect data such that each of the parameter combinations in the table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
+Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in Table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
+We collect data such that each of the parameter combinations in the Table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
 
@@
 %{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
-For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}.
+For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in Table \ref{tbl:experiment_params}.
 For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
@@
 %       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
+%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in Table \ref{tbl:experiment_params}.}
 %       \label{fig:power_diff_exp}
@@
 After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in figure \ref{fig:pval_power}.
+The theoretical justification of this is shown in Figure \ref{fig:pval_power}.
 }
@@
    \centering
-       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . p-values are shown on square root scale scale. }
+       \scalebox{0.55}{\includegraphics{p_val_definition.pdf}}
+       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition .
+Values are shown on square root scale.
+For experiment 3, $p$-values of uncontaminated part of the data are also shown.}
     %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
@@
 
-%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2.
-The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
-We see the same pattern in figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
+%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in Table \ref{tbl:pval_assumption} for both experiment 1 and 2.
+The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
+We see the same pattern in Figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
 
@@
 It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
-\green{What about other??} \blue{Not many people chose other as we see in figure \ref{fig:choice_reason}.}
+\green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}
 
@@
 \green{I find this plot really confusing!
-Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in figure \ref{fig:choice_reason} ok?
+Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in Figure \ref{fig:choice_reason} ok?
 Should we remove the left plot?}

### 2012-07-04 paper_jasa.tex
comments updated

@@
 The term  $E\left[ (1 - p_D)^{m-1}\right]$ is definitely not a function of $p_D$ since $p_D$ should be eliminated after the expectation.
-Instead, its a function of $m$ only !
-Is this interesting that under the minimum $p$-value assumption the power is a function of lineup size $m$ only?)},
+Instead, its a function of $m$ and the distribution of $p_D$ !
+If the density of $p$-value is very right skewed, the expectation term would be large.
+The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in.
+Should we add that in Equation \ref{dist_p}?
+)},
 conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 

### 2012-07-05 paper_jasa.tex
general cleanup

@@
 Some subjects performed substantially better than the conventional test, and on average the visual test was better.
-
-\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point.
-What can you do to make the dashed line visible?} \blue{The dashed line layer is placed on top of all other layers.
-Still there is trouble to see it.
-Do you think it will work?}
 
@@
 Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
-This is investigated in the paper \cite{waldo}.
+This is investigated in \citet{niladri:2012}.
 Here numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
@@
 \noindent It was possible to choose more than one reason.
-The question begged for ``visible trend'' (2) to be the reasons that observers gave for correctly choosing the actual data plot.
+The question begged for ``visible trend'' (2) to be the reason that observers gave for correctly choosing the actual data plot.
 Either reasons 1 or 3 would be indicative of Type III error.

### 2012-07-05 R/paper_jasa.R
made different plots for the reasons of choice, think these are easier to understand

@@ -770,6 +770,39 @@ qplot(prop_correct,choice_reason, geom="point", data=reason.dat3, size=I(3)) +
 
 ggsave("../images/choice_reason.pdf", height = 4, width = 6)
 
+# Di's changes
+reason.dat3.sub<-subset(reason.dat3, counts>10)
+reason.dat3$choice_reason <- factor(reason.dat3$choice_reason)
+ord<-order(reason.dat3.sub$prop_correct)
+reason.dat3$choice_reason <- factor(reason.dat3$choice_reason, levels=levels(reason.dat3$choice_reason)[ord])
+qplot(prop_correct, choice_reason, geom="point", data=reason.dat3.sub, size=counts) +
+  ylab("Reasons for choice") + xlab("Proportion correct") + xlim(c(0,1)) + scale_size("Number", limits=c(0, 500))
+ggsave("../images/choice_reason2.pdf", height = 4, width = 6)
+
+# Now focus on "marginals"
+dat3$choice_reason1 <- grepl("1", as.character(dat3$choice_reason))
+dat3$choice_reason2 <- grepl("2", as.character(dat3$choice_reason))
+dat3$choice_reason3 <- grepl("3", as.character(dat3$choice_reason))
+dat3$choice_reason4 <- grepl("4", as.character(dat3$choice_reason))
+reason.any1.dat3 <- ddply(dat3,.(choice_reason1), summarise,
+                     prop_correct = mean(response),
+                     counts = length(response))
+reason.any2.dat3 <- ddply(dat3,.(choice_reason2), summarise,
+                     prop_correct = mean(response),
+                     counts = length(response))
+reason.any3.dat3 <- ddply(dat3,.(choice_reason3), summarise,
+                     prop_correct = mean(response),
+                     counts = length(response))
+reason.any4.dat3 <- ddply(dat3,.(choice_reason4), summarise,
+                     prop_correct = mean(response),
+                     counts = length(response))
+reason.any.dat3 <- data.frame(choice_reason=c("Most different plot", "Visible trend","Clustering visible","Other"), prop_correct=c(reason.any1.dat3$prop_correct[2], reason.any2.dat3$prop_correct[2], reason.any3.dat3$prop_correct[2], reason.any4.dat3$prop_correct[2]), counts=c(reason.any1.dat3$counts[2], reason.any2.dat3$counts[2], reason.any3.dat3$counts[2], reason.any4.dat3$counts[2]))
[…]

### 2012-07-05 R/paper_jasa.R
p-value - min(p-value) plot with lineups of beta=0 removed

@@ -609,6 +609,7 @@ print(xtable(pval_sm), include.rownames=FALSE)
 # Difference between data plot p-value and minimum null plot p-value 
 
 get_pval_success <- function (dat){
+  dat <- subset(dat, abs(beta)>0)
   res <- ddply(dat, .(pic_name), summarize,      
         prop_correct=mean(response),
         p_value=p_value[1],

### 2012-07-05 paper_jasa.tex
made different plots for the reasons of choice, think these are easier to understand
working on the language of section 6.8

@@
 
-A little known error amongst statisticians is what was coined as Type III in \citet{mosteller:48}.
+A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
 This occurs when the null hypothesis is rejected for thw rong reason.
@@
 Observers were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see the association.
+But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
 For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
@@
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible choices
+Observers were provided with four possible reasons to use for their choice:
 
@@
 
-\noindent It was possible to choose more than one reason.
-The question begged for ``visible trend'' (2) to be the reason that observers gave for correctly choosing the actual data plot.
-Either reasons 1 or 3 would be indicative of Type III error.
+\noindent with the possibility to use more than one.
+The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for thier choice.
+Reasons 1 or 3 would be indicative of Type III error.
+Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
+If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason.
+
 It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
@@
 
-\green{I find this plot really confusing!
-Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in Figure \ref{fig:choice_reason} ok?
-Should we remove the left plot?}
-
 \begin{figure}[hbtp]
    \centering
-       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
-       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
+%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
+%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
+       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
+       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
        \caption{Reasons for choosing the plot in a lineup for experiment 3.

### 2012-07-05 R/paper_jasa.R
non-contaminated p-value plot removed

@@ -900,7 +900,8 @@ qplot(p_value,visual_pval, data=subset(pdat,nchar(experiment) < 13)) +
   facet_grid(.~experiment)+
   scale_x_sqrt() +scale_y_sqrt()
 
-pdat$visual_pval <- apply(pdat[,2:3],1,function(x)return(get_pval(m=20,K=x[1],x=x[2])))
+ggsave(file="../images/p_val_definition.pdf", height = 4, width = 10)
+
 qplot(p_value,visual_pval, data=pdat) +
   geom_abline(colour="grey") +
   stat_smooth(method="loess", se=F, span=.45, degree=1)+
@@ -909,7 +910,7 @@ qplot(p_value,visual_pval, data=pdat) +
   facet_grid(.~experiment)+
   scale_x_sqrt() +scale_y_sqrt()
 
-ggsave(file="../images/p_val_definition.pdf", height = 4, width = 11.5)
+
 
 
 qplot(strength,visual_pval, data=pdat) +

### 2012-07-05 paper_jasa.tex
non-contaminated p-value plot removed
finished with Type III error section

@@
 \noindent with the possibility to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for thier choice.
+The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
 Reasons 1 or 3 would be indicative of Type III error.
 Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
-If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason.
-
-It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
+If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
+This is not what we see.
+At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
+The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
+The size of the point corresponds to the number of subjects putting this as the reaosn.
+Subjects that chose all three reasons almost always chose the actual data plot.
+This was followed by using 1 and 3, and then 1 and 4.
+The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
+At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
+``Visible trend'' comes in third.
+This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
+
+%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
 Figure \ref{fig:choice_reason} examines this.
@@
 %       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
+       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
        \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
        \caption{Reasons for choosing the plot in a lineup for experiment 3.
@@
 \end{figure}
+
+For visual inference, making a Type III, is not actually a problem.
+It is only a possibility in this ecperiment because we are working with known structure.
+In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
+This does show the importance of the qualitative answers that observers give for their choices.
 

### 2012-07-05 paper_jasa.tex
caption for figure 13 updated

@@
    \centering
-       \scalebox{0.55}{\includegraphics{p_val_definition.pdf}}
+       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
        \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition .
@@
        \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Reasons for choosing the plot in a lineup for experiment 3.
-Reason 12 means both reasons 1 and 2 are chosen.
-Number of subjects choosing that reason are shown on the right hand side.
-Notice that 345 people choose only reason (3) "Clustering visible", out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
-Those who chose reason 123 got the actual plot about 97\% times.}
+       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error.
+The vertical axis shows the reasons.
+At left, all subjects choices are shown, and reason 123 means all three reasons are used.
+At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
+The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot.
+Size of the point corresponds to the number of subjects using that reason.}
        \label{fig:choice_reason}

### 2012-07-05 paper_jasa.tex
listed lineups chosen for follow-up eye-tracker experiment

@@
 \cite{*** yifan's CC needs to be made into a tech report, and referenced here.}
+Lineups chosen for the follow up eye-tracker experiment were (experiment 1) (1) $n=100, \beta=3, \sigma=5$, rep=1 (2) $n=100, \beta=5, \sigma=5$, rep=1 (3) $n=100, \beta=10, \sigma=12$, rep=1 and (4) $n=300, \beta=5, \sigma=5$, rep=1, and  (experiment 2) (5) $n=100, \beta=3.5, \sigma=12$, rep=1 (6) $n=100, \beta=4.5, \sigma=12$, rep=2 (7) $n=300, \beta=1, \sigma=5$, rep=1 and (8) $n=300, \beta=1, \sigma=5$, rep=2.
 

### 2012-07-05 paper_jasa.tex
typo fixed

--- 2012-07-05 paper_jasa.tex
@@
 After many small pilot studies with local personnel, it was clear that some care was needed to set up the human subjects experiments.
-It was best for a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
-The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the observers, for each of the three experiments.
+It was best for a observer or a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
+The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects, for each of the three experiments.
 
@@
 
-The experiment is designed to study the ability of human observers to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
+The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
 Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
@@
 
-%With multiple observers it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
+%With multiple subjects it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
 The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
-Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more observers select that plot.
+Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more subjects select that plot.
 Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.
 
-%Similary, the strength of the signal, is calculated using the proportion of observers picking actual data plot.
+%Similary, the strength of the signal, is calculated using the proportion of subjects picking actual data plot.
 It is closely related to the power of the test.
@@
 The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
-As the conventional $p$-value increases, the estimated signal strength increases -- observers find it harder to detect the actual data plot at larger $p$-values.
-The rise is very sharp for experiment 1, which indicates that observers have a harder time detecting the actual data plot for even small $p$-values.
+As the conventional $p$-value increases, the estimated signal strength increases -- subjects find it harder to detect the actual data plot at larger $p$-values.
+The rise is very sharp for experiment 1, which indicates that subjects have a harder time detecting the actual data plot for even small $p$-values.
 
@@
 This can be examined for the data collected, to see if this assumption is, indeed, reasonable.
-%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value.
+%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value.
 In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
@@
 But even when they didn't pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
-So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
+So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
 
@@
 Same experiment, when $\beta=0, n=100, sd=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the subjects were looking at in making their choices.
 \cite{*** yifan's CC needs to be made into a tech report, and referenced here.}
@@
 Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
-Rather it is assumed that the observer has the ability to select the plot with the most extreme structure.
+Rather it is assumed that the subject has the ability to select the plot with the most extreme structure.
 A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, sd=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
@@
 But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
-How this small set of null plots influences the observers choice is important for understanding the reliability of visual inference.
+How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
 If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
@@
 
-In the previous section we noticed that observers tend to pick the plot with the lowest $p$-value.
-If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an observer will pick this null plot instead of the actual data plot.
+In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
+If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
 Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
-The plots display the proportion of observers who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
+The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
 A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
-If there is a big negative difference we'd expect to see that the observers would more often make the correct choice of plot.
-When there is a large positive difference, observers fail to pick the correct plot.
+If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot.
+When there is a large positive difference, subjects fail to pick the correct plot.
 If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
@@
 Our minimum $p$-value assumption basically says this fact, right?}.
-We see that when there is a big difference between the $p$-values observers choose the actual data plot more often.
+We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.
 
@@
 
-A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
+A little known error amongst statisticians is what was coined as Type III in \citet{mosteller:48}.
 This occurs when the null hypothesis is rejected for thw rong reason.
 Experiment 3 is prone to this type of error.
-Observers were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
-For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
+Subjects were asked to identify the plot that had the largest slope, either positive or negative.
+But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see the association.
+For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the subject to detect the actual data plot by looking for the cluster instead of the slope.
 This would be considered to be a Type III error because it is not what was requested, but it still gives the correct answer.
-This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
-Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
+This is not a problem with the lineup protocol, because we want subjects to react to anything that sets the the actual data plot apart.
+Here, in the controlled but simple setting, it is of interest to see what subjects cued to in the actual data plot.
 
 Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice:
+Teasing this out of the results is possible by looking at the reasons subjects gave for their choices.
+Subjects were provided with four possible choices
 
@@
 
-\noindent with the possibility to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
-Reasons 1 or 3 would be indicative of Type III error.
-Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
-If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
-This is not what we see.
-At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
-The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
-The size of the point corresponds to the number of subjects putting this as the reaosn.
-Subjects that chose all three reasons almost always chose the actual data plot.
-This was followed by using 1 and 3, and then 1 and 4.
-The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
-At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
-``Visible trend'' comes in third.
-This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
-
-%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
+\noindent It was possible to choose more than one reason.
+The question begged for ``visible trend'' (2) to be the reason that subjects gave for correctly choosing the actual data plot.
+Either reasons 1 or 3 would be indicative of Type III error.
+It is interesting to calculate the proportion of subjects who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
 Figure \ref{fig:choice_reason} examines this.
-For observers who chose the actual data plot, the proportions for the three reasons are shown.
+For subjects who chose the actual data plot, the proportions for the three reasons are shown.
 It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
@@
 %\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
+
+\green{I find this plot really confusing!
+Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in Figure \ref{fig:choice_reason} ok?
+Should we remove the left plot?}
 
@@
 
-For visual inference, making a Type III, is not actually a problem.
-It is only a possibility in this ecperiment because we are working with known structure.
-In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
-This does show the importance of the qualitative answers that observers give for their choices.
-
 \section{Conclusions}
@@
 In practice there is no correct answer.
-In these experiments we controlled the effets that observers saw, so there was a correct choice in each case.
+In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
 In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
-And observers would provide open-ended answers for their choice of plot.
+And subjects would provide open-ended answers for their choice of plot.
 

### 2012-07-05 paper_jasa.tex

@@
 Same experiment, when $\beta=0, n=100, sd=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the subjects were looking at in making their choices.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices.
 \cite{*** yifan's CC needs to be made into a tech report, and referenced here.}
+Lineups chosen for the follow up eye-tracker experiment were (experiment 1) (1) $n=100, \beta=3, \sigma=5$, rep=1 (2) $n=100, \beta=5, \sigma=5$, rep=1 (3) $n=100, \beta=10, \sigma=12$, rep=1 and (4) $n=300, \beta=5, \sigma=5$, rep=1, and  (experiment 2) (5) $n=100, \beta=3.5, \sigma=12$, rep=1 (6) $n=100, \beta=4.5, \sigma=12$, rep=2 (7) $n=300, \beta=1, \sigma=5$, rep=1 and (8) $n=300, \beta=1, \sigma=5$, rep=2.
 
@@
 
-A little known error amongst statisticians is what was coined as Type III in \citet{mosteller:48}.
+A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
 This occurs when the null hypothesis is rejected for thw rong reason.
 Experiment 3 is prone to this type of error.
-Subjects were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see the association.
-For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the subject to detect the actual data plot by looking for the cluster instead of the slope.
+Observers were asked to identify the plot that had the largest slope, either positive or negative.
+But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
+For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
 This would be considered to be a Type III error because it is not what was requested, but it still gives the correct answer.
-This is not a problem with the lineup protocol, because we want subjects to react to anything that sets the the actual data plot apart.
-Here, in the controlled but simple setting, it is of interest to see what subjects cued to in the actual data plot.
+This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
+Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
 
 Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons subjects gave for their choices.
-Subjects were provided with four possible choices
+Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
+Observers were provided with four possible reasons to use for their choice:
 
@@
 
-\noindent It was possible to choose more than one reason.
-The question begged for ``visible trend'' (2) to be the reason that subjects gave for correctly choosing the actual data plot.
-Either reasons 1 or 3 would be indicative of Type III error.
-It is interesting to calculate the proportion of subjects who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
+\noindent with the possibility to use more than one.
+The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
+Reasons 1 or 3 would be indicative of Type III error.
+Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
+If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
+This is not what we see.
+At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
+The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
+The size of the point corresponds to the number of subjects putting this as the reason.
+Subjects that chose all three reasons almost always chose the actual data plot.
+This was followed by using 1 and 3, and then 1 and 4.
+The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
+At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
+``Visible trend'' comes in third.
+This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
+
+%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
 Figure \ref{fig:choice_reason} examines this.
-For subjects who chose the actual data plot, the proportions for the three reasons are shown.
+For observers who chose the actual data plot, the proportions for the three reasons are shown.
 It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
@@
 %\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
-
-\green{I find this plot really confusing!
-Can we just look at whether they selected choice ``clustering visible'' or not, if they were correct?} \blue{Is the right hand side plot in Figure \ref{fig:choice_reason} ok?
-Should we remove the left plot?}
 
@@
        \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Reasons for choosing the plot in a lineup for experiment 3.
-Reason 12 means both reasons 1 and 2 are chosen.
-Number of subjects choosing that reason are shown on the right hand side.
-Notice that 345 people choose only reason (3) "Clustering visible", out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error).
-Those who chose reason 123 got the actual plot about 97\% times.}
+       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error.
+The vertical axis shows the reasons.
+At left, all subjects choices are shown, and reason 123 means all three reasons are used.
+At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
+The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot.
+Size of the point corresponds to the number of subjects using that reason.}
        \label{fig:choice_reason}
 \end{figure}
+
+For visual inference, making a Type III, is not actually a problem.
+It is only a possibility in this ecperiment because we are working with known structure.
+In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
+This does show the importance of the qualitative answers that observers give for their choices.
 

### 2012-07-05 paper_jasa.tex
outlining pieces that need to be updated

@@
 The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}.
-For computation we use R \citep{R} package lme4.
+For computation we use R \citep{R} package lme4 \citep{who are the authors of this package?}.
 
 Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
+\green{What do we learn about the models from these parameter estimates?
+Explain this here too.}
 
@@
 There are some noticeable exceptions to this rule.
-In experiment 1, when $\beta=5, n=300, sd=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.
-Same experiment, when $\beta=0, n=100, sd=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices.
-\cite{*** yifan's CC needs to be made into a tech report, and referenced here.}
-Lineups chosen for the follow up eye-tracker experiment were (experiment 1) (1) $n=100, \beta=3, \sigma=5$, rep=1 (2) $n=100, \beta=5, \sigma=5$, rep=1 (3) $n=100, \beta=10, \sigma=12$, rep=1 and (4) $n=300, \beta=5, \sigma=5$, rep=1, and  (experiment 2) (5) $n=100, \beta=3.5, \sigma=12$, rep=1 (6) $n=100, \beta=4.5, \sigma=12$, rep=2 (7) $n=300, \beta=1, \sigma=5$, rep=1 and (8) $n=300, \beta=1, \sigma=5$, rep=2.
+In experiment 1, when $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.
+Same experiment, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices \citep{zhao2012}.
+(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
 
@@
 Rather it is assumed that the subject has the ability to select the plot with the most extreme structure.
-A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, sd=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
+A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, \sigma=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
 Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.
@@
 
+\green{
+\begin{itemize} \itemsep 0in
+\item Summarize the key pieces of the paper again, especially the major findings
+\item Explain caveats, nuances, advantages with the work
+\item Describe the next steps of the work, and possible ideas for future extensions
+\end{itemize}
+}
+
 The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
@@
 
-This work was funded by National Science Foundation grant DMS 1007697.
+This work was funded in part by National Science Foundation grant DMS 1007697.
 

### 2012-07-05 paper_jasa.tex
lme4 reference added
concluding remarks.

@@
 The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}.
-For computation we use R \citep{R} package lme4 \citep{who are the authors of this package?}.
+For computation we use R \citep{R} package lme4 \citep{lme4:2011}.
 
@@
 For visual inference, making a Type III, is not actually a problem.
-It is only a possibility in this ecperiment because we are working with known structure.
+It is only a possibility in this experiment because we are working with known structure.
 In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
@@
 \item Summarize the key pieces of the paper again, especially the major findings
+
+\blue{It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
+But this paper demonstrates that statistical graphics could also be used in statistical inference and establishes a formal framework to accomplish that.
+The simulation study reveals that even in a situation where conventional test has its all the convenient assumptions met, visual inference is still comparable to conventional test.
+Also, for some trained observers/subjects visual inference may produce even better power.
+On the other hand when the underlying assumptions of a conventional test do not satisfy and the test fails to produce enough power, visual inference performs extremely well.
+
+We obtain the estimation of the power of visual test in different ways and present the results in this paper.
+The experimental data suggests that observers usually pick the plot which has the minimum $p$-value or maximum signal strength in the plot.
+If we assume it to be true in general,  the power of visual test is then obtained theoretically.
+Interestingly the theoretical power of visual test is also very close to that of conventional test even though the situation of the experiment is worse for visual test and best for conventional test.}
+
 \item Explain caveats, nuances, advantages with the work
+
+\blue{One of the most critical steps in visual test would be select an appropriate test statistic for the hypothesis of interest.
+A good test statistic should show a very specific pattern if the null hypothesis is not true.
+A small difference in color template or orientation of the plot could make a difference.
+This is the part of our ongoing and future research.
+
+Even though a single observer can make a decision we recommend having multiple observer to evaluate a lineup to obtain a definitive $p$-value to make a decision confidently.
+
+The experimental subjects that we recruited came from a various level of education and not necessarily well aware of statistical plots.
+But still we achieve the power of visual test comparable to the conventional test.
+This suggests that visual statistical technique could be a smart tool for high school students to get to know the first course of statistical inference.
+The major advantage of visual statical inference is its intuitive concept of testing technique which is easy to be grasped and fun to learn.
+}
+
 \item Describe the next steps of the work, and possible ideas for future extensions
+
+\blue {When there is no conventional test available, we can conveniently use visual testing procedure due to its flexibility of having no assumption restrictions.
+Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
+The direction of future research would be to study that and apply the methods in those areas.}
 \end{itemize}

### 2012-07-05 paper_jasa.tex
comments on turk workers' skill test

@@
 Same experiment, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices \citep{zhao2012}.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices \citep{zhao:2012}.
 (Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
@@
 Along the lines of the Army tests that Susan brought to the working group.
+\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
+Thus anyone from any background could participate the experiment.
+It would be interesting to see how well the trained subjects performs.
+Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
+One of our future work would be to examine this in details.
+For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
+After being trained they may have to pass the qualification test before participating the experiment.
+Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
 

### 2012-07-05 paper_jasa.tex
conventional inference is chosen and all other terms are replaced

@@
 This changed, when \citet{buja:2009} conceptually introduced two protocols for formal tests of visual findings.
-In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against classical statistical testing  of the significance of regression model parameters.
+In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against conventional statistical testing  of the significance of regression model parameters.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
@@
 \citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
+This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
 One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
@@
 Which plot do you think is the most different from the others?
-(The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
+(The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors conventional inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
 The null hypothesis typically derives from the task at hand, or the type of plot being made.
@@
 The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
-The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
+The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
 Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
@@
 %\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis.
-
-%In this paper we compare the lineup protocol to classical statistical inference, specifically applied to the regression framework.
+This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
+
+%In this paper we compare the lineup protocol to conventional statistical inference, specifically applied to the regression framework.
 We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
@@
 
-Table \ref{tbl:compare} illustrates the lineup protocol in relation to traditional hypothesis testing.
+Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing.
 Both start from the same place, the same null hypothesis.
@@
 In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
-In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
+In conventional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
 If it is extreme on this scale then the null hypothesis is rejected.
@@
 \hline
-  & Traditional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
+  & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
 \hline
@@
 
-%This section outlines the concepts of visual inference in comparison to the procedures of classical statistical inference.
+%This section outlines the concepts of visual inference in comparison to the procedures of conventional statistical inference.
 Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
@@
 %Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
-For visual inference, unlike in the classical hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
+For visual inference, unlike in the conventional hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
 When the alternative hypothesis is true, it is expected that the plot of the actual data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
@@
 
-Power is an important consideration in deciding which test to use for solving a problem. %In classical testing, new test procedures are examined for their power under a variety of scenarios. 
+Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
 For visual inference, this corresponds to the choice of plot to use.
 Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
-This is akin to different choices of statistics in classical hypothesis testing, for example, mean, median, or trimmed mean.
+This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean.
 One is typically better than another.
@@
 The following properties are only derived for the situation of a  fully parameterized  simulation study.
-This allows us to directly compare lineup tests against their classical counterparts and also identify  properties relevant for a quality assessment of  lineups, which are typically used  in non-parametric situations.
+This allows us to directly compare lineup tests against their conventional counterparts and also identify  properties relevant for a quality assessment of  lineups, which are typically used  in non-parametric situations.
 
@@
 
-%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the classical test statistic.
+%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
 %, with non-centrality parameter $\delta$. 
@@
 To assess whether the linear assumption is appropriate we would make a plot of residuals against fitted values (case 5).
-For the purposes of comparing visual against classical inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
+For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
 
@@
 %%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in classical test be $t_{obs}$ with $p$-value $p_D$.
+For the regression slope parameter $\beta$ suppose the associated test statistic in conventional test be $t_{obs}$ with $p$-value $p_D$.
 By definition we have
@@
 To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
-This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the classical $p$-value as details given in Lemma \ref{lemma}.
+This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}.
 Thus the expected power of the visual test in this scenario is
@@
 Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-The first two are classical settings where we would not expect the lineup protocol to do better than the traditional test, but we hope it to perform favorably.
-The third experiment is a scenario where assumptions required for the traditional test are violated, and we would expect the lineup protocol to outperform the traditional test.
+The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.
+The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
 (Data and lineups used in the experiments are available in the supplementary material.)
@@
 For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
-A 1000 sets were generated for each parameter combination and the traditional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
+A 1000 sets were generated for each parameter combination and the conventional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
 The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
@@
 
-In the real setting, where visual inference is to be useful, there will be no conventional tests or traditional $p$-values.
+In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values.
 Assessing the strength of perceived structure is a critical component of visual inference.
@@
 %Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
-This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense.
+This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a conventional sense.
 Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
@@
 }
-%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
+%is equivalent to conventional test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
 %and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
@@
 The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
-We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to traditional inference.
-The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where traditional tests cannot be used.
+We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
+The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
 For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
@@
 This changed, when Buja et al.(2009) conceptually introduced two protocols for formal tests of visual findings.
-In this paper we take this a step further by comparing the lineup protocol (Buja et al.2009) against classical statistical testing  of the significance of regression model parameters.
+In this paper we take this a step further by comparing the lineup protocol (Buja et al.2009) against conventional statistical testing  of the significance of regression model parameters.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.

### 2012-07-05 paper_jasa.tex
smoothing out conclusions

--- 2012-07-05 paper_jasa.tex
@@
 \begin{table*}[hbtp]
-\caption{Comparison of visual inference with existing inference technique.  }
+\caption{Comparison of visual inference with conventional inference.}
 \centering 
@@
 
-\green{
-\begin{itemize} \itemsep 0in
-\item Summarize the key pieces of the paper again, especially the major findings
-
-\blue{It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
-But this paper demonstrates that statistical graphics could also be used in statistical inference and establishes a formal framework to accomplish that.
-The simulation study reveals that even in a situation where conventional test has its all the convenient assumptions met, visual inference is still comparable to conventional test.
-Also, for some trained observers/subjects visual inference may produce even better power.
-On the other hand when the underlying assumptions of a conventional test do not satisfy and the test fails to produce enough power, visual inference performs extremely well.
-
-We obtain the estimation of the power of visual test in different ways and present the results in this paper.
-The experimental data suggests that observers usually pick the plot which has the minimum $p$-value or maximum signal strength in the plot.
-If we assume it to be true in general,  the power of visual test is then obtained theoretically.
-Interestingly the theoretical power of visual test is also very close to that of conventional test even though the situation of the experiment is worse for visual test and best for conventional test.}
-
-\item Explain caveats, nuances, advantages with the work
-
-\blue{One of the most critical steps in visual test would be select an appropriate test statistic for the hypothesis of interest.
+%\begin{itemize} \itemsep 0in
+%\item Summarize the key pieces of the paper again, especially the major findings
+
+It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
+This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
+The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
+It is possible that, for some trained observers visual inference may produce even better power.
+The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
+
+Power of visual tests can be estimated in different ways, which was explained in this paper.
+If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
+Interestingly, the theoretical power of visual test is close to that of conventional test in the conditions tested in this paper.
+
+%\item Explain caveats, nuances, advantages with the work
+
+One of the most critical steps in visual test would be select an appropriate test statistic for the hypothesis of interest.
 A good test statistic should show a very specific pattern if the null hypothesis is not true.
@@
 The major advantage of visual statical inference is its intuitive concept of testing technique which is easy to be grasped and fun to learn.
-}
-
-\item Describe the next steps of the work, and possible ideas for future extensions
-
-\blue {When there is no conventional test available, we can conveniently use visual testing procedure due to its flexibility of having no assumption restrictions.
+
+%\item Describe the next steps of the work, and possible ideas for future extensions
+
+When there is no conventional test available, we can conveniently use visual testing procedure due to its flexibility of having no assumption restrictions.
 Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
-The direction of future research would be to study that and apply the methods in those areas.}
-\end{itemize}
-}
+The direction of future research would be to study that and apply the methods in those areas.
 

### 2012-07-05 paper_jasa.tex
finished smoothing out conclusions

@@
 
-One of the most critical steps in visual test would be select an appropriate test statistic for the hypothesis of interest.
-A good test statistic should show a very specific pattern if the null hypothesis is not true.
-A small difference in color template or orientation of the plot could make a difference.
-This is the part of our ongoing and future research.
-
-Even though a single observer can make a decision we recommend having multiple observer to evaluate a lineup to obtain a definitive $p$-value to make a decision confidently.
-
-The experimental subjects that we recruited came from a various level of education and not necessarily well aware of statistical plots.
-But still we achieve the power of visual test comparable to the conventional test.
-This suggests that visual statistical technique could be a smart tool for high school students to get to know the first course of statistical inference.
-The major advantage of visual statical inference is its intuitive concept of testing technique which is easy to be grasped and fun to learn.
+One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
+A good test statistic is expected to reveal structure if the null hypothesis is not true.
+Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
+This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
+%This is the part of our ongoing and future research.
+
+Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers.
+This ensures the ability to obtain a good estimate of the $p$-value for the test, and more confidence in the decision.
+
+Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
+Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
+Yet the results were good.
+We expect that visual inference could a useful tool for high school students to learn about randomness as a primer to statistical thinking.
+A major advantage of visual statisticcal inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.
+
+No restrictions were placed on Turk workers, in terms of abilities.
+It would be interesting to see how well the trained subjects performs.
+Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
+One of our future work would be to examine this in details.
+For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
+After being trained they may have to pass the qualification test before participating the experiment.
+Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.
 
@@
 
-When there is no conventional test available, we can conveniently use visual testing procedure due to its flexibility of having no assumption restrictions.
-Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
+When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature.
+We have successfully used the procedures informally for local projects.
+It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no exising gauges of statistical significance.
+The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
+If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
+Follow-up questions on the reasons provide qualitative insight.
+Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
+
+%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
 The direction of future research would be to study that and apply the methods in those areas.
 
-The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
+%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
 We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
@@
 
-\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
+%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
 Along the lines of the Army tests that Susan brought to the working group.
@@
 
-Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
-
-In practice there is no correct answer.
+%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
+
+%In practice there is no correct answer.
 In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
@@
 
-What are your suggestions for future experiments, and next steps of the research?
+%What are your suggestions for future experiments, and next steps of the research?
 

### 2012-07-06 paper_jasa.tex
cleaned section 6.7

@@
 If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
-
-\green{Something is definitely not right!
+That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.
+
+%\green{Something is definitely not right!
 The code looks like it does actual data p-value - min(null $p$-value).
@@
 
- \green{Is this really pct correct, or 1-pct correct?
+ %\green{Is this really pct correct, or 1-pct correct?
 \blue{This is proportion correct}.

### 2012-07-06 paper_jasa.tex
coment added wehere we need to think more

@@
 It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers.
-Type II error is harder to calculate, as is usually the case.
+Type II error is harder to calculate, as is usually the case \blue{(Is it true?
+For conventional test it is easy to compute Type II error !)}.
 In visual inference, individual abilities need to be incorporated to calculate Type II error.
@@
 This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details.
-Estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
+Estimating $p_i$ will be one of the main concerns for the remainder of the paper.} One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
 
@@
 
-\hh{\noindent An important difference between conventional testing and the lineup scenarios is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
+\hh{\noindent An important difference between conventional testing and the lineup scenarios \blue{(Can we just say 'visual test'?)} is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
 we are assuming, again, that $X$ counts the number of observers (out of $K$) who identify the data plot from the lineup.
@@
 
-Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a power of 
+Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a power of \blue{(Is this power or Type I error?)}
 \[
@@
 
-\subsection{Subject-specific abilities}
+\subsection{Subject-specific abilities}~\label{sec:model}
 Calculating the power  is intricately related to the probability to identify the data plot from the lineup.

### 2012-07-06 paper_jasa.tex
proof reading

@@
 
-\hh{In order to compare power of conventional tests and lineups side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
+\hh{In order to compare power of conventional and visual tests side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
 }
@@
 Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-\green{What do we learn about the models from these parameter estimates?
+The larger variability in the response data for experiment 3 as we see in Figure \ref{fig:power_loess_effect} contributes the random effect variance estimate be much larger compared to those of other experiments.
+Because fo the same reason we see the standard error for intercept estimate is much higher as well.
+Also notice the slope estimate for experiment three is lowest as well.
+It is because for small parameter values the power is much higher for experiment 3.
+% \green{What do we learn about the models from these parameter estimates?
 Explain this here too.}
 
-\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
+%\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
 
@@
 %This is the part of our ongoing and future research.
+
+\blue{Another critical part is the specific question that the observer needs to answer evaluating the lineup.
+The question should be more than just asking to find the most different plot.
+Selecting the appropriate word to describe the pattern we may see when alternative hypothesis is true is important to help the observers understand what pattern they should look for.
+By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
 

### 2012-07-06 paper_jasa.tex
comments on lemma 3.1

@@
 %\]
-This implies that as $p_D$, the $p$-value, becomes smaller, the probability that the observer identifies the data plot goes up and \blue{(How can we say that from this?
-The term  $E\left[ (1 - p_D)^{m-1}\right]$ is definitely not a function of $p_D$ since $p_D$ should be eliminated after the expectation.
-Instead, its a function of $m$ and the distribution of $p_D$ !
-If the density of $p$-value is very right skewed, the expectation term would be large.
+
+}
+% If the density of $p$-value is very right skewed, the expectation term would be large.
 The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in.
@@
 )},
-conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
-}
+%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
+%}
+%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$.
+
 %
@@
 
-Figure \ref{fig:pval_power} gives an overview of the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
+\blue{
+The above statement allows two immediate conclusions for the use of lineups:
+The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$: since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the data plot decreases with it.
+
+Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$; 
+if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and  less affected by an increase in $m$.
+This can also be seen in Figure \ref{fig:pval_power}, which
+visualizes the statement of lemma \ref{lemma}: }  Figure \ref{fig:pval_power} gives an overview of the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
 We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the actual data plot from null plots in a lineup of size $m=20$.

### 2012-07-06 paper_jasa.tex
small changes, and letter-casing in bib file

@@
 
-\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
+Note that no participant was shown the same lineup twice.
+%\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
 This happens when subject accidentally use browser's ``back'' button and same data is saved twice.
@@
 
-\blue{Another critical part is the specific question that the observer needs to answer evaluating the lineup.
-The question should be more than just asking to find the most different plot.
-Selecting the appropriate word to describe the pattern we may see when alternative hypothesis is true is important to help the observers understand what pattern they should look for.
-By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
+\blue{Another critical part is the  question that the observer needs to answer when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
+If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
+This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
+% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
 
@@
 We expect that visual inference could a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statisticcal inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.
+A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.
 

### 2012-07-06 paper_jasa.tex
abstract updated
small changes

@@
 In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against conventional statistical testing  of the significance of regression model parameters.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
+\blue{We extend and refine the terminology of visual inference to a more general framework to allow for comparisons with conventional tests.} A human subjects experiment is conducted using simulated data to provide controlled conditions.
 Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
@@
 It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers.
-Type II error is harder to calculate, as is usually the case \blue{(Is it true?
-For conventional test it is easy to compute Type II error !)}.
+Type II error is harder to calculate, as is usually the case.
 In visual inference, individual abilities need to be incorporated to calculate Type II error.
@@
 
-\hh{\noindent An important difference between conventional testing and the lineup scenarios \blue{(Can we just say 'visual test'?)} is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
+\hh{\noindent An important difference between conventional and visual testing  is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
 we are assuming, again, that $X$ counts the number of observers (out of $K$) who identify the data plot from the lineup.
@@
 
-Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a power of \blue{(Is this power or Type I error?)}
+Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a type I error of:
 \[

### 2012-07-06 paper_jasa.tex
tweaked abstract

@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently there were no formal visual methods in place for determining statistical significance of findings.
-This changed, when \citet{buja:2009} conceptually introduced two protocols for formal tests of visual findings.
-In this paper we take this a step further by comparing the lineup protocol \citep{buja:2009} against conventional statistical testing  of the significance of regression model parameters.
-\blue{We extend and refine the terminology of visual inference to a more general framework to allow for comparisons with conventional tests.} A human subjects experiment is conducted using simulated data to provide controlled conditions.
+Until recently, there were no formal visual methods in place for determining statistical significance of findings.
+This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings.
+In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
+A human subjects experiment is conducted using simulated data to provide controlled conditions.
 Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.

### 2012-07-07 paper_jasa.tex
changes along the discussion with Mahbub

@@
 $\tau$ & is a vector of independent real-valued random variables $\tau_{\ell i}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
-$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.\\
+$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
+Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
 \end{tabular}

### 2012-07-07 paper_jasa.tex
comment on p-value difference

--- 2012-07-06 paper_jasa.tex
@@
 
-In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
+\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
 If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
@@
 If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
-That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.
+That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.}
 

### 2012-07-10 paper_jasa.tex
finished revising the p-value difference, conclusions, and figure captions

--- 2012-07-07 paper_jasa.tex
@@
 While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$.
-This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data (Figure \ref{fig:power_loess_effect}).
+This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data.% (Figure \ref{fig:power_loess_effect}).
+*** You cannot reference the figure before you explain it!!!
 
@@
 Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-The larger variability in the response data for experiment 3 as we see in Figure \ref{fig:power_loess_effect} contributes the random effect variance estimate be much larger compared to those of other experiments.
-Because fo the same reason we see the standard error for intercept estimate is much higher as well.
+\green{The fixed effects estimates indicate that for all experiments as the effect increases the proportion correct increases.
+For experiment 3 it is very weak.
+The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.} %The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
+%contributes to the random effect variance estimate be much larger compared to those of other experiments.
+For the same reason we see the standard error for intercept estimate is much higher as well.
 Also notice the slope estimate for experiment three is lowest as well.
@@
 (Empty cells indicate no lineup shown with this combination.) Red indicates the plot with the lowest $p$-value in the lineup.
-In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
+(Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.) In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
 This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small.
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-The colored lines indicate the plot with the lowest $p$-value.
+Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
 Rows correspond to $\beta$, smallest to largest values from top to bottom.
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
-The colored lines indicate the plot with the lowest $p$-value.
+Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
 Rows correspond to $\beta$, smallest to largest values from top to bottom.
@@
 If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
-Measuring the difference between the plots is almost impossible!
-But in the controlled setting here we can again use $p$-value as a proxy for similarity and difference.
-
-In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
+Measuring the difference, generally, between the plots is almost impossible!
+But in the controlled setting here we can use $p$-value of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
+
+We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}).
+What we are interested in here is how this pick is affected by the distributin=on of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
+If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
+Figure \ref{fig:pval_difference} investigates this.
+The difference between the $p$-value of the actual data is compared with the lowest from the null plots.
+This is plotted horizontally, and the proportion correct is plotted vertically.
+Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
+In experiment 1 (boxplots) there were a lot more lineups where the actual data plot had the smaller $p$-value, but only barely.
+This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
+Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
+Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} show a little more complicated story.
+When there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as a force choose the same plot.
+when there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
+
+%\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
 If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
@@
        \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
-       \caption{Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.}
+       \caption{Investigating how does the strength of structure in the null plots affects the choice of the actual data plot.
+Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.
+Negative values indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
+Close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
        \label{fig:pval_difference}
@@
 The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-It is possible that, for some trained observers visual inference may produce even better power.
 The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
@@
 
-\blue{Another critical part is the  question that the observer needs to answer when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
+The question that is posed to observers can affect the abililty to evaluate lineups.
+In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
+In practice, the question posed will be very general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
+
+% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
 If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
@@
 
+The performance of subjects was quite varied, but consistent.
 No restrictions were placed on Turk workers, in terms of abilities.
-It would be interesting to see how well the trained subjects performs.
-Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
-One of our future work would be to examine this in details.
-For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
-After being trained they may have to pass the qualification test before participating the experiment.
-Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.
+There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test.
+ It would be interesting to see how well trained subjects might perform.
+Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference.
+Future work might explore this.
 

### 2012-07-10 R/paper_jasa.R

@@ -511,11 +511,21 @@ prr.min<-ddply(prr, "pic_name", summarise,
                    n=n[1],
                    sd=sd[1],
                    rep=rep[1])
+prr.actual<-ddply(prr, "pic_name", summarise,
+                   lpvalue=lpvalue[obs_pvalue==pvalue],
+                   rfreq=rfreq[obs_pvalue==pvalue],
+                   beta=beta[1],
+                   n=n[1],
+                   sd=sd[1],
+                   rep=rep[1])
+prr.actual <- subset(prr.actual, !(prr.actual$pic_name == prr.min$pic_name & prr.actual$lpvalue == prr.min$lpvalue))
 p <- qplot(lpvalue, rfreq, geom="point", data=prr) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0, colour="hotpink"), data=prr.min) + 
   geom_point(mapping=aes(xend=lpvalue, yend=0, colour="hotpink"), data=prr.min) + 
   xlab(expression(paste("P-value on ",log[10]," scale"))) + ylab("Relative frequency of picks") +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0, colour="turquoise"), data=prr.actual) + 
+  geom_point(mapping=aes(xend=lpvalue, yend=0, colour="turquoise"), data=prr.actual) + 
   facet_grid(beta~n+sd+rep, labeller="label_both") + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","0")) + 
   opts(legend.position="none")
 p
@@ -553,11 +563,21 @@ prr2.min<-ddply(prr2, "pic_name", summarise,
                    n=n[1],
                    sd=sd[1],
                    rep=rep[1])
+prr2.actual<-ddply(prr2, "pic_name", summarise,
+                   lpvalue=lpvalue[obs_pvalue==pvalue],
+                   rfreq=rfreq[obs_pvalue==pvalue],
+                   b=b[1],
+                   n=n[1],
+                   sd=sd[1],
+                   rep=rep[1])
+prr2.actual <- subset(prr2.actual, !(prr2.actual$pic_name == prr2.min$pic_name & prr2.actual$lpvalue == prr2.min$lpvalue))
 p <- qplot(lpvalue, rfreq, geom="point", data=prr2) +
[…]

### 2012-07-10 paper_jasa.tex
typo

--- 2012-07-07 paper_jasa.tex
@@
 For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n_{\gamma} \gamma(1-\gamma)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
+Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
 

### 2012-07-10 paper_jasa.tex
author description and keywords added

@@
 %\documentclass{article}
-\documentclass{article}
+\documentclass[12pt]{article}
 \usepackage{times}
@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook\\
-        Department of Statistics, Iowa State University}
+\author{
+{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+\thanks{Mahbubul Majumder is Graduate Student, Heike Hofmann is Associate  Professor, Dianne Cook is Full Professor, Department of
+Statistics and Statistical Laboratory, Iowa State University,
+Ames, IA 50011-1210. This research is supported in part by the
+National Science Foundation Grant \# DMS 1007697.}}
+\date{\vspace{-.7in}}
 %\date{\today}  %\today is replaced with the current date
@@
 Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
+
+{\bf Keywords: \sf visual inference, statistical graphics, lineup, non-parametric test} 
 \end {abstract}
@@
 \]
-
 \begin{eqnarray*}
-P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt =  E\left[ (1 - p_D)^{m-1}\right].
+P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
+&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
+&=&  E\left[ (1 - p_D)^{m-1}\right].
 \end{eqnarray*}
@@
 
-\paragraph{Acknowledgement:}
-
-This work was funded in part by National Science Foundation grant DMS 1007697.
+%\paragraph{Acknowledgement:}
+%This work was funded in part by National Science Foundation grant DMS 1007697.
 

### 2012-07-10 paper_jasa.tex
added key words
model covariate

@@
 
-{\bf Keywords: \sf visual inference, statistical graphics, lineup, non-parametric test} 
+{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, data mining, visualization, exploratory data analysis} 
 \end {abstract}
@@
 as well lineup-specific elements, e.g.
-difficulty level.
+effect size or difficulty level.
 \\
@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently there were no formal visual methods in place for determining statistical significance of findings.
-This changed, when Buja et al.(2009) conceptually introduced two protocols for formal tests of visual findings.
-In this paper we take this a step further by comparing the lineup protocol (Buja et al.2009) against conventional statistical testing  of the significance of regression model parameters.
+Until recently, there were no formal visual methods in place for determining statistical significance of findings.
+This changed when Buja et al.
+(2009) conceptually introduced the lineup protocol for formal tests of visual findings.
+In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results equivalent to the conventional test and for some scenarios yields better power than the conventional test.
+Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
 \end {abstract}

### 2012-07-10 paper_jasa.tex
removed appendix, and some blue color comments

@@
 Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted.
-Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the Appendix).
+Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
 For each difficulty level a specific number of lineups were randomly picked to show for evaluation.
@@
 Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
-\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
+}%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
 \begin{center}
@@
 %\end{multicols}
+
+\end{document}  %End of document.
+
 \section*{Appendix}
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
-       \caption{The overall power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for conventional test is shown in black solid lines for comparison.}
-       \label{fig:power_screening}
-\end{figure}
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
-      \caption{The subject-specific power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.
-The corresponding power curve for conventional test is shown in pink lines for comparison.}
-       \label{fig:power_screening_subject}	
-\end{figure}
-
-\end{document}  %End of document.
-
-\begin {abstract} 
-Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
+%\begin {abstract} 
+%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
 Until recently, there were no formal visual methods in place for determining statistical significance of findings.
@@
 Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
-\end {abstract}
+%\end {abstract}
 

### 2012-07-13 paper_jasa.tex
re-worked parts of section 2

@@
 \end{dfn}
-Note that for $x=0$ the $p$-value becomes --mathematically-- equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
+\noindent Note that for $x=0$ the $p$-value becomes, mathematically, equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
 being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
 
-\noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
+Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
 It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
@@
 \begin{center}
-\begin{tabular}{rrrrrrrrrrrrrrrrrrr}
-$K$ &  $x$ & $p$-value && $K$ &  $x$ & $p$-value && $K$ &  $x$ & $p$-value && $K$  & $x$ & $p$-value && $K$  & $x$ & $p$-value\\ \cline{1-3} \cline{5-7}\cline{9-11}\cline{13-15}\cline{17-19}
-1 &  1 & 0.0500 && 2 &  1 & 0.0975 && 3 & 1 & 0.1426 && 4 & 1 & 0.1855 && 5 & 1 & 0.2262 \\%\cline{1-3}
-&&&& 2 &  2 & 0.0025 && 3 & 2 & 0.0073 && 4 & 2 & 0.0140 && 5 & 2 & 0.0226 \\
-&&&& &&&& 3 & 3 & 0.0001 && 4 & 3 & 0.0005 && 5 & 3 & 0.0012 \\%\cline{1-6}
-&&&&  &    &        &   &   &        &&& 4 & 4 & $< 0.0001$ && 5 & 4 & $< 0.0001$ 
+\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
+$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline %\cline{1-3}\cline{4-6}\cline{7-9}\cline{10-12}\cline{17-19}
+1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
+&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
+&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
+&&&  &    &           &   &        && 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
 \end{tabular}
@@
 
-\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
-This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details.
-Estimating $p_i$ will be one of the main concerns for the remainder of the paper.} One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
+If there is evidence for individual skills influencing the probability $p_i$, then \green{$X_i \sim Binom_{1, p_i}$} and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
+This makes the distribution of $X$ a Poisson-Binomial by definition (see \citet{butler93} for details).
+\green{Ways to estimate $p_i$ will be be discussed in the following sections.}
+%one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
 
@@
 \begin{dfn} \label{dfn:power}
-The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
+The \textbf{power} of a visual test, $V_{\theta}$, is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
     \begin{equation*}
@@
 
-\hh{\noindent An important difference between conventional and visual testing  is that lineups always depend on observers -- this brings the number of people evaluating a particular lineup into the actual evaluation of power, i.e.
-we are assuming, again, that $X$ counts the number of observers (out of $K$) who identify the data plot from the lineup.
-Then power is calculated as
+\noindent \green{An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
+Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
 \[
@@
 \] 
-where $F_{X, \theta}$ is the distribution of $X$ -- this depends on which hypothesis is true:
-
-Under the null hypothesis $X \sim Binom_{K, 1/m}$, leading to a type I error of:
+Here $F_{X, \theta}$ is the distribution of $X$, and this depends on which hypothesis is true.
+Under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
 \[
-\text{Power}_V(\theta, K)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
+\widehat{Power}_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
 \]
-where $x_\alpha$ is defined as in (\ref{dfn:test}), as  $P(X \ge x_{\alpha}) < \alpha$.
-
-If the alternative hypothesis is true with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
-Again, assessing an individual's probability to identify the data plot  is central to the actual calculation .
-}
+where $x_\alpha$ is such that $P(X \ge x_{\alpha}) < \alpha$.
+On the other hand, if the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
+Assessing an individual's probability to identify the actual data plot will require that an individual evaluate multiple lineups, which may not always be feasible.}
+
 %\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
@@
 
-Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
-For visual inference, this corresponds to the choice of plot to use.
-Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
+\green{Power is an important consideration in deciding which test to use for solving a problem.
+%In conventional testing, new test procedures are examined for their power under a variety of scenarios.
+Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.} Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
 This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean.

### 2012-07-13 paper_jasa.tex
revised section 3, changed first person to third person language

@@
 
-\section{Effect of Lineup Size and Number of Observers}~\label{sec:size}
+\section{Effect of Observer Skills and Lineup Size}~\label{sec:size}
 
@@
 \subsection{Subject-specific abilities}~\label{sec:model}
-Calculating the power  is intricately related to the probability to identify the data plot from the lineup.
-
-Suppose each of $K$ independent observers gives evaluations on multiple lineups and responses are associated with binary random variable $X_{\ell i}$, where  $X_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$,  $\ell = 1, ..., L$, and 0 otherwise.
-
-\hh{
-We model $p_{\ell i} = E(X_{\ell i})$ as  a mixed effects logistic regression
- accommodating both for different abilities of individuals as well as differences in the difficulty of lineups:
-
-Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
-We can model $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
+%Calculating the power  is intricately related to the probability to identify the actual data plot from the lineup.
+
+Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $\ell = 1, ..., L$, and 0 otherwise.
+A mixed effects logistic regression model is used for $p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.
+
+Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
+The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
 \begin{equation} \label{eqn:mixed}
@@
 
-where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and
+\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and
 
@@
 \end{tabular}
-}
-
-Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} we obtain the estimate of subject and lineup specific probability of successful evaluation as 
+
+\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leands to the estimate of the subject and the lineup specific probability of successful evaluation as
 \begin{equation} \label{eqn:mixed_power}
@@
 
-\noindent
+%\noindent
 \subsection{Lineup size, $m$}
 
-One of the big differences between visual inference and the conventional testing situation is the use of a finite number $m-1$ of representatives of the null distribution as comparison against the test statistic.
-The choice of $m$ has an obvious impact on the test, and we will discuss this in the following paragraph.
-
-The following properties are only derived for the situation of a  fully parameterized  simulation study.
-This allows us to directly compare lineup tests against their conventional counterparts and also identify  properties relevant for a quality assessment of  lineups, which are typically used  in non-parametric situations.
-
-We need to make two crucial assumptions:
+The finite number $m-1$ of representatives of the null distribution used  as comparison against the test statistic, is a major difference between visual inference and the conventional testing.
+The choice of $m$ has an obvious impact on the test.
+
+The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper.
+They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups, for when they are used in practical settings.
+Two assumptions are critical:
 \begin{enumerate} \itemsep 0in
-\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
+\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
 \item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
 \end{enumerate}
-We will be able to show evidence for the second assumption with the help of the study discussed in Section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup.
+Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup.
 The better suited a design is for a particular task, the higher its power will be.
 
-\hh{In order to compare power of conventional and visual tests side-by-side, we will have to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
-}
+In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
 
@@
 The probability of picking the data plot from a lineup depends on the size $m$ of the lineup and the strength of the signal in the data plot. 
-Under the above assumptions, we can express the probability as:
+Under the above assumptions, the probability is expressed as:
 \[
@@
 \begin{proof}
-We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
-
- By definition we have
+%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
+
+ By definition
 
@@
 %, with non-centrality parameter $\delta$. 
-Then the distribution function of the $p$-value $p_D$ under $H_0$ is uniform, since:  
+\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
 \begin{eqnarray}\label{dist_p}
-F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
+\green{F_{p_D}}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
   &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
@@
 Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
-
-\hh{Their minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function 
+The minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function
 \[
@@
 \]
+
+\noindent Thus
 \begin{eqnarray*}
@@
 
-This proves the statement above. 
+%\noindent This proves the statement above. 
 %We further see:
@@
 
-}
 % If the density of $p$-value is very right skewed, the expectation term would be large.
@@
 
-\blue{
-The above statement allows two immediate conclusions for the use of lineups:
-The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$: since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the data plot decreases with it.
-
-Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$; 
-if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and  less affected by an increase in $m$.
-This can also be seen in Figure \ref{fig:pval_power}, which
-visualizes the statement of lemma \ref{lemma}: }  Figure \ref{fig:pval_power} gives an overview of the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
-We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the actual data plot from null plots in a lineup of size $m=20$.
-% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
+\green{The above lemma allows two immediate conclusions for the use of lineups.
+The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$.
+This may seem counterintuitive, but it makes sense.
+With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true.
+Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$.
+This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.
+Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
+It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.} % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
 %also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.

### 2012-07-13 paper_jasa.tex
some typo fixed

--- 2012-07-10 paper_jasa.tex
@@
 \begin{tabular}{llll} 
-\hline
+\hline \hline
   & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
@@
 \begin{center}
+\scalebox{.9}{
 \begin{tabular}{rrrrrrrrrrrrrrrrrrr}
@@
 \end{tabular}
+}
 \end{center}
@@
 \end{itemize}
-where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
+where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
 \end{dfn}
@@
 
-\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
+\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p_i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
 This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details.
@@
 \[
-\text{Power}_V(\theta, K)= 1 - Binom_{K, 11/m} (x_\alpha - 1),
+\text{Power}_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
 \]
-where $x_\alpha$ is defined as in (\ref{dfn:test}), as  $P(X \ge x_{\alpha}) < \alpha$.
+where $x_\alpha$ is defined as in (\ref{dfn:test}), as  $P(X \ge x_{\alpha}) \le \alpha$.
 
@@
 
-\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
+%\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
 Can someone please decide which to use, and make the appropriate changes?
@@
 
-\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
+%\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
 \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
@@
 \centering
-% centering table
+\scalebox{.9}{
 \begin{tabular}{c c c c c}
@@
 \end{tabular}
+}
 \label{tbl:experiment_params}
@@
 \begin{tabular}{ccrrcc}
-   \hline 
+   \hline \hline
  Experiment & Subject & Male & Female & Responses &Lineup\\ 
@@
 \begin{tabular}{cr@{.}lcc}
-  \hline
+  \hline \hline
  &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
@@
 Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.
+\blue{The loess curve (blue) estimates the observed power}, and grey bands show simultaneous bootstrap confidence band.
 Conventional test power is drawn as a dashed line.
@@
 %Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
- The estimated subject-specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines).
-The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
+ The estimated subject-specific power curves are shown in \blue{Figure \ref{fig:power_mixed_subject} (light grey lines).
+The blue curve is the overall} average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
 Subject-specific power is quite different for three experiments.
@@
        \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
-       \caption{Investigating how does the strength of structure in the null plots affects the choice of the actual data plot.
+       \caption{Investigating how the strength of structure in the null plots affects the choice of the actual data plot.
 Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.
@@
 Yet the results were good.
-We expect that visual inference could a useful tool for high school students to learn about randomness as a primer to statistical thinking.
+We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
 A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.

### 2012-07-13 paper_jasa.tex

--- 2012-07-13 paper_jasa.tex
@@
 \begin{center}
+\scalebox{.9}{
 \begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline

### 2012-07-17 paper_jasa.tex
typos up to page 7

@@
 
-Statistical graphics nourishes the discovery process in data analysis: revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally contrasting prevailing hypotheses.
-It is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
+Statistical graphics nourish the discovery process in data analysis by revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally by contrasting prevailing hypotheses.
+The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
 Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
@@
 This has also been addressed in recent research.
-\citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics.
+\citet{buja:2009} proposes a protocol that allows the testing of discoveries made from statistical graphics.
 This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
@@
 The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
-Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
+Testing is done based on a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
 The results are compared with those of the conventional test.
@@
 Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing.
-Both start from the same place, the same null hypothesis.
+Both start from the same place, the same set of hypotheses.
 The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
-In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
+In the lineup protocol, the test statistic is a plot of the data.
+Here, side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
 In conventional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
@@
 \begin{dfn} \label{dfn:test}
-A \textbf{visual test statistic}, $T(.)$, is a function of a sample that makes a plot.
-$T(y)$ maps the actual data to the plot, and we call this the \textbf{actual data plot}, and $T(y_0)$ maps samples drawn from the null distribution into the same plot form, and these are called \textbf{null plots}.
+A \textbf{visual test statistic}, $T(.)$, is a function of a sample that produces a plot.
+$T(y)$ maps the actual data to the plot, and we call this the \textbf{(actual) data plot}, and $T(y_0)$ maps a sample drawn from the null distribution into the same plot form.
+These type of plots are called \textbf{null plots}.
 \end{dfn}
@@
 \noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
+\\
+\noindent
 The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
@@
 
-\noindent The $(m-1)$ null plots can be members of the sampling distribution of the test statistic assuming that the null hypothesis is true.
+\noindent The $(m-1)$ null plots are % can be 
+members of the sampling distribution of the test statistic assuming that the null hypothesis is true.
 If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e.
@@
 \begin{dfn}\label{dfn:pvalue}
-The $p$-value of a lineup of size $m$ is given as
+The $p$-value of a lineup of size $m$ is  estimated as %given as
 
@@
 \begin{dfn}\label{dfn:test}
-The \textbf{visual test}, $V_{\theta}$  of size $\alpha$, is defined to be 
+The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
 \begin{itemize}\itemsep-3pt
@@
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
-\textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
+The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.\\ \noindent
+The \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
 \end{dfn}
@@
 
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$.
-For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.
-It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers.
+\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$. %, with an exact $x_{\alpha}$. 
+For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
 Type II error is harder to calculate, as is usually the case.
 In visual inference, individual abilities need to be incorporated to calculate Type II error.
-This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
-For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different.
+%This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
+Here, we need to estimate the probability that an observer sees the actual data plot as different, when it really is different.
 This involves understanding the individual's visual skills.
-Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup.
+Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
+Let $p_i$ be the probability that individual $i$ picks out the actual data plot from the lineup.
 If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
@@
 
-\noindent \green{An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
+\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
 Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
@@
 where $x_\alpha$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
-On the other hand, if the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
-Assessing an individual's probability to identify the actual data plot will require that an individual evaluate multiple lineups, which may not always be feasible.}
+If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
+Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.
 

### 2012-07-17 paper_jasa.tex
proof-read through page 14

@@
 
-\green{Power is an important consideration in deciding which test to use for solving a problem.
-%In conventional testing, new test procedures are examined for their power under a variety of scenarios.
-Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.} Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
+Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
+Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.
+Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
 This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean.
@@
 
-Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $\ell = 1, ..., L$, and 0 otherwise.
-A mixed effects logistic regression model is used for $p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.
-
-Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
+Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i} \sim  Binom_{1, p_{\ell i}}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $1 \le \ell \le L$, and 0 otherwise.
+A mixed effects logistic regression model is used for $P(X_{\ell i} = 1) =  p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.
+
+%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
 The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
@@
 $Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
-$\tau$ & is a vector of independent real-valued random variables $\tau_{\ell i}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
+$\tau$ & is a vector of independent normally distributed random variables $\tau_{\ell i}$, with  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
@@
 
-\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leands to the estimate of the subject and the lineup specific probability of successful evaluation as
+\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
 \begin{equation} \label{eqn:mixed_power}
@@
 The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper.
-They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups, for when they are used in practical settings.
+They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups when they are used in practical settings.
 Two assumptions are critical:
@@
 \begin{eqnarray}\label{dist_p}
-\green{F_{p_D}}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
+F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
   &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
@@
 
-\green{The above lemma allows two immediate conclusions for the use of lineups.
+The above lemma allows two immediate conclusions for the use of lineups.
 The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$.
-This may seem counterintuitive, but it makes sense.
-With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true.
+%This may seem counterintuitive, but it makes sense.
+%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
 Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$.
@@
 Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
-It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.} % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
+It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.
+% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
 %also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.
@@
 \citet{cook:99}).
-Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and the visual test statistic that might be used for testing these hypotheses.
-These plots should be familiar, because they are all in common use.
+Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
+%These plots should be familiar, because they are all in common use. 
 For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
-To assess whether the linear assumption is appropriate we would make a plot of residuals against fitted values (case 5).
-For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.
+To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
+For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.
 
@@
 Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are different and we fit the null model to the actual data, the resulting residuals should show a significant difference between the two groups.
-To test this boxplots of the residuals conditioned on the two levels of $X_k$ are made, as the observed test statistic.
+If the responses for the two levels of the categorical variable $X_k$ in the model are different, the  residuals from fitting the null model should show a significant difference between the two groups.
+For a visual test, we draw boxplots of the residuals conditioned on the two levels of $X_k$.
 If $\beta_k\ne 0$ the boxplots should show a vertical displacement.
@@
 
-The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
+The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
 It forms the benchmark upon which we evaluate the visual test.
@@
 
-Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk.
+Participants for all the experiments were recruited through Amazon's online web service, Mechanical Turk  \citep{turk}.
 A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}.
@@
 Gender, age, education and geographic location of each participant are also collected.
-The data was collected through the web.
+%The data was collected through the web.
 For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.
@@
 To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
-The actual data plot was generated by making a side-by-side boxplot of the residuals (Table \ref{tbl:stat_multiple}, case 3).
-The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and these were plotted in the same way as side-by-side boxplots.
+The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3).
+The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way.
 The actual data plot was randomly placed among these null data plots to produce the lineup.
@@
 Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
-
-Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted.
-Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
-For each difficulty level a specific number of lineups were randomly picked to show for evaluation.
+% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}
+
+Each subject viewed at least 10 lineups with the option to evaluate more.
+%and we allowed them to provide more evaluations if they wanted. 
+Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
+For each difficulty level a specific number of lineups was randomly picked  for evaluation.
 This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$.
@@
 To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data.
-An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.
+To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
+For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest).
+Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
 The actual data plot location is 11.
-This lineup, we think, is a difficult one.
-To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
-For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest).
-For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
+
+For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
 Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
@@
 Each subject saw a substantial number of difficult lineups.
-\hh{While this allows us a better estimation for the power in small slopes, }
-%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
-but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
-\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
+
+%\hh{While this allows us a better estimation for the power in small slopes, }
+%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
+%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
+%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
 
@@
 
-The first two simulation experiments use data generated with a normal error model, satisfying the conditions for conventional test procedures.
+The first two simulation experiments use data generated under a normal error model, satisfying the conditions for conventional test procedures.
 In these situations there exists a test, and there would, in general, be no need to use visual inference.
-The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the conventional test, will compare favorably with it.
+The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
  This third simulation is closer to the mark for the purpose of visual inference.
@@
     \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
-    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
+    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
   \end{array} \right.
 \]
-where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data.
-For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points.
-Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
+where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$.
+$n_c$ is the size of thecontaminated data.
+For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points.
+Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
 %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.

### 2012-07-17 paper_jasa.tex
figure caption
colour mentioned

@@
        \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition .
-Values are shown on square root scale.
-For experiment 3, $p$-values of uncontaminated part of the data are also shown.}
+       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . Values are shown on square root scale. }
     %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
+\blue{Red indicates the} plot with the lowest $p$-value, and \blue{blue indicates the} actual data plot, when it is different from that with the lowest $p$-value.
 Rows correspond to $\beta$, smallest to largest values from top to bottom.
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
+\blue{Red indicates the} plot with the lowest $p$-value, and \blue{blue indicates} the actual data plot, when it is different from that with the lowest $p$-value.
 Rows correspond to $\beta$, smallest to largest values from top to bottom.

### 2012-07-17 paper_jasa.tex
proof-read to page 17

@@
 %\usepackage[dvips]{graphics}
+\usepackage{hyperref}
 
@@
 $W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$.
-Covariates could include  demographic information of individuals, such as age, gender, education level etc.
-as well lineup-specific elements, e.g.
+Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g.
 effect size or difficulty level.
@@
 A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
-This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2.
-The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.
-
-Five replications of actual data sets for each slope value shown in Table \ref{tbl:experiment_params}, producing 30 lineups.
-Six groups of difficulty level based on slope are defined, difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.
-Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.
+This experiment  follows the test in Table \ref{tbl:stat_multiple}, case 2.
+The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null  plots are scatterplots of null data generated by plotting simulated residuals from $N(0, {\hat{\sigma}}^2$)  against $X_1$.
+
+Experiment three consists of a total of 30 lineups, made up of five replicates for each of the six slopes as shown in Table \ref{tbl:experiment_params}.
+We use the slope directly as a measure of difficulty, with difficulty = 0 for the largest slope and difficulty = 5 for the smallest slope.
+Subjects were exposed to a total of ten lineups, with two lineups from each of the difficulty levels 0 through 3, and one  lineup each from levels 4 and 5.
 
@@
 For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
-For the purposes of identifying these participants and cleaning the data, each block of 10 lineups that a subject received included several very easy ones.
-One of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
+For the purpose of identifying these participants and cleaning the data, we make use of one of the very easy lineups that every subject was exposed to.
+For each subject we use a randomly selected easy lineup.
+If the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
+If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
+Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
+
+%
+%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
 The data of participants who did not answer correctly was removed.
-For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
+%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
 That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.
@@
 
-Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-
 \subsection{Model fitting}
-For each parameter combination, the effect is computed as
+\noindent For each parameter combination, the effect is computed as
 
 \[
-E=\beta (\sqrt {n}/\sigma).
+E=\sqrt {n} \cdot \beta/\sigma.
 \]
 
-The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept.
-we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
+The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept., i.e. 
+ $W_{\ell i} = E_{\ell i}$ in Equation \ref{eqn:mixed}. 
+Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
 Different subjects are expected to have different skill levels in evaluating lineups.
-To account for subject specific variation we use subject as random effect.
-We do not include random intercept, instead, we use an interaction between subject and effect $E$ to allow for subject specific random slope.
-This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.
-While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$.
-This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data.% (Figure \ref{fig:power_loess_effect}).
-*** You cannot reference the figure before you explain it!!!
-
-For all the three experiments, each subject saw a lineup just once.
-This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
-The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}.
-For computation we use R \citep{R} package lme4 \citep{lme4:2011}.
+We account for this by using subject-specific random  slopes for  effect ($E$).
+
+For experiment 3 we so fit intercepts: both a fixed effect and subject-specific random intercepts, since forcing power
+%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
+to be fixed at 0.05 for $E=0$ is not justified by the data from experiment 3.% (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
+
+%For all the three experiments, each subject saw a lineup just once.
+This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
+
+The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. For computation we use R \citep{R} package lme4 \citep{lme4:2011}.
 

### 2012-07-17 paper_jasa.tex
double spacing makes page size = 45
author email added

@@
 % \usepackage{pifont} % this package is used to print check mark \checkmark
+%\linespread{2}
 
@@
 {Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student, Heike Hofmann is Associate  Professor, Dianne Cook is Full Professor, Department of
+\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Full Professor, Department of
 Statistics and Statistical Laboratory, Iowa State University,

### 2012-07-17 paper_jasa.tex
up to page 18

@@
 For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
-For the purpose of identifying these participants and cleaning the data, we make use of one of the very easy lineups that every subject was exposed to.
-For each subject we use a randomly selected easy lineup.
-If the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
+For the purpose of identifying these participants and cleaning the data, we take action based on a subject's answer to one of the very easy lineups that everybody was exposed to:
+if the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
 If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
@@
 \]
-
-The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept., i.e. 
- $W_{\ell i} = E_{\ell i}$ in Equation \ref{eqn:mixed}. 
+\noindent
+The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept, i.e. 
+ $W_{\ell i} = E_{\ell i}$.
 Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
-Different subjects are expected to have different skill levels in evaluating lineups.
-We account for this by using subject-specific random  slopes for  effect ($E$).
-
-For experiment 3 we so fit intercepts: both a fixed effect and subject-specific random intercepts, since forcing power
+Different skill levels of subjects are accounted for  by allowing subject-specific random  slopes for  effect ($E$).
+
+For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power
 %While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
-to be fixed at 0.05 for $E=0$ is not justified by the data from experiment 3.% (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
+to be fixed at 0.05 for $E=0$ is not justified by the data. % (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
 
@@
 
-The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. For computation we use R \citep{R} package lme4 \citep{lme4:2011}.
+%The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. 
+For computation we use package {\tt lme4} \citep{lme4:2011} and software R 2.15.0 \citep{R}. $p$-value calculations are based on asymptotic normality.
 
 Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-\green{The fixed effects estimates indicate that for all experiments as the effect increases the proportion correct increases.
-For experiment 3 it is very weak.
-The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.} %The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
+The fixed effects estimates indicate that for all experiments  the proportion of correct responses increases as the effect increases.
+This effect is less pronounced  for experiment 3.
+The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.
+%The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
 %contributes to the random effect variance estimate be much larger compared to those of other experiments.
@@
 
-\noindent Figure \ref{fig:power_loess_effect} shows the power against effect for the three experiments.
-Data values are represented as dots with size indicating count at this value.
-The loess curves fit the data, estimating the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands.
+\noindent Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.
+Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
+A loess  fit to the data gives an estimate of  the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
 For comparison, the dashed lines show the corresponding power curves of the conventional tests.
-The observed power for experiments 1 and 2 are lower than the power of the conventional test.
-This is not a surprise -- it is not expected that visual inference will beat conventional tests, in the scenario when assumptions are satisfied.
+Not surprisingly, the observed power for experiments 1 and 2 is lower than the power of the conventional test.
 It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
+Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
 
@@
 Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-\blue{The loess curve (blue) estimates the observed power}, and grey bands show simultaneous bootstrap confidence band.
+The loess curve (continuous line) estimates the observed power, and surrounding bands show simultaneous bootstrap confidence band.
 Conventional test power is drawn as a dashed line.
@@
 
-Results for experiment 3, are quite different.
+Results for experiment 3 are quite different.
 This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
 There is no actual conventional power in this situation, because assumptions are violated.
-It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
+The dashed line shows conventional power based on uncontaminated data, whereas the cross marks show effective power based on the coefficient estimated from the contaminated data.
+% It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
 The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
 For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
-Notice that the power structure is completely broken for contaminated data.
-For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
-
-It is really curious to see in the results of experiment 3 that the power of the visual test is really high when effect is small.
+% 
+%Note that the power structure is completely broken for contaminated data. %For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
+
+It is really curious to see in the results of experiment 3 that the power of the visual test  high independent of effect size. 
 These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
-Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data.
+Although subjects were asked to select the plot that exhibited the highest association between the two variables, they might have cued in on the cluster of contaminated data.
 This will be explored further in Section \ref{sec:TypeIII}.
@@
 For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
-
-It can also be seen that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2.
-For experiment 3 the scenario is different.
-The contamination in the data is visible which sets the lower limit of proportion correct at a higher value.
 

### 2012-07-17 paper_jasa.tex
figure 10

@@
    \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
+      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
+  %      \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-\blue{Red indicates the} plot with the lowest $p$-value, and \blue{blue indicates the} actual data plot, when it is different from that with the lowest $p$-value.
-Rows correspond to $\beta$, smallest to largest values from top to bottom.
+%\blue{Red indicates the} plot with the lowest $p$-value, and
+       \blue{blue indicates the} actual data plot, when it is different from that with the lowest $p$-value.
+The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
+       %Rows correspond to $\beta$, smallest to largest values from top to bottom.
 Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
 Empty cells indicate no lineup for that combination of parameters.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value.}
+Highest counts tend to be the plot in the lineup having the lowest $p$-value.
+       }
        \label{fig:P-val_log}

### 2012-07-17 R/paper_jasa.R

@@ -519,6 +519,7 @@ prr.actual<-ddply(prr, "pic_name", summarise,
                    sd=sd[1],
                    rep=rep[1])
 prr.actual <- subset(prr.actual, !(prr.actual$pic_name == prr.min$pic_name & prr.actual$lpvalue == prr.min$lpvalue))
+
 p <- qplot(lpvalue, rfreq, geom="point", data=prr) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0, colour="hotpink"), data=prr.min) + 
@@ -526,11 +527,44 @@ p <- qplot(lpvalue, rfreq, geom="point", data=prr) +
   xlab(expression(paste("P-value on ",log[10]," scale"))) + ylab("Relative frequency of picks") +
   geom_segment(mapping=aes(xend=lpvalue, yend=0, colour="turquoise"), data=prr.actual) + 
   geom_point(mapping=aes(xend=lpvalue, yend=0, colour="turquoise"), data=prr.actual) + 
-  facet_grid(beta~n+sd+rep, labeller="label_both") + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","0")) + 
+  facet_grid(beta~n+sd+rep, labeller="label_both") + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","1")) + 
   opts(legend.position="none")
 p
 
 ggsave(file="../images/p_val_log_counts.pdf", height=10, width=12)
+
+# alternative ordering for the previous plot
+prr$facets <- with(prr, paste("beta:", beta,"\n n:", n,"\n sd:",sd, paste=""))
+prr$effect <- with(prr, sqrt(as.numeric(as.character(n)))*as.numeric(as.character(beta))/as.numeric(as.character(sd)))
+prr$facets <- reorder(prr$facets, prr$effect, median)
+
+prr.actual <- subset(prr, pvalue==obs_pvalue)
+
+p1 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) <= 10)) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
[…]

### 2012-07-17 paper_jasa.tex
fig 10 - boundaries in y direction are 0,1

@@
 
-It is really curious to see in the results of experiment 3 that the power of the visual test  high independent of effect size. 
-These results are based on whether the subject correctly picked the actual data plot, regardless of reason.
+Results of experiment 3 are curious insofar, as  power of the visual test is  high independent of effect size.
+However, these results are based on correct identification of the actual data plot, regardless of reason.
 Although subjects were asked to select the plot that exhibited the highest association between the two variables, they might have cued in on the cluster of contaminated data.
@@
 %Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
- The estimated subject-specific power curves are shown in \blue{Figure \ref{fig:power_mixed_subject} (light grey lines).
+\noindent The estimated subject-specific power curves are shown in \blue{Figure \ref{fig:power_mixed_subject} (light grey lines).
 The blue curve is the overall} average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.

### 2012-07-17 paper_jasa.tex
color back in fig 10

@@
 %Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
-\noindent The estimated subject-specific power curves are shown in \blue{Figure \ref{fig:power_mixed_subject} (light grey lines).
-The blue curve is the overall} average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test.
-Subject-specific power is quite different for three experiments.
+\noindent 
+Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
+The thick continuous line shows overall estimated power, the thinner lines correspond to subject-specific power curves.
+For comparison, the  dashed lines show power curves of the conventional test.
+Subject-specific power is quite different between the three experiments.
 In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
@@
 As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values.
-When $p$-value is larger than 0.15, it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
+For  $p$-values  larger than 0.15 it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
 
@@
 
-From the experimental data the visual $p$-values are computed using Definition \ref{dfn:pvalue}.
-Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$-value.
-The pattern of visual $p$-value is interesting.
-If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) giving a clear indication to reject $H_0$ or not.
+From the experimental data the visual $p$-values are estimated based on Definition \ref{dfn:pvalue}.
+Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.
+The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate,
+% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
+% 
+ giving a clear indication to reject $H_0$ or not.
 This is why we do not see lot of visual $p$-values in between 0.05 and 0.8 especially for experiment 2.

### 2012-07-17 R/paper_jasa.R

@@ -548,7 +548,7 @@ p1 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) <=
   geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
   geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
   facet_grid(rep~facets) + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","1")) + 
-  opts(legend.position="none")
+  opts(legend.position="none") + scale_y_continuous(limits=c(0,1))
 p1
 ggsave(file="../images/p_val_log_counts-a.pdf", height=4, width=12)
 
@@ -560,7 +560,7 @@ p2 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) > 
   geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) > 10)) +
   geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) > 10)) +
   facet_grid(rep~facets) + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","1")) + 
-  opts(legend.position="none")
+  opts(legend.position="none") + scale_y_continuous(limits=c(0,1))
 p2
 ggsave(file="../images/p_val_log_counts-b.pdf", height=4, width=12)
 
@@ -589,6 +589,48 @@ p <- qplot(lpvalue, counts, geom="point", data=subset(prr2,lpvalue != lobs_pvalu
   facet_grid(b~n+sd+rep, scales="free_y", labeller="label_both") +
   opts(legend.position="none")
 p
+
+# alternative to previous plot with different ordering
+prr2 <- get_merged_pvalue(dat2,pval2)
+prr2$facets = with(prr2, paste("beta: ",b, "\nn:",n,"\nsd: ",sd))
[…]

### 2012-07-17 paper_jasa.tex
figure 11

@@
  giving a clear indication to reject $H_0$ or not.
-This is why we do not see lot of visual $p$-values in between 0.05 and 0.8 especially for experiment 2.
-This guide the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$.
+This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
+This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$.
 For visual test this is not common.
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-%\blue{Red indicates the} plot with the lowest $p$-value, and
-       \blue{blue indicates the} actual data plot, when it is different from that with the lowest $p$-value.
+\blue{Red indicates the} plot with the lowest $p$-value, and
+       \blue{blue indicates the} actual data plot. %, when it is different from that with the lowest $p$-value. 
        The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
@@
    \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
+%       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
+       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-a.pdf}       
+       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-b.pdf}
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
-\blue{Red indicates the} plot with the lowest $p$-value, and \blue{blue indicates} the actual data plot, when it is different from that with the lowest $p$-value.
-Rows correspond to $\beta$, smallest to largest values from top to bottom.
+\blue{Red indicates the} plot with the lowest $p$-value.
+%, and \blue{blue indicates} the actual data plot, when it is different from that with the lowest $p$-value.
+       %Rows correspond to $\beta$, smallest to largest values from top to bottom.
 Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
-Empty cells indicate no lineup for that combination of parameters.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.}
+%       
+Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
+       Empty cells indicate combination of parameters with fewer than five replicates. 
+       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
+       }
        \label{fig:P-val_log2}

### 2012-07-17 R/paper_jasa.R

@@ -539,11 +539,16 @@ prr$effect <- with(prr, sqrt(as.numeric(as.character(n)))*as.numeric(as.characte
 prr$facets <- reorder(prr$facets, prr$effect, median)
 
 prr.actual <- subset(prr, pvalue==obs_pvalue)
+prr.min<-ddply(prr, "pic_name", summarise,
+               lpvalue=min(lpvalue),
+               rfreq=rfreq[order(lpvalue)[1]],
+               facets=facets[1],
+               rep=rep[1])
 
 p1 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) <= 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
-  #  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) > 10)) + 
-  #  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) > 10)) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) <= 10)) + 
+  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) <= 10)) +
   xlab(" ") + ylab("Relative frequency of picks") +
   geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
   geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
@@ -554,8 +559,8 @@ ggsave(file="../images/p_val_log_counts-a.pdf", height=4, width=12)
 
 p2 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) > 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
-#  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) > 10)) + 
-#  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) > 10)) +
[…]

### 2012-07-18 paper_jasa.tex
figures 10 and 11 - changed layers

@@
 This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
-This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$.
-For visual test this is not common.
+This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
+ % for which researchers get confused whether to reject or 'not reject' $H_0$. 
+ For visual tests this is not common.
 
@@
 
-We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}.
-The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
+%We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}. 
+%The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
 For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2.
 This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.
-For experiment 3, we see all the visual $p$ values are almost zero no matter what the conventional $p$-values are.
-This is expected since conventional test fails to reject $H_0$ where we should reject and visual test does the job nicely.
+
+For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are.
+This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
+% and the visual test does the job nicely.
 
@@
 
-One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
+\noindent One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
 In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
@@
 The dot represents this point, and the lines connect vertically with 0 count.
-Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
+Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.
+
+%Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
 Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
-(Empty cells indicate no lineup shown with this combination.) Red indicates the plot with the lowest $p$-value in the lineup.
-(Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.) In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.
-This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small.
+(Empty cells indicate no lineup shown with this combination.)
+
+Red indicates the plot with the lowest $p$-value in the lineup.
+Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.
+In both experiments  people tended to select the plot with the lowest $p$-value with an increase in effect size.
+ %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
  The results are clearer for experiment 2, that used a continuous covariate.
-But even when they didn't pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
+But even when subjects did not pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
 So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
@@
 
-There are some noticeable exceptions to this rule.
-In experiment 1, when $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.
-Same experiment, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelming chose a plot with much larger $p$-value.
+There are some noticeable exceptions to this rule. In experiment 1, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelmingly chose a plot with much larger $p$-value, similarly, for 
+parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.
 For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices \citep{zhao:2012}.
-(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
+%(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
 

### 2012-07-18 R/paper_jasa.R

@@ -547,11 +547,11 @@ prr.min<-ddply(prr, "pic_name", summarise,
 
 p1 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) <= 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
+  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) <= 10)) + 
   geom_point(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) <= 10)) +
   xlab(" ") + ylab("Relative frequency of picks") +
-  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
-  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) <= 10)) +
   facet_grid(rep~facets) + scale_x_continuous(limits=c(-2.1,0.1), breaks=c(-2,-1,0), labels=c("0.01","0.1","1")) + 
   opts(legend.position="none") + scale_y_continuous(limits=c(0,1))
 p1
@@ -559,11 +559,11 @@ ggsave(file="../images/p_val_log_counts-a.pdf", height=4, width=12)
 
 p2 <- qplot(lpvalue, rfreq, geom="point", data=subset(prr, as.numeric(facets) > 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0)) +
+  geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) > 10)) +
+  geom_point(mapping=aes(xend=lpvalue, yend=0), colour="turquoise", data=subset(prr.actual, as.numeric(facets) > 10)) +
   geom_segment(mapping=aes(xend=lpvalue, yend=0), colour="hotpink", data=subset(prr.min, as.numeric(facets) > 10)) + 
[…]

### 2012-07-18 paper_jasa.tex
done proof-reading

@@
 \subsection{Model fitting}
-\noindent For each parameter combination, the effect is computed as
+ For each parameter combination, the effect is computed as
 
@@
 
-\noindent Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.
+ Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.
 Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
@@
 %Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
-\noindent 
 Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
@@
 
-\noindent One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
+ One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
 In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
@@
 parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices \citep{zhao:2012}.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
 %(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
 
-Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
+%Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
 Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
@@
 
-Visual inference is like other types of computationally intensive test procedures, such as permutation tests, where the statistics from the data are compared with those from null data.
+ Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
  But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
@@
 If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
-Measuring the difference, generally, between the plots is almost impossible!
-But in the controlled setting here we can use $p$-value of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
+Measuring the difference, generally, between  plots is almost impossible.
+However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
 
 We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}).
-What we are interested in here is how this pick is affected by the distributin=on of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
+What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
 If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
@@
 Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
-In experiment 1 (boxplots) there were a lot more lineups where the actual data plot had the smaller $p$-value, but only barely.
+In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
 This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
 Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} show a little more complicated story.
-When there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as a force choose the same plot.
-when there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
+Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as one force choose the same plot.
+When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
 
@@
 
-In practice, $p$-value is not going to be a valid way to compare plots.
+In practice, the $p$-value is not going to be a valid way to compare plots.
 Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
 This is investigated in \citet{niladri:2012}.
-Here numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
+Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
 
@@
 A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
-This occurs when the null hypothesis is rejected for thw rong reason.
+This occurs when the null hypothesis is correctly rejected but for the wrong reason.
 Experiment 3 is prone to this type of error.
@@
 But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
-For the human eye this cluster of points is probably as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
-This would be considered to be a Type III error because it is not what was requested, but it still gives the correct answer.
+For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
+This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
 This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
@@
 In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be very general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
+In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
 
@@
 
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers.
-This ensures the ability to obtain a good estimate of the $p$-value for the test, and more confidence in the decision.
+Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
 
@@
 Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet the results were good.
+Yet, the results were good.
 We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it is may make statistics fun to learn.
+A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
 
@@
 We have successfully used the procedures informally for local projects.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no exising gauges of statistical significance.
+It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.

### 2012-07-18 paper_jasa.tex
typo marked

@@
 Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
-In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
+In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, \blue{but only just}.
 This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
 Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as one force choose the same plot.
+Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the \blue{subjects' as one force} choose the same plot.
 When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).

### 2012-07-18 paper_jasa.tex
comments colour removed

--- 2012-07-18 paper_jasa.tex
@@
        \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
-  %      \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1.
-\blue{Red indicates the} plot with the lowest $p$-value, and
-       \blue{blue indicates the} actual data plot. %, when it is different from that with the lowest $p$-value. 
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
+Red indicates the plot with the lowest $p$-value, and
+       blue indicates the actual data plot, when it is different from that with the lowest $p$-value. 
        The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
-       %Rows correspond to $\beta$, smallest to largest values from top to bottom.
-Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
-Empty cells indicate no lineup for that combination of parameters.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value.
        }
@@
    \centering
-%       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
        \includegraphics[width=0.75\textwidth]{p_val_log_counts2-a.pdf}       
        \includegraphics[width=0.75\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2.
-\blue{Red indicates the} plot with the lowest $p$-value.
-%, and \blue{blue indicates} the actual data plot, when it is different from that with the lowest $p$-value.
-       %Rows correspond to $\beta$, smallest to largest values from top to bottom.
-Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate.
-%       
+       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2.
+Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
 Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
@@
 Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as one force choose the same plot.
+Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
 When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).

### 2012-07-20 paper_jasa.tex
discussion on chapters

@@
 Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
-Section \ref{sec:simulation} describes the results of the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and an analysis of the resulting data.
-The last section describes the next steps in this research.
+\blue{Section \ref{sec:simulation} describes the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents the results of the experiments with analysis of the resulting data.} The last section describes the next steps in this research.
 
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 
@@
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and
-       blue indicates the actual data plot, when it is different from that with the lowest $p$-value. 
-       The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
-       }
+Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
+The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
        \label{fig:P-val_log}
@@
    \centering
-       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-a.pdf}       
-       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-b.pdf}
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-a.pdf}       
+       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-b.pdf}
        \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2.

### 2012-07-22 paper_jasa.tex
typo fixed
removed tabular form of model description

@@
 
-If there is evidence for individual skills influencing the probability $p_i$, then \green{$X_i \sim Binom_{1, p_i}$} and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
+If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p_i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
 This makes the distribution of $X$ a Poisson-Binomial by definition (see \citet{butler93} for details).
-\green{Ways to estimate $p_i$ will be be discussed in the following sections.}
+Ways to estimate $p_i$ will be  discussed in the following sections.
 %one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
@@
 
-\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and
-
-\begin{tabular}{lp{5.5in}}
-$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$.
+\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
 Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g.
 effect size or difficulty level.
-\\
-$\delta$ & is a vector of parameters for fixed effects.\\
-$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
-$\tau$ & is a vector of independent normally distributed random variables $\tau_{\ell i}$, with  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$.
+%
+$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
+$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
-\end{tabular}
+Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
 
@@
 where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$.
-$n_c$ is the size of thecontaminated data.
+$n_c$ is the size of the contaminated data.
 For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points.
@@
 
-The question that is posed to observers can affect the abililty to evaluate lineups.
+The question that is posed to observers can affect the ability to evaluate lineups.
 In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.

### 2012-07-23 paper_jasa.tex
definition p-value

@@
 Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
-\blue{Section \ref{sec:simulation} describes the three simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents the results of the experiments with analysis of the resulting data.} The last section describes the next steps in this research.
+Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data.
+%The last section describes the next steps in this research.
 
@@
 \begin{dfn}\label{dfn:pvalue}
-The $p$-value of a lineup of size $m$ is  estimated as %given as
+The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as
 
@@
 \] 
-Here $F_{X, \theta}$ is the distribution of $X$, and this depends on which hypothesis is true.
-Under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
+Here $F_{X, \theta}$ is the distribution of $X$ and $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. Note that the distribution  $F_X$ depends on which hypothesis is true:
+under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
 \[
-\widehat{Power}_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
+Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
 \]
-where $x_\alpha$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
+%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
 If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.

### 2012-07-29 paper_jasa.tex
edited as suggested by Hal

@@
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
+Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
 
@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ \blue{(cases 1-4)}.
 To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
@@
 
-\subsection{Continuous covariate}
+\subsection{Continuous covariate}  \label{sec:continuous}
 
@@
 One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?}
+Which plot is the most different from the others, in the sense that there is the steepest slope?
+(The position of the actual data plot is provided in Section \ref{sec:continuous}.)}
        \label{fig:test_continuous}
@@
 
-\subsection{Contaminated data}
+\subsection{Contaminated data} \label{sec:contamination}
 
@@
 One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?}
+Which plot is the most different from the others, in the sense that there is the steepest slope?
+\blue{(The position of the actual data plot is provided in Section \ref{sec:contamination}.)}}
        \label{fig:test_contaminated}

### 2012-07-30 paper_jasa.tex
figure 5 answer reference added

@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ \blue{(cases 1-4)}.
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
 To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
@@
 
-An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? %Plot 7 is the actual data plot.
+An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? The actual data plot location is 7.
 
@@
 Which plot is the most different from the others, in the sense that there is the steepest slope?
-\blue{(The position of the actual data plot is provided in Section \ref{sec:contamination}.)}}
+(The position of the actual data plot is provided in Section \ref{sec:contamination}.)}
        \label{fig:test_contaminated}

### 2012-08-07 cover-letter.tex
cover letter

\documentclass{letter}
\usepackage{graphics}
\usepackage{fullpage}
\topmargin=-0.2in
%\headheight=0in
\setlength\oddsidemargin{-0.2in}
\textheight=8.5in
\textwidth=7in
\def\apl#1#2{
\begin{letter}{#1}
\signature{Mahbubul Majumder, Heike Hofmann and Dianne Cook\\Department of Statistics\\Iowa State University 
\\Ames, Iowa 50011-1210\\ Phone (515) 294 3440\\{\tt mahbub@iastate.edu}}
\opening{{Dear #2,}}

Please find enclosed a manuscript  ``Validation of Visual
Statistical Inference, Applied to Linear Models'' which we are
submitting for consideration for publication in the
Theory and Methods section of the Journal of American Statistical
Association.

This is the second foundational paper describing research which bridges
graphics, exploratory data analysis/data mining, and statistical
inference. The first paper by Buja, et al (2009) RSS Phil. Trans. A
set up the conceptual foundation. This paper clarifies and defines the
approach and validates it with a user study. There are two more papers
published on the topic, one that puts the ideas in the information
visualization context that won the best conference paper at InfoVis
2010, and another that explains how the method can be used to choose
the best graphic for a particular purpose, that is to be presented at
InfoVis 2012 in October.

Since theory and methods section of the journal publishes graphical
methods in statistics, we believe it is a potentially interesting
paper for the JASA readership. One concern we have is that would be
considered to be long by JASA standards (28 pages, single spacing) and
we would not find it easy to shorten it.

Thank you for your consideration of our work.

\closing{Sincerely yours,
}
\end{letter}}

\begin{document}

{{\pdfimage width 4in {letterhead.jpg}}}
\apl{Professor Xuming He \\Theory and Methods Editor\\Journal of American Statistical Association}{Professor He}

\end{document}

### 2012-08-08 cover-letter.tex
cover letter

@@
 \begin{letter}{#1}
-\signature{Mahbubul Majumder, Heike Hofmann and Dianne Cook\\Department of Statistics\\Iowa State University 
+\signature{Mahbubul Majumder\\
+with Heike Hofmann and Dianne Cook\\Department of Statistics\\Iowa State University 
 \\Ames, Iowa 50011-1210\\ Phone (515) 294 3440\\{\tt mahbub@iastate.edu}}
@@
 
-Since theory and methods section of the journal publishes graphical
-methods in statistics, we believe it is a potentially interesting
-paper for the JASA readership. One concern we have is that would be
+The JASA instructions for authors say
+
+\begin{quote}
+``Authors are reminded that, although statistical theory has roots in mathematics, the phrase theory and methods should be interpreted broadly to include all techniques relevant to statistics and probability.
+This may include computational and graphical methods as well as more traditional mathematical methods.''
+\end{quote}
+We believe this paper is of interest to  the JASA readership. One concern we have is that it may be
 considered to be long by JASA standards (28 pages, single spacing) and
@@
 {{\pdfimage width 4in {letterhead.jpg}}}
-\apl{Professor Xuming He \\Theory and Methods Editor\\Journal of American Statistical Association}{Professor He}
+\apl{Professor Xuming He and Jun Liu\\Theory and Methods Editors\\Journal of American Statistical Association}{Professors He and Liu}
 

### 2012-08-08 paper_jasa.tex
double spacing and plot of residual vs xj discussion

@@
 % \usepackage{pifont} % this package is used to print check mark \checkmark
-%\linespread{2}
+% \linespread{1.6} % factor 1.6 = double space
+
+\usepackage{setspace}
+\doublespacing
 
@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
+\blue{For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).} To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
 For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.

### 2012-08-09 paper_jasa.tex
buja reference added

--- 2012-07-30 paper_jasa.tex
@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4).
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).
 To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).

### 2012-08-09 paper_jasa blinded.tex
blinded version of the paper

--- 2012-08-09 paper_jasa.tex
@@
 %\usepackage[dvips]{graphics}
-\usepackage{hyperref}
+
+%\usepackage{hyperref} % for hyper reference
 
@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{
-{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Full Professor, Department of
-Statistics and Statistical Laboratory, Iowa State University,
-Ames, IA 50011-1210. This research is supported in part by the
-National Science Foundation Grant \# DMS 1007697.}}
+%\author{
+%{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+%\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of
+%Statistics and Statistical Laboratory, Iowa State University,
+%Ames, IA 50011-1210. This research is supported in part by the
+%National Science Foundation Grant \# DMS 1007697.}}
 \date{\vspace{-.7in}}
@@
 parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012:blinded}.
 %(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
@@
 Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
-This is investigated in \citet{niladri:2012}.
+This is investigated in \citet{niladri:2012:blinded}.
 Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.

### 2012-08-09 R/paper_jasa.R

@@ -37,9 +37,20 @@ clean_data <- function(raw_dat){
   return(subset(cleaned_dat,indx_dup))
 }
 
-dat1 <- clean_data(raw.dat1)
-dat2 <- clean_data(raw.dat2)
-dat3 <- clean_data(raw.dat3)
+indx1 <- c(3,6,7,8,9,17,18,20)
+indx2 <- c(4,7,8,9,10,17,18,20)
+
+dat1 <- clean_data(raw.dat1)[,-indx1]
+dat2 <- clean_data(raw.dat2)[,-indx1]
+dat3 <- clean_data(raw.dat3)[,-indx2]
+
+
+# saving cleaned data as suplementary materials
+write.csv(dat1, file='../submission_initial/suplementary/data/data_turk1.csv',row.names=F)
+write.csv(dat2, file='../submission_initial/suplementary/data/data_turk2.csv',row.names=F)
+write.csv(dat3, file='../submission_initial/suplementary/data/data_turk3.csv',row.names=F)
+
+
 
 
 # =================== mathematical test t distribution  ===================
@@ -331,7 +342,7 @@ get_predict_mixed <- function(dat, newdat, intercept=F){
   fit.mixed <- fit_model(dat)
   X <- newdat$effect
   if(intercept){
-    if (dat$experiment=="turk3"){
+    if (dat$experiment[1]=="turk3"){
       subject <- apply(ranef(fit.mixed)[[1]],1,paste,collapse="_")
       } else subject <- ranef(fit.mixed)[[1]][,1]
     d <- data.frame(expand.grid(effect=X, subject=subject))
@@ -758,100 +769,6 @@ ggsave(file="../images/pval_difference.pdf", height=4, width=8)
 dd <- subset(pval.dat, pic_name %in% subset(pval.diff,pval_diff >.8)[,1])
 
 
-# =================== Turk1 data analysis  ================================
-
-# ------function to calculate UMP power -----------------------
-
-# source("calculate_ump_power.R")
-calculate_ump_power1(3, 100, 5)
-
-# ---------------------- fitting loess smoother  ------------
-
-get_smooth_power <- function(dat.n, test = "Empirical") {
-    betas <- seq(0.01, 16, by = 0.2)
-    dat_smooth <- NULL
-    for (s in c(5, 12)) {
-        for (n in c(100, 300)) {
-            dats <- subset(dat.n, sigma == s & sample_size == n)
[…]

### 2012-08-15 paper_jasa blinded.tex
reference updated as suggested by JASA

--- 2012-08-09 paper_jasa.tex
@@
 
-\usepackage{setspace}
+\usepackage{setspace, url}
 \doublespacing
@@
 
+\pdfminorversion=4 % as instructed by JASA file upload
+
 \begin{document}
@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{
-{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Full Professor, Department of
-Statistics and Statistical Laboratory, Iowa State University,
-Ames, IA 50011-1210. This research is supported in part by the
-National Science Foundation Grant \# DMS 1007697.}}
+%\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+%\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
 \date{\vspace{-.7in}}
@@
 parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
+For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
 %(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)

### 2012-08-15 paper_jasa.tex
reference updated as suggested by JASA

@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{
-{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Full Professor, Department of
-Statistics and Statistical Laboratory, Iowa State University,
-Ames, IA 50011-1210. This research is supported in part by the
-National Science Foundation Grant \# DMS 1007697.}}
+\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
 \date{\vspace{-.7in}}

### 2012-08-15 submission_initial/cover-letter.tex
initially submitted materials

--- 2012-08-08 cover-letter.tex
@@
 %\headheight=0in
+\pdfminorversion=4 % as instructed by JASA file upload
 \setlength\oddsidemargin{-0.2in}
@@
 We believe this paper is of interest to  the JASA readership. One concern we have is that it may be
-considered to be long by JASA standards (28 pages, single spacing) and
+considered to be long by JASA standards (38 pages, double spacing) and
 we would not find it easy to shorten it.

### 2012-08-15 submission1_initial/paper_jasa.tex
submission after fixing reference

--- 2012-08-15 paper_jasa.tex
@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
-This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
+\author{
+{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of
+Statistics and Statistical Laboratory, Iowa State University,
+Ames, IA 50011-1210. This research is supported in part by the
+National Science Foundation Grant \# DMS 1007697.}}
 \date{\vspace{-.7in}}
@@
 parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
+For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
 %(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)

### 2012-08-15 paper_jasa.tex
updated raference and abstract
tweaked the abstract

@@
 \graphicspath{{images/}}
-\usepackage{color}
-\newcommand{\blue}[1]{{\color{blue} #1}} %MM
-\newcommand{\red}[1]{{\color{red} #1}} 
-\newcommand{\green}[1]{{\color{green} #1}} %DC
-
-\definecolor{orange}{rgb}{1,0.5,0}
-\newcommand{\hh}[1]{{\color{orange} #1}} %HH
 
@@
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently, there were no formal visual methods in place for determining statistical significance of findings.
-This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings.
-In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
+A new method, called the lineup protocol, enables determining the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.
+In this paper inferential methods for statistical graphics is developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
+This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
+Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional are violated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult of tasks.
 

### 2012-08-15 submission2_reference_updated/paper_jasa blinded.tex
submitted version after fixing reference

--- 2012-08-15 paper_jasa.tex
@@
 \title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+%\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+%\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}

### 2012-09-05 submission2_reference_updated/R/calculate_ump_power.R
complete version of submission 2

@@ -1,0 +1,59 @@
+calculate_ump_power1 <- function(beta, n, sigma) {
+  # n <- 100
+  # sigmasq <- 12^2
+  alpha <- 0.05
+  sigmasq <- sigma^2
+  beta_not <- 0
+  se_beta <- sqrt(sigmasq/(n * (0.5^2)))  # refer to docs derivation of power
+  mu <- beta/se_beta
+  alpha <- alpha/2
+  t_n <- qt(p = 1 - alpha, df = n - 3)
+  res <- pt(q = -t_n, df = n - 3, ncp = mu) - pt(q = t_n, df = n - 3, ncp = mu) + 1
+  return(res)
+}
+
+calculate_ump_power2 <- function (beta, n, sigma){
+  alpha <- 0.05
+  beta_not <- 0
+  df <- n-2 # two parameters
+  x <- subset(read.csv("../data/Xdata.csv"),N==n)[,1]
+  ssx <- sum((x-mean(x))^2)
+  se_beta <- sigma/sqrt(ssx) 
+  mu <- beta/se_beta
+  t_n <- qt(p=1-alpha/2,df=df)
+  res <- pt(q=-t_n, df=df, ncp=mu)-pt(q=t_n, df=df, ncp=mu)+1
+  return(res)
+}
+
+getx <- function(n){
+  x1 <- rnorm(n,0,1)
+  nc <- 15*n/100
+  x2 <- rnorm(n=nc,mean=-1.75, sd=1/3)
+  return(c(x2,x1))
+}
+
+calculate_ump_power3 <- function (beta, n, sigma,x){
+  alpha <- 0.05
+  sigmasq <- sigma^2
+  beta_not <- 0
+  df <- n-1 # one parameter(s)
+  se_beta <- sqrt(sigmasq/sum((x-mean(x))^2))
+  mu <- beta/se_beta
+  alpha <- alpha/2
+  t_n <- qt(p=1-alpha,df=df)
+  res <- pt(q=-t_n, df=df, ncp=mu)-pt(q=t_n, df=df, ncp=mu)+1
+  return(res)
+}
+
+get_ump_power_by_effect <- function(){
+  set.seed(2035)
+  ump <- rbind(data.frame(experiment="Experiment 1",effect=seq(0,16, by=.2),
+                pow= calculate_ump_power1(beta=seq(0,16, by=.2)/10, n=100, sigma=1)),
+               data.frame(experiment="Experiment 2",effect=seq(0,5.5, by=.2),
[…]

### 2012-09-05 paper_sociology.tex

--- 2012-08-15 paper_jasa.tex
@@
 %\documentclass{article}
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 \usepackage{times}
@@
 
-\usepackage{setspace, url}
-\doublespacing
+%\usepackage{setspace, url}
+%\doublespacing
 
@@
 % Article top matter
-\title{Validation of Visual Statistical Inference, Applied to Linear Models}
+\title{Impact of Demographics and Skills of the Observer on Visual Inference }
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
@@
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
-\date{\vspace{-.7in}}
+\date{\vspace{-.5in}}
 %\date{\today}  %\today is replaced with the current date
@@
 \begin {abstract}  
-Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-A new method, called the lineup protocol, enables determining the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.
-In this paper inferential methods for statistical graphics is developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
-This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional are violated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult of tasks.
-
-{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, data mining, visualization, exploratory data analysis} 
+Visual Inference is dependent on the careful evaluation of the lineups by individual observers.
+Each individual is different in their cognitive phycology and judiciousness which can affect the power of visual inference.
+To estimate the power of visual inference this can be controlled by getting evaluations from multiple people from a diverse population.
+But the other factor that may affect the power of visual inference is the demographics and skills of the observer.
+In this paper we examine this in details.
+The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
+
+{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, cognitive phycology, visualization, exploratory data analysis} 
 \end {abstract}
@@
 \section{Introduction}
+
+The fundamental concept of visual statistical inference is introduced by \citet{buja:2009} and later these concepts are validated by \citet{majumder:2012}.
+
+\subsection{Visual Inference} A short introduction of visual inferential procedure.
+\subsection{Estimation of Power of Visual Inference} Discussion of the methods for estimating the power of visual inference \citep{majumder:2012}.
+
+\section{Factors Affecting the Power of Visual Inference} A brief description of the factors that may affect the power of visual inference.
+
+\subsection{Choice of Visual Test Statistic} Discuss the effective feature of visual test statistics \citep{heike:2012}.
+May present some results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
+\subsection{Question that Human Observer Faces}
+\subsection{Observer Personality}
+\subsection{Visual Perception of Human Eye} Present the results we have on eye tracking experiment \citep{zhao:2012}.
+\subsection{Demographics of Observer}
+\subsection{Skills of Observer}
+
+\section{Simulation Experiment}
+
+\subsection{Experiment Setup}
+\subsection{Diversity of the Experimental Subjects}
+
+\section{Results}
+
+\subsection{Learning Trend of the Observer}
+
+\section{Conclusion}
+
+\bibliographystyle{asa}
+\bibliography{references}
+
+\end{document}
 

### 2013-02-19 paper_jasa.tex
reduce pages by restructuring figures and moving proof of lemma to supplimentary

@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional are violated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult of tasks.
+Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional test are violated, and there may be some super-individuals who yield better power than the conventional test even in the most difficult tasks.
 
@@
 
-\begin{proof}
-%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
-
- By definition
-
-$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-
-%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
-%, with non-centrality parameter $\delta$. 
-\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
-\begin{eqnarray}\label{dist_p}
-F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
-  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
-  &=&%\left\{ \begin{array}{ll}
-          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
- %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%       \end{array} \right.     
-\end{eqnarray}
-
-%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
+\begin{proof} The proof and details of the lemma is attached in the supplementary documents. 
+\end{proof}
+
+%\begin{proof}
+%%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
+%
+% By definition  
+%
+%$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
+%
+%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
+%%, with non-centrality parameter $\delta$. 
+%\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
+%\begin{eqnarray}\label{dist_p}
+%F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
+%  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
+%  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
+%  &=&%\left\{ \begin{array}{ll}
+%          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
+% %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
+%%       \end{array} \right.     
+%\end{eqnarray}
+%
+%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
 As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-
-Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots.
+%
+%
+%
+%Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots.
 Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
 The minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function
-\[
-F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
-\]
-
-\noindent Thus
-\begin{eqnarray*}
-P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
-&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
-&=&  E\left[ (1 - p_D)^{m-1}\right].
-\end{eqnarray*}
-
-%\noindent This proves the statement above. 
-%We further see:
 %\[
-%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
+%F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
 %\]
-
-% If the density of $p$-value is very right skewed, the expectation term would be large.
+%
+%\noindent Thus
+%\begin{eqnarray*}
+%P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
+%&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
+%&=&  E\left[ (1 - p_D)^{m-1}\right].
+%\end{eqnarray*}
+%
+%
+%%\noindent This proves the statement above. 
+%%We further see:
+%%\[
+%%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
+%%\]
+%
+%
+%% If the density of $p$-value is very right skewed, the expectation term would be large.
 The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in.
@@
 )},
-%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
-%}
-%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$.
-
-%
-%
-%
-%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
-%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
-%\begin{eqnarray*}
-%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
-%\end{eqnarray*}
-%which, in particular, is independent of $p_{0,i}$ for all $i$.
-%
-%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
+%%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
+%%}
+%%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$. 
+%
+%%
+%%
+%%
+%%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
+%%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
+%%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
+%%\begin{eqnarray*}
+%%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
+%%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
+%%\end{eqnarray*}
+%%which, in particular, is independent of $p_{0,i}$ for all $i$.
+%%
+%%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
 Further we will assume that all observers  have the same ability in identifying the data plot.
-%
-%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
-%
-%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
-%\[
-%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
-%\]
-\end{proof}
+%%
+%%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
+%%
+%%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
+%%\[
+%%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
+%%\]
+%\end{proof}
 
@@
 
-Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
+Figure \ref{fig:P-val_log2} shows the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
 In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically.
@@
 
+%\begin{figure*}[hbtp]
+%   \centering
+%      \includegraphics[width=0.8\textwidth]{p_val_log_counts-a.pdf}
+%       \includegraphics[width=0.8\textwidth]{p_val_log_counts-b.pdf}
+%       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
+Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
+The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
+%       \label{fig:P-val_log}
+%\end{figure*}
+
 \begin{figure*}[hbtp]
    \centering
-      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
-       \label{fig:P-val_log}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-a.pdf}       
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2.
+      \includegraphics[width=0.8\textwidth]{p_val_log_counts-a.pdf}
+       \includegraphics[width=0.8\textwidth]{p_val_log_counts-b.pdf}   
+       \includegraphics[width=0.8\textwidth]{p_val_log_counts2-a.pdf}       
+       \includegraphics[width=0.8\textwidth]{p_val_log_counts2-b.pdf}
+       \caption{ Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all individual lineups of both experiment 1 and 2.
 Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
@@
 
-We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}).
+We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
 What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
@@
 Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
+Figure \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
 When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).

### 2013-02-19 supplementary.tex
reduce pages by restructuring figures and moving proof of lemma to supplimentary

--- 2012-08-15 paper_jasa.tex
@@
 
-\usepackage{setspace, url}
+\usepackage{setspace}
 \doublespacing
@@
 
-% Article top matter
-\title{Validation of Visual Statistical Inference, Applied to Linear Models}
-\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
-This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
-\date{\vspace{-.7in}}
-%\date{\today}  %\today is replaced with the current date
-\maketitle
-
-\begin {abstract}  
-Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently, there were no formal visual methods in place for determining statistical significance of findings.
-This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings.
-In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
-
-{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, data mining, visualization, exploratory data analysis} 
-\end {abstract}
-
-%\begin{multicols}{2}
-%\twocolumn
-
-\section{Introduction}
-
-%\checkmark
-
-Statistical graphics nourish the discovery process in data analysis by revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally by contrasting prevailing hypotheses.
-The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
-Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
-
-Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years.
-Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results.
-This body of work provides a framework for evaluating new statistical graphics.
-In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots.
-\cite{wilkinson:1999} developed a grammar of graphics that is enhanced by \cite{hadley:2009}.
-These methods provide a mechanism to abstract the way data is mapped to graphical form.
-Finally, technology advances make it simple and easy for everyone to draw plots of data, and particularly the existence of software systems, such as R \citep{R}, enable making beautiful data graphics that can be tightly coupled with statistical modeling.
-
-However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance.
-This has also been addressed in recent research.
-\citet{buja:2009} proposes a protocol that allows the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
-One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
-Figure \ref{fig:test_category} shows an example lineup.
-Which plot do you think is the most different from the others?
-(The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors conventional inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
-The null hypothesis typically derives from the task at hand, or the type of plot being made.
-The alternative encompasses all possible antitheses, all types of patterns that might be detected in the actual data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
-The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.
-
-\begin{figure}[htp]
-%\begin{figurehere}
+\begin{table}[hbtp]
+%\renewcommand\thetable{}
+\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section 5.1 of the paper.} 
 \centering
-       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
-       \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the largest shift or location difference between the boxplots?
-(The position of the actual data plot is provided in Section \ref{sec:category}.)}
-       \label{fig:test_category}
-\end{figure}
-%\end{figurehere}
-
-The protocol has only been informally tested until now.
-In the work described in this paper, the lineup protocol is compared head to head with the equivalent conventional test.
-Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
-This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference.
-The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
-The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
-Testing is done based on a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
-The results are compared with those of the conventional test.
-
-%, analysis requires plots of data.
-For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
-Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.
-
-%The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
-In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS ease the production of high quality statistical plots.
-A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
-\cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
-\blue {Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
-
-%\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
-
-%In this paper we compare the lineup protocol to conventional statistical inference, specifically applied to the regression framework.
-We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
-
-The paper is organized as follows.
-Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data.
-The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}.
-Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
-Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data.
-%The last section describes the next steps in this research.
-
-\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test}
-
-Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing.
-Both start from the same place, the same set of hypotheses.
-The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
-In the lineup protocol, the test statistic is a plot of the data.
-Here, side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
-In conventional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
-If it is extreme on this scale then the null hypothesis is rejected.
-In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
-If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.
-
-\begin{table*}[hbtp]
-\caption{Comparison of visual inference with conventional inference.}
-\centering 
-\begin{tabular}{llll} 
-\hline \hline
-  & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
+\begin{tabular}{c c c c  c c c}
+\hline\hline
+Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
+ \cline{2-4}
+level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
 \hline
-  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta > 0$& $H_0: \beta=0$ vs $H_1: \beta > 0$\\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
-
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{lineup_category_small.pdf}} \end{center} \end{minipage} \\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
- Reject $H_0$ if & actual $T$ is extreme & actual plot is identifiable \\
+easy&100& 5&8 & 1& 12 & 1\\
+&100&12&16 &1&&\\
+&300& 5&5 &1&&\\
+&300&12&10 &1 &&\\
+\hline
+medium&100& 5&3 &203 & 9 &2\\
+&300& 5&2, 3 & 97, 1&&\\
+\hline
+hard&100&12&3, 8, 10 & 277, 126, 23& 18 &6\\
+&300& 5&1 & 371 &&\\
+&300&12&3, 5& 375, 74 &&\\
+\hline
+mixed&100& 5&1, 5, 0& 214, 2, 73 & 21 &1\\
+&100&12&1& 100& &\\
+&300& 5&0 & 73&&\\
+&300&12&7, 1& 2, 152&&\\
+\hline
+Total &&&&&60&10\\
 \hline
 \end{tabular}
-\label{tbl:compare}
-\end{table*}
-
-%Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
-
-%Power of a plot and $p$-value have been derived in Buja and Wickham as ...
-
-%for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
-
-%This section outlines the concepts of visual inference in comparison to the procedures of conventional statistical inference.
-Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
-
-In general, we define $\theta$ to be a population parameter of interest, with $\theta \in \Theta$, the parameter space.
-Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-A test statistic, $T(y)$, is a function that maps the sample into a numerical summary, that can be used to test the null hypothesis.
-The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
-$T(y)$ is assessed relative to null values of this statistic $T(y_0)$ which would be possible values of $T$ if $\theta \in \Theta$.
-%Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
-For visual inference, unlike in the conventional hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
-When the alternative hypothesis is true, it is expected that the plot of the actual data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
-We will call a plot with this property a {\it visual statistic} for $\theta$.
-More formally,
-
-\begin{dfn} \label{dfn:test}
-A \textbf{visual test statistic}, $T(.)$, is a function of a sample that produces a plot.
-$T(y)$ maps the actual data to the plot, and we call this the \textbf{(actual) data plot}, and $T(y_0)$ maps a sample drawn from the null distribution into the same plot form.
-These type of plots are called \textbf{null plots}.
-\end{dfn}
-
-\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
-\\
-\noindent
-The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
-
-\begin{dfn}\label{dfn:lplot}
-A \textbf{lineup} is a layout of $m$ randomly placed visual statistics, consisting of 
-\begin{itemize}\itemsep-3pt
-\item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
-\item the test statistic, $T(y)$, produced by plotting the actual data, possibly arising from $H_1$.
-\end{itemize}
-\end{dfn}
-
-\noindent The $(m-1)$ null plots are % can be 
-members of the sampling distribution of the test statistic assuming that the null hypothesis is true.
-If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e.
-the plot of the data, that makes it visually distinguishable from the null plots.
-A careful visual inspection of the lineup by %\strikethrough{an}
-independent observers follows;  observers are asked to point out the plot most different from the lineup.
-If the test statistic is identified in the lineup, this is considered as evidence against the null hypothesis.
-%If it is not identified the conclusion is to not reject the null hypothesis.
-
-This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
-For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
-Under the null hypothesis $X \sim \text{Binom}_{K, 1/m}$,  therefore:
-
-\begin{dfn}\label{dfn:pvalue}
-The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as
-
-\begin{equation}\label{binom}
-P(X \ge x) = 1 - \text{Binom}_{{K, 1/m}} (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i}
-\end{equation}
-
-\noindent with $X$ defined as above, and $x$ is the number of observers selecting the actual data plot. 
-\end{dfn}
-\noindent Note that for $x=0$ the $p$-value becomes, mathematically, equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
-being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
-
-Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
-It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
-
-\begin{table}[htp]
-\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
-\begin{center}
-\scalebox{.9}{
-\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
-$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline %\cline{1-3}\cline{4-6}\cline{7-9}\cline{10-12}\cline{17-19}
-1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
-&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
-&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
-&&&  &    &           &   &        && 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
-\end{tabular}
-}
-\end{center}
-\label{pvalue}
+\label{tbl:dist_lineup1}
 \end{table}
 
-%\begin{dfn}\label{dfn:test}
-%\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
-%The \textbf{visual test}, $V_{\theta}$, is defined to be 
-%\begin{itemize}\itemsep-3pt
-%\item \textbf{Reject} $H_0$ if `enough' observers select the actual data plot in the lineup of $m$ plots, and
-%\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
-%\end{itemize}
-%where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
-%\end{dfn}
-
-%\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
-
-%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.}
-
-\begin{dfn}\label{dfn:test}
-The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
-\begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
-\item \textbf{Fail to reject} $H_0$  otherwise. 
-\end{itemize}
-where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
-\end{dfn}
-
-\noindent
-Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:
-
-\begin{dfn}\label{dfn:error}
-The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.\\ \noindent
-The \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
-\end{dfn}
-
-%\green{Oh, we are using observed and observer, which might be confusing.
-May need to change observed data plot to something else.
-Will think about this.
-Will use ``actual'' for now.} \hh{I think that change is done}
-
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$. %, with an exact $x_{\alpha}$. 
-For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
-Type II error is harder to calculate, as is usually the case.
-In visual inference, individual abilities need to be incorporated to calculate Type II error.
-%This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
-Here, we need to estimate the probability that an observer sees the actual data plot as different, when it really is different.
-This involves understanding the individual's visual skills.
-Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
-Let $p_i$ be the probability that individual $i$ picks out the actual data plot from the lineup.
-If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
-
-If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p_i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
-This makes the distribution of $X$ a Poisson-Binomial by definition (see \citet{butler93} for details).
-Ways to estimate $p_i$ will be  discussed in the following sections.
-%one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
-
-%\green{WHY??? And how does this relate to the logistic model below?} 
-%This leads to the definition of the power of the visual test, as follows:
-
-%The lineup plot can be evaluated by one or more individuals.
-When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
-
- %Notice that when $N=1$, this $p$-value is $\frac1m$.
-
-%\subsection{Power}
-
-%Next, we will develop power theoretically and then relate it to the empirical results.
-%We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
-we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$.
-In a next step we will then relate the theoretical results to our empirical findings.
-
-%\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
-- why do we need them?
-wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?
-- $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
-%\begin{dfn} \label{dfn:power}
-%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-%    \begin{equation*}
-%      \text{Power}_V(\theta)= 
-%        \begin{cases} 
-%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
-%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-%
-%
-%\red{Alternative Power definition:}
-\begin{dfn} \label{dfn:power}
-The \textbf{power} of a visual test, $V_{\theta}$, is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-    \begin{equation*}
-      \text{Power}_V(\theta)= Pr(\text{Reject } H_0 \mid \theta) 
-%      = P(X \ge x_{\alpha}) =
-%        \begin{cases} 
-%               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
-%              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
-%        \end{cases}
-    \end{equation*}
-\end{dfn}
-
-\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
-Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
-\[
-\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
-\] 
-Here $F_{X, \theta}$ is the distribution of $X$ and $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. Note that the distribution  $F_X$ depends on which hypothesis is true:
-under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
-\[
-Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
-\]
-%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
-If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
-Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.
-
-%\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
-
-%\red{there may be a flow in this definition.
-Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power.
-I feel like we should not define power using binomial.}
-
-Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
-Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.
-Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
-This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean.
-One is typically better than another.
-For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer.
-Power is typically used to measure this characteristic of a test.
-%Power is the probability of rejecting $H_0$ when it is not true.
-We can assess power therefore both empirically based on experimental data and through theoretical considerations.
-
-\section{Effect of Observer Skills and Lineup Size}~\label{sec:size}
-
-%\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
-Can someone please decide which to use, and make the appropriate changes?
-How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above?
-It looks like this is a repeat.
-One needs to go once notation is clarified.}
-
-%\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
-\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
-\\ Should I change everything accordingly?
-I see that we have been careful not to use $\beta$ for power!
-}
-
-\subsection{Subject-specific abilities}~\label{sec:model}
-%Calculating the power  is intricately related to the probability to identify the actual data plot from the lineup.
-
-Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i} \sim  Binom_{1, p_{\ell i}}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $1 \le \ell \le L$, and 0 otherwise.
-A mixed effects logistic regression model is used for $P(X_{\ell i} = 1) =  p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.
-
-%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
-The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
-\begin{equation} \label{eqn:mixed}
-g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
-\end{equation}
-
-\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
-Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g.
-effect size or difficulty level.
-%
-$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
-$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
-$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
-
-\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
-\begin{equation} \label{eqn:mixed_power}
-\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
-\end{equation}
-
-%\red{This model is  described in more detail in equation \eqref{eqn:mixed}.}
-
-%\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability
-%\begin{equation}\label{eqn:pvalue}
-%Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
-%\end{equation}
-%where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
-%\begin{dfn} \label{dfn:pval}
-%The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
-%where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
-Since $u$ is a discrete value with boundaries 0 and $K$,
-% we can also re-formulate the $p$-value in terms of upper and lower boundaries:
-% \begin{equation*}
-% \begin{array}{rcll}
-% P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
-%( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
-% \end{array}
-% \end{equation*}
-% }
-
-%\red{HH: Tried to re-write it to answer the question below.
-%Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
-%
-%    \begin{equation*}
-%        \begin{cases} 
-%              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
-%              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-
-%\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
-%\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
-
-%\noindent
-\subsection{Lineup size, $m$}
-
-The finite number $m-1$ of representatives of the null distribution used  as comparison against the test statistic, is a major difference between visual inference and the conventional testing.
-The choice of $m$ has an obvious impact on the test.
-
-The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper.
-They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups when they are used in practical settings.
-Two assumptions are critical:
-\begin{enumerate} \itemsep 0in
-\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
-\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
-\end{enumerate}
-Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup.
-The better suited a design is for a particular task, the higher its power will be.
-
-In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
-
-\begin{lemma}~\label{lemma}
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic.
-Suppose the associated test statistic in a conventional test is observed as  $t_{obs}$ with $p$-value $p_D$.
-
-The probability of picking the data plot from a lineup depends on the size $m$ of the lineup and the strength of the signal in the data plot. 
-Under the above assumptions, the probability is expressed as:
-\[
-P(p_D < p_0) =  E\left[ (1 - p_D)^{m-1}\right]
-\]
-where $p_D$ is the $p$-value associated with the data in the test statistic, and $p_0$ is the minimum of all $p$-values in the data going into null plots.
-\end{lemma}
+\section{Proof of the Lemma}
 
@@
 
-%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the Figure on $1/m$ versus power.
-These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
+\section{Data Cleaning}
 
-The above lemma allows two immediate conclusions for the use of lineups.
-The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$.
-%This may seem counterintuitive, but it makes sense.
-%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
-Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$.
-This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.
-Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
-It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.
-% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
-%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.
+Following six approaches have been examined to clean the data.
+Table \ref{tbl:summary} displays number of subjects and there total evaluations in each screening criteria.
+For our paper we picked criteria six.
 
-%Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
-%Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
+\begin{enumerate}
+\item {\bf Include all participants} and their evaluations.
+\item Exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
+\item Exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
+\item Include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing.
+\item Include participants' records, if at least {\bf 50\% of all very easy (p-value $<$ 0.0002) lineups are correct} .
+\item Use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
+If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
+\end{enumerate}
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+The power curves obtained by fitting mixed model to the data for all the experiments applying different screening criteria are shown in Figure \ref{fig:power_screening_subject}.
+This shows how the screening criteria may affect the results.
+The corresponding power curves of conventional test are also shown.
+Notice that no mater what criteria we apply the result does not change much for experiment 2.
+% Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
+
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
+%       \caption{The overall power estimated from mixed model  for each screening criteria are shown. The corresponding power curve for conventional test is shown for comparison.}
+%       \label{fig:power_screening}
+%\end{figure}
+
+\begin{figure}[hbtp]
    \centering
-   \includegraphics[width=3in]{images/powerplot.pdf} 
-   \caption{Probability that the data plot has the smallest probability in a lineup of size $m$.
-With increasing $p$-value the probability drops -- when it reaches $1/m$ a horizontal line is drawn to emphasize insufficient sensitivity of the test due to the lineup size.
-   %Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
-The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true.
-}
-}
-   \label{fig:pval_power}
+       \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
+       \caption{The subject specific power estimated from mixed model for all the six screening criteria are shown. The corresponding power curves for conventional test is shown for comparison.}
+       \label{fig:power_screening_subject}
 \end{figure}
 
-%Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
-%Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
-Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$.
-We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
-%
-%For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
-%we get
-%\[
-%N_m\sim \left \{ 
-%\begin{array}{ll}
-%B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
-%B_{K, 1/m} & \text { under } H_0
-%\end{array}
-%\right .
-%\]
-%
-%By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
-%%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
-%as:
-%\begin{equation}\label{eqn:power_estimate}
-%%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
-%\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
-%\end{equation}
-%
-%Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
-While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
-%\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
-We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact.
-%Both of these topics are further highlighted in the application section.}
-
-%Picking out the data plot is still a measure of how different the data plot is compared to the null plots, and therefore is a test statistic of the visual test.
-
-%This allows us to compute this value independently of the parametric setting. We therefore define the {\it signal strength} of a  visual test as the following:
-%
-%\begin{dfn} \label{dfn:signal}
-%The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
-%\end{dfn}
-%
-%In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
-%\[
-%\sigma_T^{m-1} = \frac{n_m}{K}.
-%\]
-%
-%In a parametric setting, signal strength of a test coincides with the power of the test.
-
-%Theoretical power for the regression parameters is derived in the next section.
-%\subsection{Estimating Power from Lineups}
-%
-%We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
-This gives an estimated power of
-%\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
-
-%From model \ref{eqn:mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
-%the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
-%%Thus we have the overall probability of success ($\pi$ or power) 
-%for specified sample size ($n$) and  variance ($\sigma$)  as 
-%
-%\begin{equation}\label{eqn:power} 
-%Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
-%\end{equation}
-%
-%The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
-combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
-Using equation \eqref{eqn:mixed}, we then have the power as
-%
-%\begin{equation}\label{eqn:effect} 
-%Power(E) = \pi=Pr(Y=1| E) 
-%\end{equation}
-
-\section{Application to Linear Models} \label{sec:regression}
-
-To make these concepts more concrete consider how this would operate in the linear models setting. Consider a linear regression model 
-\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
-\end{equation}
-where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.
-
-In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g.
-\citet{cook:99}).
-Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
-%These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
-For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.
-
-%Note that some charts can be associated with several different testing situations. In our paper, we pick the examples for situation 2 and 3 as candidates.
-
-\begin{table*}[hbtp] 
-\caption{Visual test statistics for testing hypotheses related to the model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
-\centering 
-\begin{tabular}{m{0.5cm}m{3cm}m{2cm}m{3cm}m{5.5cm}} 
-\hline\hline 
-Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
+\begin{table}[hbtp]
+\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
+Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information. }
+\begin{center}
+\begin{tabular}{rrrrr|rrrr|rrrr}
   \hline
-1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid.
-For lineup plot, we simulate data from fitted null model.
-\\
-2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
-For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
-5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
-For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
-For lineup plot, we simulate data from standard normal.
-\\
-7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
-For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
-\\
-8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data.
-For lineup plot, we simulate data from fitted model.
-\\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
+Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
+ \cline{2-5}  \cline{6-9}   \cline{10-13}
+criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
   \hline
-\end{tabular} 
-\label{tbl:stat_multiple} 
-\end{table*}
-
-Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are different, the  residuals from fitting the null model should show a significant difference between the two groups.
-For a visual test, we draw boxplots of the residuals conditioned on the two levels of $X_k$.
-If $\beta_k\ne 0$ the boxplots should show a vertical displacement.
-
-%A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
-The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
-The test statistic, the plot containing the actual data, is randomly placed among these null plots.
-If the test statistic is identifiable the null hypothesis is rejected.
-% with a $p$-value of at most 0.05.
-
-%\subsection{Expected power}
-
-% Di: This is specific for the regression model
-%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
-%%Now consider estimating the power of the visual test. 
-%%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in conventional test be $t_{obs}$ with $p$-value $p_D$.
-By definition we have
-%%
-%%$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-%%
-%%\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
-%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
-%%\begin{eqnarray}\label{dist_p}
-%%F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-%%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
-%%  &=&\left\{ \begin{array}{ll}
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%%       \end{array} \right.     
-%%\end{eqnarray}
-%%
-%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
-Suppose  $p_0=\min_i(p_{0,i})$.
-Thus $p_0 \sim \text{Beta}(1,m-1)$.
-
-The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
-It forms the benchmark upon which we evaluate the visual test.
-To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
-This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}.
-Thus the expected power of the visual test in this scenario is
-
-\begin{equation}\label{power_exp} 
-   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
-\end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Notice that the expected power of the visual test is %almost as good as
-close to the power of the conventional test.
-
-%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
-%\begin{thm}\label{thm:power}
-% equation \eqref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
-%\end{thm}
-% 
-%{ \em proof}
-%
-%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
-%\begin{eqnarray*}
-%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
-%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
-%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
-%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
-%
-%Thus expected power in equation \eqref{power_exp} be
-%\begin{eqnarray*}
-% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
-%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
-%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
-%  &=& E(1-p_D)^{m-1}   \\
-%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{.6}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
-       \label{fig:power_expected}
-\end{figure}
-
-% \pagebreak
-
-% -------- Heike's writeup added here
-
-%\subsection{Some Power Considerations}
-
-%\subsection{Expected number of choices}
-%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
-%\[
-%E[X] = (m-1)(1-E[p_D]).
-%\]
-%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-%This should allow us to infer some of the signal strength $p_D$ as
-%\begin{equation} \label{plot_signal}
-%E[p_D] = 1 - E[X]/(m-1),
-%\end{equation}
-%i.e.
-by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
-%
-
-\section{Human Subjects Experiments with Simulated Data} \label{sec:simulation}
-
-Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.
-The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
-(Data and lineups used in the experiments are available in the supplementary material.)
-
-After many small pilot studies with local personnel, it was clear that some care was needed to set up the human subjects experiments.
-It was best for a observer or a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
-The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects, for each of the three experiments.
-
-Participants for all the experiments were recruited through Amazon's online web service, Mechanical Turk  \citep{turk}.
-A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}.
-Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice.
-Gender, age, education and geographic location of each participant are also collected.
-%The data was collected through the web.
-For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.
-
-Note that no participant was shown the same lineup twice.
-%\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
-This happens when subject accidentally use browser's ``back'' button and same data is saved twice.
-We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject.
-For this we only keep the very first response.}
-
-\subsection{Discrete covariate}\label{sec:category}
-
-The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
-Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
-The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test.
-Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
-Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
-Three replicates at each level were generated.
-These produced 60 different ``actual data sets'', and thus, 60 different lineups.
-To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
-The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3).
-The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way.
-The actual data plot was randomly placed among these null data plots to produce the lineup.
-Figure \ref{fig:test_category} is an example of one of these lineups.
-The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
-%	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
-Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
-For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
-Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
-% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}
-
-Each subject viewed at least 10 lineups with the option to evaluate more.
-%and we allowed them to provide more evaluations if they wanted. 
-Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
-For each difficulty level a specific number of lineups was randomly picked  for evaluation.
-This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$.
-To satisfy this plan we needed to recruit at least 300 subjects.
-
-\begin{table}[hbtp]
-\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
-\centering
-\scalebox{.9}{
-\begin{tabular}{c c c c c}
-% creating 10 columns
-\hline\hline
-
-% inserting double-line
-& & \multicolumn{3}{c}{Slope ($\beta$)} \\
- \cline{3-5}
-
-Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
- ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
-\\ [0.5ex]
-\hline
-% inserts single-line
-
-% Entering 1st row
-&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
-\raisebox{1.5ex}{100} &12
-& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]
-
-% Entering 2rd row
-&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
-\raisebox{1.5ex}{300} & 12
-& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
-% [1ex] adds vertical space
-\hline
-% inserts single-line
-\end{tabular}
-}
-\label{tbl:experiment_params}
-\end{table}
-
-%For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in Table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
-We collect data such that each of the parameter combinations in the Table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
-
-%\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
-This produces an empirical distribution of p-value under alternative hypothesis.
-Finally, three data sets that have 16th, 50th and 84th quantile of the p-values were taken.
-Similar procedure was used to generate 19 null plots based on the empirical distribution of p-value under null hypothesis}.
-For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
-A 1000 sets were generated for each parameter combination and the conventional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
-The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
-Additional control was applied to the 19 null plots.
-Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
-
-\subsection{Continuous covariate}  \label{sec:continuous}
-
-This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2.
-Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}.
-We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
-For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
-For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
-
-The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid.
-To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data.
-To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
-For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest).
-Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.
-This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-The actual data plot location is 11.
-
-For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
-Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
-The order that the lineups are shown to participants is randomized.
-Each subject saw a substantial number of difficult lineups.
-
-%\hh{While this allows us a better estimation for the power in small slopes, }
-%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
-%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
-%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
-%	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
-
-%{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
-For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in Table \ref{tbl:experiment_params}.
-For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
-The 10 lineups are chosen as follows;
-
-%\begin{enumerate}
-%	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
-%	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
-%	\item Finally select another lineup with difficulty level 0
-%\end{enumerate}
-
-\begin{figure}[p]
-%\begin{figurehere}
-   \centering
-       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
-       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?
-(The position of the actual data plot is provided in Section \ref{sec:continuous}.)}
-       \label{fig:test_continuous}
-\end{figure}
-
-\subsection{Contaminated data} \label{sec:contamination}
-
-The first two simulation experiments use data generated under a normal error model, satisfying the conditions for conventional test procedures.
-In these situations there exists a test, and there would, in general, be no need to use visual inference.
-The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
- This third simulation is closer to the mark for the purpose of visual inference.
-The assumptions for the conventional test are violated, by contaminating the data.
-The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.
-The data is generated from the following model:
-
-\[
-  Y_i = \left\{
-  \begin{array}{l l}
-    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
-    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
-  \end{array} \right.
-\]
-where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$.
-$n_c$ is the size of the contaminated data.
-For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points.
-Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
-%Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
-A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
-This experiment  follows the test in Table \ref{tbl:stat_multiple}, case 2.
-The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null  plots are scatterplots of null data generated by plotting simulated residuals from $N(0, {\hat{\sigma}}^2$)  against $X_1$.
-
-Experiment three consists of a total of 30 lineups, made up of five replicates for each of the six slopes as shown in Table \ref{tbl:experiment_params}.
-We use the slope directly as a measure of difficulty, with difficulty = 0 for the largest slope and difficulty = 5 for the smallest slope.
-Subjects were exposed to a total of ten lineups, with two lineups from each of the difficulty levels 0 through 3, and one  lineup each from levels 4 and 5.
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
-%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? The actual data plot location is 7.
-
-%\begin{figure}[hbt]
-%\begin{figurehere}
-%   \centering
-%       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
-%       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
-The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope.
-%}
-%       \label{fig:cont_dat}
-%\end{figure}
-
-\begin{figure}[p]
-%\begin{figurehere}
-   \centering
-       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
-       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?
-(The position of the actual data plot is provided in Section \ref{sec:contamination}.)}
-       \label{fig:test_contaminated}
-\end{figure}
-
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
-Show a sample lineup, probably the most difficult, instead of the single plot.}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_observed}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_loess1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
-%       \label{fig:power_loess2}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in Table \ref{tbl:experiment_params}.}
-%       \label{fig:power_diff_exp}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-}
-%       \label{fig:power_observed_exp3}
-%\end{figure*}
-
-\section{Results} \label{sec:results}
-
-\subsection{Data Cleaning}
-
-% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants.
-It is not clear why they did not provide the information.
-Now the concern arises whether we should keep their responses in the study.
-It is also possible that participants were not serious enough to provide feedback.
-%For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
-This could be a criteria to determine their attentiveness in the study.
-For this we consider six different data cleaning criteria as shown below;
-
-%{\bf Screening scenarios:}
-Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA.
-Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data.
-For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
-For the purpose of identifying these participants and cleaning the data, we take action based on a subject's answer to one of the very easy lineups that everybody was exposed to:
-if the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
-If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
-Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-
-%
-%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
-The data of participants who did not answer correctly was removed.
-%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
-That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.
-
-%Amazon Mechanical Turk data suffers from the problem that some participants are trying to 'game' the system, because they are being paid for their efforts, not much by the standards of the US minimum wage, but enough to encourage honest efforts.
-in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
-In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.
-%There are six suggested approach to clean the data. 
-%They are
-%\begin{enumerate}
-%\item {\bf include all participants} and their evaluations
-%\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-%\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-%\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
-%\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
-%\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
-If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-%\end{enumerate}
-
-%\begin{table}[hbtp]
-%\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-%Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
-This is due to missing demographic information.
-\hh{Do we need this table or could we make this a visual comparison?
-For me, there's too many numbers with too many digits.}}
-%\begin{center}
-%\begin{tabular}{rrrrr|rrrr|rrrr}
-%  \hline
-%Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
-% \cline{2-5}  \cline{6-9}   \cline{10-13}
-%criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-%  \hline
-%1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-%  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-%  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-%  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-%  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-%  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:summary}
-%\end{table}
-
-\begin{table}[hbtp]
-\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets.
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
-}%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
-\begin{center}
-\begin{tabular}{ccrrcc}
-   \hline \hline
- Experiment & Subject & Male & Female & Responses &Lineup\\ 
-    \hline
-1 & 239 & 121 & 107 & 2249 &  60 \\ 
-  2 & 351 & 185 & 164 & 3636 &  70 \\ 
-  3 & 155 & 103 &  52 & 1511 &  29 \\  
+1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
+  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
+  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
+  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
+  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
+  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
    \hline
@@
 
-\subsection{Model fitting}
- For each parameter combination, the effect is computed as
-
-\[
-E=\sqrt {n} \cdot \beta/\sigma.
-\]
-\noindent
-The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept, i.e. 
- $W_{\ell i} = E_{\ell i}$.
-Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
-Different skill levels of subjects are accounted for  by allowing subject-specific random  slopes for  effect ($E$).
-
-For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power
-%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
-to be fixed at 0.05 for $E=0$ is not justified by the data. % (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
-
-%For all the three experiments, each subject saw a lineup just once.
-This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
-
-%The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. 
-For computation we use package {\tt lme4} \citep{lme4:2011} and software R 2.15.0 \citep{R}. $p$-value calculations are based on asymptotic normality.
-
-Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-The fixed effects estimates indicate that for all experiments  the proportion of correct responses increases as the effect increases.
-This effect is less pronounced  for experiment 3.
-The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.
-%The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
-%contributes to the random effect variance estimate be much larger compared to those of other experiments.
-For the same reason we see the standard error for intercept estimate is much higher as well.
-Also notice the slope estimate for experiment three is lowest as well.
-It is because for small parameter values the power is much higher for experiment 3.
-% \green{What do we learn about the models from these parameter estimates?
-Explain this here too.}
-
-%\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
-
-\begin{table}[hbtp]
-\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
-\begin{center}
-\begin{tabular}{cr@{.}lcc}
-  \hline \hline
- &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
- \cline{2-4}
- Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
-  \hline
-  1 & 0&39 & 0.0094 & 0.0080 \\ 
-  2 & 1&21 &  0.0197 &  0.0443 \\ 
-  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
-     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
-     &-0&78 (correlation) & & \\
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:model_par}
-\end{table}
-
-\subsection{Power comparison}
-
- Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.
-Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
-A loess  fit to the data gives an estimate of  the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
-For comparison, the dashed lines show the corresponding power curves of the conventional tests.
-Not surprisingly, the observed power for experiments 1 and 2 is lower than the power of the conventional test.
-It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
-Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Power in comparison to effect for the three experiments.
-Points indicate subject responses, with size indicating count.
-Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-The loess curve (continuous line) estimates the observed power, and surrounding bands show simultaneous bootstrap confidence band.
-Conventional test power is drawn as a dashed line.
-For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
-Power for contaminated data is shown by cross mark.}
-       \label{fig:power_loess_effect}
-\end{figure}
-
-Results for experiment 3 are quite different.
-This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
-There is no actual conventional power in this situation, because assumptions are violated.
-The dashed line shows conventional power based on uncontaminated data, whereas the cross marks show effective power based on the coefficient estimated from the contaminated data.
-% It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
-The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
-For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
-% 
-%Note that the power structure is completely broken for contaminated data. %For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
-
-Results of experiment 3 are curious insofar, as  power of the visual test is  high independent of effect size.
-However, these results are based on correct identification of the actual data plot, regardless of reason.
-Although subjects were asked to select the plot that exhibited the highest association between the two variables, they might have cued in on the cluster of contaminated data.
-This will be explored further in Section \ref{sec:TypeIII}.
-
-%(scatter plot) is better than what we observed for experiment 1 (box plot).
-For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
-
-%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
-This shows how the screening criteria may affect the results.
-The corresponding power curves of conventional test are also shown.
-Notice that no mater what criteria we apply the result does not change much for experiment 2.
-Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
-
-%
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
-%       \caption{The overall power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
-%       \label{fig:power_mixed}
-%\end{figure}
-
-%Each participant was shown a sequence of 10 lineup plots.
-In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
-The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-%       \caption{Subject specific power obtained from equation \eqref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect_subject1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-\blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
-%       \label{fig:power_loess3}
-%\end{figure*}
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.40}{\includegraphics{power_model.pdf}}
-%       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_model}
-%\end{figure}
-
-%We fit model \eqref{eqn:mixed} to the survey data obtained from the simulation experiment.
-The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}.
-Model \ref{eqn:mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}.
-The plot includes 20 randomly selected subject-specific power curves.
-Notice that the power curve estimated for one subject is above the UMP test power curve.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
-%       \caption{Estimated subject specific power curve from model \ref{eqn:mixed} for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_subject}
-%\end{figure}
-
-\subsection{Subject-specific variation}
-
-%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
-Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
-The thick continuous line shows overall estimated power, the thinner lines correspond to subject-specific power curves.
-For comparison, the  dashed lines show power curves of the conventional test.
-Subject-specific power is quite different between the three experiments.
-In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
-In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects.
-In experiment 3 there is the most subject-specific variation.
-Some subjects performed substantially better than the conventional test, and on average the visual test was better.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
-The corresponding power curve for conventional test (dashed line) is shown for comparison.
-The overall average power curve estimated from the model is shown (light blue).}
-       \label{fig:power_mixed_subject}
-\end{figure}
-
-\subsection{Estimating the $p$-value in the real world}
-
-%\green{third experiment doesn't really have a p-value, does it?
-Need to explain this.
-Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.
-So many small p-values are coming from the fact that contamination makes the slope negatively significant.
-For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant.
-It is like the effect of influential outlier.}
-
-In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values.
-Assessing the strength of perceived structure is a critical component of visual inference.
-In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup.
-As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values.
-For  $p$-values  larger than 0.15 it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_prop_correct.pdf}}
-       \caption{Proportion of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in Figure \ref{fig:pval_power}.
-}
-       \label{fig:pval_pcorrect}
-\end{figure*}
-
-From the experimental data the visual $p$-values are estimated based on Definition \ref{dfn:pvalue}.
-Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.
-The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate,
-% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
-% 
- giving a clear indication to reject $H_0$ or not.
-This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
-This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
- % for which researchers get confused whether to reject or 'not reject' $H_0$. 
- For visual tests this is not common.
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . Values are shown on square root scale. }
-    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-       \label{fig:pval_definition}
-\end{figure*}
-
-%We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}. 
-%The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
-For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2.
-This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.
-
-For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are.
-This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
-% and the visual test does the job nicely.
-
-%With multiple subjects it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
-The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
-Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more subjects select that plot.
-Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.
-
-%Similary, the strength of the signal, is calculated using the proportion of subjects picking actual data plot.
-It is closely related to the power of the test.
-From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
-This gives the signal strength estimate as
-%\[
-%\hat p_D= (1- {u}/{K})/({m-1})
-%\]
-%where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
-For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions.
-The blue line is a loess fit to the data, and the grey line indicates equality of the two axes.
-The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
-As the conventional $p$-value increases, the estimated signal strength increases -- subjects find it harder to detect the actual data plot at larger $p$-values.
-The rise is very sharp for experiment 1, which indicates that subjects have a harder time detecting the actual data plot for even small $p$-values.
-
-%\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
-% \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
-%
-%\green{Explain why the term ``plot signal'' is introduced}
-
-%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
-For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
-As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
-When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
-We observed this for both experiment 1 and 2.
-For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
-
-%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
-
-%Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
-This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a conventional sense.
-Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
-%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.
-}  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.
-But it is true for experiment 2, right?.
-}
-%is equivalent to conventional test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
-%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
-%       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
-The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
-p-values are shown on square root scale scale.}
-%    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-%       \label{fig:pval_plot_signal}
-%\end{figure*}
-
-\subsection{Do people tend to pick the lowest $p$-value?}
-
- One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
-In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
-This can be examined for the data collected, to see if this assumption is, indeed, reasonable.
-%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value.
-In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
-
-Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
-In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically.
-The dot represents this point, and the lines connect vertically with 0 count.
-Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.
-
-%Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
-Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
-(Empty cells indicate no lineup shown with this combination.)
-
-Red indicates the plot with the lowest $p$-value in the lineup.
-Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.
-In both experiments  people tended to select the plot with the lowest $p$-value with an increase in effect size.
- %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
- The results are clearer for experiment 2, that used a continuous covariate.
-But even when subjects did not pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
-So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
-
-\begin{figure*}[hbtp]
-   \centering
-      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
-       \label{fig:P-val_log}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-a.pdf}       
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
-       Empty cells indicate combination of parameters with fewer than five replicates. 
-       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
-       }
-       \label{fig:P-val_log2}
-\end{figure*}
-
-There are some noticeable exceptions to this rule. In experiment 1, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelmingly chose a plot with much larger $p$-value, similarly, for 
-parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
-%(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
-
-%Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
-Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
-Rather it is assumed that the subject has the ability to select the plot with the most extreme structure.
-A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, \sigma=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
-Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.
-
-%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in Table \ref{tbl:pval_assumption} for both experiment 1 and 2.
-The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
-We see the same pattern in Figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value. The data used is from screening criteria 1. } 
-%\begin{center}
-%\begin{tabular}{cccc}
-%  \hline
-%screening & Experiment & Total &  minimum $p$-value\\ 
-%criteria & & lineups & picked by the most \\
-%  \hline
-%   & 1 &  60 &  37 \\ [-1ex]
-%    \raisebox{1.5ex}{1} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%    \raisebox{1.5ex}{2} & 2 &  70 &  61 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%   \raisebox{1.5ex} {3} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  41 \\  [-1ex]
-%    \raisebox{1.5ex}{4} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  39 \\  [-1ex]
-%    \raisebox{1.5ex}{5} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  40 \\  [-1ex]
-%    \raisebox{1.5ex}{6} & 2 &  70 &  60 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
-%\begin{center}
-%\begin{tabular}{ccc} \hline
-%screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
-% \cline{2-3}
-%criteria & Experiment 1 & Experiment 2 \\ 
-%  \hline
-%1 & 37 & 60 \\ 
-%2 & 38 & 61 \\  
-%3 & 38 & 60 \\ 
-%4 & 41 & 60 \\ 
-%5 & 39 & 60 \\
-%6 & 40 & 60 \\ 
-%\hline
-%Total lineups & 60 & 70 \\  
-%\hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:pval_assumption} 
-%\end{table}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
-%       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
-This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
-There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error.
-Can we?
-\green{Maybe don't need this plot}}
-%       \label{fig:P-val_rank}
-%\end{figure*}
-
-\subsection{How much do null plots affect the choice?}
-
- Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
- But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
-How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
-If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
-Measuring the difference, generally, between  plots is almost impossible.
-However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
-
-We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}).
-What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
-If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
-Figure \ref{fig:pval_difference} investigates this.
-The difference between the $p$-value of the actual data is compared with the lowest from the null plots.
-This is plotted horizontally, and the proportion correct is plotted vertically.
-Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
-In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
-This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
-Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
-When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
-
-%\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
-If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
-Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
-The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
-A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
-If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot.
-When there is a large positive difference, subjects fail to pick the correct plot.
-If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
-That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.}
-
-%\green{Something is definitely not right!
-The code looks like it does actual data p-value - min(null $p$-value).
-But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically!
-Suppose data plot $p$-value is larger than 0.8.
-Under null hypothesis, $p$-value $\sim U(0,1)$.
-Thus min(null $p$-value) could be close to zero by chance producing this large difference.
-And this has happened in both experiment 1 and 2.}
-
- %\green{Is this really pct correct, or 1-pct correct?
-\blue{This is proportion correct}.
-When there is a big diff in p-value almost no-one gets it right!
-Maybe it is just a scale issue?} \blue{In fact this is what we expect.
-If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small.
-This should lead people to pick the wrong plot.
-Our minimum $p$-value assumption basically says this fact, right?}.
-We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.
-
-%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
-Thanks!
-
-In practice, the $p$-value is not going to be a valid way to compare plots.
-Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
-This is investigated in \citet{niladri:2012}.
-Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
-       \caption{Investigating how the strength of structure in the null plots affects the choice of the actual data plot.
-Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.
-Negative values indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
-Close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
-       \label{fig:pval_difference}
-\end{figure}
-
-\subsection{Type III error}~\label{sec:TypeIII}
-
-A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
-This occurs when the null hypothesis is correctly rejected but for the wrong reason.
-Experiment 3 is prone to this type of error.
-Observers were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
-For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
-This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
-This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
-Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
-
-Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice:
-
-\begin{enumerate} \itemsep 0in
-\item Most different plot
-\item Visible trend
-\item Clustering visible
-\item Other
-\end{enumerate}
-
-\noindent with the possibility to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
-Reasons 1 or 3 would be indicative of Type III error.
-Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
-If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
-This is not what we see.
-At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
-The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
-The size of the point corresponds to the number of subjects putting this as the reason.
-Subjects that chose all three reasons almost always chose the actual data plot.
-This was followed by using 1 and 3, and then 1 and 4.
-The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
-At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
-``Visible trend'' comes in third.
-This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
-
-%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
-Figure \ref{fig:choice_reason} examines this.
-For observers who chose the actual data plot, the proportions for the three reasons are shown.
-It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
-\green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}
-
-%This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}
-
-%\green{Not sure that this goes here, but something needs to be said about Type III.}
-
-%\blue{We also see this feature in figure \ref{fig:choice_reason}}.
-%\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
-
-\begin{figure}[hbtp]
-   \centering
-%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
-%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
-       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
-       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error.
-The vertical axis shows the reasons.
-At left, all subjects choices are shown, and reason 123 means all three reasons are used.
-At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
-The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot.
-Size of the point corresponds to the number of subjects using that reason.}
-       \label{fig:choice_reason}
-\end{figure}
-
-For visual inference, making a Type III, is not actually a problem.
-It is only a possibility in this experiment because we are working with known structure.
-In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
-This does show the importance of the qualitative answers that observers give for their choices.
-
-\section{Conclusions}
-
-%\begin{itemize} \itemsep 0in
-%\item Summarize the key pieces of the paper again, especially the major findings
-
-It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
-The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-
-Power of visual tests can be estimated in different ways, which was explained in this paper.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is close to that of conventional test in the conditions tested in this paper.
-
-%\item Explain caveats, nuances, advantages with the work
-
-One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
-A good test statistic is expected to reveal structure if the null hypothesis is not true.
-Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
-This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
-%This is the part of our ongoing and future research.
-
-The question that is posed to observers can affect the ability to evaluate lineups.
-In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
-
-% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
-If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
-This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
-% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
-
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
-
-Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
-Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet, the results were good.
-We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
-
-The performance of subjects was quite varied, but consistent.
-No restrictions were placed on Turk workers, in terms of abilities.
-There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test.
- It would be interesting to see how well trained subjects might perform.
-Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference.
-Future work might explore this.
-
-%\item Describe the next steps of the work, and possible ideas for future extensions
-
-When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature.
-We have successfully used the procedures informally for local projects.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
-Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
-
-%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
-The direction of future research would be to study that and apply the methods in those areas.
-
-%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
-We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
-The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
-For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
-Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.
-
-%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
-Along the lines of the Army tests that Susan brought to the working group.
-\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
-Thus anyone from any background could participate the experiment.
-It would be interesting to see how well the trained subjects performs.
-Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
-One of our future work would be to examine this in details.
-For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
-After being trained they may have to pass the qualification test before participating the experiment.
-Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
-
-%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
-
-%In practice there is no correct answer.
-In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
-In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
-And subjects would provide open-ended answers for their choice of plot.
-
-%What are your suggestions for future experiments, and next steps of the research?
-
-%\paragraph{Acknowledgement:}
-%This work was funded in part by National Science Foundation grant DMS 1007697.
-
-%\bibliographystyle{plain}
-%\bibliographystyle{plainnat}
-\bibliographystyle{asa}
-%\bibliographystyle{ieeetr}
-\bibliography{references}
-
-%\end{multicols}
-
-\end{document}  %End of document.
-
-\section*{Appendix}
-
-\begin{table}[hbtp]
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
-\centering
-\begin{tabular}{c c c c  c c c}
-\hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
-\hline
-easy&100& 5&8 & 1& 12 & 1\\
-&100&12&16 &1&&\\
-&300& 5&5 &1&&\\
-&300&12&10 &1 &&\\
-\hline
-medium&100& 5&3 &203 & 9 &2\\
-&300& 5&2,3 & 97, 1&&\\
-\hline
-hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
-&300& 5&1 & 371 &&\\
-&300&12&3,5& 375,74 &&\\
-\hline
-mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
-&100&12&1& 100& &\\
-&300& 5&0 & 73&&\\
-&300&12&7,1& 2, 152&&\\
-\hline
-Total &&&&&60&10\\
-\hline
-\end{tabular}
-\label{tbl:dist_lineup1}
-\end{table}
-
-%\begin {abstract} 
-%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently, there were no formal visual methods in place for determining statistical significance of findings.
-This changed when Buja et al.
-(2009) conceptually introduced the lineup protocol for formal tests of visual findings.
-In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
-%\end {abstract}
-
 \end{document}

### 2013-02-19 R/generate_expected_visual_power_simulation-hh.R

@@ -1,5 +1,5 @@
 # This file computes the expected power for visual test.
-# Last modified Jun 2012, di re-formatted and ran, works
+# Last modified Feb 2013, di re-formatted and ran, works
 library(ggplot2)
 library(plyr)
 library(MASS)
@@ -45,7 +45,8 @@ calculate_theoretical_power <- function (beta, n, sigma, alpha=0.05){
   return(res)
 }
 
-powerdf <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=seq(10, 30, by=5), test="Visual"))
+# powerdf <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=seq(10, 30, by=5), test="Visual"))
+powerdf <- data.frame(expand.grid(beta=seq(0, 16, by=0.1), sigma=12, n=100, m=20, test="Visual"))
 powerdf$power <- ldply(1:nrow(powerdf), function(x) {
   mean(replicate(10, with(powerdf, generate_visual_power(n[x],
     beta=beta[x], sigma=sigma[x], m=m[x]))))
@@ -60,10 +61,20 @@ powerdf <- rbind(powerdf, power2)
 levels(powerdf$test)
 powerdf$emphasize <- 0.5 + ((powerdf$m==20) | (powerdf$test=="Conventional"))
 
-ggplot(aes(beta, power, group=m, linetype=test, size=emphasize), data=powerdf) + 
-  geom_smooth(aes(colour=m), alpha=0.5, fill=NA, method="loess", span=0.1) + 
-  scale_colour_gradient("Lineup size m", breaks=c(10,20,30)) + 
+# ggplot(aes(beta, power, group=m, linetype=test, size=emphasize), data=powerdf) + 
+#   geom_smooth(aes(colour=m), alpha=0.5, fill=NA, method="loess", span=0.1) + 
+#   scale_colour_gradient("Lineup size m", breaks=c(10,20,30)) + 
+#   scale_size_identity() + ylab("Power") +
+#   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
+
+ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdf) + 
+  geom_smooth( method="loess", span=0.1) + 
   scale_size_identity() + ylab("Power") +
   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
 
 ggsave("../images/power_expected.pdf", width=7, height=5, units="in")
+
+
+
+
+

### 2013-02-19 R/paper_jasa.R

@@ -78,6 +78,36 @@ for (K in 1:5){
 	}
 }
 
+
+# The following codes demonstrate that for same proportion of correctness
+# the p-values get smaller with larger values of K 
+Kdat <- NULL 
+for (K in c(1, 5,10,15)){
+  x <- 0:K
+  prop_correct <- x/K
+  pvals <- sapply(x,get_pval, m=20,K=K)
+  Kdat <- rbind(Kdat,data.frame(prop_correct, pvals, K))
+}
+
+qplot(prop_correct, pvals, geom=c("point","line"), data=Kdat, linetype=factor(K), color=factor(K)) +
+  xlab("Proportion correct (x/K)") + ylab("Visual test p-value") +
+  scale_linetype_discrete("K") + scale_color_discrete("K")
+
+ggsave("../images/K-pvals.pdf", height=5, width=6)
+
+
+Kdat1 <- rbind(data.frame(K=c(2,8,10,14), prop_correct=1/2),
+               data.frame(K=c(3,9,12,15), prop_correct=1/3),
+               data.frame(K=c(4,8,12,16), prop_correct=3/4))
+Kdat1$x <- with(Kdat1, prop_correct*K)
+Kdat1$pvals <- apply(Kdat1[,c("K","x")],1,function(t){get_pval(m=20, K=t[1],x=t[2])})
+
+qplot(K, pvals, geom=c("point","line"), data=round(Kdat1,2), linetype=factor(prop_correct), color=factor(prop_correct)) +
+  xlab("number of observer (K)") + ylab("Visual test p-value") +
+  scale_linetype_discrete("prop_correct") + scale_color_discrete("prop_correct")
+ggsave("../images/K-pvals1.pdf", height=5, width=6)
+
+
 # === plotting test stat and lineup for categorical variable
 
 # generating data
@@ -471,9 +501,9 @@ get_merged_pvalue <- function(dat,pval){
   p_values <- melt(pval, id="pic_name")
   pr <- merge(p_values,response_count, by=c("pic_name","variable"))
   prr <- ddply(pr, .(pic_name),summarise,
+               rfreq=counts[order(value)]/sum(counts),
                counts=counts[order(value)],
-               rfreq=counts[order(value)]/sum(counts[order(value)]),
-               pvalue=round(value,4)[order(value)],
[…]

### 2013-02-21 paper_jasa.tex
previous lineups were replaced by actual lineups

@@
 \graphicspath{{images/}}
+
+\usepackage{color}
+\newcommand{\blue}[1]{{\color{blue} #1}} %MM
+\newcommand{\red}[1]{{\color{red} #1}} 
+\newcommand{\green}[1]{{\color{green} #1}} %DC
+
+\definecolor{orange}{rgb}{1,0.5,0}
+\newcommand{\hh}[1]{{\color{orange} #1}} %HH
 
@@
    \centering
-       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
+%       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
+       \includegraphics[width=0.95\textwidth]{plot_turk1_300_10_12_1.png}
        \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
@@
 The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
-Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
+Data is simulated using a range of values of $\beta_2$ or slopes as shown in Table \ref{tbl:experiment_params}, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
 The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test.
@@
 The actual data plot was randomly placed among these null data plots to produce the lineup.
-Figure \ref{fig:test_category} is an example of one of these lineups.
-The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
+\blue {Figure \ref{fig:test_category} is an example of one of these lineups.
+The actual data plot is number 15 which was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
+For this lineup, 15 out of 16 observers correctly pick the data plot.
+}
 
@@
 We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
-For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
-For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
+\blue{For each combination of parameters, at least three different actual data sets are produced, yielding a total of 70 lineups.
+}
+
+%For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
 
@@
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-The actual data plot location is 11.
+\blue{The actual data plot location is 5.
+None of the 65 observers could detect the actual plot.}
 
@@
    \centering
-       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
+%       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
+       \includegraphics[width=0.95\textwidth]{plot_turk2_100_25_5_3.png}
        \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous.
@@
 
-An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? The actual data plot location is 7.
+An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.
+Can you pick which plot is different?
+\blue{The actual data plot location is 1.
+Out of 31 observers 13 observers could detect the actual plot.}
 
@@
    \centering
-       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
+%       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
+       \includegraphics[width=0.95\textwidth]{plot_turk3_100_40_5_5.png}
        \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure.
@@
 Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
+\blue {A summary of all the data along with cleaned data are added as the supplementary material.}
 
@@
        \includegraphics[width=0.8\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{ Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all individual lineups of both experiment 1 and 2.
+       \caption{ \blue {Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all individual lineups of both experiment 1 and 2.
 Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
@@
 Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
-       }
+    }   }
        \label{fig:P-val_log2}

### 2013-02-21 paper_jasa.tex
pdf version of lineup for experiment 1 and 2 added

@@
 %       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
-       \includegraphics[width=0.95\textwidth]{plot_turk1_300_10_12_1.png}
+       \includegraphics[width=0.95\textwidth]{plot_turk1_300_10_12_1.pdf}
        \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
@@
 The actual data plot is number 15 which was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
-For this lineup, 15 out of 16 observers correctly pick the data plot.
+For this lineup, 15 out of 16 observers picked the actual data plot.
 }
@@
 \blue{The actual data plot location is 5.
-None of the 65 observers could detect the actual plot.}
+None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup as we can see in Figure \ref{fig:P-val_log2} for Experiment 2.}
 
@@
 %       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
-       \includegraphics[width=0.95\textwidth]{plot_turk2_100_25_5_3.png}
+       \includegraphics[width=0.95\textwidth]{plot_turk2_100_25_5_3.pdf}
        \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous.
@@
 \blue{The actual data plot location is 1.
-Out of 31 observers 13 observers could detect the actual plot.}
+Out of 31 observers 13 observers picked the actual plot.}
 
@@
  For each parameter combination, the effect is computed as 
-
 \[
@@
 
-Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+%Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
 Observers were provided with four possible reasons to use for their choice:
-
-\begin{enumerate} \itemsep 0in
-\item Most different plot
-\item Visible trend
-\item Clustering visible
-\item Other
-\end{enumerate}
-
-\noindent with the possibility to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
+%
+%
+%\begin{enumerate} \itemsep 0in
+%\item Most different plot
+%\item Visible trend
+%\item Clustering visible
+%\item Other
+%\end{enumerate}
+%
+%
+%\noindent with the possibility to use more than one.
+
+\blue{Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
+Observers were provided with four possible reasons to use for their choice: (1) Most different plot (2) Visible trend (3) Clustering visible (4) Other, with the option to use more than one.} The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
 Reasons 1 or 3 would be indicative of Type III error.

### 2013-02-21 R/paper_jasa.R

@@ -107,6 +107,40 @@ qplot(K, pvals, geom=c("point","line"), data=round(Kdat1,2), linetype=factor(pro
   scale_linetype_discrete("prop_correct") + scale_color_discrete("prop_correct")
 ggsave("../images/K-pvals1.pdf", height=5, width=6)
 
+# Information about the actual lineups used in the paper
+
+dd1 <- subset(dat1, pic_name=="plot_turk1_300_10_12_1.png")
+table(dd1$response_no)
+dd2 <- subset(dat2, pic_name=="plot_turk2_100_25_5_3.png")
+table(dd2$response_no)
+dd3 <- subset(dat3, pic_name=="plot_turk3_100_40_5_5.png")
+table(dd3$response_no)
+
+# The actual lineups are regenerated from the original simulated data
+# to be used in the paper
+
+# regenrating lineup for categorical variable (experiment1)
+datp1 <- read.table("../data/dat_turk1_300_10_12_1.txt", header=T)[,c(1:20,22)]
+colnames(datp1) <- c(1:20,"grp")
+m.dat1 <- melt(datp1,id="grp")
+p <- qplot(grp,value,colour=grp, xlab="Group",geom="boxplot", ylab="Residual",data=m.dat1)
+p <- p + facet_wrap(~variable)
+ggsave(filename="../images/plot_turk1_300_10_12_1.pdf",plot=p, ,height=7,width=7.5)
+
+# regenrating lineup for continuous variable (experiment2)
+datp2 <- read.table("../data/dat_turk2_100_25_5_3.txt", header=T)
+colnames(datp2) <- c("X",1:20)
+m.dat2 <- melt(datp2,id=c("X"))
+p <- qplot( X, value, data=m.dat2, xlab="X",geom="point", ylab="Y", alpha=I(.2))
+p <- p + geom_smooth(method="lm", se=F, size=1) + facet_wrap(~variable)
+ggsave(filename="../images/plot_turk2_100_25_5_3.pdf",plot=p, ,height=7,width=7.5)
+
+# regenrating lineup for contaminated data (experiment3)
+# we can regenerate only the observed plot. Null plot data was not saved
+datp3 <- read.table("../data/dat_turk3_100_40_5_5.txt", header=T)
+p <- qplot( x, y, data=datp3, xlab="X",geom="point", ylab="Y", alpha=I(.3))
+
+
 
 # === plotting test stat and lineup for categorical variable
 

### 2013-02-21 R/paper_sociology.R

@@ -1,1105 +1,0 @@
-# The complete R code for JASA Paper
-
-library(ggplot2)
-library(plyr)
-library(reshape)
-library(xtable)
-library(lme4)
-library(stringr)
-
-# getting the data and some common functions
-
-raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
-raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
-raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
-
-source("calculate_ump_power.R") # functions to compute power
-
-# cleaning the data based on criteria 6 and removing the duplicated data
-# (id,pic_id) should be unique since no person gets a lineup more than once.
-clean_data <- function(raw_dat){
-  easy_dat <- subset(raw_dat, p_value < 0.0002)
-  if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
-  d <- ddply(subset(easy_dat),.(id),summarise,
-             easy_cnt = length(response),
-             percent_correct = mean(response)*100,
-             include_id = response[1],
-             excluded_lineup = paste(id[1],"_",pic_id[1],sep=""),
-             excluded_lineup_p_value = round(p_value[1],4)
-             )
-  included_id <- d$id[d$include_id]
-  indx <- raw_dat$id %in% included_id
-  excluded_lineup <- paste(raw_dat$id,"_",raw_dat$pic_id,sep="") %in% d$excluded_lineup
-  indx[excluded_lineup] <- FALSE
-  cleaned_dat <- subset(raw_dat,indx)
-  indx_dup <- with(cleaned_dat, !duplicated(data.frame(id,pic_id))) #duplication index
-  # now returning cleaned data removing duplication
-  return(subset(cleaned_dat,indx_dup))
-}
-
-indx1 <- c(3,6,7,8,9,17,18,20)
-indx2 <- c(4,7,8,9,10,17,18,20)
-
-dat1 <- clean_data(raw.dat1)[,-indx1]
-dat2 <- clean_data(raw.dat2)[,-indx1]
-dat3 <- clean_data(raw.dat3)[,-indx2]
-
-
-# saving cleaned data as suplementary materials
[…]

### 2013-03-04 paper_jasa.tex
power curves updated

@@
    \centering
-       \scalebox{.6}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
+       \scalebox{.6}{\includegraphics{power_expected_k.pdf}}
+       \caption{Expected power of visual test  for different number of observers ($K$) and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
        \label{fig:power_expected}
@@
    \centering
-       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
+       \scalebox{0.55}{\includegraphics{power_loess_effect_k.pdf}}
        \caption{Power in comparison to effect for the three experiments.
@@
    \centering
-       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
+       \scalebox{0.60}{\includegraphics{power_mixed_subject_k.pdf}}
        \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
@@
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice: (1) Most different plot (2) Visible trend (3) Clustering visible (4) Other, with the option to use more than one.} The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
+Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.} The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
 Reasons 1 or 3 would be indicative of Type III error.

### 2013-03-04 R/generate_expected_visual_power_simulation-hh.R

@@ -67,12 +67,12 @@ powerdf$emphasize <- 0.5 + ((powerdf$m==20) | (powerdf$test=="Conventional"))
 #   scale_size_identity() + ylab("Power") +
 #   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
 
-ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdf) + 
-  geom_smooth( method="loess", span=0.1) + 
-  scale_size_identity() + ylab("Power") +
-  scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
-
-ggsave("../images/power_expected.pdf", width=7, height=5, units="in")
+# ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdf) + 
+#   geom_smooth( method="loess", span=0.1) + 
+#   scale_size_identity() + ylab("Power") +
+#   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta))
+# 
+# ggsave("../images/power_expected.pdf", width=7, height=5, units="in")
 
 
 

### 2013-03-04 R/paper_jasa.R

@@ -64,20 +64,26 @@ ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
 # ===== computation of visual p-values for diffrent combination of K and x 
 
 get_pval <- function(m,K,x){
-	p <- pbinom(size=K,prob=1/m,q=x)
+	p <- pbinom(size=K,prob=1/m,q=x) # gives P(X<=x)
 	px <- dbinom(size=K,prob=1/m,x=x)
 	pval <- 1-p + px
 	return(pval)
 }
 
 pvals <- NULL
-for (K in 1:5){
+for (K in 1:15){
 	for (x in 0:K){
-		pval <- get_pval(m=20,K=K,x=x)
-		pvals <- rbind(pvals,c(m=20,K=K,x=x,pval=round(pval,4)))
+    m <- 20
+		pval <- get_pval(m=m,K=K,x=x)
+		pvals <- rbind(pvals,c(m=m,K=K,x=x,pval=round(pval,4)))
 	}
 }
 
+
+qplot(x,pval, data=subset(as.data.frame(pvals),K==15), geom=c("point","line")) +
+  geom_hline(yintercept = 0.05) + scale_y_sqrt() +
+  geom_text(aes(3,0, label="x")) +
+  ylab("p-value in square root scale") + xlab(expression(x[alpha]))
 
 # The following codes demonstrate that for same proportion of correctness
 # the p-values get smaller with larger values of K 
@@ -515,6 +521,66 @@ ggsave( file="../images/power_screening_subject.pdf",height=8,width=10 )
 # 
 # qplot(effect, pred, group=subject, data=subset(newdata, subject %in% sample(size=2, x=unique(newdata$subject))), geom="line")
 
+
+
+# ====== getting power using binomial sums and proportion
+# This is done using definition of power. The proportion
+# p in this computation is estimated by proportion correct.
+
+get_power <- function(K,p){
+  pvals <- sapply(1:K,get_pval,m=20,K=K)  
+  xa <- min(which(pvals<=0.05))
+  res <- sum(dbinom(x=xa:K,size=K,prob=p))
+  return(res)
+}
+
+# ========== computing power for different values of K
+# the intial proportion correct is the power for K=1
+
+get_power(.1, K=5)
+ps <- seq(.05,1,by=.01)
+pows <- sapply(ps, get_power, K=5)
+qplot(ps,pows)
+
+pi_effect$k5 <- sapply(pi_effect$pred, get_power, K=5)
[…]

### 2013-03-05 paper_jasa.tex
actual plot location hidden

@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-A new method, called the lineup protocol, enables determining the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.
-In this paper inferential methods for statistical graphics is developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
+A new method, called the lineup protocol, \red{enables determining} the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.
+In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
 This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional test are violated, and there may be some super-individuals who yield better power than the conventional test even in the most difficult tasks.
+\blue{Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when signal is statistically significant or data is contaminated, a scenario where assumptions required for performing a conventional test are violated, and there may be some super-individuals who yield better power than the conventional test even in the most difficult tasks.}
 
@@
 Statistical graphics nourish the discovery process in data analysis by revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally by contrasting prevailing hypotheses.
-The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
+The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
 Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
 
-Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years.
+Statistical graphics research and development is \red{agreeably established} now, and has been cemented by several new developments in recent years.
 Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results.
@@
 This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference.
-The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
+The intended use of lineups is where there is no existing test, \blue{and unlikely ever to be any numerical test}.
 The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
@@
 
-Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing.
-Both start from the same place, the same set of hypotheses.
+\blue{An illustration of the lineup protocol in relation to conventional hypothesis testing is presented in Table \ref{tbl:compare}.} Both methods start from the same place, the same set of hypotheses.
 The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
@@
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
+Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without \red{loosing} independence.
 
@@
 
-In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
+In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
 
@@
 \blue {Figure \ref{fig:test_category} is an example of one of these lineups.
-The actual data plot is number 15 which was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
+It was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
+The actual data plot location is the integer number nearest to $log_e(3269017)$.
 For this lineup, 15 out of 16 observers picked the actual data plot.
@@
 Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
+Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected} with $K=1$) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
 For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
@@
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-\blue{The actual data plot location is 5.
+\blue{The actual data plot location is the integer value nearest to $log_e(148.4132)$.
 None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup as we can see in Figure \ref{fig:P-val_log2} for Experiment 2.}
 
-For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
+For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, giving 5 lineups.
 Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
@@
 Can you pick which plot is different?
-\blue{The actual data plot location is 1.
-Out of 31 observers 13 observers picked the actual plot.}
+\blue{The actual data plot location is the integer number nearest to $log_e (log_e (15.15426))$ and 13 out of 31 observers picked the actual plot.}
 
@@
 %{\bf Screening scenarios:}
-Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA.
+Amazon Mechanical Turk workers are paid for their efforts, not substantially, but on the scale of the minimum wage of the USA.
 Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data.
@@
 Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-\blue {A summary of all the data along with cleaned data are added as the supplementary material.}
+\blue {A summary of all the data along with cleaned data are added to the supplementary material.}
 
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject_k.pdf}}
-       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
+       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} for $K=5$.
 The corresponding power curve for conventional test (dashed line) is shown for comparison.
@@
 
-In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values.
+In the real setting, where visual inference is to be useful, there will be no conventional test $p$-values.
 Assessing the strength of perceived structure is a critical component of visual inference.
@@
        \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
-       \caption{Investigating how the strength of structure in the null plots affects the choice of the actual data plot.
-Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.
-Negative values indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
-Close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
+       \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
+Negative differences indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
+Difference close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
        \label{fig:pval_difference}
@@
 
-When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature.
-We have successfully used the procedures informally for local projects.
+When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
+We have successfully used the procedure informally for local projects.
 It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
-Follow-up questions on the reasons provide qualitative insight.
+\blue{If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.} Follow-up questions on the reasons provide qualitative insight.
 Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.

### 2013-03-05 R/paper_jasa.R

@@ -52,7 +52,6 @@ write.csv(dat3, file='../submission_initial/suplementary/data/data_turk3.csv',ro
 
 
 
-
 # =================== mathematical test t distribution  ===================
 
 t <- seq(-4, 4, by = 0.1)

### 2013-03-05 paper_jasa.tex
Figure caption updated

@@
 Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-The loess curve (continuous line) estimates the observed power, and surrounding bands show simultaneous bootstrap confidence band.
+The loess curve (continuous line) estimates the observed proportion correct (power for $K=1$), and surrounding bands show simultaneous bootstrap confidence band.
+Observed proportion is used to obtain power for $K=5$.
 Conventional test power is drawn as a dashed line.
@@
        \scalebox{0.60}{\includegraphics{power_mixed_subject_k.pdf}}
-       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} for $K=5$.
+       \caption{Subject-specific  power for $K=5$ obtained using the subject-specific proportion correct estimated from model \ref{eqn:mixed}.
 The corresponding power curve for conventional test (dashed line) is shown for comparison.

### 2013-03-06 paper_jasa.tex
discussion on power updated

@@
 
-\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
+\noindent The inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation by a single observer as
 \begin{equation} \label{eqn:mixed_power}
@@
 This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}.
-Thus the expected power of the visual test in this scenario is
+\blue {Thus the expected probability to reject by a single observer ($K=1$) in this scenario is
 
 \begin{equation}\label{power_exp} 
-   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
+%   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
+ p(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
 \end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Notice that the expected power of the visual test is %almost as good as
-close to the power of the conventional test.
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test for different number of observers ($K$) obtained using $p(\beta)$ from Equation \ref{power_exp}.
+Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger.
+}
 
@@
 Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.
-The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
+\red{The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.} The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
 (Data and lineups used in the experiments are available in the supplementary material.)
@@
 For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.
-
 Note that no participant was shown the same lineup twice.
@@
 
-The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
+The number of evaluations required for each lineup to provide reasonable estimates of the \blue{ proportion correct ($\hat p$) is determined by the variance of $p$}.
 Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected} with $K=1$) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
+Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected} with $K=1$) \blue{we consider $\gamma =p$}.
 For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
@@
 Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
-A loess  fit to the data gives an estimate of  the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
+\blue {A loess  fit to the data gives an estimate of  the observed proportion correct $\hat p(E)$ for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
+$\hat p(E)$ is considered to be the power for $K=1$ and it is used to obtain power for $K=5$.
 For comparison, the dashed lines show the corresponding power curves of the conventional tests.
-Not surprisingly, the observed power for experiments 1 and 2 is lower than the power of the conventional test.
-It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
+It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and exceeds the power of conventional test as $K$ increases.}
 Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
@@
 %Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
+\blue{Subject-specific proportion correct $\hat p_i(E)$ is obtained using Equation \ref{eqn:mixed_power} and it is used to obtain power for $K=5$}.
 Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
@@
 Subject-specific power is quite different between the three experiments.
-In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
-In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects.
+In experiment 2 subjects performed similarly, and substantially better than the conventional test.
+In experiment 1 there is more variability between subjects, with some doing better than the conventional test on large effects.
 In experiment 3 there is the most subject-specific variation.

### 2013-03-07 paper_jasa.tex
di approves

@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-A new method, called the lineup protocol, \red{enables determining} the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.
+A new method, called the lineup protocol, enables statistical significance testing of visual findings, which bridges a big gulf between exploratory and inferential statistics.
 In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-\blue{Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when signal is statistically significant or data is contaminated, a scenario where assumptions required for performing a conventional test are violated, and there may be some super-individuals who yield better power than the conventional test even in the most difficult tasks.}
-
-{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, data mining, visualization, exploratory data analysis} 
+\blue{Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional test are violated.
+Surprisingly, visual tests have higher power than the conventional tests when the effect size is large.
+And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks.}
+
+{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, data mining, visualization, exploratory data analysis, practical significance, effect size} 
 \end {abstract}
@@
 
-Statistical graphics research and development is \red{agreeably established} now, and has been cemented by several new developments in recent years.
+%Statistical graphics research and development is \red{agreeably established} now, and has been cemented by s
+Several new developments in graphics research have been achieved in recent years.
 Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results.
-This body of work provides a framework for evaluating new statistical graphics.
+The research by \cite{heer:2010} used subjects recruited from Amazon's Mechanical Turk \citep{turk} for their studies.
+This body of work provides a contemporary framework for evaluating new statistical graphics.
 In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots.
@@
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without \red{loosing} independence.
+Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without losing independence.
 
@@
 \end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test for different number of observers ($K$) obtained using $p(\beta)$ from Equation \ref{power_exp}.
-Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger.
+Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test for different $K$ (number of observers), obtained using $p(\beta)$ from Equation \ref{power_exp}.
+Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger, and when $\beta$ gets larger.
 }
@@
        \scalebox{.6}{\includegraphics{power_expected_k.pdf}}
-       \caption{Expected power of visual test  for different number of observers ($K$) and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
+       \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20. }
        \label{fig:power_expected}
@@
 
-Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-\red{The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.} The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
+Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression setting.
+The first two experiments have conventional ideal scenarios where we would not expect the lineup protocol to do better than the conventional test.
+The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
 (Data and lineups used in the experiments are available in the supplementary material.)
@@
 \blue{The actual data plot location is the integer value nearest to $log_e(148.4132)$.
-None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup as we can see in Figure \ref{fig:P-val_log2} for Experiment 2.}
+None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup.}
 
@@
 For comparison, the dashed lines show the corresponding power curves of the conventional tests.
-It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and exceeds the power of conventional test as $K$ increases.}
+It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, in experiments 1 and 2.
+In experiment 3 the power of the visual test exceeds that for the conventional test, as expected.
+*** Needs more work - explain separately for each experiment, and for other $K$ than shown.}
 Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
@@
 Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
-Empty cells indicate combination of parameters with fewer than five replicates.
-Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
+Empty cells indicate combination of parameters that were not tested.
+Highest counts tend to be the plot in the lineup having the lowest $p$-value, more so for experiment 2 than 1.
     }   }
@@
 
-\blue{Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+\blue{Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.

### 2013-03-11 paper_jasa.tex
power discussion

@@
 Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger, and when $\beta$ gets larger.
+} \hh{Conversely, visual power is below conventional power for parameter values close to the null hypothesis.
+This is even more pronounced for large number of observers.
+At the same time, the point of intersection between visual and conventional power approaches the value of the null hypothesis as the number of observers approaches infinity, leading to an asymptotically perfect power curve of zero in the null hypothesis and one for any alternative value.
+This dichotomy of visual power can also be seen in more detail in figure \ref{fig:power_loess_effect}.
 }
@@
        \scalebox{.6}{\includegraphics{power_expected_k.pdf}}
-       \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20. }
+       \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20.
+\hh{We should not use color here - let's use black for all he lines, but make the legend a bit larger so we can distinguish the line types based on that.
+{\tt legend.key.width = unit(1.5, "cm")} will do it}.
+       \hh{Could you standardize the $x$ axis to $\beta/\sigma$? That will give a better insight to how close we need to be to the null hypothesis to have visual power dip below conventional power.}}
        \label{fig:power_expected}
@@
 % \pagebreak
-
-% -------- Heike's writeup added here
 

### 2013-03-11 paper_jasa.tex
lineup numbers

@@
 The actual data plot location is the integer number nearest to $log_e(3269017)$.
-For this lineup, 15 out of 16 observers picked the actual data plot.
+\hh{could you exchange that for a number that we can calculate without a calculator?
+$4^2-1$ would be completely fine - it's not supposed to be a real stumbling block, it's just supposed to not be immediately recognizable.
+}  For this lineup, 15 out of 16 observers picked the actual data plot.
 }

### 2013-03-14 paper_jasa.tex
polishing text

--- 2013-03-07 paper_jasa.tex
@@
 \noindent Note that for $x=0$ the $p$-value becomes, mathematically, equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
-being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
-
+being larger than $P(X \ge 1)$ in this situation.
+If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
 Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
@@
 
-%\begin{dfn}\label{dfn:test}
-%\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
-%The \textbf{visual test}, $V_{\theta}$, is defined to be 
-%\begin{itemize}\itemsep-3pt
-%\item \textbf{Reject} $H_0$ if `enough' observers select the actual data plot in the lineup of $m$ plots, and
-%\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
-%\end{itemize}
-%where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
-%\end{dfn}
-
-%\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
-
-%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.}
-
 \begin{dfn}\label{dfn:test}
@@
 \begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
+\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot, and
 \item \textbf{Fail to reject} $H_0$  otherwise. 
 \end{itemize}
-where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
+where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}|H_0) \le \alpha$. 
 \end{dfn}
@@
 
-%\green{Oh, we are using observed and observer, which might be confusing.
-May need to change observed data plot to something else.
-Will think about this.
-Will use ``actual'' for now.} \hh{I think that change is done}
-
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$. %, with an exact $x_{\alpha}$. 
+\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$.  
 For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
@@
 \noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
-Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
+Thus $X$, the number of observers who identify the actual data plot from the lineup, affects the estimation of power  and the power is estimated by
 \[
@@
 Power is typically used to measure this characteristic of a test.
-%Power is the probability of rejecting $H_0$ when it is not true.
-We can assess power therefore both empirically based on experimental data and through theoretical considerations.
 
@@
 In experiment 3 the power of the visual test exceeds that for the conventional test, as expected.
-*** Needs more work - explain separately for each experiment, and for other $K$ than shown.}
+For larger values of $K$ estimated power exceeds the power of conventional test.}
 Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
@@
 Conventional test power is drawn as a dashed line.
-For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
+For experiment 3, conventional power is based on the slopes of the non-contaminated part of the data.
 Power for contaminated data is shown by cross mark.}
@@
 The corresponding power curve for conventional test (dashed line) is shown for comparison.
-The overall average power curve estimated from the model is shown (light blue).}
+The overall estimated average power curve is shown (light blue).}
        \label{fig:power_mixed_subject}
@@
    \centering
-       \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
+       \scalebox{0.60}{\includegraphics{pval_difference.pdf}}
        \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
@@
        \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error.
-The vertical axis shows the reasons.
-At left, all subjects choices are shown, and reason 123 means all three reasons are used.
+       \caption{Reasons of plot choices vs  proportion of times the subjects correctly chose the actual data plot for experiment 3 that examines the occurrence of Type III error.
+At left, all subjects' choices are shown, and reason 123 means all three reasons are used.
 At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
-The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot.
 Size of the point corresponds to the number of subjects using that reason.}
@@
 If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is close to that of conventional test in the conditions tested in this paper.
+Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
 

### 2013-03-14 paper_jasa.tex
expected power Figure updated

@@
        \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20.  
-\hh{We should not use color here - let's use black for all he lines, but make the legend a bit larger so we can distinguish the line types based on that.
-{\tt legend.key.width = unit(1.5, "cm")} will do it}.
        \hh{Could you standardize the $x$ axis to $\beta/\sigma$? That will give a better insight to how close we need to be to the null hypothesis to have visual power dip below conventional power.}}
@@
 It was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
-The actual data plot location is the integer number nearest to $log_e(3269017)$.
-\hh{could you exchange that for a number that we can calculate without a calculator?
-$4^2-1$ would be completely fine - it's not supposed to be a real stumbling block, it's just supposed to not be immediately recognizable.
-}  For this lineup, 15 out of 16 observers picked the actual data plot.
+The actual data plot location is ($4^2-1$).
+For this lineup, 15 out of 16 observers picked the actual data plot.
 }
@@
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-\blue{The actual data plot location is the integer value nearest to $log_e(148.4132)$.
+\blue{The actual data plot location is $(2^2+1)$.
 None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup.}
@@
 Can you pick which plot is different?
-\blue{The actual data plot location is the integer number nearest to $log_e (log_e (15.15426))$ and 13 out of 31 observers picked the actual plot.}
+\blue{The actual data plot location is $(3^2-2^3)$ and 13 out of 31 observers picked the actual plot.}
 

### 2013-03-14 R/paper_jasa.R

@@ -6,6 +6,7 @@ library(reshape)
 library(xtable)
 library(lme4)
 library(stringr)
+library(grid)
 
 # getting the data and some common functions
 
@@ -929,13 +930,13 @@ powerdfk$test <- factor(powerdfk$test)
 powerdfk$test <- factor(powerdfk$test, levels(powerdfk$test)[c(1,2,4,3)])
 
 
-ggplot(aes(beta, power, linetype=test, size=emphasize, colour=test), data=powerdfk) + 
+ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdfk) + 
   geom_smooth( method="loess", span=0.1) + 
   scale_size_identity() + ylab("Power") +
   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta)) +
-  scale_color_discrete("Test") 
+  theme(legend.key.width = unit(1.5, "cm"))
 
-ggsave("../images/power_expected_k.pdf", width=7, height=5, units="in")
+ggsave("../images/power_expected_k.pdf", width=8, height=5, units="in")
 
 # ====================== Turk2 conventional power calculations  
 

### 2013-03-20 paper_jasa.tex
expected power curve changed

@@
 \noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
-\\
-\noindent
 The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
@@
 The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as 
-
-\begin{equation}\label{binom}
+$$
 P(X \ge x) = 1 - \text{Binom}_{{K, 1/m}} (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i}
-\end{equation}
+$$
 
@@
 
-\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
+\noindent An important difference between conventional and visual testing is that lineups will depend on observers' evaluation.
 Thus $X$, the number of observers who identify the actual data plot from the lineup, affects the estimation of power  and the power is estimated by
@@
 
-%\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
-
-%\red{there may be a flow in this definition.
-Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power.
-I feel like we should not define power using binomial.}
-
 Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
@@
 
-%\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
-Can someone please decide which to use, and make the appropriate changes?
-How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above?
-It looks like this is a repeat.
-One needs to go once notation is clarified.}
-
-%\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
-\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
-\\ Should I change everything accordingly?
-I see that we have been careful not to use $\beta$ for power!
-}
-
 \subsection{Subject-specific abilities}~\label{sec:model}
-%Calculating the power  is intricately related to the probability to identify the actual data plot from the lineup.
 
@@
 \end{figure}
-
-%Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
-%Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
-Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$.
-We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
-%
-%For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
-%we get
-%\[
-%N_m\sim \left \{ 
-%\begin{array}{ll}
-%B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
-%B_{K, 1/m} & \text { under } H_0
-%\end{array}
-%\right .
-%\]
-%
-%By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
-%%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
-%as:
-%\begin{equation}\label{eqn:power_estimate}
-%%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
-%\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
-%\end{equation}
-%
-%Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
-While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
-%\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
-We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact.
-%Both of these topics are further highlighted in the application section.}
-
-%Picking out the data plot is still a measure of how different the data plot is compared to the null plots, and therefore is a test statistic of the visual test.
-
-%This allows us to compute this value independently of the parametric setting. We therefore define the {\it signal strength} of a  visual test as the following:
-%
-%\begin{dfn} \label{dfn:signal}
-%The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
-%\end{dfn}
-%
-%In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
-%\[
-%\sigma_T^{m-1} = \frac{n_m}{K}.
-%\]
-%
-%In a parametric setting, signal strength of a test coincides with the power of the test.
-
-%Theoretical power for the regression parameters is derived in the next section.
-%\subsection{Estimating Power from Lineups}
-%
-%We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
-This gives an estimated power of
-%\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
-
-%From model \ref{eqn:mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
-%the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
-%%Thus we have the overall probability of success ($\pi$ or power) 
-%for specified sample size ($n$) and  variance ($\sigma$)  as 
-%
-%\begin{equation}\label{eqn:power} 
-%Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
-%\end{equation}
-%
-%The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
-combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
-Using equation \eqref{eqn:mixed}, we then have the power as
-%
-%\begin{equation}\label{eqn:effect} 
-%Power(E) = \pi=Pr(Y=1| E) 
-%\end{equation}
 
@@
        \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20.  
-       \hh{Could you standardize the $x$ axis to $\beta/\sigma$? That will give a better insight to how close we need to be to the null hypothesis to have visual power dip below conventional power.}}
+       }
        \label{fig:power_expected}
@@
 It was best for a observer or a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
-The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects, for each of the three experiments.
+The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects.
 
@@
 
-The number of evaluations required for each lineup to provide reasonable estimates of the \blue{ proportion correct ($\hat p$) is determined by the variance of $p$}.
+The number of evaluations required for each lineup to provide reasonable estimates of the \blue{ proportion correct ($\hat p$) is determined by the variance of the number of correct evaluations}.
 Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
@@
 \begin{table}[hbtp]
-\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
+\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for the simulation experiments.} % title name of the table
 \centering
@@
 \caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets.
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
+Note that in some of the lineups the number of male and female participants does not add up to the total number of participants due to missing demographic information.
 }%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
@@
 
-For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power
-%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
-to be fixed at 0.05 for $E=0$ is not justified by the data. % (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
-
-%For all the three experiments, each subject saw a lineup just once.
-This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
-
-%The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. 
+For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power to be fixed at 0.05 for $E=0$ is not \blue{ required by the experimental design}.
+
 For computation we use package {\tt lme4} \citep{lme4:2011} and software R 2.15.0 \citep{R}. $p$-value calculations are based on asymptotic normality.
@@
 The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.
-%The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
-%contributes to the random effect variance estimate be much larger compared to those of other experiments.
-For the same reason we see the standard error for intercept estimate is much higher as well.
-Also notice the slope estimate for experiment three is lowest as well.
-It is because for small parameter values the power is much higher for experiment 3.
-% \green{What do we learn about the models from these parameter estimates?
-Explain this here too.}
-
-%\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
 
@@
 The dashed line shows conventional power based on uncontaminated data, whereas the cross marks show effective power based on the coefficient estimated from the contaminated data.
-% It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
-The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
-For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
-% 
-%Note that the power structure is completely broken for contaminated data. %For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
 
@@
 
-%(scatter plot) is better than what we observed for experiment 1 (box plot).
-For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
-
-%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
-This shows how the screening criteria may affect the results.
-The corresponding power curves of conventional test are also shown.
-Notice that no mater what criteria we apply the result does not change much for experiment 2.
-Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
-
-%
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
-%       \caption{The overall power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
-%       \label{fig:power_mixed}
-%\end{figure}
-
-%Each participant was shown a sequence of 10 lineup plots.
-In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
-The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-%       \caption{Subject specific power obtained from equation \eqref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect_subject1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-\blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
-%       \label{fig:power_loess3}
-%\end{figure*}
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.40}{\includegraphics{power_model.pdf}}
-%       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_model}
-%\end{figure}
-
-%We fit model \eqref{eqn:mixed} to the survey data obtained from the simulation experiment.
-The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}.
-Model \ref{eqn:mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}.
-The plot includes 20 randomly selected subject-specific power curves.
-Notice that the power curve estimated for one subject is above the UMP test power curve.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
-%       \caption{Estimated subject specific power curve from model \ref{eqn:mixed} for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_subject}
-%\end{figure}
-
 \subsection{Subject-specific variation}
 
-%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
 \blue{Subject-specific proportion correct $\hat p_i(E)$ is obtained using Equation \ref{eqn:mixed_power} and it is used to obtain power for $K=5$}.
@@
 
-%\green{third experiment doesn't really have a p-value, does it?
-Need to explain this.
-Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.
-So many small p-values are coming from the fact that contamination makes the slope negatively significant.
-For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant.
-It is like the effect of influential outlier.}
-
 In the real setting, where visual inference is to be useful, there will be no conventional test $p$-values.
@@
 The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate,
-% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
-% 
+
  giving a clear indication to reject $H_0$ or not.
@@
 This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
- % for which researchers get confused whether to reject or 'not reject' $H_0$. 
+
  For visual tests this is not common.
@@
        \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . Values are shown on square root scale. }
-    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
        \label{fig:pval_definition}
@@
 
-%We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}. 
-%The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
-For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2.
-This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.
-
 For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are.
 This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
-% and the visual test does the job nicely.
-
-%With multiple subjects it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
-The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
-Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more subjects select that plot.
-Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.
-
-%Similary, the strength of the signal, is calculated using the proportion of subjects picking actual data plot.
-It is closely related to the power of the test.
-From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
-This gives the signal strength estimate as
-%\[
-%\hat p_D= (1- {u}/{K})/({m-1})
-%\]
-%where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
-For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions.
-The blue line is a loess fit to the data, and the grey line indicates equality of the two axes.
-The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
-As the conventional $p$-value increases, the estimated signal strength increases -- subjects find it harder to detect the actual data plot at larger $p$-values.
-The rise is very sharp for experiment 1, which indicates that subjects have a harder time detecting the actual data plot for even small $p$-values.
-
-%\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
-% \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
-%
-%\green{Explain why the term ``plot signal'' is introduced}
-
-%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
-For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
-As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
-When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
-We observed this for both experiment 1 and 2.
-For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
-
-%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
-
-%Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
-This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a conventional sense.
-Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
-%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.
-}  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.
-But it is true for experiment 2, right?.
-}
-%is equivalent to conventional test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
-%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
-%       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
-The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
-p-values are shown on square root scale scale.}
-%    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-%       \label{fig:pval_plot_signal}
-%\end{figure*}
 
@@
 Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.
-
-%Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
-Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
-(Empty cells indicate no lineup shown with this combination.)
 
@@
 So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%      \includegraphics[width=0.8\textwidth]{p_val_log_counts-a.pdf}
-%       \includegraphics[width=0.8\textwidth]{p_val_log_counts-b.pdf}
-%       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
-%       \label{fig:P-val_log}
-%\end{figure*}
 

### 2013-03-20 R/paper_jasa.R

@@ -936,7 +936,7 @@ ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdfk) +
   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta)) +
   theme(legend.key.width = unit(1.5, "cm"))
 
-ggsave("../images/power_expected_k.pdf", width=8, height=5, units="in")
+ggsave("../images/power_expected_k.pdf", width=7.5, height=5, units="in")
 
 # ====================== Turk2 conventional power calculations  
 

### 2013-03-28 paper_jasa.tex
some small changes to reduce size

@@
 A human subjects experiment is conducted using simulated data to provide controlled conditions.
-\blue{Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional test are violated.
+Results suggest that the lineup protocol performs comparably with the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional test are violated.
 Surprisingly, visual tests have higher power than the conventional tests when the effect size is large.
-And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks.}
+And, interestingly, there may be some super-visual individuals who yield better performance and power than the conventional test even in the most difficult tasks.
 
@@
 This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference.
-The intended use of lineups is where there is no existing test, \blue{and unlikely ever to be any numerical test}.
+The intended use of lineups is where there is no existing test, and unlikely ever to be any numerical test.
 The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
@@
 
-\blue{An illustration of the lineup protocol in relation to conventional hypothesis testing is presented in Table \ref{tbl:compare}.} Both methods start from the same place, the same set of hypotheses.
+An illustration of the lineup protocol in relation to conventional hypothesis testing is presented in Table \ref{tbl:compare}.
+Both methods start from the same place, the same set of hypotheses.
 The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
@@
 \begin{tabular}{llll} 
-\hline \hline
+\hline % \hline
   & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
@@
 \end{table*}
-
-%Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
-
-%Power of a plot and $p$-value have been derived in Buja and Wickham as ...
-
-%for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
-
-%This section outlines the concepts of visual inference in comparison to the procedures of conventional statistical inference.
-Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
 
@@
 The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
-$T(y)$ is assessed relative to null values of this statistic $T(y_0)$ which would be possible values of $T$ if $\theta \in \Theta$.
-%Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
+$T(y)$ is assessed relative to null values of this statistic $T(y_0)$, the possible values of $T$ if $\theta \in \Theta$.
+
 For visual inference, unlike in the conventional hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
@@
 
-\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
+\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type and specification of aesthetics, necessary for complete reproducibility.
 The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
@@
 A \textbf{lineup} is a layout of $m$ randomly placed visual statistics, consisting of 
-\begin{itemize}\itemsep-3pt
+\begin{itemize}\itemsep-10pt
 \item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
@@
 If the test statistic is identified in the lineup, this is considered as evidence against the null hypothesis.
-%If it is not identified the conclusion is to not reject the null hypothesis.
-
 This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
@@
 being larger than $P(X \ge 1)$ in this situation.
-If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
+By increasing either $m$ or $K$, the value at a higher precision can be determined.
 Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
@@
 The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
-\begin{itemize}\itemsep-3pt
+\begin{itemize}\itemsep-10pt
 \item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot, and
@@
 \noindent
-Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:
+Associated with any test there is the risk of  Type I or II errors, which for visual inference are defined as follows:
 
 \begin{dfn}\label{dfn:error}
-The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.\\ \noindent
+The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
 The \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
@@
 Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
-Let $p_i$ be the probability that individual $i$ picks out the actual data plot from the lineup.
-If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
+Let $p_i$ be the probability that individual $i$ picks out the actual data plot.
+If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot.
 
@@
 \[
-\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
+\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1).
 \] 
@@
 \[
-Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
+Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1).
 \]
-%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
 If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
@@
 
-Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
-Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.
+Power is an important consideration in deciding which test to use for solving a problem.
+Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in choosing plots to use.
 Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
@@
 
-%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
-The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
+The model can be  fit as:
 \begin{equation} \label{eqn:mixed}
@@
 
-%\red{This model is  described in more detail in equation \eqref{eqn:mixed}.}
-
-%\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability
-%\begin{equation}\label{eqn:pvalue}
-%Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
-%\end{equation}
-%where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
-%\begin{dfn} \label{dfn:pval}
-%The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
-%where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
-Since $u$ is a discrete value with boundaries 0 and $K$,
-% we can also re-formulate the $p$-value in terms of upper and lower boundaries:
-% \begin{equation*}
-% \begin{array}{rcll}
-% P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
-%( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
-% \end{array}
-% \end{equation*}
-% }
-
-%\red{HH: Tried to re-write it to answer the question below.
-%Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
-%
-%    \begin{equation*}
-%        \begin{cases} 
-%              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
-%              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-
-%\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
-%\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
-
-%\noindent
 \subsection{Lineup size, $m$}
@@
 \begin{lemma}~\label{lemma}
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic.
-Suppose the associated test statistic in a conventional test is observed as  $t_{obs}$ with $p$-value $p_D$.
+Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic. Suppose the associated test statistic is observed as  $t_{obs}$ with $p$-value $p_D$.
 
@@
 
-\begin{proof} The proof and details of the lemma is attached in the supplementary documents. 
+\begin{proof} The proof and the details of the lemma are attached to the supplementary documents. 
 \end{proof}
@@
 This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}.
-\blue {Thus the expected probability to reject by a single observer ($K=1$) in this scenario is
+Thus the expected probability to reject by a single observer ($K=1$) in this scenario is
 
@@
 Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger, and when $\beta$ gets larger.
-} \hh{Conversely, visual power is below conventional power for parameter values close to the null hypothesis.
+Conversely, visual power is below conventional power for parameter values close to the null hypothesis.
 This is even more pronounced for large number of observers.
@@
 This dichotomy of visual power can also be seen in more detail in figure \ref{fig:power_loess_effect}.
-}
 
@@
 The actual data plot was randomly placed among these null data plots to produce the lineup.
-\blue {Figure \ref{fig:test_category} is an example of one of these lineups.
+Figure \ref{fig:test_category} is an example of one of these lineups.
 It was generated for $n$=300, $\beta_2$=10 and $\sigma$=12.
@@
 For this lineup, 15 out of 16 observers picked the actual data plot.
-}
 
@@
 
-The number of evaluations required for each lineup to provide reasonable estimates of the \blue{ proportion correct ($\hat p$) is determined by the variance of the number of correct evaluations}.
+The number of evaluations required for each lineup to provide reasonable estimates of the proportion correct ($\hat p$) is determined by the variance of the number of correct evaluations.
 Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected} with $K=1$) \blue{we consider $\gamma =p$}.
+Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected} with $K=1$) we consider $\gamma =p$.
 For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
@@
 
-%For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in Table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
-We collect data such that each of the parameter combinations in the Table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
-
-%\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
-This produces an empirical distribution of p-value under alternative hypothesis.
-Finally, three data sets that have 16th, 50th and 84th quantile of the p-values were taken.
-Similar procedure was used to generate 19 null plots based on the empirical distribution of p-value under null hypothesis}.
-For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
-A 1000 sets were generated for each parameter combination and the conventional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
-The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
-Additional control was applied to the 19 null plots.
-Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
-
 \subsection{Continuous covariate}  \label{sec:continuous}
@@
 We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
-\blue{For each combination of parameters, at least three different actual data sets are produced, yielding a total of 70 lineups.
-}
-
-%For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
+For each combination of parameters, at least three different actual data sets are produced, yielding a total of 70 lineups.
 
@@
 This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-\blue{The actual data plot location is $(2^2+1)$.
-None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup.}
+The actual data plot location is $(2^2+1)$.
+None of the 65 observers picked the actual plot while 46 observers picked plot 18 which has the lowest $p$-value among all the plots in this lineup.
 
@@
 Each subject saw a substantial number of difficult lineups.
-
-%\hh{While this allows us a better estimation for the power in small slopes, }
-%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
-%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
-%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
-%	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
-
-%{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
-For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in Table \ref{tbl:experiment_params}.
-For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
-The 10 lineups are chosen as follows;
-
-%\begin{enumerate}
-%	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
-%	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
-%	\item Finally select another lineup with difficulty level 0
-%\end{enumerate}
 
@@
 Can you pick which plot is different?
-\blue{The actual data plot location is $(3^2-2^3)$ and 13 out of 31 observers picked the actual plot.}
+The actual data plot location is $(3^2-2^3)$ and 13 out of 31 observers picked the actual plot.
 
@@
 Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-\blue {A summary of all the data along with cleaned data are added to the supplementary material.}
-
-%
-%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
-The data of participants who did not answer correctly was removed.
-%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
-That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.
-
-%Amazon Mechanical Turk data suffers from the problem that some participants are trying to 'game' the system, because they are being paid for their efforts, not much by the standards of the US minimum wage, but enough to encourage honest efforts.
-in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
-In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.
-%There are six suggested approach to clean the data. 
-%They are
-%\begin{enumerate}
-%\item {\bf include all participants} and their evaluations
-%\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-%\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-%\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
-%\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
-%\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
-If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-%\end{enumerate}
-
-%\begin{table}[hbtp]
-%\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-%Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
-This is due to missing demographic information.
-\hh{Do we need this table or could we make this a visual comparison?
-For me, there's too many numbers with too many digits.}}
-%\begin{center}
-%\begin{tabular}{rrrrr|rrrr|rrrr}
-%  \hline
-%Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
-% \cline{2-5}  \cline{6-9}   \cline{10-13}
-%criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-%  \hline
-%1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-%  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-%  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-%  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-%  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-%  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:summary}
-%\end{table}
+A summary of all the data along with cleaned data are added to the supplementary material.
 
 \begin{table}[hbtp]
-\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets.
+\caption{Number of subjects, gender, total lineups seen and distinct lineups for all three experimental data sets.
 Note that in some of the lineups the number of male and female participants does not add up to the total number of participants due to missing demographic information.
-}%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
+}
 \begin{center}
@@
 
-For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power to be fixed at 0.05 for $E=0$ is not \blue{ required by the experimental design}.
+For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power to be fixed at 0.05 for $E=0$ is not required by the experimental design.
 
@@
 Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
-\blue {A loess  fit to the data gives an estimate of  the observed proportion correct $\hat p(E)$ for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
+A loess  fit to the data gives an estimate of  the observed proportion correct $\hat p(E)$ for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
 $\hat p(E)$ is considered to be the power for $K=1$ and it is used to obtain power for $K=5$.
@@
 In experiment 3 the power of the visual test exceeds that for the conventional test, as expected.
-For larger values of $K$ estimated power exceeds the power of conventional test.}
+For larger values of $K$ estimated power exceeds the power of conventional test.
 Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
@@
 
-\blue{Subject-specific proportion correct $\hat p_i(E)$ is obtained using Equation \ref{eqn:mixed_power} and it is used to obtain power for $K=5$}.
+Subject-specific proportion correct $\hat p_i(E)$ is obtained using Equation \ref{eqn:mixed_power} and it is used to obtain power for $K=5$.
 Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
@@
 Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.
-The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate,
-
- giving a clear indication to reject $H_0$ or not.
+The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate, giving a clear indication to reject $H_0$ or not.
 This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
@@
        \includegraphics[width=0.8\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{ \blue {Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all individual lineups of both experiment 1 and 2.
+       \caption{ Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all individual lineups of both experiment 1 and 2.
 Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
@@
 Highest counts tend to be the plot in the lineup having the lowest $p$-value, more so for experiment 2 than 1.
-    }   }
+    }   
        \label{fig:P-val_log2}
@@
 For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
-%(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
-
-%Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
-Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
-Rather it is assumed that the subject has the ability to select the plot with the most extreme structure.
-A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, \sigma=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
-Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.
-
-%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in Table \ref{tbl:pval_assumption} for both experiment 1 and 2.
-The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
-We see the same pattern in Figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value. The data used is from screening criteria 1. } 
-%\begin{center}
-%\begin{tabular}{cccc}
-%  \hline
-%screening & Experiment & Total &  minimum $p$-value\\ 
-%criteria & & lineups & picked by the most \\
-%  \hline
-%   & 1 &  60 &  37 \\ [-1ex]
-%    \raisebox{1.5ex}{1} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%    \raisebox{1.5ex}{2} & 2 &  70 &  61 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%   \raisebox{1.5ex} {3} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  41 \\  [-1ex]
-%    \raisebox{1.5ex}{4} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  39 \\  [-1ex]
-%    \raisebox{1.5ex}{5} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  40 \\  [-1ex]
-%    \raisebox{1.5ex}{6} & 2 &  70 &  60 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
-%\begin{center}
-%\begin{tabular}{ccc} \hline
-%screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
-% \cline{2-3}
-%criteria & Experiment 1 & Experiment 2 \\ 
-%  \hline
-%1 & 37 & 60 \\ 
-%2 & 38 & 61 \\  
-%3 & 38 & 60 \\ 
-%4 & 41 & 60 \\ 
-%5 & 39 & 60 \\
-%6 & 40 & 60 \\ 
-%\hline
-%Total lineups & 60 & 70 \\  
-%\hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:pval_assumption} 
-%\end{table}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
-%       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
-This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
-There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error.
-Can we?
-\green{Maybe don't need this plot}}
-%       \label{fig:P-val_rank}
-%\end{figure*}
 
@@
 
-%\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
-If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
-Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
-The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
-A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
-If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot.
-When there is a large positive difference, subjects fail to pick the correct plot.
-If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
-That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.}
-
-%\green{Something is definitely not right!
-The code looks like it does actual data p-value - min(null $p$-value).
-But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically!
-Suppose data plot $p$-value is larger than 0.8.
-Under null hypothesis, $p$-value $\sim U(0,1)$.
-Thus min(null $p$-value) could be close to zero by chance producing this large difference.
-And this has happened in both experiment 1 and 2.}
-
- %\green{Is this really pct correct, or 1-pct correct?
-\blue{This is proportion correct}.
-When there is a big diff in p-value almost no-one gets it right!
-Maybe it is just a scale issue?} \blue{In fact this is what we expect.
-If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small.
-This should lead people to pick the wrong plot.
-Our minimum $p$-value assumption basically says this fact, right?}.
-We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.
-
-%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
-Thanks!
-
 In practice, the $p$-value is not going to be a valid way to compare plots.
@@
 
-%Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice:
-%
-%
-%\begin{enumerate} \itemsep 0in
-%\item Most different plot
-%\item Visible trend
-%\item Clustering visible
-%\item Other
-%\end{enumerate}
-%
-%
-%\noindent with the possibility to use more than one.
-
-\blue{Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.} The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
+Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.
+The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
 Reasons 1 or 3 would be indicative of Type III error.
@@
 
-%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
-Figure \ref{fig:choice_reason} examines this.
-For observers who chose the actual data plot, the proportions for the three reasons are shown.
-It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
-\green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}
-
-%This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}
-
-%\green{Not sure that this goes here, but something needs to be said about Type III.}
-
-%\blue{We also see this feature in figure \ref{fig:choice_reason}}.
-%\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
-
 \begin{figure}[hbtp]
    \centering
-%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
-%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
        \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
@@
 
-%\begin{itemize} \itemsep 0in
-%\item Summarize the key pieces of the paper again, especially the major findings
-
 It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
@@
 The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-\blue{If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.} Follow-up questions on the reasons provide qualitative insight.
+If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
+Follow-up questions on the reasons provide qualitative insight.
 Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.

### 2013-04-04 paper_jasa.tex
section 6.7 and 6.8 reduced

@@
 However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
-
-We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
+If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
+The detailed discussion is attached to the supplementary material.
+
+%We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
 What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
@@
 
-In practice, the $p$-value is not going to be a valid way to compare plots.
+%In practice, the $p$-value is not going to be a valid way to compare plots.
 Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{pval_difference.pdf}}
-       \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.60}{\includegraphics{pval_difference.pdf}}
+%       \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
 Negative differences indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
 Difference close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
-       \label{fig:pval_difference}
-\end{figure}
+%       \label{fig:pval_difference}
+%\end{figure}
 
@@
 This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
-This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
+
+%This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
 Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
 
-Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+%Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
@@
 The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
-At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
+
+%At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
 ``Visible trend'' comes in third.
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
-       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Reasons of plot choices vs  proportion of times the subjects correctly chose the actual data plot for experiment 3 that examines the occurrence of Type III error.
+%\begin{figure}[hbtp]
+%   \centering
+%       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
+%       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
+%       \caption{Reasons of plot choices vs  proportion of times the subjects correctly chose the actual data plot for experiment 3 that examines the occurrence of Type III error.
 At left, all subjects' choices are shown, and reason 123 means all three reasons are used.
@@
 Size of the point corresponds to the number of subjects using that reason.}
-       \label{fig:choice_reason}
-\end{figure}
+%       \label{fig:choice_reason}
+%\end{figure}
 
@@
 This does show the importance of the qualitative answers that observers give for their choices.
+The detailed discussion on this is attached to the supplementary documents.
 

### 2013-04-04 supplementary.tex
section 6.7 and 6.8 reduced

@@
 
+\section{How much do null plots affect the choice?}
+
+ Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
+ But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
+How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
+If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
+Measuring the difference, generally, between  plots is almost impossible.
+However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
+
+We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
+What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
+If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
+Figure \ref{fig:pval_difference} investigates this.
+The difference between the $p$-value of the actual data is compared with the lowest from the null plots.
+This is plotted horizontally, and the proportion correct is plotted vertically.
+Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
+In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
+This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
+Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
+Figure \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
+When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
+
+In practice, the $p$-value is not going to be a valid way to compare plots.
+Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
+This is investigated in \citet{niladri:2012}.
+Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
+
+\begin{figure}[hbtp]
+   \centering
+       \scalebox{0.60}{\includegraphics{pval_difference.pdf}}
+       \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
+Negative differences indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
+Difference close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
+       \label{fig:pval_difference}
+\end{figure}
+
+\section{Type III error}~\label{sec:TypeIII}
+
+A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
+This occurs when the null hypothesis is correctly rejected but for the wrong reason.
+Experiment 3 is prone to this type of error.
+Observers were asked to identify the plot that had the largest slope, either positive or negative.
+But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
+For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
+This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
+This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
+Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
+
+Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
+Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.
+The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
+Reasons 1 or 3 would be indicative of Type III error.
+Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
+If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
+This is not what we see.
+At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
+The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
+The size of the point corresponds to the number of subjects putting this as the reason.
+Subjects that chose all three reasons almost always chose the actual data plot.
+This was followed by using 1 and 3, and then 1 and 4.
+The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
+At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
+``Visible trend'' comes in third.
+This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
+
+\begin{figure}[hbtp]
+   \centering
+       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
+       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
+       \caption{Reasons of plot choices vs  proportion of times the subjects correctly chose the actual data plot for experiment 3 that examines the occurrence of Type III error.
+At left, all subjects' choices are shown, and reason 123 means all three reasons are used.
+At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
+Size of the point corresponds to the number of subjects using that reason.}
+       \label{fig:choice_reason}
+\end{figure}
+
+For visual inference, making a Type III, is not actually a problem.
+It is only a possibility in this experiment because we are working with known structure.
+In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
+This does show the importance of the qualitative answers that observers give for their choices.
+
 \section{Data Cleaning}

### 2013-04-04 R/paper_jasa.R

@@ -931,7 +931,7 @@ powerdfk$test <- factor(powerdfk$test, levels(powerdfk$test)[c(1,2,4,3)])
 
 
 ggplot(aes(beta, power, linetype=test, size=emphasize), data=powerdfk) + 
-  geom_smooth( method="loess", span=0.1) + 
+  geom_smooth( method="loess", span=0.1, color="Black") + 
   scale_size_identity() + ylab("Power") +
   scale_linetype_discrete("Test") + scale_x_continuous(expression(beta)) +
   theme(legend.key.width = unit(1.5, "cm"))

### 2013-04-04 paper_jasa.tex
tiny change
conclusion is reduced

@@
 However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
-If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
+If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we found that subjects have a harder time detecting the actual data plot.
 The detailed discussion is attached to the supplementary material.
@@
 
-It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
 This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
+For this, specific terminologies were defined.
+The methods of obtaining $p$-value and estimating power of visual tests were introduced.
+If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
+Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
 The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
 The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-
-Power of visual tests can be estimated in different ways, which was explained in this paper.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
-
-%\item Explain caveats, nuances, advantages with the work
-
-One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
+Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure enough power.
+
+When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
+Lineups are successfully used to evaluate the power of competing graphical designs in \cite{heike:2012}.
+This method is also applied in \citet{niladri:2012}.
+It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
+The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
+If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
+Follow-up questions on the reasons provide qualitative insight.
+Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
+
+%One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
 A good test statistic is expected to reveal structure if the null hypothesis is not true.
@@
 This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
-%This is the part of our ongoing and future research.
-
-The question that is posed to observers can affect the ability to evaluate lineups.
-In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
-
-% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
-If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
-This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
-% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
-
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
-
-Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
-Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet, the results were good.
-We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
 
@@
 
+%\item Explain caveats, nuances, advantages with the work
+
+ %This is the part of our ongoing and future research.
+
+%The question that is posed to observers can affect the ability to evaluate lineups.
+In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
+In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
+
+% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
+If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
+This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
+
+% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
+
+%Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
+Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
+Yet, the results were good.
+We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
+A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
+
 %\item Describe the next steps of the work, and possible ideas for future extensions
-
-When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
-We have successfully used the procedure informally for local projects.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
-Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
 
@@
 \end{document}
+
+\section{Old Conclusions}
+
+It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
+This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
+The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
+The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
+
+Power of visual tests can be estimated in different ways, which was explained in this paper.
+If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
+Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
+
+%\item Explain caveats, nuances, advantages with the work
+
+One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
+A good test statistic is expected to reveal structure if the null hypothesis is not true.
+Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
+This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
+%This is the part of our ongoing and future research.
+
+The question that is posed to observers can affect the ability to evaluate lineups.
+In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
+In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
+
+% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
+If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
+This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
+
+% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
+
+Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
+
+Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
+Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
+Yet, the results were good.
+We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
+A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
+
+The performance of subjects was quite varied, but consistent.
+No restrictions were placed on Turk workers, in terms of abilities.
+There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test.
+ It would be interesting to see how well trained subjects might perform.
+Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference.
+Future work might explore this.
+
+%\item Describe the next steps of the work, and possible ideas for future extensions
+
+When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
+We have successfully used the procedure informally for local projects.
+It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
+The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
+If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
+Follow-up questions on the reasons provide qualitative insight.
+Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
+
+%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
+The direction of future research would be to study that and apply the methods in those areas.
+
+%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
+We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
+The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
+For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
+Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.
+
+%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
+Along the lines of the Army tests that Susan brought to the working group.
+\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
+Thus anyone from any background could participate the experiment.
+It would be interesting to see how well the trained subjects performs.
+Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
+One of our future work would be to examine this in details.
+For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
+After being trained they may have to pass the qualification test before participating the experiment.
+Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
+
+%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
+
+%In practice there is no correct answer.
+In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
+In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
+And subjects would provide open-ended answers for their choice of plot.
+
+%What are your suggestions for future experiments, and next steps of the research?

### 2013-04-04 paper_jasa.tex

@@
 How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
-If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
+If the actual data plot is very different from all of the null plots, then the null plots should not have much influence on the choice.
 Measuring the difference, generally, between  plots is almost impossible.
 However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
-If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we found that subjects have a harder time detecting the actual data plot.
+If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
 The detailed discussion is attached to the supplementary material.

### 2013-04-11 paper_jasa.tex
couple of typos up to page 13

@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-A new method, called the lineup protocol, enables statistical significance testing of visual findings, which bridges a big gulf between exploratory and inferential statistics.
-In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
+The lineup protocol enables statistical significance testing of visual findings, bridging the  gulf between exploratory and inferential statistics.
+In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests in scenarios when a conventional test exists.
 This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
@@
 If it is extreme on this scale then the null hypothesis is rejected.
-In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
+In contrast in visual inference, the plot of the data is compared with a set of plots of  samples drawn from the null distribution.
 If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.
@@
 Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
-It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
+%It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
 
@@
 
-\begin{dfn}\label{dfn:test}
+\begin{dfn}\label{dfn:visualtest}
 The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
@@
 Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
-Let $p_i$ be the probability that individual $i$ picks out the actual data plot.
+Let $p_i$ be the probability that individual~$i$ picks out the actual data plot.
 If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot.
@@
 
-\begin{proof} The proof and the details of the lemma are attached to the supplementary documents. 
+\begin{proof} The proof and the details of the lemma are attached in the supplementary documents. 
 \end{proof}
@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
-For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot residuals obtained from fitting the model without $X_j$ against $X_j$ or for a single covariate we may plot $Y$ against $X_j$ (cases 1-4 in Table~\ref{tbl:stat_multiple}).
+To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5 in Table~\ref{tbl:stat_multiple}).
+For the purpose of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable, respectively.
 
@@
 1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid.
-For lineup plot, we simulate data from fitted null model.
+For null plots   we simulate data from fitted null model.
 \\
 2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+For null plots   we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+3 & $H_0: \beta_k=0$ (for binary $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+For null plots   we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
-For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
+4 & $H_0: \beta_k=0$ (interaction of continuous and  binary $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
+For null plots   we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
 5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
-For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
+For null plots   we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
 6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
-For lineup plot, we simulate data from standard normal.
+For null plots   we simulate data from standard normal.
 \\
 7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
-For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
+For null plots   we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
 \\
 8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data.
-For lineup plot, we simulate data from fitted model.
+For null plots   we simulate data from fitted model.
 \\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
+9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For null plots   we simulate data with correlation $\rho$.\\
 \hline 
@@
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test for different $K$ (number of observers), obtained using $p(\beta)$ from Equation \ref{power_exp}.
-Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger, and when $\beta$ gets larger.
+Notice that the expected power of the visual test exceeds the power of the conventional test as $K$  increases, and when $\beta$ gets larger.
 Conversely, visual power is below conventional power for parameter values close to the null hypothesis.
@@
 At the same time, the point of intersection between visual and conventional power approaches the value of the null hypothesis as the number of observers approaches infinity, leading to an asymptotically perfect power curve of zero in the null hypothesis and one for any alternative value.
-This dichotomy of visual power can also be seen in more detail in figure \ref{fig:power_loess_effect}.
+We observe this dichotomy of visual power  in the data collected from user experiments, too.
+It features prominently in  figure \ref{fig:power_loess_effect}.
 
@@
 Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression setting.
-The first two experiments have conventional ideal scenarios where we would not expect the lineup protocol to do better than the conventional test.
+The first two experiments have  ideal scenarios for conventional testing, where we would not expect the lineup protocol to do better than the conventional test.
 The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.

### 2013-04-11 paper_jasa.tex
working on conclusions

--- 2013-04-04 paper_jasa.tex
@@
 
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
-For this, specific terminologies were defined.
-The methods of obtaining $p$-value and estimating power of visual tests were introduced.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
-The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure enough power.
-
-When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
-Lineups are successfully used to evaluate the power of competing graphical designs in \cite{heike:2012}.
-This method is also applied in \citet{niladri:2012}.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
+This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009}.
+Specific terminologies were defined, and methods for obtaining the $p$-value and estimating the power of visual tests were introduced.
+It was assumed that observers select the plot having the strongest signal in the lineup, in order to calculate the theoretical power,  and the experimental data was consistent with this assumption.
+The simulation study showed that the results obtained by visual inference are comparable to those obtained by conventional inference, in the setting where conventional testing can be done.
+Visual inference is intended to provide valid tests where no conventional test exists, and our experiments in a controlled scenario suggest that it will perform as expected in the intended applications.
+The power of a visual test increases with the number of observers, which interestingly, leads to a result that the theoretical power of visual test can be better than that of conventional test.
+
+The lineup protocol operates analogously to statistical tests that have broad alternative hypotheses.
 If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
@@
 Future work might explore this.
+
+Lineups have been successfully applied to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
+It is hoped that this proves to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 

### 2013-04-12 paper_jasa.tex
typos

--- 2013-04-11 paper_jasa.tex
@@
 For example, to examine the effect of variable $X_j$ on $Y$, we would plot residuals obtained from fitting the model without $X_j$ against $X_j$ or for a single covariate we may plot $Y$ against $X_j$ (cases 1-4 in Table~\ref{tbl:stat_multiple}).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5 in Table~\ref{tbl:stat_multiple}).
+To assess whether the  assumptions of linearity is appropriate we would draw a plot of residuals against fitted values (case 5 in Table~\ref{tbl:stat_multiple}).
 For the purpose of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable, respectively.
@@
 \\
-7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
+7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of residuals obtained by fitting partial regression.
 For null plots   we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
@@
 \\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For null plots   we simulate data with correlation $\rho$.\\
+9 & For  $p=1$ only: $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid.
+For null plots   we simulate data with correlation $\rho$.\\
 \hline 
@@
 At the same time, the point of intersection between visual and conventional power approaches the value of the null hypothesis as the number of observers approaches infinity, leading to an asymptotically perfect power curve of zero in the null hypothesis and one for any alternative value.
-We observe this dichotomy of visual power  in the data collected from user experiments, too.
+We observe this dichotomy of visual power  in power estimates based on the data collected from user experiments, too.
 It features prominently in  figure \ref{fig:power_loess_effect}.
@@
 Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
-Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
-Three replicates at each level were generated.
-These produced 60 different ``actual data sets'', and thus, 60 different lineups.
-To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
-The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3).
+Three data sets were generated for each of the parameter values shown in Table \ref{tbl:experiment_params} resulting in 60 different ``actual data sets'', and thus, 60 different lineups.
+For each lineup, the null model was fit to the actual data set to obtain residuals and parameter estimates.
+The actual data plot was drawn as  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3).
 The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way.
@@
 % Entering 1st row
-&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
+&  \phantom{0}5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
 \raisebox{1.5ex}{100} &12
@@
 % Entering 2rd row
-&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
+&  \phantom{0}5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
 \raisebox{1.5ex}{300} & 12
@@
 
-This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2.
+This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case~2.
 Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}.
-We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
+We arbitrarily set $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
 For each combination of parameters, at least three different actual data sets are produced, yielding a total of 70 lineups.
@@
 Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.
-This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
+This lineup is generated using a sample size ($n$) of 100, slope ($\beta$) of 1.25 and and error SD ($\sigma$) of 5.
 The actual data plot location is $(2^2+1)$.
@@
 
-For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, giving 5 lineups.
-Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
-The order that the lineups are shown to participants is randomized.
-Each subject saw a substantial number of difficult lineups.
+For each combination of sample size and standard deviation, each participant is given five randomly selected lineups, one of each difficulty level.
+Another set of four lineups is chosen from a second tier of selected combinations of sample size and standard deviation, with difficulty levels 0 to 3.
+A last lineup was randomly selected from a set of  lineups with difficulty level 0.
+The order in which the lineups are shown to participants is randomized.
+%Each subject saw a substantial number of difficult lineups.
 
@@
 In these situations there exists a test, and there would, in general, be no need to use visual inference.
-The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
+The simulation is conducted in the hope that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
  This third simulation is closer to the mark for the purpose of visual inference.
-The assumptions for the conventional test are violated, by contaminating the data.
+The assumptions for the conventional test are violated by contaminating the data.
 The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.
@@
 %{\bf Screening scenarios:}
-Amazon Mechanical Turk workers are paid for their efforts, not substantially, but on the scale of the minimum wage of the USA.
+Amazon Mechanical Turk workers are paid for their efforts, not substantially, but on the scale of the minimum wage in the USA.
 Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data.
 For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
-For the purpose of identifying these participants and cleaning the data, we take action based on a subject's answer to one of the very easy lineups that everybody was exposed to:
-if the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
-If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
-Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
+For the purpose of identifying these participants and cleaning the data, we use one of the very easy lineups that everybody was exposed to as a {\it reference lineup} and  take action based on a subject's answer to this reference:
+if the subject failed to identify the actual data plot on the reference lineup, we remove all of this subject's data from the analysis.
+If the answer on the reference lineup is correct, we remove the answer for this lineup from the analysis, but keep all of the remaining answers.
+Table~\ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
 A summary of all the data along with cleaned data are added to the supplementary material.
@@
 \subsection{Model fitting}
- For each parameter combination, the effect is computed as 
-\[
-E=\sqrt {n} \cdot \beta/\sigma.
-\]
+ For each parameter combination, {\it effect} $E$ is derived as 
+%\[
+$E=\sqrt {n} \cdot \beta/\sigma.$
+%\]
 \noindent
@@
        \scalebox{0.55}{\includegraphics{power_loess_effect_k.pdf}}
-       \caption{Power in comparison to effect for the three experiments.
+       \caption{\hh{Make the legend for the lines longer - just as in fig, 3} Power in comparison to effect for the three experiments.
 Points indicate subject responses, with size indicating count.
@@
 For experiment 3, conventional power is based on the slopes of the non-contaminated part of the data.
-Power for contaminated data is shown by cross mark.}
+Power of the conventional test for contaminated data is shown by cross marks.}
        \label{fig:power_loess_effect}
@@
 
-Results of experiment 3 are curious insofar, as  power of the visual test is  high independent of effect size.
+Results of experiment 3 are curious insofar, as  power of the visual test is  largely independent of effect size.
 However, these results are based on correct identification of the actual data plot, regardless of reason.
@@
        \scalebox{0.6}{\includegraphics{p_val_prop_correct.pdf}}
-       \caption{Proportion of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
+       \caption{Proportion of correct responses decreases rapidly with  increasing $p$-values.
+For $p$-values above 0.15 it becomes very unlikely that observers identify the actual plot.
 The theoretical justification of this is shown in Figure \ref{fig:pval_power}.
@@
 The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate, giving a clear indication to reject $H_0$ or not.
-This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
-This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
-
+This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment~2.
+This guides the researcher to make decision confidently while conventional tests with marginal $p$-values make the decision whether to reject or not harder.
 For visual tests this is not common.
@@
 For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are.
-This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
+This is expected as the conventional test loses its power to reject $H_0$ even when the alternative is true, whereas the visual test performs well.
 
@@
 
- One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
+ One assumption made in order to evaluate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
 In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
-This can be examined for the data collected, to see if this assumption is, indeed, reasonable.
+We examine  the data collected from the first two experiments, to see if this assumption is, indeed, reasonable.
 %In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value.
@@
 
-Figure \ref{fig:P-val_log2} shows the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
-In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically.
+\hh{Figure \ref{fig:P-val_log2} gives an overview of all selections in all lineups of experiments 1 and 2.
+Each panel of the figure corresponds to a single lineup.
+Each `pin' -- a short line topped by a dot -- corresponds to one plot in the lineup.
+The $x$-location of the pin shows the plot's $p$-value on a log scale, its height  is given by the number of observer choosing this plot.
+}
+%Figure \ref{fig:P-val_log2} shows the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
+In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the number of subjects who selected this plot is displayed vertically.
 The dot represents this point, and the lines connect vertically with 0 count.
@@
 Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.
-In both experiments  people tended to select the plot with the lowest $p$-value with an increase in effect size.
+In both experiments  people tended to select the plot with the lowest $p$-value.
+% with an increase in effect size.
  %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
@@
 Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data.
-But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
-How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
+Unlike randomization tests visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
+How this small set of null plots influences the subjects' choice is important for understanding the reliability of visual inference.
 If the actual data plot is very different from all of the null plots, then the null plots should not have much influence on the choice.
@@
 If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
-The detailed discussion is attached to the supplementary material.
+The detailed discussion is attached in the supplementary material.
 
@@
 A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
-This occurs when the null hypothesis is correctly rejected but for the wrong reason.
+Type III errors are defined as the probability of  correctly rejecting  the null hypothesis but for the wrong reason.
 Experiment 3 is prone to this type of error.
-Observers were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
-For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
-This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
+Participants were asked to identify the plot with the largest absolute slope.
+But the actual data plot featured a cluster of points, the contamination that made the conventional test  fail to see any trend.
+For the human eye this cluster of points is  as visible as the association between the remaining points, enabling the observer to identify the actual data plot by looking for the cluster instead of the slope.
+This would be considered  a Type III error because it leads to a correct rejection of the null hypothesis, but is not related to the value of the slope parameter.
 
@@
 In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
-This does show the importance of the qualitative answers that observers give for their choices.
-The detailed discussion on this is attached to the supplementary documents.
+However, this highlights the importance of getting qualitative reasoning from observers  for their choices.
+A more detailed discussion can be found in the supplementary documents.
 
@@
 
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
-For this, specific terminologies were defined.
-The methods of obtaining $p$-value and estimating power of visual tests were introduced.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
-The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure enough power.
-
-When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
-Lineups are successfully used to evaluate the power of competing graphical designs in \cite{heike:2012}.
-This method is also applied in \citet{niladri:2012}.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
+This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009}.
+Specific terminology was defined, and methods for obtaining the $p$-value and estimating the power of visual tests were introduced.
+It was assumed that observers select the plot having the strongest signal in the lineup, in order to calculate the theoretical power,  and the experimental data proved to be consistent with this assumption.
+The simulation study showed that  results obtained by visual inference are comparable to those obtained by conventional inference, in the setting where %conventional testing can be done.
+Visual inference is intended to provide valid tests where
+no conventional test exists, and our experiments in a controlled scenario suggest that it will perform as expected in the intended applications.
+The power of a visual test increases with the number of observers, which interestingly, leads to a result that the theoretical power of visual test can be better than that of conventional tests.
+
+The lineup protocol operates similarly to statistical tests that have broad alternative hypotheses.
 If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
 Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
+In conventional testing, multiple comparisons are often done to refine and understand the test results, and perhaps some similar approaches might be developed for visual inference.
 
@@
 Future work might explore this.
+
+Visual inference has been successfully used in two practical applications: to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
+It is hoped that the lineup protocol will prove to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 

### 2013-04-12 paper_jasa.tex
finished with conclusions

--- 2013-04-11 paper_jasa.tex
@@
 
-The lineup protocol operates analogously to statistical tests that have broad alternative hypotheses.
+The lineup protocol operates similarly to statistical tests that have broad alternative hypotheses.
 If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
 Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
+In conventional testing, multiple comparisons are often done to refine and understand the test results, and perhaps some similar approaches might be developed for visual inference.
 
@@
 
-Lineups have been successfully applied to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
-It is hoped that this proves to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
+Visual inference has been successfully used in two practical applications: to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
+It is hoped that the lineup protocol will prove to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 

### 2013-04-12 paper_jasa.tex
small typos

--- 2013-04-12 paper_jasa.tex
@@
 In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g.
-\citet{cook:99}).
+\citealt{cook:99}).
 Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
@@
        \scalebox{.6}{\includegraphics{power_expected_k.pdf}}
-       \caption{Comparison of the expected power of the visual test for different $K$ (number of observers) with the power of the conventional test, for $n =100$, $\sigma = 12$ and $m$=20.  
+       \caption{Comparison of the expected power of a visual test of size $m=20$ for different $K$ (number of observers) with the power of the conventional test, for $n =100$ and $\sigma = 12$.  
        }

### 2013-04-12 paper_jasa.tex
re-worked conclusions

@@
 Specific terminology was defined, and methods for obtaining the $p$-value and estimating the power of visual tests were introduced.
-It was assumed that observers select the plot having the strongest signal in the lineup, in order to calculate the theoretical power,  and the experimental data proved to be consistent with this assumption.
-The simulation study showed that  results obtained by visual inference are comparable to those obtained by conventional inference, in the setting where %conventional testing can be done.
+In order to calculate the theoretical power, it was assumed that observers will select the plot having the strongest signal in the lineup, and the experimental data suggests that for most observers this assumption holds.
+Results from visual inference in the controlled setting of the simulation study are comparable to those obtained by conventional inference.
+%, in the setting where %conventional testing can be done.
 Visual inference is intended to provide valid tests where 

### 2013-04-16 paper_jasa.tex
supplementary materials and reference updated

--- 2013-04-12 paper_jasa.tex
@@
 Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-The lineup protocol enables statistical significance testing of visual findings, bridging the  gulf between exploratory and inferential statistics.
-In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests in scenarios when a conventional test exists.
+A new method, called the lineup protocol, enables statistical significance testing of visual findings, which bridges a big gulf between exploratory and inferential statistics.
+In this paper inferential methods for statistical graphics are developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists.
 This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
@@
 If it is extreme on this scale then the null hypothesis is rejected.
-In contrast in visual inference, the plot of the data is compared with a set of plots of  samples drawn from the null distribution.
+In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
 If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.
@@
 Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
-%It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
+It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
 
@@
 
-\begin{dfn}\label{dfn:visualtest}
+\begin{dfn}\label{dfn:test}
 The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
@@
 Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
-Let $p_i$ be the probability that individual~$i$ picks out the actual data plot.
+Let $p_i$ be the probability that individual $i$ picks out the actual data plot.
 If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot.
@@
 
-\begin{proof} The proof and the details of the lemma are attached in the supplementary documents. 
+\begin{proof} The proof and the details of the lemma are attached to the supplementary documents. 
 \end{proof}
@@
 %These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot residuals obtained from fitting the model without $X_j$ against $X_j$ or for a single covariate we may plot $Y$ against $X_j$ (cases 1-4 in Table~\ref{tbl:stat_multiple}).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5 in Table~\ref{tbl:stat_multiple}).
-For the purpose of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable, respectively.
+For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).
+To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
+For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.
 
@@
 1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid.
-For null plots   we simulate data from fitted null model.
+For lineup plot, we simulate data from fitted null model.
 \\
 2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots.
-For null plots   we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-3 & $H_0: \beta_k=0$ (for binary $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
-For null plots   we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
+3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
+For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
-4 & $H_0: \beta_k=0$ (interaction of continuous and  binary $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
-For null plots   we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
+4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
+For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
 5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
-For null plots   we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
+For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
 \\
 6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
-For null plots   we simulate data from standard normal.
+For lineup plot, we simulate data from standard normal.
 \\
 7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
-For null plots   we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
+For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
 \\
 8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data.
-For null plots   we simulate data from fitted model.
+For lineup plot, we simulate data from fitted model.
 \\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For null plots   we simulate data with correlation $\rho$.\\
+9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
 \hline 
@@
 Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test for different $K$ (number of observers), obtained using $p(\beta)$ from Equation \ref{power_exp}.
-Notice that the expected power of the visual test exceeds the power of the conventional test as $K$  increases, and when $\beta$ gets larger.
+Notice that the expected power of the visual test exceeds the power of the conventional test as $K$ gets larger, and when $\beta$ gets larger.
 Conversely, visual power is below conventional power for parameter values close to the null hypothesis.
@@
 At the same time, the point of intersection between visual and conventional power approaches the value of the null hypothesis as the number of observers approaches infinity, leading to an asymptotically perfect power curve of zero in the null hypothesis and one for any alternative value.
-We observe this dichotomy of visual power  in the data collected from user experiments, too.
-It features prominently in  figure \ref{fig:power_loess_effect}.
+This dichotomy of visual power can also be seen in more detail in figure \ref{fig:power_loess_effect}.
 
@@
 Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression setting.
-The first two experiments have  ideal scenarios for conventional testing, where we would not expect the lineup protocol to do better than the conventional test.
+The first two experiments have conventional ideal scenarios where we would not expect the lineup protocol to do better than the conventional test.
 The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
@@
 
-\subsection{Data Cleaning}
+\subsection{Data Cleaning}  \label{sec:data_cleaning}
 
@@
 
-\subsection{How much do null plots affect the choice?}
+\subsection{How much do null plots affect the choice?}\label{sec:null_choice}
 
@@
 
-\subsection{Type III error}~\label{sec:TypeIII}
+\subsection{Type III error}~\label{sec:TypeIII}\label{sec:typeIII_error}
 
@@
 
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009}.
-Specific terminologies were defined, and methods for obtaining the $p$-value and estimating the power of visual tests were introduced.
-It was assumed that observers select the plot having the strongest signal in the lineup, in order to calculate the theoretical power,  and the experimental data was consistent with this assumption.
-The simulation study showed that the results obtained by visual inference are comparable to those obtained by conventional inference, in the setting where conventional testing can be done.
-Visual inference is intended to provide valid tests where no conventional test exists, and our experiments in a controlled scenario suggest that it will perform as expected in the intended applications.
-The power of a visual test increases with the number of observers, which interestingly, leads to a result that the theoretical power of visual test can be better than that of conventional test.
-
-The lineup protocol operates similarly to statistical tests that have broad alternative hypotheses.
+This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
+For this, specific terminologies were defined.
+The methods of obtaining $p$-value and estimating power of visual tests were introduced.
+If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
+Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
+The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
+The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
+Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure enough power.
+
+When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
+Lineups are successfully used to evaluate the power of competing graphical designs in \cite{heike:2012}.
+This method is also applied in \citet{niladri:2012}.
+It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
+The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
 If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
 Follow-up questions on the reasons provide qualitative insight.
-In conventional testing, multiple comparisons are often done to refine and understand the test results, and perhaps some similar approaches might be developed for visual inference.
+Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
 
@@
 Future work might explore this.
-
-Visual inference has been successfully used in two practical applications: to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
-It is hoped that the lineup protocol will prove to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
 
@@
 %\bibliographystyle{plainnat}
+%\bibliographystyle{ieeetr}
 \bibliographystyle{asa}
-%\bibliographystyle{ieeetr}
 \bibliography{references}

### 2013-04-16 supplementary.tex
supplementary materials and reference updated

@@
 %\usepackage{hyperref} % for hyper reference
+
+% for external reference
+\usepackage{xr}
+\externaldocument[p-]{paper_jasa}
 
@@
 
+\title{Validation of Visual Statistical Inference, Applied to Linear Models (supplementary materials)}
+%\author{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
+\date{\vspace{-.7in}}
+
 \begin{document}
+
+\maketitle
+
+\vspace{1.5cm}
+
+The materials in this document supplement the information presented in the manuscript ``Validation of Visual Statistical Inference, Applied to Linear Models".
+Section \ref{sec:lineup_selection} describes how the lineups are presented to the subjects for evaluation, section \ref{sec:proof_lemma} presents the proof of the Lemma \ref{p-lemma} in the manuscript.
+The detailed discussion on how much the choice of null plots affect is included in Section \ref{sec:null_choice} which supplement the information in Section \ref{p-sec:null_choice} of the manuscript.
+Section \ref{sec:typeIII_error} includes the detailed discussion about the Type-III error supplementing Section  \ref{p-sec:typeIII_error}.
+The data cleaning process and how that may affect the result are discussed in Section \ref{sec:data_cleaning} which supplements Section \ref{p-sec:data_cleaning} of the manuscript.
+
+\section{Selection of Lineups for each subject} \label{sec:lineup_selection}
+Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript .
+The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
+The number of lineups randomly chosen for each difficulty level is shown in the table as well.
+Note that every subject saw at least one easy lineup in the pool of 10 lineups shown.
+Some subjects saw more than one easy lineup that comes from the mixed group.
+The number of total evaluations required for each lineup is also shown in the table and a total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
 
@@
 %\renewcommand\thetable{}
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section 5.1 of the paper.} 
+\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{p-sec:category} of the paper.} 
 \centering
@@
 
-\section{Proof of the Lemma}
+\section{Proof of the Lemma} \label{sec:proof_lemma}
+
+The proof of the Lemma \ref{p-lemma} in the manuscript is shown below;
 
@@
 
-\section{How much do null plots affect the choice?}
+\section{How much do null plots affect the choice?} \label{sec:null_choice}
 
@@
 
-We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
+We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{p-fig:P-val_log2} in the mauscript).
 What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
@@
 Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figure \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
+Figure \ref{p-fig:P-val_log2} in the manuscript shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
 When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
@@
 
-\section{Type III error}~\label{sec:TypeIII}
+\section{Type III error}~\label{sec:TypeIII}\label{sec:typeIII_error}
 
@@
 
-Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Figure \ref{p-fig:power_loess_effect} in the manuscript indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
@@
 
-\section{Data Cleaning}
+\section{Data Cleaning}\label{sec:data_cleaning}
 
@@
        \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
-       \caption{The subject specific power estimated from mixed model for all the six screening criteria are shown. The corresponding power curves for conventional test is shown for comparison.}
+       \caption{The subject specific power ($K=1$) estimated from mixed model for all the six screening criteria are shown.
+The corresponding power curves for conventional test is shown for comparison.}
        \label{fig:power_screening_subject}
@@
 
+\bibliographystyle{asa}
+\bibliography{references}
+
 \end{document}

### 2013-04-16 paper_jasa.tex
image legend and supplementary reference updated

--- 2013-04-12 paper_jasa.tex
@@
        \scalebox{0.55}{\includegraphics{power_loess_effect_k.pdf}}
-       \caption{\hh{Make the legend for the lines longer - just as in fig, 3} Power in comparison to effect for the three experiments.
+       \caption{Power in comparison to effect for the three experiments.
 Points indicate subject responses, with size indicating count.
@@
 
-\hh{Figure \ref{fig:P-val_log2} gives an overview of all selections in all lineups of experiments 1 and 2.
+Figure \ref{fig:P-val_log2} gives an overview of all selections in all lineups of experiments 1 and 2.
 Each panel of the figure corresponds to a single lineup.
@@
 The $x$-location of the pin shows the plot's $p$-value on a log scale, its height  is given by the number of observer choosing this plot.
-}
 %Figure \ref{fig:P-val_log2} shows the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.

### 2013-04-16 supplementary.tex
image legend and supplementary reference updated

@@
 
-Figure \ref{p-fig:power_loess_effect} in the manuscript indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
+Figure \ref{p-fig:power_loess_effect} in the manuscript indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
 Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.
+Observers were provided with four possible reasons to use for their choice:
+
+\begin{enumerate} \itemsep 0in
+\item Most different plot
+\item Visible trend
+\item Clustering visible
+\item Other
+\end{enumerate}
+
+\noindent with the possibility to use more than one.
 The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.

### 2013-04-16 R/paper_jasa.R

@@ -1269,7 +1269,8 @@ ggplot()+
   geom_point(aes(effect,pow), data=effect_hat.dat, shape=4) +
   facet_grid(.~experiment, scales="free") + 
   scale_colour_gradient("Test",limits=c(10,30), guide="none") + 
-  ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses")
+  ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses") +
+  theme(legend.key.width = unit(1.5, "cm"))
 
 ggsave(filename = "../images/power_loess_effect_k.pdf", height = 4.5,width = 12)
 

### 2013-04-16 supplementary.tex
screening power curves removed from supplementary

@@
 
-Following six approaches have been examined to clean the data.
-Table \ref{tbl:summary} displays number of subjects and there total evaluations in each screening criteria.
-For our paper we picked criteria six.
+Following six approaches have been examined while cleaning the data.
+Table \ref{tbl:summary} displays number of subjects and their total evaluations in each screening criteria.
+The criteria 1 includes all the participants hence showing number of total subjects participated in the survey for all the experiments.
+For our paper we used criteria six to clean the data and the results presented in the paper are based on the cleaned data.
 
@@
 
-The power curves obtained by fitting mixed model to the data for all the experiments applying different screening criteria are shown in Figure \ref{fig:power_screening_subject}.
+%The power curves obtained by fitting mixed model to the data for all the experiments applying different screening criteria are shown in Figure \ref{fig:power_screening_subject}.
 This shows how the screening criteria may affect the results.
@@
 Notice that no mater what criteria we apply the result does not change much for experiment 2.
-% Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
 
@@
 
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
-       \caption{The subject specific power ($K=1$) estimated from mixed model for all the six screening criteria are shown.
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
+%       \caption{The subject specific power ($K=1$) estimated from mixed model for all the six screening criteria are shown.
 The corresponding power curves for conventional test is shown for comparison.}
-       \label{fig:power_screening_subject}
-\end{figure}
+%       \label{fig:power_screening_subject}
+%\end{figure}
 

### 2013-04-16 R/paper_jasa.R

@@ -1272,7 +1272,7 @@ ggplot()+
   ylab("Power") + xlab("Effect") + scale_size_continuous("# Responses") +
   theme(legend.key.width = unit(1.5, "cm"))
 
-ggsave(filename = "../images/power_loess_effect_k.pdf", height = 4.5,width = 12)
+ggsave(filename = "../images/power_loess_effect_k.pdf", height = 4,width = 12)
 
 
 # examining dat3

### 2013-04-18 supplementary.tex
final versions

@@
 
-The materials in this document supplement the information presented in the manuscript ``Validation of Visual Statistical Inference, Applied to Linear Models".
-Section \ref{sec:lineup_selection} describes how the lineups are presented to the subjects for evaluation, section \ref{sec:proof_lemma} presents the proof of the Lemma \ref{p-lemma} in the manuscript.
+The materials in this document supplement the information presented in the manuscript ``Validation of Visual Statistical Inference, Applied to Linear Models''.
+Section \ref{sec:proof_lemma} presents the proof of the Lemma \ref{p-lemma} in the manuscript.
+Section \ref{sec:lineup_selection} describes how the lineups are presented to the subjects for evaluation.
+The data cleaning process and how that may affect the result are discussed in Section \ref{sec:data_cleaning} which supplements Section \ref{p-sec:data_cleaning} of the manuscript.
 The detailed discussion on how much the choice of null plots affect is included in Section \ref{sec:null_choice} which supplement the information in Section \ref{p-sec:null_choice} of the manuscript.
 Section \ref{sec:typeIII_error} includes the detailed discussion about the Type-III error supplementing Section  \ref{p-sec:typeIII_error}.
-The data cleaning process and how that may affect the result are discussed in Section \ref{sec:data_cleaning} which supplements Section \ref{p-sec:data_cleaning} of the manuscript.
-
-\section{Selection of Lineups for each subject} \label{sec:lineup_selection}
-Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript .
-The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
-The number of lineups randomly chosen for each difficulty level is shown in the table as well.
-Note that every subject saw at least one easy lineup in the pool of 10 lineups shown.
-Some subjects saw more than one easy lineup that comes from the mixed group.
-The number of total evaluations required for each lineup is also shown in the table and a total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
-
-\begin{table}[hbtp]
-%\renewcommand\thetable{}
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{p-sec:category} of the paper.} 
-\centering
-\begin{tabular}{c c c c  c c c}
-\hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
- \cline{2-4}
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
-\hline
-easy&100& 5&8 & 1& 12 & 1\\
-&100&12&16 &1&&\\
-&300& 5&5 &1&&\\
-&300&12&10 &1 &&\\
-\hline
-medium&100& 5&3 &203 & 9 &2\\
-&300& 5&2, 3 & 97, 1&&\\
-\hline
-hard&100&12&3, 8, 10 & 277, 126, 23& 18 &6\\
-&300& 5&1 & 371 &&\\
-&300&12&3, 5& 375, 74 &&\\
-\hline
-mixed&100& 5&1, 5, 0& 214, 2, 73 & 21 &1\\
-&100&12&1& 100& &\\
-&300& 5&0 & 73&&\\
-&300&12&7, 1& 2, 152&&\\
-\hline
-Total &&&&&60&10\\
-\hline
-\end{tabular}
-\label{tbl:dist_lineup1}
-\end{table}
 
@@
 \end{proof}
+
+\section{Selection of Lineups for each subject} \label{sec:lineup_selection}
+Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript .
+The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
+The number of lineups randomly chosen for each difficulty level is shown in the table as well.
+Note that every subject saw at least one easy lineup in the pool of 10 lineups shown.
+Some subjects saw more than one easy lineup that comes from the mixed group.
+The number of total evaluations required for each lineup is also shown in the table and a total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
+
+\begin{table}[hbtp]
+%\renewcommand\thetable{}
+\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{p-sec:category} of the paper.} 
+\centering
+\begin{tabular}{c c c c  c c c}
+\hline\hline
+Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
+ \cline{2-4}
+level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
+\hline
+easy&100& 5&8 & 1& 12 & 1\\
+&100&12&16 &1&&\\
+&300& 5&5 &1&&\\
+&300&12&10 &1 &&\\
+\hline
+medium&100& 5&3 &203 & 9 &2\\
+&300& 5&2, 3 & 97, 1&&\\
+\hline
+hard&100&12&3, 8, 10 & 277, 126, 23& 18 &6\\
+&300& 5&1 & 371 &&\\
+&300&12&3, 5& 375, 74 &&\\
+\hline
+mixed&100& 5&1, 5, 0& 214, 2, 73 & 21 &1\\
+&100&12&1& 100& &\\
+&300& 5&0 & 73&&\\
+&300&12&7, 1& 2, 152&&\\
+\hline
+Total &&&&&60&10\\
+\hline
+\end{tabular}
+\label{tbl:dist_lineup1}
+\end{table}
+
+\section{Data Cleaning}\label{sec:data_cleaning}
+
+Following six approaches have been examined while cleaning the data.
+Table \ref{tbl:summary} displays number of subjects and their total evaluations in each screening criteria.
+The criteria 1 includes all the participants hence showing number of total subjects participated in the survey for all the experiments.
+For our paper we used criteria six to clean the data and the results presented in the paper are based on the cleaned data.
+
+\begin{enumerate}
+\item {\bf Include all participants} and their evaluations.
+\item Exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
+\item Exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
+\item Include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing.
+\item Include participants' records, if at least {\bf 50\% of all very easy (p-value $<$ 0.0002) lineups are correct} .
+\item Use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
+If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
+\end{enumerate}
+
+%The power curves obtained by fitting mixed model to the data for all the experiments applying different screening criteria are shown in Figure \ref{fig:power_screening_subject}.
+This shows how the screening criteria may affect the results.
+The corresponding power curves of conventional test are also shown.
+Notice that no mater what criteria we apply the result does not change much for experiment 2.
+
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
+%       \caption{The overall power estimated from mixed model  for each screening criteria are shown. The corresponding power curve for conventional test is shown for comparison.}
+%       \label{fig:power_screening}
+%\end{figure}
+
+%\begin{figure}[hbtp]
+%   \centering
+%       \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
+%       \caption{The subject specific power ($K=1$) estimated from mixed model for all the six screening criteria are shown.
+The corresponding power curves for conventional test is shown for comparison.}
+%       \label{fig:power_screening_subject}
+%\end{figure}
+
+\begin{table}[hbtp]
+\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
+Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information. }
+\begin{center}
+\begin{tabular}{rrrrr|rrrr|rrrr}
+  \hline
+Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
+ \cline{2-5}  \cline{6-9}   \cline{10-13}
+criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
+  \hline
+1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
+  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
+  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
+  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
+  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
+  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:summary}
+\end{table}
 
@@
 
-\section{Data Cleaning}\label{sec:data_cleaning}
-
-Following six approaches have been examined while cleaning the data.
-Table \ref{tbl:summary} displays number of subjects and their total evaluations in each screening criteria.
-The criteria 1 includes all the participants hence showing number of total subjects participated in the survey for all the experiments.
-For our paper we used criteria six to clean the data and the results presented in the paper are based on the cleaned data.
-
-\begin{enumerate}
-\item {\bf Include all participants} and their evaluations.
-\item Exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-\item Exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-\item Include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing.
-\item Include participants' records, if at least {\bf 50\% of all very easy (p-value $<$ 0.0002) lineups are correct} .
-\item Use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
-If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-\end{enumerate}
-
-%The power curves obtained by fitting mixed model to the data for all the experiments applying different screening criteria are shown in Figure \ref{fig:power_screening_subject}.
-This shows how the screening criteria may affect the results.
-The corresponding power curves of conventional test are also shown.
-Notice that no mater what criteria we apply the result does not change much for experiment 2.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
-%       \caption{The overall power estimated from mixed model  for each screening criteria are shown. The corresponding power curve for conventional test is shown for comparison.}
-%       \label{fig:power_screening}
-%\end{figure}
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.70}{\includegraphics{power_screening_subject.pdf}}
-%       \caption{The subject specific power ($K=1$) estimated from mixed model for all the six screening criteria are shown.
-The corresponding power curves for conventional test is shown for comparison.}
-%       \label{fig:power_screening_subject}
-%\end{figure}
-
-\begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information. }
-\begin{center}
-\begin{tabular}{rrrrr|rrrr|rrrr}
-  \hline
-Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
- \cline{2-5}  \cline{6-9}   \cline{10-13}
-criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-  \hline
-1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:summary}
-\end{table}
-
 \bibliographystyle{asa}

### 2013-04-19 paper_jasa.tex
final versions
cleaning latex codes

@@
 
-%\begin{multicols}{2}
-%\twocolumn
-
 \section{Introduction}
-
-%\checkmark
 
@@
 
-%Statistical graphics research and development is \red{agreeably established} now, and has been cemented by s
 Several new developments in graphics research have been achieved in recent years.
@@
 The results are compared with those of the conventional test.
-
-%, analysis requires plots of data.
-For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
-Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.
-
-%The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
-In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS ease the production of high quality statistical plots.
-A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
-\cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
-\blue {Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
-
-%\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
-
-%In this paper we compare the lineup protocol to conventional statistical inference, specifically applied to the regression framework.
-We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
 
@@
 For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
-Type II error is harder to calculate, as is usually the case.
-In visual inference, individual abilities need to be incorporated to calculate Type II error.
-%This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
+Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. 
 Here, we need to estimate the probability that an observer sees the actual data plot as different, when it really is different.
@@
 Ways to estimate $p_i$ will be  discussed in the following sections.
-%one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
-
-%\green{WHY??? And how does this relate to the logistic model below?} 
-%This leads to the definition of the power of the visual test, as follows:
-
-%The lineup plot can be evaluated by one or more individuals.
-When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
-
- %Notice that when $N=1$, this $p$-value is $\frac1m$.
-
-%\subsection{Power}
-
-%Next, we will develop power theoretically and then relate it to the empirical results.
-%We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
-we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$.
-In a next step we will then relate the theoretical results to our empirical findings.
-
-%\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
-- why do we need them?
-wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?
-- $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
-%\begin{dfn} \label{dfn:power}
-%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-%    \begin{equation*}
-%      \text{Power}_V(\theta)= 
-%        \begin{cases} 
-%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
-%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-%
-%
-%\red{Alternative Power definition:}
+
 \begin{dfn} \label{dfn:power}
@@
       \text{Power}_V(\theta)= Pr(\text{Reject } H_0 \mid \theta) 
-%      = P(X \ge x_{\alpha}) =
-%        \begin{cases} 
-%               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
-%              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
-%        \end{cases}
     \end{equation*}
@@
 
-%\begin{proof}
-%%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
-%
-% By definition  
-%
-%$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-%
-%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
-%%, with non-centrality parameter $\delta$. 
-%\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
-%\begin{eqnarray}\label{dist_p}
-%F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-%  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
-%  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
-%  &=&%\left\{ \begin{array}{ll}
-%          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
-% %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%%       \end{array} \right.     
-%\end{eqnarray}
-%
-%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-%
-%
-%
-%Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots.
-Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
-The minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function
-%\[
-%F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
-%\]
-%
-%\noindent Thus
-%\begin{eqnarray*}
-%P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
-%&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
-%&=&  E\left[ (1 - p_D)^{m-1}\right].
-%\end{eqnarray*}
-%
-%
-%%\noindent This proves the statement above. 
-%%We further see:
-%%\[
-%%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
-%%\]
-%
-%
-%% If the density of $p$-value is very right skewed, the expectation term would be large.
-The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in.
-Should we add that in Equation \ref{dist_p}?
-)},
-%%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
-%%}
-%%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$. 
-%
-%%
-%%
-%%
-%%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
-%%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-%%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
-%%\begin{eqnarray*}
-%%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-%%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
-%%\end{eqnarray*}
-%%which, in particular, is independent of $p_{0,i}$ for all $i$.
-%%
-%%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
-Further we will assume that all observers  have the same ability in identifying the data plot.
-%%
-%%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
-%%
-%%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
-%%\[
-%%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
-%%\]
-%\end{proof}
-
-%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the Figure on $1/m$ versus power.
-These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
-
 The above lemma allows two immediate conclusions for the use of lineups.
 The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$.
-%This may seem counterintuitive, but it makes sense.
-%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
 Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$.
@@
 It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.
-% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
-%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.
-
-%Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
-%Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
 
@@
 With increasing $p$-value the probability drops -- when it reaches $1/m$ a horizontal line is drawn to emphasize insufficient sensitivity of the test due to the lineup size.
-   %Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
-The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true.
-}
 }
@@
 
-%A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
-The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
-The test statistic, the plot containing the actual data, is randomly placed among these null plots.
-If the test statistic is identifiable the null hypothesis is rejected.
-% with a $p$-value of at most 0.05.
-
-%\subsection{Expected power}
-
-% Di: This is specific for the regression model
-%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
-%%Now consider estimating the power of the visual test. 
-%%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in conventional test be $t_{obs}$ with $p$-value $p_D$.
-By definition we have
-%%
-%%$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-%%
-%%\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
-%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
-%%\begin{eqnarray}\label{dist_p}
-%%F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-%%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
-%%  &=&\left\{ \begin{array}{ll}
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%%       \end{array} \right.     
-%%\end{eqnarray}
-%%
-%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
-Suppose  $p_0=\min_i(p_{0,i})$.
-Thus $p_0 \sim \text{Beta}(1,m-1)$.
-
 The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
@@
 
-%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
-%\begin{thm}\label{thm:power}
-% equation \eqref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
-%\end{thm}
-% 
-%{ \em proof}
-%
-%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
-%\begin{eqnarray*}
-%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
-%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
-%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
-%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
-%
-%Thus expected power in equation \eqref{power_exp} be
-%\begin{eqnarray*}
-% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
-%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
-%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
-%  &=& E(1-p_D)^{m-1}   \\
-%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%
 \begin{figure}[hbtp]
@@
 \end{figure}
-
-% \pagebreak
-
-%\subsection{Some Power Considerations}
-
-%\subsection{Expected number of choices}
-%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
-%\[
-%E[X] = (m-1)(1-E[p_D]).
-%\]
-%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-%This should allow us to infer some of the signal strength $p_D$ as
-%\begin{equation} \label{plot_signal}
-%E[p_D] = 1 - E[X]/(m-1),
-%\end{equation}
-%i.e.
-by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
-%
 
@@
 Note that no participant was shown the same lineup twice.
+
 %\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
@@
 
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
-%	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
 The number of evaluations required for each lineup to provide reasonable estimates of the proportion correct ($\hat p$) is determined by the variance of the number of correct evaluations.
@@
 Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
-% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}
 
@@
 The order in which the lineups are shown to participants is randomized.
-%Each subject saw a substantial number of difficult lineups.
 
@@
 
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
-%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
 An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.
@@
 The actual data plot location is $(3^2-2^3)$ and 13 out of 31 observers picked the actual plot.
-
-%\begin{figure}[hbt]
-%\begin{figurehere}
-%   \centering
-%       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
-%       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
-The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope.
-%}
-%       \label{fig:cont_dat}
-%\end{figure}
 
@@
 
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
-Show a sample lineup, probably the most difficult, instead of the single plot.}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_observed}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_loess1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
-%       \label{fig:power_loess2}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in Table \ref{tbl:experiment_params}.}
-%       \label{fig:power_diff_exp}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-}
-%       \label{fig:power_observed_exp3}
-%\end{figure*}
-
 \section{Results} \label{sec:results}
@@
 
-%{\bf Screening scenarios:}
 Amazon Mechanical Turk workers are paid for their efforts, not substantially, but on the scale of the minimum wage in the USA.
@@
 We examine  the data collected from the first two experiments, to see if this assumption is, indeed, reasonable.
-%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value.
-In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
 
@@
 The $x$-location of the pin shows the plot's $p$-value on a log scale, its height  is given by the number of observer choosing this plot.
-%Figure \ref{fig:P-val_log2} shows the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
-In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the number of subjects who selected this plot is displayed vertically.
-The dot represents this point, and the lines connect vertically with 0 count.
 Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.
@@
 In both experiments  people tended to select the plot with the lowest $p$-value.
-% with an increase in effect size.
- %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
 The results are clearer for experiment 2, that used a continuous covariate.
@@
 If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
-The detailed discussion is attached in the supplementary material.
-
-%We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figure \ref{fig:P-val_log2}).
-What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
-If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
-Figure \ref{fig:pval_difference} investigates this.
-The difference between the $p$-value of the actual data is compared with the lowest from the null plots.
-This is plotted horizontally, and the proportion correct is plotted vertically.
-Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
-In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
-This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
-Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figure \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
-When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
-
-%In practice, the $p$-value is not going to be a valid way to compare plots.
-Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
-This is investigated in \citet{niladri:2012}.
-Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.60}{\includegraphics{pval_difference.pdf}}
-%       \caption{Scatter plot of difference between the data plot's $p$-value and the  smallest $p$-value of the null plots vs proportion correct.
-Negative differences indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
-Difference close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
-%       \label{fig:pval_difference}
-%\end{figure}
 
@@
 
-%This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
-Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
-
-%Figure \ref{fig:power_loess_effect} indicated that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice: (1) Most different plot, (2) Visible trend, (3) Clustering visible and (4) Other, with the option to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
-Reasons 1 or 3 would be indicative of Type III error.
-Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
-If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
-This is not what we see.
-At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
-The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
-The size of the point corresponds to the number of subjects putting this as the reason.
-Subjects that chose all three reasons almost always chose the actual data plot.
-This was followed by using 1 and 3, and then 1 and 4.
-The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
-
-%At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
-``Visible trend'' comes in third.
-This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
-%       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-%       \caption{Reasons of plot choices vs  proportion of times the subjects correctly chose the actual data plot for experiment 3 that examines the occurrence of Type III error.
-At left, all subjects' choices are shown, and reason 123 means all three reasons are used.
-At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
-Size of the point corresponds to the number of subjects using that reason.}
-%       \label{fig:choice_reason}
-%\end{figure}
-
 For visual inference, making a Type III, is not actually a problem.
@@
 However, this highlights the importance of getting qualitative reasoning from observers  for their choices.
-A more detailed discussion can be found in the supplementary documents.
 
@@
 
-%One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
-A good test statistic is expected to reveal structure if the null hypothesis is not true.
-Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
-This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
-
 The performance of subjects was quite varied, but consistent.
@@
 
-%\item Explain caveats, nuances, advantages with the work
-
- %This is the part of our ongoing and future research.
-
-%The question that is posed to observers can affect the ability to evaluate lineups.
-In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
-
-% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
-If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
-This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
-
-% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
-
-%Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
-Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet, the results were good.
-We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
-
-%\item Describe the next steps of the work, and possible ideas for future extensions
-
-%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
-The direction of future research would be to study that and apply the methods in those areas.
-
-%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
-We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
-The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
-For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
-Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.
-
-%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
-Along the lines of the Army tests that Susan brought to the working group.
-\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
-Thus anyone from any background could participate the experiment.
-It would be interesting to see how well the trained subjects performs.
-Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
-One of our future work would be to examine this in details.
-For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
-After being trained they may have to pass the qualification test before participating the experiment.
-Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
-
-%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
-
-%In practice there is no correct answer.
-In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
-In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
-And subjects would provide open-ended answers for their choice of plot.
-
-%What are your suggestions for future experiments, and next steps of the research?
-
 %\paragraph{Acknowledgement:}
 %This work was funded in part by National Science Foundation grant DMS 1007697.
+
+\paragraph{Supplementary Material:} Proof of Lemma \ref{lemma}, details of data collection and cleaning, longer discussion of effect of null plots and Type III error.
 
@@
 \end{document}  %End of document.
-
-\section*{Appendix}
-
-\begin{table}[hbtp]
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
-\centering
-\begin{tabular}{c c c c  c c c}
-\hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
-\hline
-easy&100& 5&8 & 1& 12 & 1\\
-&100&12&16 &1&&\\
-&300& 5&5 &1&&\\
-&300&12&10 &1 &&\\
-\hline
-medium&100& 5&3 &203 & 9 &2\\
-&300& 5&2,3 & 97, 1&&\\
-\hline
-hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
-&300& 5&1 & 371 &&\\
-&300&12&3,5& 375,74 &&\\
-\hline
-mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
-&100&12&1& 100& &\\
-&300& 5&0 & 73&&\\
-&300&12&7,1& 2, 152&&\\
-\hline
-Total &&&&&60&10\\
-\hline
-\end{tabular}
-\label{tbl:dist_lineup1}
-\end{table}
-
-%\begin {abstract} 
-%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently, there were no formal visual methods in place for determining statistical significance of findings.
-This changed when Buja et al.
-(2009) conceptually introduced the lineup protocol for formal tests of visual findings.
-In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
-%\end {abstract}
-
-\end{document}
-
-\section{Old Conclusions}
-
-It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
-The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-
-Power of visual tests can be estimated in different ways, which was explained in this paper.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is better than that of conventional test in the conditions tested in this paper.
-
-%\item Explain caveats, nuances, advantages with the work
-
-One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
-A good test statistic is expected to reveal structure if the null hypothesis is not true.
-Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
-This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
-%This is the part of our ongoing and future research.
-
-The question that is posed to observers can affect the ability to evaluate lineups.
-In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
-
-% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
-If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
-This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
-
-% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
-
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
-
-Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
-Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet, the results were good.
-We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
-
-The performance of subjects was quite varied, but consistent.
-No restrictions were placed on Turk workers, in terms of abilities.
-There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test.
- It would be interesting to see how well trained subjects might perform.
-Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference.
-Future work might explore this.
-
-%\item Describe the next steps of the work, and possible ideas for future extensions
-
-When there is no conventional test available, visual testing procedure can be used due to its few assumptions and non-parametric nature.
-We have successfully used the procedure informally for local projects.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-If the null hypothesis is rejected, generally we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
-Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
-
-%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
-The direction of future research would be to study that and apply the methods in those areas.
-
-%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
-We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
-The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
-For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
-Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.
-
-%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
-Along the lines of the Army tests that Susan brought to the working group.
-\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
-Thus anyone from any background could participate the experiment.
-It would be interesting to see how well the trained subjects performs.
-Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
-One of our future work would be to examine this in details.
-For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
-After being trained they may have to pass the qualification test before participating the experiment.
-Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
-
-%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
-
-%In practice there is no correct answer.
-In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
-In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
-And subjects would provide open-ended answers for their choice of plot.
-
-%What are your suggestions for future experiments, and next steps of the research?

### 2013-04-23 supplementary.tex
supplementary section 6.7 and 6.8 cleaned

@@
 The detailed discussion on how much the choice of null plots affect is included in Section \ref{sec:null_choice} which supplement the information in Section \ref{p-sec:null_choice} of the manuscript.
-Section \ref{sec:typeIII_error} includes the detailed discussion about the Type-III error supplementing Section  \ref{p-sec:typeIII_error}.
+Section \ref{sec:typeIII_error} includes the detailed discussion about the Type-III error supplementing Section  \ref{p-sec:typeIII_error} of the manuscript.
 
@@
 \section{Selection of Lineups for each subject} \label{sec:lineup_selection}
-Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript .
+Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript.
 The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
@@
 Some subjects saw more than one easy lineup that comes from the mixed group.
-The number of total evaluations required for each lineup is also shown in the table and a total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
+The number of total evaluations required for each lineup is also shown in the table.
+For example,  a lineup with sample size 100, standard error 5 and slope parameter 3 requires 203 evaluations so that the proportion correct can be estimated with margin of error 0.05 as described in the manuscript.
+A total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
+We see in Table \ref{tbl:summary} that the total number of subjects actually participated in experiment 1 is 520 which is much higher than 300.
+The goal was to get total number of subject close to 300.
 
@@
 The criteria 1 includes all the participants hence showing number of total subjects participated in the survey for all the experiments.
-For our paper we used criteria six to clean the data and the results presented in the paper are based on the cleaned data.
+As per the experimental design we always show at least one easy lineup and that lineup is used to process the payment of the subject.
+Thus, for our paper we used criteria six to clean the data and the results presented in the paper are based on the cleaned data.
 
@@
 
- Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
- But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
-How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
-If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
-Measuring the difference, generally, between  plots is almost impossible.
-However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
+It is discussed in the Section \ref{p-sec:null_choice} of the manuscript that $p$-values can be used to quantify the similarity of the visual pattern in the plots used for the simulation experiments.
+Based on this, we explore more details on how much the null plots affect the choice made by the subjects.
 
@@
 
-A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
+%A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
 This occurs when the null hypothesis is correctly rejected but for the wrong reason.
@@
 Figure \ref{p-fig:power_loess_effect} in the manuscript indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice:
+Teasing this out of the results is possible by looking at the reasons participants gave for their choices.
+Participants were provided with four possible reasons to use for their choice:
 
@@
 
-For visual inference, making a Type III, is not actually a problem.
+%For visual inference, making a Type III, is not actually a problem.
 It is only a possibility in this experiment because we are working with known structure.

### 2013-04-23 supplementary.tex
cleaning supplementary

@@
 Section \ref{sec:lineup_selection} describes how the lineups are presented to the subjects for evaluation.
-The data cleaning process and how that may affect the result are discussed in Section \ref{sec:data_cleaning} which supplements Section \ref{p-sec:data_cleaning} of the manuscript.
-The detailed discussion on how much the choice of null plots affect is included in Section \ref{sec:null_choice} which supplement the information in Section \ref{p-sec:null_choice} of the manuscript.
-Section \ref{sec:typeIII_error} includes the detailed discussion about the Type-III error supplementing Section  \ref{p-sec:typeIII_error} of the manuscript.
+The data cleaning process is described in Section \ref{sec:data_cleaning}, supplementing Section \ref{p-sec:data_cleaning} of the manuscript.
+A detailed discussion on how much the sample of null plots might affect the observer's choice is in Section \ref{sec:null_choice}, supplementing a summary given in Section \ref{p-sec:null_choice} of the manuscript.
+Section \ref{sec:typeIII_error} contains more discussion about Type-III error, and supplements Section  \ref{p-sec:typeIII_error} of the manuscript.
 
@@
 \section{Selection of Lineups for each subject} \label{sec:lineup_selection}
-Table \ref{tbl:dist_lineup1} shows how the lineups are presented to the subjects for evaluation for the simulation experiment with categorical covariate described in Section \ref{p-sec:category} of the manuscript.
+Table \ref{tbl:dist_lineup1} shows the selection process of the lineups for the subjects, across the experimental design parameters of experiment 1 (Section \ref{p-sec:category} of the manuscript), as required to obtain a margin of error of 0.05.
 The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
-The number of lineups randomly chosen for each difficulty level is shown in the table as well.
-Note that every subject saw at least one easy lineup in the pool of 10 lineups shown.
-Some subjects saw more than one easy lineup that comes from the mixed group.
-The number of total evaluations required for each lineup is also shown in the table.
-For example,  a lineup with sample size 100, standard error 5 and slope parameter 3 requires 203 evaluations so that the proportion correct can be estimated with margin of error 0.05 as described in the manuscript.
-A total of 300 subjects would provide a total of 3000 evaluations that should roughly satisfy that plan.
-We see in Table \ref{tbl:summary} that the total number of subjects actually participated in experiment 1 is 520 which is much higher than 300.
-The goal was to get total number of subject close to 300.
+The number of evaluations, along with the number of lineups, and the number of lineups from each category that a single subject would get, are shown in the table.
+Note that, every subject saw at least one easy lineup in the pool of 10 lineups they were given, with some subjects  seeing two easy lineups if they get one from the mixed group.
+These ideal sample sizes were generated using a goal of obtaining a margin of error no bigger than 0.05.
+For example,  a lineup with sample size = 100, standard error  = 5 and slope parameter = 3 requires 203 evaluations so that the proportion correct can be estimated with margin of error 0.05 as described in the manuscript.
+A total of 300 subjects would provide a total of 3000 evaluations with this plan.
+Table \ref{tbl:summary} shows the number of subjects actually participating in experiment 1 is 520 which is much higher than 300.
 
@@
 \hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
+Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & Number of lineups\\
  \cline{2-4}

### 2013-04-26 supplementary.tex
data cleaning section in supplementary updated

@@
 A total of 300 subjects would provide a total of 3000 evaluations with this plan.
-Table \ref{tbl:summary} shows the number of subjects actually participating in experiment 1 is 520 which is much higher than 300.
+Table \ref{tbl:summary} shows the number of subjects actually participating in experiment 1 is 424 which is much higher than 300.
+The goal was to get final accepted data for 300 subjects.
 
@@
 
-Following six approaches have been examined while cleaning the data.
+As per the experimental design each subject is shown at least one easy lineup.
+The subject was informed that an easy lineup would be used to accept their evaluations.
+Thus once we have all the evaluations we cleaned the data based on the easy lineups shown.
+The lineup having p-value $<$ 0.0002 is considered an easy lineup.
+We sample one easy lineup from a subject's records.
+If that lineup is evaluated {\bf correctly}, we include all (other) lineups of that subject, otherwise we exclude all lineup evaluations by this participant.
+Table \ref{tbl:summary} displays number of subjects and their total evaluations before and after cleaning the data.
+
+%Following six approaches have been examined while cleaning the data.
 Table \ref{tbl:summary} displays number of subjects and their total evaluations in each screening criteria.
@@
 
-\begin{enumerate}
-\item {\bf Include all participants} and their evaluations.
-\item Exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-\item Exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-\item Include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing.
-\item Include participants' records, if at least {\bf 50\% of all very easy (p-value $<$ 0.0002) lineups are correct} .
-\item Use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
+%\begin{enumerate}
+%\item {\bf Include all participants} and their evaluations.
+%\item Exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
+%\item Exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
+%\item Include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing.
+%\item Include participants' records, if at least {\bf 50\% of all very easy (p-value $<$ 0.0002) lineups are correct} .
+%\item Use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
 If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-\end{enumerate}
+%\end{enumerate}
 
@@
   \hline
-Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
+Data &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
  \cline{2-5}  \cline{6-9}   \cline{10-13}
-criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
+cleaning & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
   \hline
-1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
+before & 424 & 226 & 180 & 4516 & 386 & 199 & 182 & 4330 & 242 & 158 &  79 & 2565 \\ 
+after  & 239 & 121 & 107 & 2249 & 351 & 185 & 164 & 3636 & 155 & 103 &  52 & 1511 \\ 
    \hline
@@
 \end{table}
+
+After cleaning the data we did not have 300 subjects for experiment 1 that we planned for.
+But we did not go for collecting more data as the proportion correct could be estimated with margin of error not much higher than 0.05 even with 239 subjects.
+We can see this from the bootstrap confidence band in Figure \ref{p-fig:power_loess_effect} in the manuscript.
+For experiment 2 and 3 we also have sufficient data even after cleaning.
 

### 2013-04-26 R/paper_jasa.R

@@ -16,12 +16,47 @@ raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
 
 source("calculate_ump_power.R") # functions to compute power
 
+
+# Getting actual raw data where 10 lineups were evaluated as per plan or
+# at least evaluated an easy lineup and got it correct.
+# The rest of the data are not considered since not completed as per 
+# experimental design.
+
+if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
+
+get_actual_raw_data <- function(raw_dat){
+  d <- ddply(raw_dat,.(id),summarise,
+             total_response = length(response),
+             percent_correct = mean(response)*100,
+             easy_correct = ifelse(experiment[1]=='turk3',
+                                   response[difficulty==0][1],
+                                   response[p_value < 0.0002][1])
+  )
+  included_id <- d$id[((d$total_response > 9) | (d$easy_correct))]
+  indx <- raw_dat$id %in% included_id
+  actual_raw_dat <- subset(raw_dat,indx)
+  indx_dup <- with(actual_raw_dat, !duplicated(data.frame(id,pic_id))) #duplication index
+  # now returning actual raw data removing duplication
+  return(subset(actual_raw_dat,indx_dup))
+}
+
+actual_raw_dat1 <- get_actual_raw_data(raw.dat1)
+actual_raw_dat2 <- get_actual_raw_data(raw.dat2)
+actual_raw_dat3 <- get_actual_raw_data(raw.dat3)
+
+get_data_info <- function(dat){
+  d <- ddply(dat, .(gender), summarize, subject=length(unique(id)))
+  return(c(sum(d$subject), d$subject[1:2], nrow(dat)))
+}
+
+get_data_info(actual_raw_dat1)
+
 # cleaning the data based on criteria 6 and removing the duplicated data
 # (id,pic_id) should be unique since no person gets a lineup more than once.
 clean_data <- function(raw_dat){
   easy_dat <- subset(raw_dat, p_value < 0.0002)
   if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
-  d <- ddply(subset(easy_dat),.(id),summarise,
+  d <- ddply(easy_dat,.(id),summarise,
              easy_cnt = length(response),
              percent_correct = mean(response)*100,
              include_id = response[1],
@@ -47,10 +82,9 @@ dat3 <- clean_data(raw.dat3)[,-indx2]
 
 
[…]

### 2013-04-26 supplementary.tex
updated sections 2,3 of supplementary

@@
 Table \ref{tbl:dist_lineup1} shows the selection process of the lineups for the subjects, across the experimental design parameters of experiment 1 (Section \ref{p-sec:category} of the manuscript), as required to obtain a margin of error of 0.05.
-The lineups are divided into four groups such as easy, medium, hard and mixed based on the parameter combinations shown in the table.
+The lineups are divided into four groups -- easy, medium, hard and mixed -- based on the parameter combinations shown in the table.
 The number of evaluations, along with the number of lineups, and the number of lineups from each category that a single subject would get, are shown in the table.
-Note that, every subject saw at least one easy lineup in the pool of 10 lineups they were given, with some subjects  seeing two easy lineups if they get one from the mixed group.
+Note that, every subject saw a block of 10 lineups, selected across these groups, including at least 1 easy lineup, and possibly 2 if one was drawn from the mixed group.
 These ideal sample sizes were generated using a goal of obtaining a margin of error no bigger than 0.05.
-For example,  a lineup with sample size = 100, standard error  = 5 and slope parameter = 3 requires 203 evaluations so that the proportion correct can be estimated with margin of error 0.05 as described in the manuscript.
+For example,  a lineup with sample size = 100, standard error  = 5 and slope parameter = 3 requires 203 evaluations so that the proportion correct can be estimated with margin of error of 0.05 following the procedures described in the manuscript.
 A total of 300 subjects would provide a total of 3000 evaluations with this plan.
-Table \ref{tbl:summary} shows the number of subjects actually participating in experiment 1 is 424 which is much higher than 300.
-The goal was to get final accepted data for 300 subjects.
+Table \ref{tbl:summary} shows the number of subjects actually participating in experiment 1 is 424 which is much higher than 300, but the number after cleaning was 239.
 
@@
 %\renewcommand\thetable{}
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{p-sec:category} of the paper.} 
-\centering
+\caption{Ideal numbers for different experimental design parameters for exaperiment 1 (Section \ref{p-sec:category} of manuscript) in order to obtain a margin of error of 0.05.
+These numbers are used to choose a sample of 10 lineups for each subject.
+}
+\begin{center}
 \begin{tabular}{c c c c  c c c}
@@
  \cline{2-4}
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
+level & $n$ & $\sigma$ & $\beta$ &required~($n_{\gamma}$) & of lineups & randomly shown \\
 \hline
@@
 \end{tabular}
+\end{center}
 \label{tbl:dist_lineup1}
@@
 
-As per the experimental design each subject is shown at least one easy lineup.
+Amazon Turk is a relatively new source of subjects for experiments.
+The workers (``turkers'') are paid, minimal amounts for their efforts, on par with conventional human subject experiments.
+Most turkers make an effort to complete tasks as requested, but some turkers do nottake the task seriously.
+Our procedure for ensuring that reliable data was available for analysis was to provide one very easy lineup, one in which the observed data plot stands out as being very different from the null plots.
 The subject was informed that an easy lineup would be used to accept their evaluations.
-Thus once we have all the evaluations we cleaned the data based on the easy lineups shown.
-The lineup having p-value $<$ 0.0002 is considered an easy lineup.
+%Thus once we have all the evaluations we cleaned the data based on the easy lineups shown.
+%The lineup having p-value $<$ 0.0002 is considered an easy lineup.
 We sample one easy lineup from a subject's records.
@@
 \begin{table}[hbtp]
-\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information. }
+\caption{Number of unique subjects and their total feedbacks before and after data cleaning. 
+Note that the number of male and female participants may not add up to the number of subjects, due to some participants declining to provide demographic information. }
 \begin{center}
@@
 
-After cleaning the data we did not have 300 subjects for experiment 1 that we planned for.
-But we did not go for collecting more data as the proportion correct could be estimated with margin of error not much higher than 0.05 even with 239 subjects.
-We can see this from the bootstrap confidence band in Figure \ref{p-fig:power_loess_effect} in the manuscript.
-For experiment 2 and 3 we also have sufficient data even after cleaning.
+After cleaning the data, for experiment 1, we did not have 300 subjects that we planned for.
+It was decided that more data was not needed, though, because the estimated with margin of error  with the 239 subjects was close to 0.05.
+This can be seen from the bootstrap confidence band in Figure \ref{p-fig:power_loess_effect} of the manuscript.
+With experiments 2 and 3 there was sufficient data even after cleaning.
+Part of the success with the later experiments comes from the researchers developing a reputation on Amazon for providing a good task and reliable payment, which means the reliable turkers look specifically for these tasks.
 

### 2013-04-29 submission3_after_review/paper_jasa_blinded.tex
submission after review

--- 2013-04-19 paper_jasa.tex
@@
 Table~\ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-A summary of all the data along with cleaned data are added to the supplementary material.
+
+% A summary of all the data along with cleaned data are added to the supplementary material.
 

### 2013-04-29 submission3_after_review/R/paper_jasa.R

@@ -1,0 +1,1338 @@
+# The complete R code for JASA Paper
+
+library(ggplot2)
+library(plyr)
+library(reshape)
+library(xtable)
+library(lme4)
+library(stringr)
+library(grid)
+
+# getting the data and some common functions
+
+raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
+raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
+raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
+
+source("calculate_ump_power.R") # functions to compute power
+
+
+# Getting actual raw data where 10 lineups were evaluated as per plan or
+# at least evaluated an easy lineup and got it correct.
+# The rest of the data are not considered since not completed as per 
+# experimental design.
+
+if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
+
+get_actual_raw_data <- function(raw_dat){
+  d <- ddply(raw_dat,.(id),summarise,
+             total_response = length(response),
+             percent_correct = mean(response)*100,
+             easy_correct = ifelse(experiment[1]=='turk3',
+                                   response[difficulty==0][1],
+                                   response[p_value < 0.0002][1])
+  )
+  included_id <- d$id[((d$total_response > 9) | (d$easy_correct))]
+  indx <- raw_dat$id %in% included_id
+  actual_raw_dat <- subset(raw_dat,indx)
+  indx_dup <- with(actual_raw_dat, !duplicated(data.frame(id,pic_id))) #duplication index
+  # now returning actual raw data removing duplication
+  return(subset(actual_raw_dat,indx_dup))
+}
+
+actual_raw_dat1 <- get_actual_raw_data(raw.dat1)
+actual_raw_dat2 <- get_actual_raw_data(raw.dat2)
+actual_raw_dat3 <- get_actual_raw_data(raw.dat3)
+
+get_data_info <- function(dat){
+  d <- ddply(dat, .(gender), summarize, subject=length(unique(id)))
[…]

### 2013-05-19 paper_jasa.tex
large p small n paper reference updated

@@
-%\documentclass{article}
 \documentclass[12pt]{article}
 \usepackage{times}
-%\usepackage{natbib}
-%\usepackage{multicol}
+
 \RequirePackage{natbib}
-\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
-%\usepackage[dvips]{graphics}
-
-%\usepackage{hyperref} % for hyper reference
+\usepackage{amsmath, amssymb, amsthm, array, graphicx,asa}
 
 \graphicspath{{images/}}
-
-\usepackage{color}
-\newcommand{\blue}[1]{{\color{blue} #1}} %MM
-\newcommand{\red}[1]{{\color{red} #1}} 
-\newcommand{\green}[1]{{\color{green} #1}} %DC
-
-\definecolor{orange}{rgb}{1,0.5,0}
-\newcommand{\hh}[1]{{\color{orange} #1}} %HH
-
-% \usepackage{pifont} % this package is used to print check mark \checkmark
-% \linespread{1.6} % factor 1.6 = double space
 
@@
 \newtheorem{lemma}[thm]{Lemma}
-
-%\topmargin -0.10in   % when making pdf
-%\textheight 9.15in  % when making pdf
 
@@
 \begin{figure}[htp]
-%\begin{figurehere}
    \centering
-%       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
        \includegraphics[width=0.95\textwidth]{plot_turk1_300_10_12_1.pdf}
@@
 \end{figure}
-%\end{figurehere}
 
@@
 Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data.
-%The last section describes the next steps in this research.
 
@@
 \begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
-$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline %\cline{1-3}\cline{4-6}\cline{7-9}\cline{10-12}\cline{17-19}
+$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline 
 1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+\begin{figure}[htbp]  
    \centering
@@
 Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
-%These plots should be familiar, because they are all in common use. 
 For example, to examine the effect of variable $X_j$ on $Y$, we would plot residuals obtained from fitting the model without $X_j$ against $X_j$ or for a single covariate we may plot $Y$ against $X_j$ (cases 1-4 in Table~\ref{tbl:stat_multiple}).
@@
 For the purpose of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable, respectively.
-
-%Note that some charts can be associated with several different testing situations. In our paper, we pick the examples for situation 2 and 3 as candidates.
 
@@
 \begin{equation}\label{power_exp} 
-%   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
  p(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
@@
 
-%\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
-This happens when subject accidentally use browser's ``back'' button and same data is saved twice.
-We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject.
-For this we only keep the very first response.}
-
 \subsection{Discrete covariate}\label{sec:category}
@@
 Each subject viewed at least 10 lineups with the option to evaluate more.
-%and we allowed them to provide more evaluations if they wanted. 
 Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
@@
 \begin{table}[hbtp]
-\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for the simulation experiments.} % title name of the table
+\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for the simulation experiments.} 
 \centering
@@
 \begin{figure}[p]
-%\begin{figurehere}
    \centering
-%       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
        \includegraphics[width=0.95\textwidth]{plot_turk2_100_25_5_3.pdf}
@@
 Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
-%Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
 A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
@@
 \begin{figure}[p]
-%\begin{figurehere}
    \centering
-%       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
        \includegraphics[width=0.95\textwidth]{plot_turk3_100_40_5_5.png}
@@
 
-% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants.
-It is not clear why they did not provide the information.
-Now the concern arises whether we should keep their responses in the study.
-It is also possible that participants were not serious enough to provide feedback.
-%For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
-This could be a criteria to determine their attentiveness in the study.
-For this we consider six different data cleaning criteria as shown below;
-
 Amazon Mechanical Turk workers are paid for their efforts, not substantially, but on the scale of the minimum wage in the USA.
@@
 Table~\ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-A summary of all the data along with cleaned data are added to the supplementary material.
 
@@
 Results from visual inference in the controlled setting of the simulation study are comparable to those obtained by conventional inference.
-%, in the setting where %conventional testing can be done.
-Visual inference is intended to provide valid tests where 
-no conventional test exists, and our experiments in a controlled scenario suggest that it will perform as expected in the intended applications.
+Visual inference is intended to provide valid tests where no conventional test exists, and our experiments in a controlled scenario suggest that it will perform as expected in the intended applications.
 The power of a visual test increases with the number of observers, which interestingly, leads to a result that the theoretical power of visual test can be better than that of conventional tests.
@@
 
-Visual inference has been successfully used in two practical applications: to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2012}.
+Visual inference has been successfully used in two practical applications: to evaluate the power of competing graphical designs \citep{heike:2012}, and to detecting signal presence in large $p$, small $n$ data \citep{niladri:2011}.
 It is hoped that the lineup protocol will prove to be  valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-
-%\paragraph{Acknowledgement:}
-%This work was funded in part by National Science Foundation grant DMS 1007697.
 
@@
 
-%\bibliographystyle{plain}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{ieeetr}
 \bibliographystyle{asa}
@@
 
-%\end{multicols}
-
-\end{document}  %End of document.
+\end{document}
